---
title: "Data Curation: HiCSC 2025"
format:
  html:
    toc: true
    eval: false #default to not evaluating the chunks, override this the template is completed
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - HiCSC2025.bib #contains citation that provide background on the project
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  - id: ktoddbrown #use the github name as an id
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role: 
      - curation: lead #review
  - id: semantic-nomad #use the github name as an id
    name:
      given: Brandon
      family: Whitehead
    orcid: 0000-0002-0337-8610
    affiliation:
      - ref: uf-ees
    role: 
      - curation: lead #review

affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
timelog:
  - activity: 
      description: Drafted initial template
      who: semantic-nomad #github username
      #when you started and ended, can be just a year or down to the minute
      end: 2026-01-20T0123
      #start: YYYY-MM-DD THH:MM
      #alternatively duration can be combined with a being/end
      #duration: P1H15M
---


```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false


library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# Bibliograph files
refernceCitation.file <- 'HiCSC2025.bib'

# Data sets
rescueDirectory <- '../../01_DataRescue'
harmonizeDirectory <- '../../02_DataHarmonization'

dataID.ls <- list(Ares2001 = 'Ares2001')

#semantic files for level 1
of_variable.file <- '../../semantics/SoilDRaHVocabulary_of_variable.csv'
is_type.file <- '../../semantics/SoilDRaHVocabulary_is_type.csv'

```

# Purpose

The [Hawaiʻi Climate Smart Partnership](https://climatesmarthawaii.org/) seeks to accelerate the implementation of climate-smart practices centered on Hawaiʻi-based producers and ancestral practitioners who are supported by advocacy, incentives, technical assistance, and market development. As part of this project, we need to build a data base of soil carbon stocks found on different Hawaiian soil types under specific land uses and practices.

# Variables of interest for this collection include:

<!--- Briefly describe what you want the data rescuers to look for when digitizing a data resource.---->

### Location
  * region (restricted to the Hawaiian Islands)
  * latitude/longitude and associated datum
  * observation year
  * depth of the top and bottom of the layer from surface

### Soil carbon stock
  * soil organic carbon fraction (any method)
  * soil carbon density (any method)
  * soil bulk density (any method)
  * soil coarse fraction (any method)
  * soil inorganic carbon fraction (any method)
  * soil loss on ignition

### Soil type
  * soil taxonomy (suborder)
  * soil mineralogy (high activity clay, low activity clay, sandy, poorly/noncrystalline)
  * soil moisture regime (ustic vs udic)
  
### Land use or practice
  * current land use (cropland, unmanaged, plantation, pasture, forest, agroforest, silvopasture, etc.)
  * site history (intensive pineapple or sugarcane plantation agriculture, intensive ranching)


# Data sources

Specifically we would like to start by redo-ing the data rescue for the [Hawaii Soil Carbon database](https://osf.io/hmtv6/).

<!---- Move here to final table when data curation in process. ---->

## Data Harmonized - tupled

<!---- Move to this table when the data tuple harmonization is complete. --->

| AuthorYYYY | citation | Data Rescue - GitHub Issue | Folder link | Data Harmonization - GitHub Issue |
|------------|-----------|----------|----------|--------|
| Ares2001 | \@ares2001 | [77](https://github.com/ktoddbrown/SoilDRaH/issues/77) | [01_DataRescue/Ares2001](../../01_DataRescue/Ares2001) |

## Data Rescue - digitized

<!---- Move to this table when the data rescue is complete and it is ready for tuple harmonization. --->

| AuthorYYYY | citation | Data Rescue - GitHub Issue | Folder link |
|------------|-----------|----------|----------|


## In process

<!---- move files into this table when someone takes lead on the data rescue, update to add in git hub issue when it is create --->

| AuthorYYYY | citation | Data Rescue - GitHub Issue | lead point of contact|
|------------|----------|-----------|----------|
| Austin1998 | \@austin1998 | [102](https://github.com/ktoddbrown/SoilDRaH/issues/102) | @SavaScott |
| Austin2010 | \@austin2010 | [119](https://github.com/ktoddbrown/SoilDRaH/issues/119) | @AshBonner310 |
| Bashkin1998 | \@bashkin1998 | [79](https://github.com/ktoddbrown/SoilDRaH/issues/79) | @laylakalandjian |
| Beilman2019 | \@beilman2019 | [146](https://github.com/ktoddbrown/SoilDRaH/issues/146) | @semantic-nomad |
| Binkley1999 | \@binkley1999 | [76](https://github.com/ktoddbrown/SoilDRaH/issues/76) | @hannguyenuf |
| Binkley2004 | \@binkley2004 | [78](https://github.com/ktoddbrown/SoilDRaH/issues/78) | @SavaScott |
| Burke2003 | \@burke2003 | [80](https://github.com/ktoddbrown/SoilDRaH/issues/80) | @Misheph |
| Crews1995 | \@crews1995 | [87](https://github.com/ktoddbrown/SoilDRaH/issues/87) | @laylakalandjian |
| Crow2016 | \@crow2016 | [123](https://github.com/ktoddbrown/SoilDRaH/issues/123) | @waldera |
| Crow2020 | \@crow2020 | [138](https://github.com/ktoddbrown/SoilDRaH/issues/138) | @laylakalandjian |
| Elmore2006 | \@elmore2006 | [93](https://github.com/ktoddbrown/SoilDRaH/issues/93) | @SavaScott |
| Giardina2014 | \@giardina2014 | [171](https://github.com/ktoddbrown/SoilDRaH/issues/171) | @ChaseSaxour |
| Idol2007 |  \@idol2007 | [83](https://github.com/ktoddbrown/SoilDRaH/issues/83) | @waldera |
| Kramer2012 | \@kramer2012 | [95](https://github.com/ktoddbrown/SoilDRaH/issues/95) | @SavaScott |
| Kramer2016 | \@kramer2016 | [137](https://github.com/ktoddbrown/SoilDRaH/issues/137) | @laylakalandjian |
| Li2010 | \@li2010 | [134](https://github.com/ktoddbrown/SoilDRaH/issues/134) | @Misheph |
| Litton2008 | \@litton2008 | [91](https://github.com/ktoddbrown/SoilDRaH/issues/91) | @laylakalandjian |
| Nusslein1999 | \@nusslein1999 | [105](https://github.com/ktoddbrown/SoilDRaH/issues/105) | @Misheph |
| Perez2001 | \@perez2001 | [92](https://github.com/ktoddbrown/SoilDRaH/issues/92) | @SavaScott |
| Scowcroft2004 | \@scowcroft2004 | [117](https://github.com/ktoddbrown/SoilDRaH/issues/117) | @SavaScott |
| Stewart2011 | \@stewart2011 | [114](https://github.com/ktoddbrown/SoilDRaH/issues/114) | @Misheph |
| Tirado-corbala2015 | \@tirado-corbala2015 | [110](https://github.com/ktoddbrown/SoilDRaH/issues/110) | @SavaScott |
| Townsend1995 | \@townsend1995 | [74](https://github.com/ktoddbrown/SoilDRaH/issues/74) | @ktoddbrown |
| Townsend1997 | \@townsend1997 | [167](https://github.com/ktoddbrown/SoilDRaH/issues/167) | @SavaScott |
| Veldkamp1994 | \@veldkamp1994 | [94](https://github.com/ktoddbrown/SoilDRaH/issues/94) | @Misheph |
| Youkhana2009 | \@youkhana2009 | [84](https://github.com/ktoddbrown/SoilDRaH/issues/84) | @SavaScott |

## Imported to bib

<!---- Move here when data source bib entry has been made ---->

| AuthorYYYY | citation | 
|------------|------------|
| ares2001 | @ares2001 |
| austin1998 | @austin1998 |
| austin2010 | @austin2010 |
| bashkin1998 | @bashkin1998 |
| beilman2019 | @beilman2019 |
| binkley1999 | @binkley1999 |
| binkley2004 | @binkley2004 |
| burke2003 | @burke2003 |
| crews1995 | @crews1995 |
| crow2016 | @crow2016 |
| crow2020 | @crow2020 |
| elmore2006 | @elmore2006 |
| giardina2014 | @giardina2014 |
| idol2007 |  @idol2007 |
| kao-kniffin2008 | @kao-kniffin2008 |
| kramer2012 | @kramer2012 |
| kramer2016 | @kramer2016 |
| krueger2021 | @krueger2021 |
| li2010 | @li2010 |
| litton2008 | @litton2008 |
| meulemans2016 | @meulemans2016 |
| nusslein1999 | @nusslein1999 |
| perez2001 | @perez2001 |
| riley1995 | @riley1995 |
| scowcroft2004 | @scowcroft2004 |
| stewart2011 | @stewart2011 |
| tirado-corbala2015 | @tirado-corbala2015 |
| townsend1995 | @townsend1995 |
| townsend1997 | @townsend1997 |
| veldkamp1994 | @veldkamp1994 |
| youkhana2009 | @youkhana2009 |

## Identified

 + [kao-kniffin2008](https://doi.org/10.1007/s00248-007-9323-1)
 + [krueger2021](http://hdl.handle.net/10125/75973)
 + [meulemans2016](http://hdl.handle.net/10125/51342)
 + [riley1995](https://doi.org/10.2307/1940650)
 + soilsurvey2020 -- see Susan Crow
 

# HiSOC data construction

```{r}
#We have study, site, and layer tables

HISOC_sourceVar <- c("source_id",
                     "citation", 
                     "doi")
HISOC_siteVar <- c("site_id",
                   "source_id",
                   "land_use_id",
                   "region", 
                   "observation_time",
                   'elevation',
                   'mean_air_temperature',
                   'total_rainfall',
                   "soil_class")
HISOC_site_history <- c( "land_use_id",
                         "source_id",
                         "land_use")
HISOC_layerVar <- c("layer_id",
                    "layer",
                    "soil_organic_carbon",
                    "soil_nitrogen",
                    "soil_phosphorus",
                    "soil_ph")

```

```{r}
#| label: constructDatabase

HiCSC.ls <- list(source = tibble(),
                 site = tibble(),
                 site_history = tibble(),
                 layer = tibble())

#Use a for-loop because we want to append multiple tables
for(dataset.str in dataID.ls){
  dataset.str <- dataID.ls[[1]]
  dataset.ls <- do.call(paste0('read', dataset.str), 
       list(dataDir = file.path(rescueDirectory, dataset.str),
       dataLevel = 'level2-HiSOC')) 
  HiCSC.ls$source <- bind_rows(HiCSC.ls$source,
                               dataset.ls$source)
  HiCSC.ls$site <- bind_rows(HiCSC.ls$site,
                             dataset.ls$site)
  HiCSC.ls$site_history <- bind_rows(HiCSC.ls$site_history,
                                     dataset.ls$site_history)
  HiCSC.ls$layer <- bind_rows(HiCSC.ls$layer,
                              dataset.ls$layer)
}
```

```{r}

#Data viz checks
```

# References

<!---- This will be populated when you render the document.---->
