---
title: "Data Rescue: Burke 2003"
format:
  html:
    toc: true
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - Burke2003_Methods.bib
  - Burke2003.bib
author:
  - id: ktoddbrown
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    note: https://github.com/ktoddbrown
    affiliations:
      - id: ESS UF
        name: University of Florida
        department: Environmental Engineering Sciences
        city: Gainesville
        state: FL
        country: USA
    role: 
      - "Transcription"
      - "Review"
  - id: Misheph
    name:
      given: Mirah
      family: Blanchard
    affiliation:
      - id: UF
        name: University of Florida
        city: Gainesville
        state: FL
        country: USA
    role: "Transcription"
    note: https://github.com/Misheph
---

```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false
#| eval: true


library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# data files for level 0
methods.file <- "Burke2003_Methods.md"
table1.file <- 'Burke2003_Table1.csv'

# Bibliography files
primaryCitation.file <- 'Burke2003.bib'
methodsCitation.file <- 'Burke2003_Methods.bib'

# read function; you will need to create this function after you have your data
#... rescue reviewed and before you finish your data harmonization.
readsource.file <- '../../02_DataHarmonization/readBurke2003.R'
if(file.exists(readsource.file)){
  source(readsource.file)
}

#semantic files for level 1
of_variable.file <- '../../semantics/SoilDRaHVocabulary_of_variable.csv'
is_type.file <- '../../semantics/SoilDRaHVocabulary_is_type.csv'

```

This documents the data transcription for @Burke2003.
For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/80.

> Burke, R.A., Molina, M., Cox, J.E., Osher, L.J. and Piccolo, M.C. (2003), Stable Carbon Isotope Ratio and Composition of Microbial Fatty Acids in Tropical Soils. J. Environ. Qual., 32: 198-206. https://doi.org/10.2134/jeq2003.1980

# Data Summary

@Burke2003 investigates the effects of soil type and land use on soil microbial community structure and orgnaic carbon stocks.
This is done by analyzing soil phospholipid fatty acid profiles (PLFA) and stable carbon isotope ratios as indicators of microbial activity.

## Fit for purpose: HiCSC

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the HiCSC.

- Location:  No latitude and longitude coordinates are listed, but regional descriptions are given in methods section. 
- Soil carbon stock: Soil organic carbon percentage is in Table 1. Bulk density and stock values do not appear to be persent
- Soil type: Classification is given in the methods section (Typic Distrandepts).
- Land use: Land use history for site given in methods.

# Tables and figures

- Table 
  1) delta 13C, organic C, PLFA, soil depth
  2) PCA of top three components by region and management

- Figure 
  1) PCA plot of PLFA data
  2) PLFA 13C by time since conversion
  3) 13C PLFA vs 13C SOM
  4) 13C PLFA by site
  
  
### Table 1

Table 1 was modified by moving the sub-tables to columns, lifting the units into the header, and down filling empty rows. 

```{r table1}
#| code-fold: true
#| message: false

read_csv(file = table1.file,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = table1.file, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```


{{< include Burke2003_Methods.md >}}

# Transcription comparison

```{gitbash eval=FALSE, echo=FALSE}
# Transcription comparison
diff --strip-trailing-cr --suppress-common-lines -y Burke2003_Methods_KTB.md Burke2003_Methods_MB.md

# Transcription comparison
diff --strip-trailing-cr --suppress-common-lines -y Burke2003_Table1_KTB.csv Burke2003_Table2_MB.csv

```

# Files

These files are in the @Burke2003 data rescue.

- [Readme](README_Burke2003.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Burke2003.bib](Burke2003.bib)
  + citation for article transcribed
- [Burke2003_Methods.bib](Burke2003_Methods.bib)
  + citations for the methods section of the article
- [Burke2003_Methods.md](Burke2003_Methods.md)
  + methods transcribed from primary article
- [Burke2003_Table1.csv](Burke2003_Table1.csv)
  + Table 1 from primary article with organic carbon fraction.

# Working notes 

## Data Rescue Plan

- [x] describe files in human readable form inside ReadMe
- [x] description with Git issue with link
- [x] add contribution yaml
- [x] create excel sheet with Table 1
- [x] export excel table to csv
- [x] copy over methods section into markdown
- [x] pull down citations in methods section to bib file
- [x] add in citation links to methods
- [x] cross check with second transcriber for tables and methods
- [x] submit to independent review
- [ ] archive on SoilDRaH

<!----
## Citation notes from Methods

Below are the citation notes from the methods section.
Some citations are missing, those are noted here and how best guess were generated.

- [x] Dean and Dixon, 1951
  + Dean, R.B., and W.J. Dixon. 1951. Simplified statistics for small numbers of observations. Anal. Chem. 23:636–638.
  + citation from table 1 caption
  + direct export
- [x] Osher 1997
  + Osher, L J 1997 Sequestration and turnover of soil carbon: The roles of minerology and land use change. PhD diss. University of California, Berkeley.
  + direct export
- [x] Townsend et al., 1995
  + Townsend, AR, PM Vitousek, and SE Trumbore. 1995. Soil organic matter dyanmics along gradients in temperature an dland use on the island of Hawaii. Ecology 76:721-733.
  + direct export
- [x] Bashkin and Binkley, 1998
  + Bashkin, MA, and D Binkley 1998. Changes in soil carbon following afforestration in hawaii. Ecology 79:828-833.
  + direct export
- [x] Vitorello et al., 1989
  + Vitorello, V.A., C.C. Cerri, F. Andreaux, C. Feller, and R.L. Victoria. 1989. Organic matter and natural carbon-13 distribution in forested and cultivated Oxisols. Soil Sci. Soc. Am. J. 53:773–778.
  + direct export
- [x] Rhoades et al., 2000
  + Rhoades. CC, GE Eckert, and DC Coleman. 2000. Soil carbon differences among forest, agriculture, and secondry vegitation in lower montane Ecuador. Ecol. Appl. 10:497-505
  + direct export
- [x] Bligh and Dyer, 1959
  + Bligh, E.G. A rapid method of total lipid extraction and purification. J. Biochem. Physiol. 1959 37 911–917
  + direct export
- [x] Zelles (1999)
  + Zelles, L. Fatty acid patterns of phospholipids and lipopolysaccharides in the characterisation of microbial communities in soil: A review. Biol. Fertil. Soils 1999 29 111–129.
  + direct export
- [x] Zelles and Bai, 1993
  + Zelles, L. Fractionation of fatty acids derived from soil lipids by solid phase extraction and their quantitative analysis by GC–MS. Soil Biol. Biochem. 1993 25 495–507.
  + direct export
- [x] White et al., 1979
  + White, D.C. Determination of sedimentary microbial biomass by extractible lipid phosphate. Oecologia 1979 40:51–62.
  + direct export
- [x] Grogan and Cronan, 1997
  + Grogan, D. Cyclopropane ring formation in membrane lipids of bacteria. Microbiol. Mol. Biol. Rev. 1997 61 429–441.
  + direct export
- [x] Nichols et al., 1986
  + Nichols, P.D. Determination of monounsaturated fatty acids double bond position and geometry for microbial monocultures and complex consortia by capillary GC–MS of their dimethyl disulfide adducts. J. Microbiol. Methods 1986 5 49–55.
  + direct export
- [x] Number Cruncher Statistical Systems, 2000
  + Number Cruncher Statistical Systems. 2000. NCSS 2000. NCSS, Kaysville, UT.
  + direct export
- [x] O'Malley et al., 1997
  + O'Malley, V.P. Using GC–MS/Combustion/IRMS to determine the 13C/12C ratios of individual hydrocarbons produced from the combustion of biomass materials—Application to biomass burning. Org. Geochem. 1997 27 567–581.
  - direct export
- [x] Abraham et al. (1998)
  + Abraham, W.-R. Ratios of carbon isotopes in microbial lipids as an indicator of substrate usage. Appl. Environ. Microbiol. 1998 64 4202–4209.
  + direct export
- [x] Rieley, 1994
  + Rieley, G. Derivatization of organic compounds prior to gas chromatographic–combustion–isotope ratio mass spectrometric analysis: Identification of isotope fractionation processes. Analyst 1994 119 915–919.
  + direct export

---->
# Level 0 Data

```{r readLevel0}

data.lvl0.df <- list(citation = 
                    list(primary = 
                           read.bib(file = primaryCitation.file), 
                         methods = 
                           read.bib(file = methodsCitation.file)
                    ),
                  method = read_lines(file = methods.file),
                  data = list(Table1 = 
                                list(caption = 
                                       read_csv(file = table1.file,
                                                col_types = cols(.default = col_character()),
                                                n_max = 1, col_names = FALSE)$X1[1],
                                     primary = 
                                       read_csv(file = table1.file,
                                                col_types = cols(.default = col_character()),
                                                skip = 1)
                                )))
```

# Level 1

```{r Level1}

#### Methods section only ####

studyMeta <- tribble(~region_id, ~site_id, ~of_variable, ~is_type, ~with_entry, ~from_source,
  # Hawaii
  'Hawaii', 'all Hawaii sites','region', 'site', 'Hamakua Coast, Hawaii, northeast flank of Mauna Kea', paste('Method ln7:', data.lvl0.df$method[7]),
  'Hawaii','all Hawaii sites','region', 'state', 'HI', paste('Method ln7:', data.lvl0.df$method[7]),
  'Hawaii','all Hawaii sites', 'elevation', 'value', '700', paste('Method ln7:', data.lvl0.df$method[7]),
  'Hawaii','all Hawaii sites', 'elevation', 'unit', 'm', paste('Method ln7:', data.lvl0.df$method[7]),
  'Hawaii','all Hawaii sites','soil_class', 'value', 'Typic Distrandepts', paste('Method ln8:', data.lvl0.df$method[8]),
  'Hawaii','all Hawaii sites','soil_age', 'maximum', '15000', paste('Method ln8:', data.lvl0.df$method[8]),
  'Hawaii', 'all Hawaii sites','soil_age', 'minimum', '20000', paste('Method ln8:', data.lvl0.df$method[8]),
  'Hawaii', 'all Hawaii sites','soil_age', 'unit', 'yr', paste('Method ln8:', data.lvl0.df$method[8]),
  'Hawaii', 'all Hawaii sites','rainfall', 'maximum', '2500', paste('Method ln9:', data.lvl0.df$method[9]),
  'Hawaii','all Hawaii sites', 'rainfall', 'minimum', '4000', paste('Method ln9:', data.lvl0.df$method[9]),
  'Hawaii', 'all Hawaii sites','rainfall', 'unit', 'mm', paste('Method ln9:', data.lvl0.df$method[9]),
  'Hawaii', 'all Hawaii sites','air_temperature', 'value', '22', paste('Method ln10:', data.lvl0.df$method[10]),
  'Hawaii', 'all Hawaii sites','air_temperature', 'unit', '°C', paste('Method ln10:', data.lvl0.df$method[10]),
  'Hawaii','all Hawaii sites', 'stand_age', 'value', '90', paste('Method ln12:', data.lvl0.df$method[12]),
  'Hawaii', 'all Hawaii sites','stand_age', 'unit', 'yr', paste('Method ln12:', data.lvl0.df$method[12]),
  'Hawaii', 'all Hawaii sites','observation_time', 'value', '1997', paste('Method ln17:', data.lvl0.df$method[17]),

  # Brazil
  'Brazil', 'all Brazil sites','region', 'municiplaity', 'Piracicaba', paste('Method ln23:', data.lvl0.df$method[23]),
  'Brazil', 'all Brazil sites','region', 'city', 'São Paulo', paste('Method ln23:', data.lvl0.df$method[23]),
  'Brazil','all Brazil sites', 'region', 'country', 'Brazil', paste('Method ln23:', data.lvl0.df$method[23]),
  'Brazil', 'all Brazil sites','elevation', 'value', '575', paste('Method ln23:', data.lvl0.df$method[23]),
  'Brazil', 'all Brazil sites','elevation', 'unit', 'm', paste('Method ln23:', data.lvl0.df$method[23]),
  'Brazil','all Brazil sites', 'rainfall', 'value', '1400', paste('Method ln24:', data.lvl0.df$method[24]),
  'Brazil','all Brazil sites', 'rainfall', 'unit', 'mm', paste('Method ln24:', data.lvl0.df$method[24]),
  'Brazil', 'all Brazil sites','air_temperature', 'value', '20', paste('Method ln24:', data.lvl0.df$method[24]),
  'Brazil','all Brazil sites', 'air_temperature', 'unit', '°C', paste('Method ln24:', data.lvl0.df$method[24]),
  'Brazil', 'all Brazil sites','soil_class', 'value', 'isothermic Typic Haplorthox', paste('Method ln25:', data.lvl0.df$method[25]),
  'Brazil', 'all Brazil sites','stand_age', 'value', '27', paste('Method ln28:', data.lvl0.df$method[28]),
  'Brazil', 'all Brazil sites','stand_age', 'value', '65', paste('Method ln28:', data.lvl0.df$method[28]),
  'Brazil', 'all Brazil sites','stand_age', 'unit', 'yr', paste('Method ln28:', data.lvl0.df$method[28]),
  'Brazil','all Brazil sites', 'observation_time', 'value', '1998', paste('Method ln28:', data.lvl0.df$method[28]),
  'Brazil', 'all Brazil sites','mineralogy', 'value', 'clayey, kalonitic', paste('Method ln25:', data.lvl0.df$method[25]),

  # Ecuador
  'Ecuador', 'all Ecuador sites','region', 'value', 'Western slope of Ecuadorian Andes', paste('Method ln38:', data.lvl0.df$method[38]),
  'Ecuador', 'all Ecuador sites','elevation', 'value', '1400', paste('Method ln38:', data.lvl0.df$method[38]),
  'Ecuador', 'all Ecuador sites','elevation', 'unit', 'm', paste('Method ln38:', data.lvl0.df$method[38]),
  'Ecuador','all Ecuador sites', 'rainfall', 'value', '3200', paste('Method ln39:', data.lvl0.df$method[39]),
  'Ecuador', 'all Ecuador sites','rainfall', 'unit', 'mm', paste('Method ln39:', data.lvl0.df$method[39]),
  'Ecuador', 'all Ecuador sites','air_temperature', 'minimum', '17', paste('Method ln39:', data.lvl0.df$method[39]),
  'Ecuador','all Ecuador sites', 'air_temperature', 'maximum', '26', paste('Method ln39:', data.lvl0.df$method[39]),
  'Ecuador', 'all Ecuador sites','air_temperature', 'unit', '°C', paste('Method ln39:', data.lvl0.df$method[39]),
  'Ecuador', 'all Ecuador sites','soil_class', 'value', 'Andic Humitropepts', paste('Method ln40:', data.lvl0.df$method[40]),
  'Ecuador','all Ecuador sites', 'soil_age', 'value', '2500', paste('Method ln40:', data.lvl0.df$method[40]),
  'Ecuador', 'all Ecuador sites','soil_age', 'unit', 'yr', paste('Method ln40:', data.lvl0.df$method[40]),
  'Ecuador', 'all Ecuador sites','stand_age', 'value', '25', paste('Method ln43:', data.lvl0.df$method[43]),
  'Ecuador', 'all Ecuador sites','stand_age', 'unit', 'yr', paste('Method ln43:', data.lvl0.df$method[43]),
  'Ecuador', 'all Ecuador sites','mineralogy', 'value', 'andesitic volcanic ash', paste('Method ln40:', data.lvl0.df$method[40]),
  'Ecuador', 'all Ecuador sites','observation_time', 'value', '1998', paste('Method ln41:', data.lvl0.df$method[41])
)

citationMeta <- tribble(~of_variable, ~is_type, ~with_entry, ~from_source,
            'citation', 'value', format(data.lvl0.df$citation$primary), 
            'journal_citation', 'doi', 'value', data.lvl0.df$citation$primary$doi, 'journal citation')            

land_use.df <- tribble(~land_use_id, ~region_id, ~site_id, ~of_variable, ~is_type, ~with_entry, ~from_source,
   # LU1 is Hawaii's Kalopa Site
    'LU1', 'Hawaii', 'Kalopa', 'land_use', 'description', 'sugarcane', paste('Method ln13:', data.lvl0.df$method[13]),
    'LU1', 'Hawaii', 'Kalopa', 'land_use', 'duration', 'P50Y', paste('Method ln13:', data.lvl0.df$method[13]),
   
   # LU2 is Hawaii's Humuula site
    'LU2', 'Hawaii', 'Humnula','land_use', 'description', 'sugarcane', paste('Method ln13:', data.lvl0.df$method[13]),
    'LU2', 'Hawaii', 'Humnula', 'land_use', 'duration', 'P90Y', paste('Method ln13:', data.lvl0.df$method[13]),
   
   # LU3 is Brazil's Forest site
   'LU3', 'Brazil', 'Forest', 'land_use', 'description', 'Forest', paste('Method ln23:', data.lvl0.df$method[23]),
   
   #LU4 is Brazil's Sugarcane P27Y
   'LU4', 'Brazil', 'Sugarcane_27Y', 'land_use', 'description', 'extensive sugarcane cultivation', paste('Method ln23:', data.lvl0.df$method[23]),
   'LU4', 'Brazil', 'Sugarcane_27Y', 'land_use', 'duration', 'P27Y', paste('Method ln28:', data.lvl0.df$method[28]),
   
   #LU5 is Brazil's Sugarcane P65Y
   'LU5', 'Brazil', 'Sugarcane_65Y', 'land_use', 'description', 'extensive sugarcane cultivation', paste('Method ln23:', data.lvl0.df$method[23]),
   'LU5', 'Brazil', 'Sugarcane_65Y', 'land_use', 'duration', 'P65Y', paste('Method ln28:', data.lvl0.df$method[28]),
   
   #LU6 is Ecuador's Forest Site
   'LU6', 'Ecuador', 'Forest', 'land_use', 'description', 'a closed canopy and a healthy, solid understory of ferns and forbs', paste('Method ln45:', data.lvl0.df$method[45]),
   'LU6', 'Ecuador', 'Forest', 'land_use', 'duration', 'P25Y', paste('Method ln45:', data.lvl0.df$method[45]),
   
   #LU7 is Ecuador's Pasture Site
  'LU7', 'Ecuador', 'Pasture','land_use', 'description', 'African bristlegrass [*Setaria sphacelata* (Schumacher) Moss] pasture sites', paste('Method ln41:', data.lvl0.df$method[41]),
   'LU7', 'Ecuador', 'Pasture', 'land_use', 'duration', 'P25Y', paste('Method ln43:', data.lvl0.df$method[43]),
  
  #LU8 is Ecuador's Sugarcane Site
  'LU8', 'Ecuador', 'Sugarcane', 'land_use', 'description', 'Sugarcane', paste('Method ln42:', data.lvl0.df$method[42]),
   'LU8', 'Ecuador', 'Sugarcane', 'land_use', 'duration', 'P25Y', paste('Method ln43:', data.lvl0.df$method[43])
   )

#### Table 1 #### - Code Broken

Table1Primary <- data.lvl0.df$data$Table1$primary |>
 mutate(row_id = paste0('R', 1:n())) |>
   pivot_longer(cols = -c(row_id),
               names_to = 'column_name', values_to = 'with_entry',
               values_drop_na = TRUE) |>
    mutate(of_variable = case_when(
      column_name == 'Location' ~ 'region',
      column_name == 'Land use' ~ 'site_id',
      column_name == 'Soil depth (cm)' ~ 'layer',
      column_name == 'PFLA (mg kg-1)' ~ 'phospholipid_acid_content',
      column_name == 'Organic C content (%)' ~ 'organic_carbon',
      column_name == 'δ13C-SOM (per mil)' ~ 'delta_13_carbon',
      .default = 'Bad Match'
    )) |>
    
 
mutate(
  value = case_when(
    column_name == "Location" ~ with_entry,
    column_name == "Land use" ~ with_entry,
    column_name == "PFLA (mg kg-1)" & str_detect(with_entry, '^\\d+(\\.\\d+)?$') ~ with_entry,
    TRUE ~ NA_character_
     ),
  lowerbound = case_when(
    str_detect(with_entry, '^\\d+\\s*-\\s*\\d+') ~ str_extract(with_entry, "\\d+(?=-)"),
    TRUE ~ NA_character_
),
  upperbound = case_when(
   str_detect(with_entry, '^\\d+\\s*-\\s*\\d+') ~ str_extract(with_entry, "(?<=-)\\d+"),
     TRUE ~ NA_character_
    ),
  mean = as.character(
      str_extract(with_entry, '^-?\\d*\\.?\\d+(?=\\s*±)')),
  `standard error` = as.character(
    str_extract(with_entry, '(?<=\\±)\\s*\\d*\\.?\\d+')),
    ) |>
  select(-with_entry) |>
  
  pivot_longer(cols = c('lowerbound','upperbound','mean','standard error', 'value'),
               names_to = 'is_type',
               values_to = 'with_entry',
               values_drop_na = TRUE) |>

  mutate(from_source = 'Table 1')

Table1Meta <- Table1Primary |>
  select(column_name, of_variable) |>
  unique() |>
  
  mutate(unit = str_extract(column_name, pattern = '(?<=\\().*(?=\\))'),
         species = case_when(
           str_detect(column_name, 'sugarcane') ~ 'Saccharum spp.', 
           TRUE ~ NA_character_
         ),
      from_source = 'Table 1 column names.') |>
  filter(!is.na(unit) | !is.na(species))

bind_rows(
  tribble(~of_variable, ~method, ~from_source,
  'region', paste0(data.lvl0.df$method[c(7,24,38)], collapse = ' '), 'Methods ln7;24;28',
  'site_id', paste0(data.lvl0.df$method[c(11:13,28,41:42)], collapse = ' '), 'Methods ln11-13,28,41-42',
  'layer', paste0(data.lvl0.df$method[c(17:19,34,47)], collapse = ' '), 'Methods ln17-19,34,47',
  'phospholipid_acid_content', paste0(data.lvl0.df$method[51:67], collapse = ' '), 'Methods ln51-67',
  'delta_13_carbon', paste0(data.lvl0.df$method[c(89:90,103)], collapse = ' '), 'Methods ln89-90,103'
    )
  ) |>
bind_rows(
  tribble(~of_variable, ~control_vocabulary, ~from_source,
    #struggling to find info on the USDA soil classification
    'stand_type', '*Saccharum spp.*: sugarcane', 'Abstract ln2')
 |>
         left_join(Table1Primary |>
                  select(column_name, of_variable) |>
                  unique(),
                by = join_by(of_variable))) |>
  pivot_longer(cols = c(unit, species, method, control_vocabulary),
               names_to = 'is_type',
               values_drop_na = TRUE,
               values_to = 'with_entry')


#### Create level 1

data.lvl1.ls <- list(
  study = studyMeta,
  primary_meta = dplyr::bind_rows(Table1Meta),
  primary = bind_rows(Table1Primary)|>
    mutate(age_id = with_entry[of_variable == 'stand_age'],
           is_type = 'value',
           .by = row_id) |>
    arrange(row_id, age_id, column_name,
            of_variable, is_type, with_entry, from_source)
  )
```

