---
title: "Data Rescue: Li 2010"
format:
  html:
    toc: true
    eval: true
date: 2026-03-01
date-format: YYYY-MM-DD
bibliography:
  - Li2010_Methods.bib #contains the citations in the methods section
  - Li2010.bib #contains citation for the data source, generally a single citation of the primary paper
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  - id: ktoddbrown 
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - uf-ees
    role: #TODO define these roles and add reference
      - transcription: review
      #- standardization: #lead #review
      #- curation: #lead #review
  - id: SavaScott
    name:
      given: Savannah
      family: Scott
    affiliation:
      - uf-ees
    role:
      - transcription: secondary
      #- standardization:
      #- curation:
  - id: Misheph
    name:
      given: Mirah
      family: Blanchard
    affiliation:
      - uf-mae
    role: 
      - transcription: lead
      #- standardization: 
      #- curation: 


affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
  - id: uf-mae
    name: University of Florida
    department: Mechanical and Aerospace Engineering
    city: Gainesville
    state: Florida
    county: USA
    address: "P.O Box 116250"
    postal-code: 32611
    url: https://mae.ufl.edu/
    
timelog:
  - activity: 
      description: Revise based on feedback
      who: Misheph #github username
      start: 2025-11-01
      end: 2026-3-01
      duration: 15H
---


```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false


library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# data files for level 0
methods.file <- "Li2010_Methods.md"
figure2.file <- 'Li2010_Figure2.csv'

# Bibliograph files
primaryCitation.file <- 'Li2010.bib'
methodsCitation.file <- 'Li2010_Methods.bib'

# read function
#readsource.file <- '../../../R/readAuthorYYYY.R'
#source('readsource.file')

#semantic files for level 1
#of_variable.file <- '../../semantics/SoilDRaHVocabulary_of_variable.csv'
#is_type.file <- '../../semantics/SoilDRaHVocabulary_is_type.csv'

```


# Data Summary

> Li, Y., Mathews, B.W. Effect of conversion of sugarcane plantation to forest and pasture on soil carbon in Hawaii. Plant Soil 335, 245–253 (2010). https://doi.org/10.1007/s11104-010-0412-4

For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/134.

This data examines how changes in land use affects soil C storage. The researchers looked at how former sugarcane plantations that were now either forests or pastures compared in soil C, soil chemistry, and stable C isotope ratios. Soil samples were collected at different soil depths then analyzed for physical properties and chemical properties using methods such as colorimetric analysis and ICP. Researchers found that forests stored more total soil carbon than pastures, and that reforestation led to greater accumulation and stabilization of SOC than conversion to pasture.

## Tables and figures

- Table 
  1) Average δ¹³C, %SOC₄, and %SOC₃
  2) Soil pH, exchangeable cations, ECEC, Al saturation, P, and N

- Figure 
  1) Monthly precipitation and temperature at study site
  2) Total soil carbon (TC) and SOC₃ (secondary forests) and SOC₄ (cane and grass) proportions
  3) Oxalate-extractable Al in pasture vs forest

## Fit for purpose: HiCSC

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the HiCSC.

  - Location: location described and coordinates provided in methods section under study sites
  - Soil carbon stock: SOC fractions provided in table 1 and figure 2, total soil C (Mg ha^-1) provided in figure 2
  - Soil type: Soil classification (Acrudoxic Hydrudands of the Akaka and Kaiwiki series) given in methods section under study sites
  - Land use: Land use and history detailed in methods section under study sites
  
## Files

These files are in the Li2010 data rescue.

- [Readme](README_AUTHORLi2010.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Li2010.bib](Li2010.bib)
  + citation for article transcribed
- [Li2010_Methods.bib](Li2010_Methods.bib)
  + citations for the methods section of the article
- [Li2010_Methods.md](Li2010_Methods.md)
  + methods transcribed from primary article
- [Li2010_Figure2.csv](Li2010_Figure2.csv)
  + figure 2 from primary article with total soil carbon (TC), SOC₃ (secondary forests), SOC₄ (cane and grass) proportions
- [temp/](temp/)
    + scratch folder that will not be archived on GitHub, it should include local copies of the data resource (likely a PDF article) and any spreadsheets or other tools you used for the transcriptions.

# Data Rescue working notes 

## Data Rescue Plan

+ Lead
  - [x] Read the article or data resource and the discription of the target project
  - [x] Create a fresh fork of the primary repository https://github.com/ktoddbrown/SoilDRaH/tree/main, it might be useful to rename it 'SoilDRaH_AuthorYYYY'
  - [x] Create a copy of this template folder and rename it to AuthorYYYY
  - [x] Fill out Data summary section (except for the Files), the Data Rescue Plan (here!), and update the author section of the yml to reflect who is working on the rescue
  - [x] Create a GitHub issue from the Data Rescue template and get feedback from a reviewer on the project (Dr Todd-Brown or a grad student typically).
  - [x] Add the reviewer to the author list in the yml
+ Do
  - [x] Extract points from Figure2 and export it to csv
  - [x] Transcribe or copy the method section
  - [x] Insert references to bib entries
  - [x] reformat equations and mathematical symbols as [LaTex](https://en.wikibooks.org/wiki/LaTeX/Mathematics)
  - [x] reformat text using [Markdown](https://www.markdownguide.org/cheat-sheet/)
  - [x] Create a bib file with the citations from the methods section
  - [x] Push to your local fork and get a second transcription pushed to your local fork by someone else on the project
+ Measure
  - [x] cross check with second transcriber for tables and methods, and resolve any differences
  - [x] compile ReadMe with all chunks up through the Level 0 data rescue section with `eval: true`
  - [x] push to your local fork and create a pull request to the main branch of the primary repository
  - [ ] address any revisions suggested in your pull request review
  - [ ] celebrate!

## Figure 2

```{r figure2, fig.height=8, fig.width=6}

fig2 <- read_csv(
  figure2.file,
  skip = 1, # skip first line
  col_names = TRUE,
  show_col_types = FALSE
)

fig2 <- fig2 %>% # modify fig2 and pass back in
  mutate(
    y = as.numeric(y),
    `error bar` = as.numeric(`error bar`),
    `graph label` = factor(
      `graph label`,
      levels = c("A", "B"), # A first and then B, top to bottom
      labels = c("0–10 cm", "10–25 cm")
    ),
    `x label` = factor(
      `x label`,
      levels = c("TC", "SOC4", "SOC3") # control order of bars
    ),
    ecosystem = factor(
      ecosystem,
      levels = c("S. Forest", "Pastures") # control fill colors and legend order
    )
  )

# plot figure
ggplot(fig2, aes(x = `x label`, y = y, fill = ecosystem)) + # initialize bar chart with height and fill according to ecosystem
  geom_col(
    position = position_dodge(width = 0.8),
    width = 0.7
  ) +
  geom_errorbar(
    aes(ymin = y, ymax = `error bar`), # start error bar at y value, extend to error bar value
    position = position_dodge(width = 0.8),
    width = 0.2
  ) +
  geom_text(
    aes(label = footnote, y = `error bar` + 2), # annotate above each error bar
    position = position_dodge(width = 0.8),
    vjust = 0
  ) +
  facet_wrap(~ `graph label`, ncol = 1) +
  scale_y_continuous(
    breaks = c(0, 20, 40, 60, 80, 100), # set y axis values
    limits = c(0, 100)
  ) +
  scale_fill_manual(
    values = c("S. Forest" = "red", "Pastures" = "blue"), # set color of ecosystem
    name = ""
  ) +
  labs(
    x = "", # no x label
    y = expression(Soil~C~pools~(Mg~ha^{-1})) # exponent formatting
  ) +
  theme_classic() +
  theme(
    legend.position = "top",
    strip.text = element_blank()
  )
```

## Methods comparison

```{bash eval=FALSE}
diff --strip-trailing-cr --suppress-common-lines -y temp/Li2010_Methods_SRS.md Li2010_Methods.md
```

## Methods Section Transcription

{{< include Li2010_Methods.md >}}

## Citation notes from Methods

Below are the citation notes from the methods section.
Some citations are missing, those are noted here and how best guess were generated.

- [x] (Parfitt et al. 1988) @Parfitt1988 
  + Parfitt RL, Childs CW, Eden DN (1988) Ferrihydrite and allophone in four andepts from Hawaii and implications for their classification.    Geoderma 41:223–241
  + direct export
- [x] (Patrick Niemeyer, USDA-NRCS) @PatrickPersComm
  + Personal Communication with member of USDA-NRCS
  + manual entry
- [x] (Bashkin and Binkley 1998) @Bashkin1988
  + Bashkin MA, Binkley D (1998) Changes in soil carbon following afforestation in Hawaii. Ecology 79:828–833
  + direct export
- [x] (Hue et al. 2000) @Hue2000
  + Hue NV, Uchida R, Ho MC (2000) Sampling and analysis of soils and plant tissues: how to take representative samples, how the samples are tested. In: Silva JA, Uchida RS (eds) Plant nutrient management in Hawaii's soils: approaches for tropical and subtropical agriculture. CTAHR, Univ. of Hawaii, Manoa, pp 23–30
  + direct export
- [x] (Shuman and Duncan 1990) @Shuman1990 
  + Shuman LM, Duncan RR (1990) Soil exchangeable cations and aluminum measured by ammonium chloride, potassium chloride, and ammonium acetate. Commun Soil Sci Plant Anal 21:1217–1228
  + direct export
- [x] (Parfitt 1989) @Parfitt1989 
  + Parfitt RL (1989) Optimum conditions for extraction of Al, Fe, and Si from soils with acid oxalate. Commun Soil Sci Plant Anal 20:801–816
  + direct export
- [x] (Resh et al. 2002) @Resh2002
  + Resh SC, Binkley D, Parrotta JA (2002) Greater soil carbon sequestration under nitrogen-fixing trees compared with eucalyptus species. Ecosystems 5:217–231
  + direct export

# Level 0 Data rescue

This level 0 data read parses the transcribed files with minimal cleaning or transformations.

```{r readLevel0}
#This chunk has two purposes. 
#...1) Check the formatting by reading in everything 
#...2) Create a list of everything to process later in level 1

data.lvl0.ls <- list(
  #Read in a list of all the bib files
  citation = list(
    #Citation for the article transcriptions are pulled from
    primary = read.bib(file = primaryCitation.file), 
    #Citations for all referenced articles
    methods = read.bib(file = methodsCitation.file)
  ),
  #Read in the text transcription of the article's methods section
  method = read_lines(file = methods.file),
  #Read in the results as tables or figure transcriptions. This includes
  #...the caption as well as the tables themselves
  data = list(
    Figure2 = list(
      #Read the caption as a text string. Captions are the first cell on 
      #...the first row.
      caption = read_csv(file = figure2.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      #Read in all the data, skipping the first row with the caption and read
      #...in the table as character. This element is a tibble (data.frame).
      primary = read_csv(file = figure2.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    ) #End 'Figure2' element
    #Repeat for all tables and figures...
   ) #end 'data' element
)

```

```{r}
#| label: checkReadFunction_lvl0
#| eval: false

datafunction.lvl0.ls <- readLi2010(dataDir = '.',
                                    dataLevel = 'level0')

if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
  print('Read function matches code here.')
}else{
  print('There is a mismatch in the data objects between the code here and the read function.')
}

```


# References

<!---- This space will be filled with the citations when rendered. Leave this blank. ---->

