---
title: "Data Rescue: Burke 2003"
format:
  html:
    toc: true
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - Burke2003_Methods_KTB.bib
  - Burke2003.bib
author:
  - name:
      given: Katherine
      family: Todd-Brown
      orcid: 0000-0002-3109-8130
      note: https://github.com/ktoddbrown
      affiliations:
        - id: ESS UF
          name: University of Florida
          department: Environmental Engineering Sciences
          city: Gainesville
          state: FL
          country: USA
      role: 
        - "Transcription"
        - "Review"
  - name:
      given: Mirah
      family: Blanchard
      affiliation:
        - id: UF
          name: University of Florida
          city: Gainesville
          state: FL
          country: USA
      role: "Transcription"
      note: https://github.com/Misheph
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(kableExtra)
```

This documents the data transcription for  .
For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/80.

> Burke, R.A., Molina, M., Cox, J.E., Osher, L.J. and Piccolo, M.C. (2003), Stable Carbon Isotope Ratio and Composition of Microbial Fatty Acids in Tropical Soils. J. Environ. Qual., 32: 198-206. https://doi.org/10.2134/jeq2003.1980

# Data Summary

@Burke2003 investigates the effects of soil type and land use on soil microbial community structure and orgnaic carbon stocks.
This is done by analyzing soil phospholipid fatty acid profiles (PLFA) and stable carbon isotope ratios as indicators of microbial activity.

## Fit for purpose: HiCSC

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the HiCSC.

- Location:  No latitude and longitude coordinates are listed, but regional descriptions are given in methods section. 
- Soil carbon stock: Soil organic carbon percentage is in Table 1. Bulk density and stock values do not appear to be persent
- Soil type: Classification is given in the methods section (Typic Distrandepts).
- Land use: Land use history for site given in methods.

# Tables and figures

- Table 
  1) delta 13C, organic C, PLFA, soil depth
  2) PCA of top three components by region and management

- Figure 
  1) PCA plot of PLFA data
  2) PLFA 13C by time since conversion
  3) 13C PLFA vs 13C SOM
  4) 13C PLFA by site
  
  
### Table 1

Table 1 was modified by moving the sub-tables to columns, lifting the units into the header, and down filling empty rows. 

```{r table1}
#| code-fold: true
#| message: false

filename <- 'Burke2003_Table1_KTB.csv'
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```


{{< include Burke2003_Methods_KTB.md >}}

# Transcription comparison

```{gitbash eval=FALSE}
diff --strip-trailing-cr --suppress-common-lines -y Burke2003_Methods_KTB.md Burke2003_Methods_MB.md
```

# Files

These files are in the Burke2003 data rescue.

- [Readme](README_Burke2003.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Burke2003.bib](Burke2003.bib)
  + citation for article transcribed
- [Burke2003_Methods.bib](Burke2003_Methods.bib)
  + citations for the methods section of the article
- [Burke2003_Methods.md](Burke2003_Methods.md)
  + methods transcribed from primary article
- [Burke2003_Table1.csv](Burke2003_Table1.csv)
  + table 1 from primary article with organic carbon fraction.
- [temp/](temp/)
    + scratch folder that will not be archived includes excel and article pdf document

# Working notes 

## Data Rescue Plan

- [ ] describe files in human readable form inside ReadMe
- [x] description with Git issue with link
- [x] add contribution yaml
- [x] create excel sheet with Table 1
- [x] export excel table to csv
- [x] copy over methods section into markdown
- [x] pull down citations in methods section to bib file
- [x] add in citation links to methods
- [ ] cross check with second transcriber for tables and methods
- [ ] submit to independent review
- [ ] archive on SoilDRaH


## Citation notes from Methods

Below are the citation notes from the methods section.
Some citations are missing, those are noted here and how best guess were generated.

- [x] Dean and Dixon, 1951
    + citation from table 1 caption
    + direct export
- [x] Osher 1997
  + Osher, L J 1997 Sequestration and turnover of soil carbon: The roles of minerology and land use change. PhD diss. University of California Berkley
- [x] Townsend et al., 1995
  + Townsend, AR, PM Vitousek, and SE Trumbore. 1995. Soil organic matter dyanmics along gradients in temperature an dland use on the island of Hawaii. Ecology 76:721-733
  + direct export
- [x] Bashkin and Binkley, 1998
  + Bashkin, MA, and D Binkley 1998. Changes in soil carbon following afforestration in hawaii. Ecology 79:828-833
- [x] Vitorello et al., 1989
  + Vitorello, VA, CC Cerri, F Andreaudx, C. Feller, and RL Victoria 1989. Organic matter and natural carbon-13 districution in forested and cultivated Oxisols. Soil Sci. Soc. Am. J. 53:773-778.
- [x] Rhoades et al., 2000
  + Rhoades. CC, GE Eckert, and DC Coleman. 2000. Soil carbon differences among forest, agriculture, and secondry vegitation in lower montane Ecuador. Ecol. Appl. 10:497-505
- [x] Bligh and Dyer, 1959
  + Bligh, E.G. A rapid method of total lipid extraction and purification. J. Biochem. Physiol. 1959 37 911–917
- [x] Zelles (1999)
  + Zelles, L. Fatty acid patterns of phospholipids and lipopolysaccharides in the characterisation of microbial communities in soil: A review. Biol. Fertil. Soils 1999 29 111–129. https://doi.org/10.1007/s003740050533
- [x] Zelles and Bai, 1993
  + Zelles, L. Fractionation of fatty acids derived from soil lipids by solid phase extraction and their quantitative analysis by GC–MS. Soil Biol. Biochem. 1993 25 495–507. https://doi.org/10.1016/0038-0717(93)90075-M
- [x] White et al., 1979
  + White, D.C. Determination of sedimentary microbial biomass by extractible lipid phosphate. Oecologia 1979 40 51–62. https://doi.org/10.1007/BF00388810
- [x] Grogan and Cronan, 1997
  + Grogan, D. Cyclopropane ring formation in membrane lipids of bacteria. Microbiol. Mol. Biol. Rev. 1997 61 429–441.
- [x] Nichols et al., 1986
  + Nichols, P.D. Determination of monounsaturated fatty acids double bond position and geometry for microbial monocultures and complex consortia by capillary GC–MS of their dimethyl disulfide adducts. J. Microbiol. Methods 1986 5 49–55. https://doi.org/10.1016/0167-7012(86)90023-0
- [x] Number Cruncher Statistical Systems, 2000
  + Number Cruncher Statistical Systems. 2000. NCSS 2000. NCSS, Kaysville, UT.
- [x] O'Malley et al., 1997
  + O'Malley, V.P. Using GC–MS/Combustion/IRMS to determine the 13C/12C ratios of individual hydrocarbons produced from the combustion of biomass materials—Application to biomass burning. Org. Geochem. 1997 27 567–581. https://doi.org/10.1016/S0146-6380(97)00087-9
- [x] Abraham et al. (1998)
  + Abraham, W.-R. Ratios of carbon isotopes in microbial lipids as an indicator of substrate usage. Appl. Environ. Microbiol. 1998 64 4202–4209.
- [x] Rieley, 1994
  + Rieley, G. Derivatization of organic compounds prior to gas chromatographic–combustion–isotope ratio mass spectrometric analysis: Identification of isotope fractionation processes. Analyst 1994 119 915–919. https://doi.org/10.1039/an9941900915




