---
title: "Data Rescue: Townsend1995"
format:
  html:
    toc: true
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - Townsend1995_Methods.bib
  - Townsend1995.bib
author:
  - name:
      given: Katherine
      family: Todd-Brown
      orcid: 0000-0002-3109-8130
      note: https://github.com/ktoddbrown
      affiliations:
        - id: ESS UF
          name: University of Florida
          department: Environmental Engineering Sciences
          city: Gainesville
          state: FL
          country: USA
      role: 
        - "Transcription"
        - "Review"
  - name:
      given: Layla
      family: Kalandjian
      affiliation:
        - id: UF
          name: University of Florida
          city: Gainesville
          state: FL
          country: USA
      role: "Transcription"
      note: https://github.com/laylakalandjian
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(kableExtra)
library(bibtex)

methods.file <- 'Townsend1995_Methods.md'
table1.file <- 'Townsend1995_Table1.csv'
table2.file <- 'Townsend1995_Table2.csv'
primaryCitation.file <- 'Townsend1995.bib'
methodsCitation.file <- 'Townsend1995_Methods.bib'
```

This documents the data transcription for  .
For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/74

> Full citation here

# Data Summary

This paper looks at soil organic carbon stock and radiocarbon age across a gradient of three elevations of two land uses (pasture and forested). They use this data to fit soil carbon pools in a three pool first order linear model.

## Fit for purpose: HiCSC

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the HiCSC.

- Location: The methods section in this paper does not have specific latitude/longitude of the sites but does give detailed regional descriptions which could be used to infer specific latitude and longitude. Year of observation is given in text.
- Soil carbon stock: Depth of sample and area-density of soil organic carbon is given, unclear how coarse fraction was corrected for or bulk density assessed.
- Soil type: Detailed classification provided in methods section.
- Land use: Detailed land use and land use history in the methods section.

# Tables and figures

- Table 
  1) Soil carbon, respiration, biomass, radiocarbon, pasture age, and forest-derived SOC
  2) Seasonal variation in soil respiration, temperature, and moisture
  3) Model performance summary
  4) Modeled pool size
  5) Soil incubation cumulative totals
  6) Estimated passive pool
  7) Model sensitivity results

- Figure 
  1) Soil pH values
  2) Pasture age vs radiocarbon
  3) Soil depth vs age
  4) Pasture site vs forest-derived carbon
  5) Modeled vs observed 13C
  6) Half-life of intermediate SOM vs mean annual temperature
  
### Table 1

Table 1 was modified by replacing delta symbols with text as well as removing superscripts *a*, *b*, and *c* denoting values used in the results section for an equation.

```{r table1}
#| code-fold: true
#| message: false

filename <- table1.file
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```

### Table 2

Table 2 was modified by combining grouped headings into column/header names.

```{r table1}
#| code-fold: true
#| message: false

filename <- table2.file
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```

{{< include Townsend1995_Methods.md >}}

# Transcription comparison

```{bash eval=FALSE}
diff --strip-trailing-cr --suppress-common-lines -y temp/Townsend1995_Methods_KTB.md Townsend1995_Methods_LK.Rmd
```

# Files

These files are in the Townsend1995 data rescue.

- [Readme](README_Townsend1995.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Townsend1995.bib](Townsend1995.bib)
  + citation for article transcribed
- [Townsend1995_Methods.bib](Townsend1995_Methods.bib)
  + citations for the methods section of the article
- [Townsend1995_Methods.md](Townsend1995_Methods.md)
  + methods transcribed from primary article
- [Townsend1995_Table1.csv](Townsend1995_Table1.csv)
- [Townsend1995_Table2.csv](Townsend1995_Table2.csv)
  + table 1 from primary article.
- [temp/](temp/)
    + scratch folder that will not be archived includes excel and article pdf document

# Working notes 

## Data Rescue Plan

- [x] copy decription from Git issue with link into readme
- [x] create excel sheet with table 1 and 2
- [x] figure out how to get caption and link filed together (?readme yaml?)
- [x] export excel table to csv
- [x] copy over methods section into markdown
- [x] pull down citations in methods section to bib file
- [x] add in citation links to methods section
- [x] add contribution yaml
- [x] describe filed in human readable form inside ReadMe


## Citation notes from Methods

Below are the citation notes from the methods section.
Some citations are missing, those are noted here and how best guess were generated.

Pending: 7 techincal reports and book chapters to manually enter

- [x] Peterson, D. W., and R. B. Moore. 1987. Geologic history and evolution of geologic concepts, island of Hawaii. Pages 149-189 in Volcanism in Hawaii. United States Geological Survey Professional paper 1350. https://hilo.hawaii.edu/maunakea/library/ref/621
    + manual entry as '@incollection'
- [x] Juvik et al. 1978
    + Juvik, J. 0., D. C. Singleton, and G. G. Clarke. 1978. Climate and water balance on the island of Hawaii. Pages 129-139 in Mauna Loa Observatory, a 20th Anniversary report. National Oceanic and Atmospheric Administration Special Report, Silver Spring, Maryland, USA.
    + manual entry as '@incollection'
- [x] Fieldes and Perrott 1965
    + Fields, M., and K. W. Perrott. 1965. The nature of allophane in soils. Part 3. Rapid field and laboratory test for allophane. New Zealand Journal of Science 9:623-629.
    + added to bib from journal export
- [x] Uehara and Gilman 1981
    + Uehara, G., and G. Gillman. 1981. The mineralogy, chemistry, and physics of tropical soils with variable charge clays. Westview, Boulder, Colorado, USA.
    + manual entry as '@book'
- [x] Raich et al. 1990
    + no citaiton
    + given citation appears in soil respiration section, it's likely this paper: Raich, J.W., Bowden, R.D. and Steudler, P.A. (1990), Comparison of Two Static Chamber Techniques for Determining Carbon Dioxide Efflux from Forest Soils. Soil Science Society of America Journal, 54: 1754-1757. https://doi.org/10.2136/sssaj1990.03615995005400060041x
    + added to bib from journal export
- [x] Trumbore 1993
    + Trumbore, S. E. 1993. Comparison of carbon dynamics in tropical and temperate soils using radiocarbon measurements. Global Biogeochemical Cycles 7:275-290.
    + added to bib from journal export
- [x] Vogel et al. 1987
    + Vogel, J. S., D. E. Nelson, and J. R. Southon. 1987. 14C background levels in an accelerator mass spectrometry system. Radiocarbon 29:323-333.
    + added to bib from journal export
- [x] Stuvier and Polach 1977
    + Stuiver, M., and H. Polach. 1977. Reporting of 14C data. Radiocarbon 19:355-363.
    + added to bib from journal export
- [x] Jenkinson and Powlson 1976
    + Jenkinson, D.S., and D. S. Powlson. 1976. The effect of biocidal treatments on metabolism in soil. V. A method for measuring soil biomass. Soil Biology and Biochemistry 8: 209-213.
    + added to bib from journal export
- [x] Coleman et al. 1978
    + Coleman, D. C., R. V. Anderson, C. V. Cole, E. T. Elliot, L. Woods, and M. K. Campion. 1978. Trophic interactions in soils as they affect energy and nutrient dynamics. IV. Flows of metabolic and biomass carbon. Microbial Ecology 4:373-380.
    + added to bib from journal export
    + year incorrectly cited
- [x] Chaussod and Nicolardot (1982)
    + no citation
    + given appearence in microbial biomass section, it is likely this citation: Chaussod,R. and Nicolardot,B., 19831980666, French, Journal article, 19, (4), Revue db
    
# Level 0: raw data reads

The Level 0 data in this package include the tables above (Table 1 and 3) as well as the methods section.
Here we draft the read function that will be transferred to the R scripts.
In the future this draft will be echoed from the R scripts.

```{r readLevel0}

data.lvl0.df <- list(citation = 
                    list(primary = 
                           read.bib(file = primaryCitation.file), 
                         methods = 
                           read.bib(file = methodsCitation.file)
                    ),
                  method = read_lines(file = methods.file),
                  data = list(Table1 = 
                                list(caption = 
                                       read_csv(file = table1.file,
                                                col_types = cols(.default = col_character()),
                                                n_max = 1, col_names = FALSE)$X1[1],
                                     primary = 
                                       read_csv(file = table1.file,
                                                col_types = cols(.default = col_character()),
                                                skip = 1)
                                ), 
                              Table2 = 
                                list(caption = 
                                       read_csv(file = table2.file,
                                                col_types = cols(.default = col_character()),
                                                n_max = 1, col_names = FALSE)$X1[1],
                                     primary = 
                                       read_csv(file = table2.file,
                                                col_types = cols(.default = col_character()),
                                                skip = 1)
                                )))
```

