---
title: "Carbon in Peat on EArth through Time"
author: "Vaasuki Marupaka"
date: "Summer 2024"
output: html_document
---

The purpose of this document is to summarize the portions of the Carbon in Peat on EArth through Time (C-PEAT) Database that are relevant to data collections in the SoilDRaH project and walk through the data ingestion.
Here you will find links to documentation from the data provider, links to where you can access the data, a description of how the data was processed, and visuals for the collection relevant variables.

# What is C-PEAT

The version of the Carbon in Peat on EArth through Time (C-PEAT)Database described here is the public facing version of the PANGAEA repository (https://www.pangaea.de).

The annotations table draws heavily from the parameters list of the original data and also from the point of contact for C-PEAT collection of datasets (Dr. Julie Loisel). 

# Data processing

```{r setup, echo=TRUE, warning=FALSE, message=FALSE}

library(readr) # read in the csv tables
library(tibble) # use tibbles instead of a data frame
library(plyr) # transform a list into a data frame for the bind
library(dplyr) # work with data tables filters/joins/reframes
library(tidyr) # work with data table pivots
library(ggplot2) # make plots
library(stringr) # extract text from descriptions
library(knitr) # make prettier tables
library(kableExtra) #make tables scrollable

source("../SoilDRaH/R/temp.R", local = knitr::knit_global())

#locate the data locally
dataDir <- '../SoilDRaH/data/test/'
dataAnnotations <- '../SoilDRaH/data/CPEAT_annotations.csv'
```

```{r, echo=FALSE}
#Change this file to run locally for you
dataDir  <- '~/Documents/GitHub/SoilDRaH/data/test/'
```

To read in the data base use the `test_readCPEAT` function.

```{r, warning=FALSE}
CPEAT.ls <- readCPEAT(dataDir, 
                  annotationFilename = dataAnnotations, 
                  verbose = FALSE, #when first running, switch this to true 
                  format = 'long')

CPEAT.original <- readCPEAT(dataDir, 
                  annotationFilename = dataAnnotations, 
                  verbose = FALSE, #when first running, switch this to true 
                  format = 'original')

CPEAT.ls$long %>%
  slice_head(n=100) %>%
  knitr::kable() %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "300px")
```

```{r, eval=FALSE}
# testing for any duplicates in the data frame
check <- unique(CPEAT.ls$long) # actually this will not help as it removes the decription after once established for the column variables!
```


There are two tables currently annotated in the C-PEAT data collection.
This represents the data directly relevant for layer-level soil carbon measurements including bulk density, organic carbon fraction, coarse fraction, depth of sample, sampling time, and sample location.
Below are the descriptions of data columns organized by table.

Please note that there is quite a bit more data in the C-PEAT and we welcome interested parties extending these annotations.

## STUDY TABLE 

```{r}
knitr::kable(CPEAT.ls$annotations %>%
               filter(of_type == 'description',
                      table_id == 'study') %>%
               select(column_id, of_variable, description = with_entry)) %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "300px")
```

## CORE TABLE

```{r}
knitr::kable(CPEAT.ls$annotations %>%
               filter(of_type %in% c('description', 'unit'),
                      table_id == 'core') %>%
               select(column_id, of_variable, description = with_entry)) %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "300px")

test <- CPEAT.ls$annotations %>%
               filter(of_type %in% c('description', 'unit'),
                      table_id == 'core') %>%
               select(column_id, of_variable, description = with_entry)
```

# Subseting for soil carbon

There is a lot of data here but not all of it is of interest for all purposes.
In this example we are interested in layer-level soil carbon related measurements, there geolocation.
Note that we are NOT checking any of the quality control flags here which you would want to do if using this for research purposes.

```{r}
# this function reads much better!
CPEAT_soc_cores <- CPEAT.ls$core %>% 
  select(doi, `Depth sed [m]`, `Depth top [m]`, `Depth bot [m]`, 
         `Depth sed [m] (Sample midpoint depth)`, 
         `Depth sed [m] (Sample midpoint depth (rounded))`, 
         `Depth sed [m] (Sample midpoint depth of OxCa...)`, 
         `Depth sed [m] (Sample midpoint depth of Baco...)`,
         `Depth sed [m] (LOI sample depth)`,
         `Samp thick [cm]`, `DBD [g/cm**3]`, `OM [%]`, `OM dens [g/cm**3]`, `Corg dens [g/cm**3]`,
         `TC [%]`, `TOC [%]`, `C [%]`, `LOI [%]`, `LOI [%] (weight-%)`, `TIC [%]`) %>% 
    dplyr::full_join(CPEAT.ls$study,
                     by = dplyr::join_by(doi)) %>% 
    dplyr::group_by(doi) %>%
    dplyr::mutate(row_number = row_number()) %>% 
    dplyr::ungroup() %>%
    tidyr::pivot_longer(cols = -row_number, 
                        names_to = 'column_id', 
                        values_to = 'with_entry', 
                        values_drop_na = TRUE) %>% 
    left_join(CPEAT.ls$annotations, 
              by = join_by(column_id),
              suffix = c('.data', ''),
              relationship = "many-to-many") %>% 
    mutate(with_entry = dplyr::if_else((with_entry == "--") | is.na(with_entry),
                                  with_entry.data, with_entry)) %>%
    select(-with_entry.data) %>% 
    select(table_id, column_id, of_variable, is_type = of_type, with_entry, row_number) # 5,546,485 obs, 6 total columns


```

## Casting values

Finally make sure that the numbers are actually numbers.

```{r}

soil_carbon.df <- CPEAT.ls$core %>% 
  select(doi, 
         contains('Depth sed'),
         #`Depth sed [m]`, 
         # `Depth sed [m] (Sample midpoint depth)`, 
         # `Depth sed [m] (Sample midpoint depth (rounded))`, 
         # `Depth sed [m] (Sample midpoint depth of OxCa...)`, 
         # `Depth sed [m] (Sample midpoint depth of Baco...)`,
         # `Depth sed [m] (LOI sample depth)`,
         `Depth top [m]`, `Depth bot [m]`, 
         `Samp thick [cm]`, `DBD [g/cm**3]`, `OM [%]`, `OM dens [g/cm**3]`, `Corg dens [g/cm**3]`,
         `TC [%]`, `TOC [%]`, `C [%]`, `LOI [%]`, `LOI [%] (weight-%)`, `TIC [%]`) %>% 
 dplyr::mutate(across(-doi, as.numeric))
  
soil_carbon.df %>%
  slice_head(n=100) %>%
  knitr::kable() %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "300px")

```

## Make small tables

Often this data is presented in three tables: a site level table with lat/lon and year, a layer level table with the depth and physiochemical information, and a meta table with information on the units and descriptions.
We given an example below of how this might be organized.

```{r}
site.df <- CPEAT.ls$study %>%
  select(doi, core_name, LATITUDE, LONGITUDE, LOCATION, ELEVATION, ELEVATION.START, ELEVATION.END) %>%
  unique()

layer.df <- soil_carbon.df %>%
  left_join(CPEAT.ls$study %>% 
              select(doi, core_name, citation),
            by = join_by(doi))

meta.df <- CPEAT.ls$annotations %>%
  #identify the variables that we are interested in
  filter(of_variable %in% c('data_doi',
             'site', 'longtitude', 'latitude', 'location',
             'layer_center', 'layer_thickness', 'layer_midpoint', 'layer_depth_top', 'layer_depth_bottom',
             'bulk_density',
             'organic_matter_fraction', 'organic_matter_density', 'organic_carbon_density',
             'total_carbon', 'total_organic_carbon', 'loss_on _ignition', 'percent_carbon',
             'depth_loi', 'LOI', 'Total_inorganic_carbon', 'depth'),
         table_id %in% c('study', 'core'))

```