---
title: "Data Rescue: Qasemian 2014"
format:
  html:
    toc: true
    eval: true 
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - Qasemian2014_Methods.bib #contains the citations in the methods section
  - Qasemian2014.bib #contains citation for the data source, generally a single citation of the primary paper
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  #remove ktoddbrown if they did not work on the rescue directly
  - id: ktoddbrown #use the github name as an id
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role: #TODO define these roles and add reference
      - transcription: review
    #  - standardization: lead #review
     # - curation: lead #review
  - id: SavaScott
    name:
      given: Savannah
      family: Scott
    affiliation:
      - ref: uf-ees
    role:
      - transcription: lead
    #  - standardization: lead #review
     # - curation: lead #review
  - id: JennavieveK
    name:
      given: Jenna
      family: Keefer
    affiliation:
      - ref: uf
    role:
      - transcription: secondary
    #  - standardization: lead #review
     # - curation: lead #review
  - id: sierrav-droid
    name:
      given: Sierra
      family: Velazquez
    affiliation:
      - ref: uf
    role:
      - transcription: secondary
    #  - standardization: lead #review
     # - curation: lead #review
    
affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
  - id: uf
    name: University of Florida
    city: Gainesville
    state: Florida
    country: USA
    postal-code: 32611
    url: https://www.ufl.edu/
    
timelog:
  - activity: 
      description: First Transcription
      who: SavaScott #github username
      #when you started and ended, can be just a year or down to the minute
      end: 2026-08-21
      start: 2026-07-16
      duration: PT3H
  - activity: 
      description: ReadMe Update and Bib files
      who: SavaScott #github username
      #when you started and ended, can be just a year or down to the minute
      end: 2026-09-12
      start: 2026-08-21
      duration: PT1H30M
  - activity: 
      description: Coding of Figures and Methods CC
      who: SavaScott #github username
      #when you started and ended, can be just a year or down to the minute
      end: 2026-09-12
      start: 2026-09-01
      duration: PT2H30M
  - activity: 
      description: Figures and and Tables CC
      who: SavaScott #github username
      #when you started and ended, can be just a year or down to the minute
      end: 2026-09-12
      start: 2026-09-12
      duration: PT1H
  - activity: 
      description: Second Transcription of Figures and Tables
      who: JennavieveK #github username
      #when you started and ended, can be just a year or down to the minute
      end: 2026-09-11
      start: 2026-09-1
      duration: PT3H
  - activity: 
      description: Second Transcription of Methods
      who: sierrav-droid #github username
      #when you started and ended, can be just a year or down to the minute
      end: 2026-07-24
      start: 2026-07-21
      duration: PT45M
---


```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false
#| eval: true


library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# data files for level 0
methods.file <- "Qasemian2014_Methods.md"
table1.file <- 'Qasemian2014_Table1.csv'
figure2.file <- 'Qasemian2014_Figure2.csv'
figure3.file <- 'Qasemian2014_Figure3.csv'
figure4.file <- 'Qasemian2014_Figure4.csv'

# Bibliography files
primaryCitation.file <- 'Qasemian2014.bib'
methodsCitation.file <- 'Qasemian2014_Methods.bib'

readsource.file <- '../../02_DataHarmonization/readQasemian2014.R'
if(file.exists(readsource.file)){
  source(readsource.file)
}

#semantic files for level 1
of_variable.file <- '../../semantics/SoilDRaHVocabulary_of_variable.csv'
is_type.file <- '../../semantics/SoilDRaHVocabulary_is_type.csv'

```


# Data Summary

> Qasemian, L., Guiral, D., & Farnet, A. M. (2014). How do microlocal environmental variations affect microbial activities of a Pinus halepensis litter in a Mediterranean coastal area?. Science of the Total Environment, 496, 198-205. https://doi.org/10.1016/j.scitotenv.2014.07.054



For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/280.

@Qasemian2014 primarily measures microbial activities that occurred in a *Pinus halepensis* litter collected from two areas on the French Mediterranean coast.
They use this data to look at how the microbial indicators changed due to proximity to the sea, finding that they intensified when they were moved closer to the coast.

## Tables and figures

- Tables 
  1) Minimum and maximum of the environmental parameters monitored in litters over a one-year period in C10 (10 m from the sea shore) and C300 (300 m from the sea shore) areas
  2) Three-way ANOVA test on enzyme activities

- Figures 
  1) Difference in i) temperature (°C), ii) humidity (percentage), iii) electrical conductivity (μS cm$^{−1}$) and iv) chloride concentration (ppm) measured in *P. halepensis* litters
  2) C:N ratio (a) and Aromaticity ratio (b) calculated from litterbags after 2 (June), 5 (September), and 13 (May) months in litterbags
  3) Temporal variations in microbial enzyme activities in Pinus halepensis litters measured after 2 (June), 5 (September), and 13 (May) months in litterbags.
  4) Basal Respiration (a), Metabolic quotient qCO$_2$ (b) and AWCD calculated from Biolog after 45 h after 2 (June), 5 (September), and 13 (May) months.
  5) PCA from Eco plate substrate utilization profile of litterbags after 2, 5 and 13 months. 
  6) PCA on microbial markers measured in litterbags after 2, 5 and 13 months. 

## Fit for purpose: DeAD

This data is identified as a data source for DeAD
The Decomposition Across Drylands (DeAD) data collection is a new data effort targeting dryland decomposition studies.
The purpose of this data collection is to develop better litter decay models for dryland ecosystems. 

### Main Variables of Interest

- Location: Study sites were from the French Mediterranean coast (Massif de Marseilleveyre, Marseille; 43° 12′ 34″ N, 5° 21′ 36″ E), 10 m and 300 m from the coast. This information is found with 2.1. Litter sampling.
- Soil carbon stock: The C:N ratio and the aromaticity ratio is found within Figure 2. 
- Duration: The litterbags were tested after 2 months (June 2009), 5 months (September 2009), and 13 (May 2010) months after being placed in situ
- Mass loss: 2.1. Litter Sampling mentions that they measured the litter mass loss, but I cannot determine if or where it is recorded in the paper.
- Solar exposure: Solar exposure was not measured in this paper.

### Other Relevant Variables 

  - Litter species: *Pinus halepensis*
  - Microsite data: Maximum and minimum of the environmental parameters that the litters were exposed to are found within Table 1, including temperature and EC. 
  - Litter chemical composition: Figure 2 has the C:N and Aromaticity ratio calculated from the litterbags. Figure 4 may also have some litter chemical compositions needed.
  - Extracellular enzyme: Figure 3 has data on the microbial enzyme activities of the litterbags.
  - Microclimate relative humidity: Humidity percentage (max and min) were included in Table 1. 
  - Litter bag materials and method: The litter was sieved through a >2 mm mesh size before being placed in litterbags, 40 g per bag. The litterbags were 20 $\times$ 25 cm, made of polypropylene with a 1 mm mesh size. 
  - Treatments: No treatments were mentioned in the methods section

## Files

These files are in the Qasemian2014 data rescue.

- [Readme](README_Qasemian2014.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Qasemian2014.bib](Qasemian2014.bib)
  + citation for article transcribed
- [Qasemian2014_Methods.bib](Qasemian2014_Methods.bib)
  + citations for the methods section of the article
- [Qasemian2014_Methods.md](Qasemian2014_Methods.md)
  + methods transcribed from primary article
- [Qasemian2014_Table1.csv](Qasemian2014_Table1.csv)
  + table 1 from primary article with minimum and maximum of the site conditions
- [Qasemian2014_Figure2.csv](Qasemian2014_Figure2.csv)
  + figure 2 from primary article with C:N ratio and Aromaticity ratio
- [Qasemian2014_Figure3.csv](Qasemian2014_Figure3.csv)
  + figure 3 from primary article with temporal variations in the microbial enzyme activities
- [Qasemian2014_Figure4.csv](Qasemian2014_Figure4.csv)
  + figure 4 from primary article with basal respiration, metabolic quotient and AWCD
- [temp/](temp/)
    + scratch folder that will not be archived on GitHub, it should include local copies of the data resource (likely a PDF article) and any spreadsheets or other tools you used for the transcriptions.

# Data Rescue working notes 

## Data Rescue Plan

+ Lead
  - [x] Read the article or data resource and the discription of the target project
  - [x] Create a fresh fork of the primary repository https://github.com/ktoddbrown/SoilDRaH/tree/main, it might be useful to rename it 'SoilDRaH_Qasemian2014'
  - [x] Create a copy of this template folder and rename it to Qasemian2014
  - [x] Fill out Data summary section (except for the Files), the Data Rescue Plan (here!), and update the author section of the yml to reflect who is working on the rescue
  - [x] Create a GitHub issue from the Data Rescue template and get feedback from a reviewer on the project (Dr Todd-Brown or a grad student typically).
  - [x] Add the reviewer to the author list in the yml
+ Do
  - [x] Create a spreadsheet with Table1 and export it to csv
  - [x] Extract points from Figure2-4 and export it to csv
  - [x] Transcribe or copy the method section
  - [x] Insert references to bib entries
  - [x] reformat equations and mathematical symbols as [LaTex](https://en.wikibooks.org/wiki/LaTeX/Mathematics)
  - [x] reformat text using [Markdown](https://www.markdownguide.org/cheat-sheet/)
  - [x] Create a bib file with the citations from the methods section
  - [x] Push to your local fork and get a second transcription pushed to your local fork by someone else on the project
+ Measure
  - [x] cross check with second transcriber for tables and methods, and resolve any differences
  - [x] compile or render this ReadMe with all chunks up through the Level 0 data rescue section with `eval: true`
  - [ ] push to your local fork and create a pull request to the main branch of the primary repository
  - [ ] address any revisions in your pull request review
  - [ ] celebrate!

## Figure 2

Figure 2 was modified during transcription to include "Months surveyed" as a x-axis title.

```{r}
#| label: figure2Check
#| code-fold: true
#| eval: true

fig2Caption <- read_csv(file = figure2.file, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption

fig2Data <- read_csv(file = figure2.file,
                   skip = 1,
                   show_col_types = FALSE)

fig2DataCN <- fig2Data %>% filter(`y label` == "C:N ratio")
fig2DataAromaticity <- fig2Data %>% filter(`y label` == "Aromaticity Ratio")

# C:N ratio column graph

fig2DataCN$`x label` <- factor(fig2DataCN$`x label`,
                                  level = c("June 2009", "September 2009", "May 2010"))
fig2DataCN$`litterbag type` <- factor(fig2DataCN$`litterbag type`,
                                  level = c("C10/C10", "C10/C300", "C300/C300", "C300/C10"))

fig2DataCN$`y high SE value` <- as.numeric(fig2DataCN$`y high SE value`)
fig2DataCN <- fig2DataCN %>%
  mutate(error_var = `y high SE value` - `y value`,
         ymin = `y value` - error_var,
         ymax = `y value` + error_var)

ggplot(fig2DataCN, aes(x = `x label`, y = `y value`, fill = `litterbag type`, group = `litterbag type`)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) + 
  theme_classic() +
  scale_fill_manual(values = c("bisque3", "azure2", "azure4", "bisque1")) +
  geom_text(aes(label = footnotes),
    position = position_dodge(width = 0.9),
    vjust = -1.95, size = 2, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), 
    position = position_dodge(width = 0.9), 
    width = 0.3, color = "black", 
    linewidth = 0.3) + labs(
       x = "Month Surveyed",
       y = "C:N ratio",
       caption = fig2Caption)
  
# Aromaticity Ratio column graph

fig2DataAromaticity$`x label` <- factor(fig2DataAromaticity$`x label`,
                                  level = c("June 2009", "September 2009", "May 2010"))
fig2DataAromaticity$`litterbag type` <- factor(fig2DataAromaticity$`litterbag type`,
                                  level = c("C10/C10", "C10/C300", "C300/C300", "C300/C10"))

fig2DataAromaticity$`y high SE value` <- as.numeric(fig2DataAromaticity$`y high SE value`)
fig2DataAromaticity <- fig2DataAromaticity %>%
  mutate(error_var = `y high SE value` - `y value`,
         ymin = `y value` - error_var,
         ymax = `y value` + error_var)

ggplot(fig2DataAromaticity, aes(x = `x label`, y = `y value`, fill = `litterbag type`, group = `litterbag type`)) +
  scale_y_continuous(breaks = c(0.1, 0.12, 0.14, 0.16, 0.18, 0.2, 0.22, 0.24, 0.26, 0.28, 0.30)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) + 
  theme_classic() +
  scale_fill_manual(values = c("bisque3", "azure2", "azure4", "bisque1")) +
  geom_text(aes(label = footnotes),
    position = position_dodge(width = 1),
    vjust = -1.75, size = 2.5, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), 
    position = position_dodge(width = 0.9), 
    width = 0.3, color = "black", 
    linewidth = 0.25) + labs(
       x = "Month Surveyed",
       y = "Aromaticity Ratio",
       caption = fig2Caption)
  

```

And is crosschecked with the secondary transcription (expected to be empty)

```{bash}
#| label: compareFigure2
#| eval: false

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.
#... Note that the template should not have a conflict here.

diff --strip-trailing-cr --suppress-common-lines -y temp/Qasemian2014_Figure2_JK.csv Qasemian2014_Figure2.csv
```


## Figure 3

Figure 3 was modified during transcription to include "Months surveyed" as a x-axis title.

```{r}
#| label: figure3Check
#| code-fold: true
#| eval: true

fig3Caption <- read_csv(file = figure3.file, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption

fig3Data <- read_csv(file = figure3.file,
                   skip = 1,
                   show_col_types = FALSE)

fig3Data$`x label` <- factor(fig3Data$`x label`,
                                  level = c("June 2009", "September 2009", "May 2010"))
fig3Data$`litterbag type` <- factor(fig3Data$`litterbag type`,
                                  level = c("C10/C10", "C10/C300", "C300/C300", "C300/C10"))

fig3Data$`high SE y value` <- as.numeric(fig3Data$`high SE y value`)
fig3Data <- fig3Data %>%
  mutate(error_var = `high SE y value` - `U gdW$^{-1}$`,
         ymin = `U gdW$^{-1}$` - error_var,
         ymax = `U gdW$^{-1}$` + error_var)

fig3DataLaccase <- fig3Data %>% filter(`graph label` == "a) Laccase")
fig3DataCellulase <- fig3Data %>% filter(`graph label` == "b) Cellulase")
fig3DataBgluc <- fig3Data %>% filter(`graph label` == "c) β-glucosidase")
fig3DataPhos <- fig3Data %>% filter(`graph label` == "d) Phosphatase")

# Figure a) Laccase

ggplot(fig3DataLaccase, aes(x = `x label`, y = `U gdW$^{-1}$`, fill = `litterbag type`, group = `litterbag type`)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) + 
  theme_classic() + 
  scale_y_continuous(breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0)) +
  scale_fill_manual(values = c("bisque3", "azure2", "azure4", "bisque1")) +
  geom_text(aes(label = footnotes),
    position = position_dodge(width = 0.70),
    vjust = -1.4, size = 2, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), 
    position = position_dodge(width = 0.9), 
    width = 0.3, color = "black", 
    linewidth = 0.3) + labs(
       x = "Month Surveyed",
       y = "U gdW${-1}$",
       caption = fig3Caption,
       title = "a) Laccase")

# Figure b) Cellulase 

ggplot(fig3DataCellulase, aes(x = `x label`, y = `U gdW$^{-1}$`, fill = `litterbag type`, group = `litterbag type`)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) + 
  theme_classic() + 
  scale_y_continuous(breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8)) +
  scale_fill_manual(values = c("bisque3", "azure2", "azure4", "bisque1")) +
  geom_text(aes(label = footnotes),
    position = position_dodge(width = 0.75),
    vjust = -2.0, size = 2.2, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), 
    position = position_dodge(width = 0.9), 
    width = 0.3, color = "black", 
    linewidth = 0.3) + labs(
       x = "Month Surveyed",
       y = "U gdW${-1}$",
       caption = fig3Caption,
       title = "b) Cellulase")

# Figure c) β-glucosidase

ggplot(fig3DataBgluc, aes(x = `x label`, y = `U gdW$^{-1}$`, fill = `litterbag type`, group = `litterbag type`)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) + 
  theme_classic() + 
  scale_y_continuous(breaks = c(0, 0.1,  0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0)) +
  scale_fill_manual(values = c("bisque3", "azure2", "azure4", "bisque1")) +
  geom_text(aes(label = footnotes),
    position = position_dodge(width = 0.75),
    vjust = -2.5, size = 2.4, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), 
    position = position_dodge(width = 0.9), 
    width = 0.3, color = "black", 
    linewidth = 0.3) + labs(
       x = "Month Surveyed",
       y = "U gdW${-1}$",
       caption = fig3Caption,
       title = "c) β-glucosidase")

# Figure d) Phosphatase

ggplot(fig3DataBgluc, aes(x = `x label`, y = `U gdW$^{-1}$`, fill = `litterbag type`, group = `litterbag type`)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) + 
  theme_classic() + 
  scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2, 2.5)) +
  scale_fill_manual(values = c("bisque3", "azure2", "azure4", "bisque1")) +
  geom_text(aes(label = footnotes),
    position = position_dodge(width = 0.75),
    vjust = -2.5, size = 2.4, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), 
    position = position_dodge(width = 0.9), 
    width = 0.3, color = "black", 
    linewidth = 0.3) + labs(
       x = "Month Surveyed",
       y = "U gdW${-1}$",
       caption = fig3Caption,
       title = "d) Phosphatase")

```

And is crosschecked with the secondary transcription (expected to be empty)

```{bash}
#| label: compareFigure3
#| eval: false

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.
#... Note that the template should not have a conflict here.

diff --strip-trailing-cr --suppress-common-lines -y temp/Qasemian2014_Figure3_JK.csv Qasemian2014_Figure3.csv
```


## Figure 4

Figure 4 was modified during transcription to include "Months surveyed" as a x-axis title.

```{r}
#| label: figure4Check
#| code-fold: true
#| eval: true

#need to fix, y-axis labels are different for each graph

fig4Caption <- read_csv(file = figure4.file, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption

fig4Data <- read_csv(file = figure4.file,
                   skip = 1,
                   show_col_types = FALSE)

fig4Data$`x label` <- factor(fig4Data$`x label`,
                                  level = c("June 2009", "September 2009", "May 2010"))
fig4Data$`litterbag type` <- factor(fig4Data$`litterbag type`,
                                  level = c("C10/C10", "C10/C300", "C300/C300", "C300/C10"))

fig4Data$`SE high y value` <- as.numeric(fig4Data$`SE high y value`)
fig4Data <- fig4Data %>%
  mutate(error_var = `SE high y value` - `y value`,
         ymin = `y value` - error_var,
         ymax = `y value` + error_var)

fig4DataA <- fig4Data %>% filter(`graph label` == "a")
fig4DataB <- fig4Data %>% filter(`graph label` == "b")
fig4DataC <- fig4Data %>% filter(`graph label` == "c")

# Figure a

ggplot(fig4DataA, aes(x = `x label`, y = `y value`, fill = `litterbag type`, group = `litterbag type`)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) + 
  theme_classic() + 
  scale_y_continuous(breaks = c(0, 20, 40, 60, 80, 100, 120)) +
  scale_fill_manual(values = c("bisque3", "azure2", "azure4", "bisque1")) +
  geom_text(aes(label = footnotes),
    position = position_dodge(width = 0.70),
    vjust = -1.4, size = 2, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), 
    position = position_dodge(width = 0.9), 
    width = 0.3, color = "black", 
    linewidth = 0.3) + labs(
       x = "Month Surveyed",
       y = "μg CO2-C gdw$^{-1}$",
       caption = fig4Caption,
       title = "a")

# Figure b

ggplot(fig4DataB, aes(x = `x label`, y = `y value`, fill = `litterbag type`, group = `litterbag type`)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) + 
  theme_classic() + 
  scale_y_continuous(breaks = c(0, 1, 2, 3, 4, 5, 6, 7, 8)) +
  scale_fill_manual(values = c("bisque3", "azure2", "azure4", "bisque1")) +
  geom_text(aes(label = footnotes),
    position = position_dodge(width = 0.70),
    vjust = -1.9, size = 2, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), 
    position = position_dodge(width = 0.9), 
    width = 0.3, color = "black", 
    linewidth = 0.3) + labs(
       x = "Month Surveyed",
       y = "μg CO2-C gdw$^{-1}$",
       caption = fig4Caption,
       title = "b")

# Figure c

ggplot(fig4DataC, aes(x = `x label`, y = `y value`, fill = `litterbag type`, group = `litterbag type`)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) + 
  theme_classic() + 
  scale_y_continuous(breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1.0, 1.2)) +
  scale_fill_manual(values = c("bisque3", "azure2", "azure4", "bisque1")) +
  geom_text(aes(label = footnotes),
    position = position_dodge(width = 0.70),
    vjust = -2.2, size = 2, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), 
    position = position_dodge(width = 0.9), 
    width = 0.3, color = "black", 
    linewidth = 0.3) + labs(
       x = "Month Surveyed",
       y = "μg CO2-C gdw$^{-1}$",
       caption = fig4Caption,
       title = "c")

```

And is crosschecked with the secondary transcription (expected to be empty)

```{bash}
#| label: compareFigure4
#| eval: false

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.
#... Note that the template should not have a conflict here.

diff --strip-trailing-cr --suppress-common-lines -y temp/Qasemian2014_Figure4_JK.csv Qasemian2014_Figure4.csv
```


## Table 1

<!--- Common table modifications include: inserting an extra column to pivot sub-headers, filling in NA to identify blanks, and adding Latex or Markdown to denote formatting.---->

Table 1 was modified by adding in dashes into empty columns and combining a subcolumn into the main column headers. 

```{r}
#| label: table1Check
#| code-fold: false
#| eval: true


read_csv(file = table1.file, 
         skip = 1, 
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = table1.file, 
                           n_max = 1, 
                           col_names = 'caption',
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE
                           )$caption
  )
```

And is crosschecked with the secondary transcription (expected to be empty)

```{bash}
#| label: compareTable1
#| eval: false

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.
#... Note that the template has an example of an unresolved conflict.

diff --strip-trailing-cr --suppress-common-lines -y temp/Qasemian2014_Table1_JK.csv Qasemian2014_Table1.csv
```

## Methods

Any different to the cross check for the methods section will be output after this line.

```{bash}
#| label: compareMethods
#| eval: false
# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.
#... Note the template example does NOT have a conflict.

diff --strip-trailing-cr --suppress-common-lines -y temp/Qasemian2014_Methods_SV.md Qasemian2014_Methods.md
```

### Methods  - Text

The final text transcription from the Methods section is rendered below.

```{r}
#| label: readMethods
#| output: asis

# This revised section will append two ### to any existing headers here so that it renders nicely with the other headers in the document.

methodsLines <- readLines(methods.file) |>
  str_replace_all(pattern = '^#', replacement = '####')

cat(paste0(methodsLines, collapse = '\n'))

```

<!----
## Citation notes from Methods

Below are the citation notes from the methods section.
This generally doesn't read well in the final report so commenting it out here but it is very useful for tracking citations when you are doing the transcriptions.
Some citations are missing, those are noted here and how best guess were generated.
Think of this as a lab notebook documenting how you found each citation.

- [ ] short citation
  + copy from bib
  + manually entered or direct export
  
- [x] Beare, M. H., Neely, C. L., Coleman, D. C., & Hargrove, W. L. (1990). A substrate-induced respiration (SIR) method for measurement of fungal and bacterial biomass on plant residues. Soil Biology and Biochemistry, 22(5), 585-594.
  + copy from bib

- [x] Anderson, T. H., & Domsch, K. H. (1985). Determination of ecophysiological maintenance carbon requirements of soil microorganisms in a dormant state. Biology and fertility of soils, 1(2), 81-89.
  + copy from bib

- [x] Criquet, S., Tagger, S., Vogt, G., Iacazio, G., & Le Petit, J. (1999). Laccase activity of forest litter. Soil Biology and Biochemistry, 31(9), 1239-1244.
  + copy from bib

- [x] Farnet, A. M., Qasemian, L., Guiral, D., & Ferré, E. (2010). A modified method based on arsenomolybdate complex to quantify cellulase activities: application to litters. Pedobiologia, 53(2), 159-160.
  + copy from bib
  
- [x] Garland, J. L. (1997). Analysis and interpretation of community-level physiological profiles in microbial ecology. FEMS microbiology ecology, 24(4), 289-300.
  + copy from bib
  
- [x] Mathers, N. J., Jalota, R. K., Dalal, R. C., & Boyd, S. E. (2007). 13C-NMR analysis of decomposing litter and fine roots in the semi-arid Mulga Lands of southern Queensland. Soil Biology and Biochemistry, 39(5), 993-1006.
  + copy from bib
  
- [x] Massiot, D., Fayon, F., Capron, M., King, I., Le Calvé, S., Alonso, B., ... & Hoatson, G. (2002). Modelling one‐and two‐dimensional solid‐state NMR spectra. Magnetic resonance in chemistry, 40(1), 70-76.
  + copy from bib
---->

# Level 0 Data rescue

This level 0 data read parses the transcribed files with minimal cleaning or transformations.

```{r}
#| label: readLevel0
#| eval: true
#This chunk has two purposes. 
#...1) Check the formatting by reading in everything 
#...2) Create a list of everything to process later in level 1

data.lvl0.ls <- list(
  #Read in a list of all the bib files
  citation = list(
    #Citation for the article transcriptions are pulled from
    primary = read.bib(file = primaryCitation.file), 
    #Citations for all referenced articles
    methods = read.bib(file = methodsCitation.file)
  ),
  #Read in the text transcription of the article's methods section
  method = read_lines(file = methods.file),
  #Read in the results as tables or figure transcriptions. This includes
  #...the caption as well as the tables themselves
  data = list(
    Table1 = list(
      #Read the caption as a text string. Captions are the first cell on 
      #...the first row.
      caption = read_csv(file = table1.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      #Read in all the data, skipping the first row with the caption and read
      #...in the table as character. This element is a tibble (data.frame).
      primary = read_csv(file = table1.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    ), #End 'Table1' element
    Figure2 = list(
      caption = read_csv(file = figure2.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      primary = read_csv(file = figure2.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    ), 
    Figure3 = list(
      caption = read_csv(file = figure3.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      primary = read_csv(file = figure3.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    ),
        Figure4 = list(
      caption = read_csv(file = figure4.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      primary = read_csv(file = figure4.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    )
    #Repeat for all tables and figures...
   ) #end 'data' element
)

```

The code is also in `readsource.file` and returned with the `dataLevel = 'level0'` argument.
Below verifies that the data from the function matches what is coded above.

```{r}
#| label: checkReadFunction_lvl0
#| eval: true

if(file.exists(readsource.file)){
  datafunction.lvl0.ls <- readAuthorYYYY(dataDir = '.',
                                         dataLevel = 'level0')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function matches code here.')
  }else{
    print('There is a mismatch in the data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}
```

# Level 1 data - CURATION PROJECT

This data is identified as a data source for the ...
Broad classes of information that we are looking for include:

<!--- Include description of where the data is located in the source.--->
- Location: 
- Soil carbon stock: 
- Soil type: 
- Land use: 

Information from the methods and other sections are pulled into tables with line by line source citations.
This information can be values, units, methods, or other useful notes and is manually extracted from the transcriptions loaded into the level 0 data above.

The final format of the data frame here will include the following 

  + grouping ids
  + `of_variable`
  + `is_type`
  + `with_entry`
  + `from_source`

The `of_variable` being used here are:

 + citation
 
 <!--- include other variables being used, this should match vocabulary in the SoilDRaHVocabulary resources --->

```{r}
#| label: createLvl1
#| eval: false

#### Method-section only data ####
#Create a table for the data in the method-section that is not directly 
#...associated with any table. Often this is study level information.

studyMeta <- tribble(~of_variable, ~is_type, ~with_entry, ~from_source,
                   'region', 'value', 'Honaunau Forest on the southwestern slopes of Mauna Loa, island of Hawaii', paste('Method ln5:', paste(data.lvl0.ls$method[5], collapse = ' ')))

#### Table/Figure N ####
# Process the primary data from Table/Figure M by maintaining data groups and 
#...standardizing the variable name so that we can link this to the study 
#...metadata and data in the other tables when needed. Link in the methods 
#...and other information from the method-section as needed.

Table1Primary <- data.lvl0.ls$data$Table1$primary |>
  #Add a row id preserving row grouping -- not unique
  mutate(row_id = paste0('R', 1:n())) |>
  #Transform the table to long data by row group and column names.
  #...This allows us to then create the cross links between the tables.
  pivot_longer(cols = -row_id,
               names_to = 'column_name', values_to = 'with_entry',
               values_drop_na = TRUE) |> #drop missing values
  #Link the variable names to the column names using a switch statement
  mutate(of_variable = case_when(
           #Soil PH is given as an example below
           column_name == 'SOC' ~ 'soil_organic_carbon')) |>
  #Flag the source of this data as Table 1, this table is merged later with
  #...other data so this keeps track of where the data came from
  mutate(from_source = 'Table 1')

## Create the metadata for table one from the header information and cited methods
Table1Meta <- Table1Primary |>
  select(column_name, of_variable) |>
  unique() |>
  #Grab everything between the parentheses as units and attribute the source as the column names.
  mutate(unit = str_extract(column_name, pattern = '(?<=\\().*(?=\\))'),
         from_source = 'Table 1 column names.') |>
  #If there aren't units then drop the row
  filter(!is.na(unit)) |>
  #Don't keep the column name now that you have the units, cross link via the variable
  select(-column_name) |>
  #Stack into the table the methods for each variable that we can find
  bind_rows(
  tribble(~of_variable, ~method, ~from_source,
            'soil_organic_carbon', paste0(data.lvl0.ls$method[72:73], collapse = ' '), 'Methods ln72-73') ) |>
  #Stack onto the tables the controlled vocabulary used
  bind_rows(
    tribble(~of_variable, ~control_vocabulary, ~from_source,
            'soil_class', 'Histosol:USDA classification for histosol soil type|Andisols:USDA classification for andisol soil type', 'expert informed')
  ) |>
  # Push the dimensions of the variable into a single column. Making this a long table.
  pivot_longer(cols = c(unit, method, control_vocabulary),
               names_to = 'is_type',
               values_drop_na = TRUE,
               values_to = 'with_entry')

#### Create level 1 #####
## Pull everything together into by stacking the meta and primary data tables
data.lvl1.ls <- list(
  meta = bind_rows(studyMeta, Table1Meta),
  primary = bind_rows(Table1Primary)|>
    arrange(row_id, column_name,
            of_variable, is_type, with_entry, from_source)
  )

```

## Check for new variables

`of_variable` in the data files should be defined in the SoilDRaH Vocabulary files.

```{r}
#| label: check_of_variable
#| eval: false

#Read in the list of current 'of_variable' entries
SoilDRaH_variable_list <- read_csv(file =of_variable.file,
                                   show_col_types = FALSE)

#compare the variables in the level 1 tables to the SoilDRaH variables
new_variables <- lapply(data.lvl1.ls, function(xx){
  setdiff(unique(xx$of_variable), 
          unique(SoilDRaH_variable_list$of_variable))
}) |>
  unlist()

#Check that there are no new variables
if(length(new_variables) == 0){
  print('All of_variables are in the SoilDRaH variable list. No additions required')
}else{
  cat(paste('Please start a issue for new variables [',
               new_variables, ']', sep = '', collapse = '\n'))
}
```

## Check for is_type

`is_type` in the data files should be defined in the SoilDRaH Vocabulary files.

```{r}
#| label: check_is_type
#| eval: false

#read in the current types and what variable they are associated with
SoilDRaH_type_list <- read_csv(file = is_type.file,
                                   show_col_types = FALSE) |>
  #remove the project definition and only keep the variable and type
  select(of_variable, is_type) |>
  #variables are pushed together in the type defined table. Separate the
  #...entries here
  separate_longer_delim(cols = 'of_variable', delim = ',')

#Pull all the of_variable-is_type pairs from the source data
source_pairs <- tibble()

#If there are a large number of entries, consider replacing this for-loop.
#...However this avoids having to install the plyr or other libraries, and 
#...there are generally no more then 3-5 tables
for(dataTable in data.lvl1.ls){
  source_pairs <- dataTable |>
    select(of_variable, is_type) |> 
    unique() |>
    bind_rows(source_pairs)
}

undefined_data_type <- source_pairs |>
  #subset for variables that are not 'incomplete'
  semi_join(SoilDRaH_variable_list |>
              filter(project_status != 'incomplete'),
            by = join_by(of_variable)) |>
  #check for of_variable-is_type tuples that are not in the SoilDRaH type definitions
  anti_join(SoilDRaH_type_list,
            by = join_by(of_variable, is_type))

#Tell the user what you found.
if(nrow(undefined_data_type) == 0){
  print('All of_variables-is_types are in the SoilDRaH type list. No additions required')
}else{
  print(paste0('Please start a issue for new of_variable-is_type [', paste(paste(undefined_data_type$of_variable, undefined_data_type$is_type, sep = '-'), collapse = ','), ']'))
}
```

## Check function match

Check for an exact match with the read function.

```{r}
#| label: checkReadFunction_lvl1

if(file.exists(readsource.file)){
  datafunction.lvl1.ls <- readAuthorYYYY(dataDir = '.',
                                         dataLevel = 'level1')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function output matches ReadMe code for level 1.')
  }else{
    print('There is a mismatch in the level 1 data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}

```


# References

<!---- This space will be filled with the citations when rendered. Leave this blank. ---->
