---
title: "Data Curation: ISCN 2025"
format:
  html:
    toc: true
    eval: false #default to not evaluating the chunks, override this the template is completed
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - ../../01_DataRescue/CPEAT2025/CPEAT2025.bib #CPEAT citation
  - ../../01_DataRescue/CPEAT2025/CPEAT2025_Methods.bib #CPEAT source
  - ../../01_DataRescue/ISCN3_2015/ISCN3_2015.bib #ISCN3 citation
  - ../../01_DataRescue/ISCN3_2015/ISCN3_Contributors.bib #ISCN3 sources
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  - id: ktoddbrown #use the github name as an id
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role: #TODO define these roles and add reference
      #- transcription: lead #secondary, review
      - standardization: lead #review
      - curation: lead #review

affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
timelog:
    - activity: 
      description: Sort through migration conflicts for issue 191
      who: ktoddbrown #github username
      #when you started and ended, can be just a year or down to the minute
      start: 2026-02-05T14:15
      end: 2026-02-05T14:45
      #alternatively duration can be combined with a start or end
      #duration: P1H15M
  - activity: 
      description: Draft inital readme
      who: ktoddbrown #github username
      #when you started and ended, can be just a year or down to the minute
      start: 2025-11-27T11:00
      end: 2025-11-27T11:38
      #alternatively duration can be combined with a start or end
      #duration: P1H15M
---


```{r}
#| label: setup


library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# Bibliograph files
refernceCitation.file <- 'ISCN2025.bib'

source('../../02_DataHarmonization/readCPEAT2025.R')
source('../../02_DataHarmonization/readISCN3.R')
```

# Purpose

The purpose ISCN 2025 (aka ISCN4) is to create a data product aggregating data on geo-located soil organic carbon profiles.
With this data we will examine the assumption of steady state on field surveyed organic carbon.
We have closed contributions for this project and are working on analysis.

# Variables of interest

<!--- Briefly describe what you want the data rescuers to look for when digitizing a data resource.---->

For this analysis we are interested in:

  * **Location**: latitude/longitude, associated datum, geo-spatial method, precision
  * **Soil organic carbon stocks**: organic carbon fraction, bulk density, coarse fraction, soil organic carbon by area, soil organic carbon by volume
  * **Layer depth**: Depth of surveyed soil layer

# Data sources

For this project, we have identified the following data sources by US federal monitoring databases and direct network contributions.
These include:

+ CPEAT (@Loisel2021, @CPEAT_web)
  - This is a collection of peat cores (@PCPEAT)
  - Status: Level 0 complete. Finalizing Level 1
+ FIA
  - This is a US Forest Service database.
  - Status: Reprocessing into the new repository structure
+ NCSS
  - This is a US NRCS database.
  - Status: Reprocessing into the new repository structure
+ ISCN3 (@ISCN3)
  - This database was originally compiled in 2015, we want to update the old NCSS data.
  - Status: Level 0 and 1 are complete



# Data Level 1 

```{r}
#| label: readLevel1

ISCN3.ls <- readISCN3(dataDir = '../../01_DataRescue/ISCN3_2015',
                      dataLevel = 'level1')

#CPEAT throws a warning but is going to be hell to debug
CPEAT.ls <- readCPEAT2025(dataDir = '../../01_DataRescue/CPEAT2025',
                          dataLevel = 'level1')
```

# Data Level 2

We want to pull

* Location, geolocation and region
* Depth
* Soil organic carbon: organic carbon density, organic carbon fraction, bulk density, and coarse fraction

as well as a bib file and citation/acknowledgement md.

```{r}
#| label: pullLocation

CPEAT.ls$data$core$of_variable_DRaH |> table()
CPEAT.ls$data$primary$of_variable_DRaH |> table()

CPEAT_primary.df <- CPEAT.ls$data$primary |>
  filter(of_variable_DRaH %in% 
           c('carbon_percent', 
             'dry_bulk_density', 
             #'field_water_percent', 
             'inorganic_carbon_percent', 
             'layer_bottom', 'layer_middle', 'layer_thickness', 'layer_top',
             'loss_on_ignition', 
             'organic_carbon_density', 
             'organic_carbon_percent', 'organic_matter_density', 
             'organic_matter_percent', 'total_carbon_percent')) |>
  filter(any(c( 'organic_carbon_density', 'organic_matter_density', 'dry_bulk_density') %in% of_variable_DRaH),
         .by = c(bibKey, row_id)) |>
  filter(is_type == 'value', !is.na(with_entry)) |>
  select(bibKey, row_id, of_variable_DRaH, with_entry) |>
  unique() |>
  #deal with 10 double layer observations by randomly picking one of two
  dplyr::reframe(with_entry = sample(with_entry, size = 1),
                 .by = c(bibKey, row_id, of_variable_DRaH)) |>
  pivot_wider(names_from = of_variable_DRaH, values_from = with_entry)
  #calculate upper/lower from mid
  

organic_matter_to_carbon <- 0.5

layer.df <- CPEAT_primary.df |>
  mutate(layer_lower = as.numeric(layer_middle) * 100 + #convert from m to cm
           0.5 * as.numeric(layer_thickness),
         layer_upper = as.numeric(layer_middle) * 100 -
           0.5 * as.numeric(layer_thickness),
         carbon_organic_density = 
           case_when(
             #do we have a carbon density?
             !is.na(organic_carbon_density) ~ 
               as.numeric(organic_carbon_density),
             !is.na(organic_matter_density) ~ #conver from matter if needed
               organic_matter_to_carbon * as.numeric(organic_matter_density),
             #if we don't have bulk density then we can't move forward
             is.na(dry_bulk_density) ~ NA_real_,
             #start with organic carbon percent
             !is.na(organic_carbon_percent) ~ as.numeric(dry_bulk_density) * 
               as.numeric(organic_carbon_percent) / 100,
             #then assume peat soils are mostly organic
             !is.na(carbon_percent) ~ as.numeric(dry_bulk_density) * 
               as.numeric(carbon_percent) / 100,
             !is.na(total_carbon_percent) ~ as.numeric(dry_bulk_density) * 
               as.numeric(total_carbon_percent) / 100,
             !is.na(loss_on_ignition) ~ as.numeric(dry_bulk_density) * 
               organic_matter_to_carbon * as.numeric(loss_on_ignition)/100,
             !is.na(organic_matter_percent) ~ as.numeric(dry_bulk_density) * 
               organic_matter_to_carbon * as.numeric(organic_matter_percent)/100,
             .default = NA_real_)) |>
  select(bibKey, #row_id, #layer_middle, layer_thickness,
         layer_lower, layer_upper, carbon_organic_density) |>
  filter(is.finite(layer_lower + layer_upper + organic_matter_to_carbon))

site.df <- CPEAT.ls$data$core |>
  filter(of_variable_DRaH %in% c(#'citation', #in bibkey
                            #'core_comment', 
                            #'data_license', 
                            #'elevation', 'elevation_end', 'elevation_start',
                            'latitude', 'longitude', 
                            'observation_year', 
                            #'peatland_class', 
                            #'principal_investigator', 
                            #'projects', 
                            'region', 
                            #'related_research', 
                            #'sampling_method', 
                            'site_name')) |>
  select(bibKey, of_variable_DRaH, with_entry) |>
  filter(bibKey %in% layer.df$bibKey) |>
  pivot_wider(names_from = of_variable_DRaH, values_from = with_entry) |>
  mutate(across(.cols = c(latitude, longitude, observation_year), as.numeric),
         study_id = 'CPEAT')

bib.ls <- c(CPEAT.ls$citation$methods[unique(site.df$bibKey, layer.df$bibKey)],
            CPEAT.ls$citation$primary)

CPEAT.ls$data$primary$of_variable_DRaH |> table()

```

```{r}
#Data viz checks
```

# References

<!---- This will be populated when you render the document.---->
