---
title: "Data Curation: ISCN 2025"
format:
  html:
    toc: true
    eval: false #default to not evaluating the chunks, override this the template is completed
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - ../../01_DataRescue/CPEAT2025/CPEAT2025.bib #CPEAT citation
  - ../../01_DataRescue/CPEAT2025/CPEAT2025_Methods.bib #CPEAT source
  - ../../01_DataRescue/ISCN3_2015/ISCN3_2015.bib #ISCN3 citation
  - ../../01_DataRescue/ISCN3_2015/ISCN3_Contributors.bib #ISCN3 sources
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  - id: ktoddbrown #use the github name as an id
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role: #TODO define these roles and add reference
      #- transcription: lead #secondary, review
      - standardization: lead #review
      - curation: lead #review

affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
timelog:
    - activity: 
      description: Sort through migration conflicts for issue 191
      who: ktoddbrown #github username
      #when you started and ended, can be just a year or down to the minute
      start: 2026-02-05T14:15
      end: 2026-02-05T14:45
      #alternatively duration can be combined with a start or end
      #duration: P1H15M
  - activity: 
      description: Draft inital readme
      who: ktoddbrown #github username
      #when you started and ended, can be just a year or down to the minute
      start: 2025-11-27T11:00
      end: 2025-11-27T11:38
      #alternatively duration can be combined with a start or end
      #duration: P1H15M
---


```{r}
#| label: setup


library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# Bibliograph files
refernceCitation.file <- 'ISCN2025.bib'

source('../../02_DataHarmonization/readCPEAT2025.R')
source('../../02_DataHarmonization/readISCN3.R')

#conversion factor between `organic matter` or `loss on ignition` and `organic carbon`
organic_matter_to_carbon <- 0.5
```

# Purpose

The purpose ISCN 2025 (aka ISCN4) is to create a data product aggregating data on geo-located soil organic carbon profiles.
With this data we will examine the assumption of steady state on field surveyed organic carbon.
We have closed contributions for this project and are working on analysis.

# Variables of interest

<!--- Briefly describe what you want the data rescuers to look for when digitizing a data resource.---->

For this analysis we are interested in:

  * **Location**: latitude/longitude, associated datum, geo-spatial method, precision
  * **Soil organic carbon stocks**: organic carbon fraction, bulk density, coarse fraction, soil organic carbon by area, soil organic carbon by volume
  * **Layer depth**: Depth of surveyed soil layer

# Data sources

For this project, we have identified the following data sources by US federal monitoring databases and direct network contributions.
These include:

+ CPEAT (@Loisel2021, @CPEAT_web)
  - This is a collection of peat cores (@PCPEAT)
  - Status: Level 2 in progress
+ FIA
  - This is a US Forest Service database.
  - Status: Reprocessing into the new repository structure
+ NCSS
  - This is a US NRCS database.
  - Status: Reprocessing into the new repository structure
+ ISCN3 (@ISCN3)
  - This database was originally compiled in 2015, we want to update the old NCSS data.
  - Status: Level 2 in progress



# Data Level 1 

```{r}
#| label: readLevel1
#| warning: false

ISCN3.ls <- readISCN3(dataDir = '../../01_DataRescue/ISCN3_2015',
                      dataLevel = 'level1')

#CPEAT throws a warning but is going to be hell to debug
CPEAT.ls <- readCPEAT2025(dataDir = '../../01_DataRescue/CPEAT2025',
                          dataLevel = 'level1')
```

# Data Level 2

We want to pull

* Location, geolocation and region
* Depth
* Soil organic carbon: organic carbon density, organic carbon fraction, bulk density, and coarse fraction

as well as a bib file and citation/acknowledgement md.

**TODO**
+ move over the vocabulary to density being mass /volume and stocks being mass / area
+ Cheange CPEAT naming conventions to lead with `carbon`

```{r}
#| label: level2_CPEAT

#CPEAT.ls$data$core$of_variable_DRaH |> table()
#CPEAT.ls$data$primary$of_variable_DRaH |> table()

CPEAT_layer.df <- CPEAT.ls$data$primary |>
  #### Subset the primary data table from CPEAT ####
  filter(of_variable_DRaH %in% 
           c('carbon_percent', 
             'dry_bulk_density', 
             #'field_water_percent', 
             'inorganic_carbon_percent', 
             'layer_bottom', 'layer_middle', 'layer_thickness', 'layer_top',
             'loss_on_ignition', 
             'organic_carbon_density', 
             'organic_carbon_percent', 'organic_matter_density', 
             'organic_matter_percent', 'total_carbon_percent')) |>
  #make sure we have some sense of the soil density either as a SOC (that we want) 
  #...or as a bulk density that we can use to calculate SOC
  filter(any(c( 'organic_carbon_density', 'organic_matter_density', 
                'dry_bulk_density') %in% of_variable_DRaH),
         .by = c(bibKey, row_id)) |> #for each observation aka row
  filter(is_type == 'value', #only look at values for now
         !is.na(with_entry)) |> #drop missing values
  select(bibKey, row_id, of_variable_DRaH, with_entry) |> #drop extra columns
  unique() |> #TODO: go back and investigate 160 repeats
  #TODO: go back and investigate. 10 double layer observations by randomly picking one of two
  dplyr::reframe(with_entry = sample(with_entry, size = 1),
                 .by = c(bibKey, row_id, of_variable_DRaH)) |>
  #expand to human readable format, remember we are only looking at values
  pivot_wider(names_from = of_variable_DRaH, values_from = with_entry) |>
  mutate(across(.cols = c('carbon_percent', 
             'dry_bulk_density', 
             #'field_water_percent', 
             'inorganic_carbon_percent', 
             #'layer_bottom', 'layer_top', #filtered out this above
             'layer_middle', 'layer_thickness', 
             'loss_on_ignition', 
             'organic_carbon_density', 
             'organic_carbon_percent', 'organic_matter_density', 
             'organic_matter_percent', 'total_carbon_percent'),
             .fns = as.numeric)) |>
  mutate(
    ####Reconstruct the CPEAT layer bounds ####
    layer_lower = layer_middle * 100 + #convert from m to cm
           0.5 * layer_thickness,
         layer_upper = layer_middle * 100 -
           0.5 * layer_thickness,
    ####Reconstruct the CPEAT ORganic carbon density ####
         carbon_organic_density = 
           case_when(
             #do we have a carbon density?
             !is.na(organic_carbon_density) ~ 
               organic_carbon_density,
             !is.na(organic_matter_density) ~ #convert from matter if needed
               organic_matter_to_carbon * organic_matter_density,
             #if we don't have bulk density then we can't move forward
             is.na(dry_bulk_density) ~ NA_real_,
             #start with organic carbon percent, 
             #... assume organic soils with no coarse fraction because these are peats
             !is.na(organic_carbon_percent) ~ dry_bulk_density * 
               organic_carbon_percent / 100,
             #then assume peat soils have negligable inorganic carbon
             !is.na(carbon_percent) ~ dry_bulk_density * 
               carbon_percent / 100,
             !is.na(total_carbon_percent) ~ dry_bulk_density * 
               total_carbon_percent / 100,
             !is.na(loss_on_ignition) ~ dry_bulk_density * 
               organic_matter_to_carbon * loss_on_ignition/100,
             !is.na(organic_matter_percent) ~ dry_bulk_density * 
               organic_matter_to_carbon * organic_matter_percent/100,
             .default = NA_real_),
         carbon_organic_density_formula = 
           case_when(
             !is.na(organic_carbon_density) ~ 'provided SOC',
             !is.na(organic_matter_density) ~ paste('organic matter * correction factor', organic_matter_to_carbon),
             #if we don't have bulk density then we can't move forward
             is.na(dry_bulk_density) ~ 'missing bulk density',
             #start with organic carbon percent
             !is.na(organic_carbon_percent) ~ 'Bulk density unspecified * organic carbon',
             #then assume peat soils are mostly organic
             !is.na(carbon_percent) ~ 'Bulk density unspecified * carbon unspecified',
             !is.na(total_carbon_percent) ~ 'Bulk density unspecified * carbon total',
             !is.na(loss_on_ignition) ~ paste('Bulk density unspecified * loss on ignition * correction factor', organic_matter_to_carbon),
             !is.na(organic_matter_percent) ~ paste('Bulk density unspecified * organic matter * correction factor', organic_matter_to_carbon),
             .default = NA_character_) |> as.factor()
         ) |>
  #### Clean up the CPEAT layer table####
  select(site_bib = bibKey, #row_id, #layer_middle, layer_thickness,
         layer_lower, layer_upper, 
         bulk_density = dry_bulk_density,
         carbon_organic_percent = organic_carbon_percent,
         carbon_percent = carbon_percent, carbon_total_percent= total_carbon_percent,
         loss_on_ignition_percent = loss_on_ignition,
         #organic_matter_percent, #all NA
         carbon_organic_density, carbon_organic_density_formula) |>
  #only keep layers where we can estimate organic carbon
  filter(is.finite(carbon_organic_density),
         is.finite(layer_lower), is.finite(layer_upper)) |>
  #Add in the identifiers for the study
  mutate(study_bib = 'Loisel2021', study_name = 'CPEAT')

#### Pull the site table for CPEAT####
CPEAT_site.df <- CPEAT.ls$data$core |>
  filter(of_variable_DRaH %in% c(#'citation', #in bibkey
                            #'core_comment', 
                            #'data_license', 
                            #'elevation', 'elevation_end', 'elevation_start',
                            'latitude', 'longitude', 
                            'observation_year', 
                            #'peatland_class', 
                            #'principal_investigator', 
                            #'projects', 
                            'region', 
                            #'related_research', 
                            #'sampling_method', 
                            'site_name')) |>
  select(site_bib = bibKey, of_variable_DRaH, with_entry) |>
  filter(site_bib %in% layer.df$site_bib) |>
  pivot_wider(names_from = of_variable_DRaH, values_from = with_entry) |>
  mutate(across(.cols = c(latitude, longitude, observation_year), as.numeric))|>
           mutate(study_bib = 'Loisel2021', study_name = 'CPEAT')

###Pull relevant references for CPEAT ###
bib.ls <- c(CPEAT.ls$citation$methods[unique(CPEAT_site.df$site_bib, CPEAT_layer.df$site_bib)],
            CPEAT.ls$citation$primary)

```

### CPEAT Figures

```{r}
ggplot(CPEAT_layer.df |>
         pivot_longer(cols = where(is.numeric),
                      names_to = 'variable', values_to = 'numerical_value')) +
  geom_histogram(aes(x=numerical_value)) +
  facet_wrap(~variable, scales = 'free') +
  theme_bw() +
  labs(title = 'Numerical values of CPEAT')

ggplot(CPEAT_layer.df) +
  geom_point(aes(x=carbon_organic_density / bulk_density * 100,
                 y = bulk_density)) +
  theme_bw() +
  labs(title = 'Bulk density vs organic carbon fraction [g cm-3]',
       x = 'Organic carbon fraction (infered from SOC/BD) [percent]')

ggplot(CPEAT_layer.df |>
         mutate(carbon_organic_density_formula = 
                  str_wrap(carbon_organic_density_formula, width = 30))) +
  geom_histogram(aes(x=carbon_organic_density)) +
  facet_wrap(~carbon_organic_density_formula, scale = 'free') +
  theme_bw() +
  labs(title = 'SOC estimates by methods', 
         x = 'Soil organic carbon [g cm-3]')
```

```{r}
#| label: ISCN3_level2

temp <- ISCN3.ls$layer |>
  filter(
    #remove prior calculations so that we can keep our decision tree consistent
    !(dataset_name_soc %in% c('ISCN SOC stock computation',
                                "AK DSC Project SOC stock computation" ) &
        of_variable %in% c('carbon_organic_area_density')),
    #remove NRCS, adding back in an updated version
    !(dataset_name_sub %in% c('NRCS Sept/2014')),
    of_variable %in% c("layer", 
                       "bulk_density", 
                       "bulk_density_estimated_fine_earth", #not in once nrcs removed
                       "bulk_density_fine_earth",
                       "bulk_density_whole_soil", 
                       "carbon", "carbon_inorganic", "carbon_organic_fraction", 
                       "carbon_organic_area_density", 
                       "carbon_total", "coarse_fraction",
                       "loss_on_ignition")) |>
# nrow(temp) #413374 observations 
  # ISCN3 calculated SOC resulted in 879 double entries of non-soc columns
  #... drop them here.
  reframe(with_entry = unique(with_entry),
          .by = -c(row_name, with_entry)) #nrow(temp) #412495

ISCN_layer.df <- temp |>
  #filter(of_variable %in% 'layer') |>
  select(contains('name'), of_variable, is_type, with_entry) |>
  #drop 488 poorly identified observations
  #TODO go back and try to clean this up when you have time
  filter(n() == 1, .by = c(contains('name'), of_variable, is_type)) |>
  pivot_wider(names_from = c(of_variable, is_type), names_sep = '_', 
              values_from = with_entry) |> #nrow(ISCN_layer.df) #63066
  mutate(across(ends_with('_value'), as.numeric),
         across(ends_with('_method'), as.factor),
         across(ends_with('_note'), as.factor)) |>
  mutate(across(ends_with('_value'), function(xx){if_else(xx < 0, NA, xx)})) |>
  #Dropping dataset_name_soc which is either dataset_name_sub or an ISCN flag
  select(-dataset_name_soc) |>
  #match the dataset name across the table
  rename(dataset_name = dataset_name_sub) |>
  #### Harmonize explicit ISCN3 methods #####
  #mutate(across(ends_with('_value'), is.finite)) |>
  #Drop carbon_organic_area_density_method which only contains 'no fill' or NA
  select(-carbon_organic_area_density_method) |>
  mutate(
    bulk_density_method = case_when(
      #bd_samp values are means of multiple observations
      #took average of similar horizons types
      str_detect(bulk_density_method, 'multiple observations') |
        str_detect(bulk_density_method, 'average of .* horizons') ~ 'averaged: composit',
      #core volume/mass
      str_detect(bulk_density_method, 'core volume/mass') ~ 'measured: core volume and unspecified mass',
      #Calculated as oven-dry mass of core, <2mm fraction, divided by core volume.
      #g of <2mm fraction divided by total volume of material (often includes roots)
      str_detect(bulk_density_method, 
                 'Calculated as oven-dry mass of core, <2mm fraction, divided by core volume.') |
        str_detect(bulk_density_method, 
                   'g of <2mm .* by total volume') ~ 'measured: sampled volume and fine earth mass',
      #predicted from %C
      #BD was estimated based on C % using a power-based regression model based on individual soil samples where both C % (x) and DBD (y) were available (n= 831, R2= 0.73): y= 1.2559x-0.6831. In samples with permafrsot, the regression-derived BD was corrected for volumetric ice-content
      str_detect(bulk_density_method, 'predict') |
        str_detect(bulk_density_method, 
                   'BD was estimated based on C %') ~ 'calculated: organic carbon',
      #calculated based on soil texture, rock fragment, and organic matter content
      str_detect(bulk_density_method, 'calculat') ~ 'calculated: texture, rock, and organic matter',
      #bd = (1.45*0.111)/((loi/100)*1.45+(1-(loi/100))*0.111) 
      str_detect(bulk_density_method, 'bd = .*loi') ~ 'calculated: loss on igintion',
      #estimated from weight per unit area 
      str_detect(bulk_density_method, 'estimat') ~ 'estimated: weight per area',
      #gap-filled with value from nearby sample, gap-filled from horizon type
      str_detect(bulk_density_method, 'gap-filled') |
        # interpolated from adjacent horizons
        # gap-filled from horizon type
        str_detect(bulk_density_method, 'interpolated')~ 'estimated: gap filled',
      #TODO Pull NRCS codes M and G
      str_detect(bulk_density_method, '^[GM]$') ~ NA_character_,
      #bd = dry weight/volume, measured as weight per unit volume, 
    .default = NA_character_),
    #check bulk density methods
    #ISCN_layer.df |> select(bulk_density_method, bulk_density_method_new) |> unique() |> view()
    carbon_method = case_when(
      carbon_method == "MIR" ~ 'measured: MIR',
      #"dry combustion, LECO analyzer", "dry combustion", "dry combustion (CHN) of oven-dried, root-free sample, <2mm fraction",
      #"mean of ≥3 samples, dry combustion, Perkin-Elmer \d{4}",
      str_detect(carbon_method, 'dry combustion') |
        #"Perkin Elmer CHN (model 2400).",
        str_detect(carbon_method, 'Perkin Elmer') |
        #"Determined total %C from high temp combustion + elemental analyzer (on <2mm fine earth fraction)", "Element Analyzer.  Loss of Ignition", "Carlo Erba CN Elemental Analyzer.", "all mineral samples fumigated in 12M HCl (48h) to remove carbonates (Hedges and Stern 1984) then run by combustion analyzer re Manies et al, 2006",
        str_detect(carbon_method, '[eE]lement(al)? [Aa]nalyzer') |
        #"Leco TruSpec CHN (model 630-100-400).",
        str_detect(carbon_method, 'Leco') |
        #"assumed to have no inorganic carbonate ; Combustion analyser 1200 deg C  (Manies et al, 2006)", "Fumigated in 12 M HCl (48 h) to remove carbonates (Hedges and Stern, 1984);.freeze-dried and kept in a desiccator; elemental analyzer (Carlo Erba 2500) used to determine C% and N% content (all samples were we have N% values).",
        str_detect(carbon_method, 'analy[zs]er') ~ 'measured: elemental analyzer',
      #"wet oxidation",
      str_detect(carbon_method, 'wet oxidation') ~ 'measured: wet oxidation',
      #"manometric",
      str_detect(carbon_method, 'manometric') ~ 'measured: manometric',
      #"Site vegetation reflects pH; acidic vegetation sites do not have carbonates, but sites with non-tussock vegetation could also contain carbonates within the %C measurement reported here",
      #"Loss on Ignition (LOI, weight %) at 550° C (for 6 h) was used to determine organic content, and at 950° C (for 2h) to determine carbonate content (Dean, 1974; Heiri et al., 2001).  A 3rd order polynomial regression model based on individual soil samples where both LOI and C% (from elemental analyser) were available (n= 171, R2= 0.98) was used to predict C% (y) for the soil samples where only LOI% was available: y= (-0.00005x3)+(0.0059x2)+(0.362x)"
      #TODO pull codes from NRCS
      str_detect(bulk_density_method, '^[GMLAT]$') ~ NA_character_,
      # "8",
      .default = NA_character_),
    coarse_fraction_method = case_when(
      str_detect(coarse_fraction_method, 'measured') ~ 'measured',
      str_detect(coarse_fraction_method, 'estimated') |
        str_detect(coarse_fraction_method, 'assumed') ~ 'estimated',
      str_detect(coarse_fraction_method, 'observed in profile') ~ 'estimated: field observation',
      str_detect(coarse_fraction_method, 'weight of >2mm') ~ NA_character_,
      .default = NA_character_
      #ISCN_layer.df |> select(coarse_fraction_method, coarse_fraction_method_new) |> unique() |> view()
    )) |>
  # reframe(count = n(),
  #         .by = c(contains('method'), ends_with('note'), ends_with('value'))) |>
  
  mutate(
    #deal with organic carbon fraction in the study level metadata
    #coarse_fraction_method"              "coarse_fraction_value"
    #Deal with bulk density: "bulk_density_fine_earth_value", 
    #..."bulk_density_method", "bulk_density_value", "bulk_density_whole_soil_value"
    bulk_density_method_new = case_when(
      #Make a stock bulk density that does not need a coarse fraction correction
      str_detect(bulk_density_method, 
                 "measured: sampled volume and fine earth mass") ~ 
        paste('bulk density for stocks:', bulk_density_method),
      #coarse fraction correction available
      is.finite(coarse_fraction_value) & 
        #... with whole soil
        is.finite(bulk_density_whole_soil_value) ~
        paste('bulk density of whole soil:', bulk_density_method),
      is.finite(coarse_fraction_value) & 
       #... with fine earth
       is.finite(bulk_density_fine_earth_value) ~
        paste('bulk density of fine earth:', bulk_density_method),
      #Without coarse fraction, what do we prioritize
      is.finite(bulk_density_value) ~ #assume data providers knew we wanted to calculate stocks
        paste('bulk density unspecified:', bulk_density_method),
      is.finite(bulk_density_fine_earth_value) ~ #under-estimate stocks
        paste('bulk density of fine earth:', bulk_density_method),
      is.finite(bulk_density_whole_soil_value) ~ #over-estimate stocks
        paste('bulk density of whole soil:', bulk_density_method),
      .default = NA_character_
    ),
    carbon_organic_area_density_method_new = case_when(
      is.finite(carbon_organic_area_density_value) &
        is.na(carbon_organic_area_density_flag) ~
        'unspecified formula',
      is.finite(carbon_organic_area_density_value) &
        !is.na(carbon_organic_area_density_flag) ~
        paste('forumla: ', carbon_organic_area_density_flag),
      .default = NA_character_
    ),
    carbon_organic_fraction_method_new = case_when(
      is.finite(carbon_organic_fraction_value) ~ paste('carbon fraction organic:', carbon_method),
      is.finite(carbon_total_value - carbon_inorganic_value) ~ paste('carbon fraction organic: calculated total - inorganic: ', carbon_method),
      .default = NA_character_
    ),
    carbon_total_fraction_method_new = case_when(
      is.finite(carbon_total_value) ~ paste('carbon fraction total:', carbon_method),
      .default = NA_character_
    )
  ) |>
  mutate(bulk_density_new_value = case_when(
    #first take the bulk density for stocks if it's there 
      str_detect(bulk_density_method_new, 
                 "measured: sampled volume and fine earth mass") &
        is.finite(bulk_density_fine_earth_value) ~ 
        bulk_density_fine_earth_value, #240 layers
      str_detect(bulk_density_method_new, 
                 "measured: sampled volume and fine earth mass") &
        is.finite(bulk_density_value) ~ 
        bulk_density_value, #3588 layers
      is.finite(coarse_fraction_value) & 
        #... with whole soil
        is.finite(bulk_density_whole_soil_value) ~
        bulk_density_whole_soil_value,
      is.finite(coarse_fraction_value) & 
       #... with fine earth
       is.finite(bulk_density_fine_earth_value) ~
        bulk_density_fine_earth_value,
      #Without coarse fraction, what do we prioritize
      is.finite(bulk_density_value) ~ #assume data providers knew we wanted to calculate stocks
        bulk_density_value,
      is.finite(bulk_density_fine_earth_value) ~ #under-estimate stocks
        bulk_density_fine_earth_value,
      is.finite(bulk_density_whole_soil_value) ~ #over-estimate stocks
        bulk_density_whole_soil_value,
      .default = NA_real_
    ),
    carbon_organic_area_density_new_value = case_when(
      is.finite(carbon_organic_area_density_value) &
        is.na(carbon_organic_area_density_flag) ~
        carbon_organic_area_density_value,
      is.finite(carbon_organic_area_density_value) &
        !is.na(carbon_organic_area_density_flag) ~
        carbon_organic_area_density_value,
      .default = NA_real_
    ),
    carbon_organic_fraction_new_value = case_when(
      is.finite(carbon_organic_fraction_value) ~ carbon_organic_fraction_value,
      is.finite(carbon_total_value - carbon_inorganic_value) ~
        carbon_total_value - carbon_inorganic_value,
      .default = NA_real_
    ),
    carbon_total_fraction_new_value = case_when(
      is.finite(carbon_total_value) ~ carbon_total_value,
      .default = NA_real_
    )
    ) |>
  #### Merge method notes with column name information
  # select(contains('carbon')) |>
  # mutate(across(ends_with('value'), is.finite)) |>
  # reframe(across(ends_with('value'), sum),
  #         .by = c(contains('flag'), contains('method'), contains('note'))) |>
  select(contains('name'),
         contains('layer'),
         #provided SOC estimations
         carbon_organic_area_density_value = carbon_organic_area_density_new_value,
         carbon_organic_area_density_method = carbon_organic_area_density_method_new,
         #bulk density estimations
         bulk_density_value = bulk_density_new_value,
         bulk_density_method = bulk_density_method_new,
         #carbon fractions
         carbon_organic_fraction_value = carbon_organic_fraction_new_value,
         carbon_total_fraction_value = carbon_total_fraction_new_value,
         #coarse fraction
         coarse_fraction_value, coarse_fraction_method
         )

ISCN_site.df <- ISCN3.ls$site |>
  semi_join(ISCN_layer.df,
            by = join_by(dataset_name_sub == dataset_name, site_name, profile_name)) |>
  select(contains('name'), with_entry, of_variable, is_type) |>
  pivot_wider(names_from = c(of_variable, is_type),
              names_sep = '_', values_from = with_entry) |>
  mutate(observation_date_date = as_date(as.numeric(observation_date_date), 
                                         origin = ymd('1900-01-01')),
         across(c(geolocation_latitude, geolocation_longitude),
                  .fns = as.numeric),
         across(c(geolocation_datum, region_country, region_state, depth_zero_reference),
                .fns = as.factor)) |>
  select(-dataset_name_soc) |>
  rename(dataset_name = dataset_name_sub)

```

```{r}

ggplot(ISCN_layer.df) +
  geom_histogram(aes(x = carbon_organic_fraction_value)) +
  geom_vline(xintercept = 55, color = 'red') +
  scale_x_log10()


ggplot(ISCN_layer.df) +
  geom_histogram(aes(x = bulk_density_value)) +
  #density of quartz
  geom_vline(xintercept = 2.65, color = 'red') +
  scale_x_log10()

ggplot(ISCN_layer.df) +
  geom_histogram(aes(x = carbon_organic_area_density_value / (as.numeric(layer_lower) - as.numeric(layer_upper)) )) +
  geom_vline(xintercept = 55/100 * 2.65, color = 'red') +
  scale_x_log10()

ggplot(ISCN_layer.df) +
  geom_histogram(aes(x = coarse_fraction_value)) +
  geom_vline(xintercept = 100, color = 'red') +
  scale_x_log10()

ggplot(ISCN_layer.df) +
  geom_point(aes(x = carbon_organic_fraction_value, y=bulk_density_value)) 

ggplot(ISCN_layer.df |>
         filter(carbon_organic_area_density_value < 10)) +
  geom_point(aes(x = carbon_organic_area_density_value / (as.numeric(layer_lower) - as.numeric(layer_upper)), 
                 y = bulk_density_value * carbon_organic_fraction_value / 100 )) +
  scale_x_log10() + scale_y_log10() +
  geom_abline(color = 'grey')

ggplot(ISCN_layer.df |>
         filter(carbon_organic_area_density_value < 10)) +
  geom_histogram(aes(x = (carbon_organic_area_density_value / (as.numeric(layer_lower) - as.numeric(layer_upper)) -  bulk_density_value * carbon_organic_fraction_value / 100 ) / (carbon_organic_area_density_value / (as.numeric(layer_lower) - as.numeric(layer_upper))))) 
```

```{r}
#Data viz checks
```

# References

<!---- This will be populated when you render the document.---->
