---
title: "Data Curation: ISCN 2025"
format:
  html:
    toc: true
    eval: false #default to not evaluating the chunks, override this the template is completed
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - ISCN2025.bib #contains citation that provide background on the project
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  - id: ktoddbrown #use the github name as an id
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role: #TODO define these roles and add reference
      - transcription: lead #secondary, review
      - standardization: lead #review
      - curation: lead #review

affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
timelog:
  - activity: 
      description: Draft inital readme
      who: ktoddbrown #github username
      #when you started and ended, can be just a year or down to the minute
      start: 2025-11-27T11:00
      end: 2025-11-27T11:38
      #alternatively duration can be combined with a start or end
      #duration: P1H15M
---


```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false


library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# Bibliograph files
refernceCitation.file <- 'ISCN2025.bib'

```

# Purpose

The purpose ISCN 2025 (aka ISCN4) is to create a data product aggregating data on geo-located soil organic carbon profiles.
With this data we will examine the assumption of steady state on field surveyed organic carbon.
We have closed contributions for this project and are working on analysis.

# Variables of interest

<!--- Briefly describe what you want the data rescuers to look for when digitizing a data resource.---->

For this analysis we are interested in:

  * **Location**: latitude/longitude, associated datum, geo-spatial method, precision
  * **Soil organic carbon stocks**: organic carbon fraction, bulk density, coarse fraction, soil organic carbon by area, soil organic carbon by volume
  * **Layer depth**: Depth of surveyed soil layer

# Data sources

For this project, we have identified the following data sources by US federal monitoring databases and direct network contributions.
These include:

+ CPEAT (@Loisel2021, @CPEAT_web)
  - This is a collection of peat cores (@PCPEAT)
  - Status: Reprocessing into the new repository structure
+ FIA
  - This is a US Forest Service database.
  - Status: Reprocessing into the new repository structure
+ NCSS
  - This is a US NRCS database.
  - Status: Reprocessing into the new repository structure
+ ISCN3 (@ISCN3)
  - This database was orginally compiled in 2015, we want to update the NCSS data.
  - Status: Reprocessing into the new repository structure



# Data Level 1 

```{r}
#Source and append all the level 1 tuples

```

# Data Level 2

```{r}
#Map the comparable methods and align any needed vocabulary
```

```{r}
#Data viz checks
```

# References

<!---- This will be populated when you render the document.---->
