---
title: "Data Rescue: FIADB March 2026"
format:
  html:
    toc: true
    code-fold: true
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - FIA2024.bib  #contains the citations for contributed datasets by authors paper
nocite: |
  @* #add all citations
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  - id: ktoddbrown #use the github name as an id
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role: #TODO define these roles and add reference
      - transcription: lead #secondary, review
      - standardization: lead #review
      - curation: lead #review
  - id: vaasukimarupaka
    name:
      given: Vaasuki
      family: Marupaka
    orcid: 0000-0002-3854-1764
    affiliation:
      - ref: uf-ees
    role:
      - transcription: lead #secondary, review
      - standardization: lead
      - curation: lead #review
affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
    
timelog:
  - activity: 
      description: Lead Transcription and ReadMe Update
      who: Vaasuki Marupaka
      end: 
      start: 
      duration:
      timelog:
---

```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false
#| eval: true

# libraries
library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R
library(stringr) # regular expressions

# Download url for the entire FIA database
downloadUrl <- "https://apps.fs.usda.gov/fia/datamart/CSV/CSV_FIADB_ENTIRE.zip"
# name the zip file
CSV_FIADB_ENTIRE_filename <- "CSV_FIADB_ENTIRE.zip"

#Download data directory
#Change this file to run locally for you
dataDir  <- '~/Documents/Research/ISCN4/FIADB_2026_pull/CSV_FIADB_ENTIRE/'
#dataDownload.dir <- 'temp'

# The below csv tables are used to construct annotations for FIADB
annotationFilenames <- c(FIA_Plot_Annotations = 'FIA2024_TableEntirePLOT.csv',
                   #ENTIRE_SOILS_VISIT Table appears in datamart download 
                   #...but is not documented in manual
                   FIA_Survey_Annotations = 'FIA2024_TableSurvey.csv',
                   FIA_Subplot_Annotations = 'FIA2024_TableSubPlot.csv',
                   FIA_Subplot_SampleLoc_Annotations =
                     'FIA2024_TableSubPlot_Soil_Sample_Location.csv',
                   #ENTIRE_SOILS_EROSION Table appears in datamart download 
                   #...but is not documented in manual
                   FIA_Subplot_Lab_Annotations =
                     'FIA2024_TableSubplot_Soil_Sample_Layer.csv',
                   FIA_Tree_Annotations = 'FIA2024_TableTREE.csv'
)

#Pulled from nested table in the manual
controlVocabulary <- 'FIA2024_Entire_Control_Vocabulary.csv'

# Bibliography files
primaryCitation.file <- 'FIA2024.bib'

# read function
readsource.file <- 'readFIA.R'
if(file.exists(readsource.file)){
  source(readsource.file)
}

#semantic files for level 1
of_variable.file <- '../../02_DataHarmonization/SoilDRaHVocabulary_of_variable.csv'
is_type.file <- '../../02_DataHarmonization/SoilDRaHVocabulary_is_type.csv'

```

The purpose of this document is to summarize the portions of the Forest Inventory Analysis Database that are relevant to data collections in the SoilDRaH project and walk through the data ingestion.
Here you will find links to documentation from the data provider, links to where you can access the data, a description of how the data was processed, and visuals for the collection relevant variables.

# What is the FIADB

The version of the US Forest Service's Forest Inventory Analysis Database described here is the public facing version of the Forest Inventory and Analysis program.

> **Forest Inventory and Analysis**
For more than 80 years, the FIA program has been recognized as a world leader in conducting national-scale forest inventories. FIA information is widely used to address local and regional issues related to trends in forest extent, health and productivity; land cover and land use change; and the changing demographics of private forest landowners.

Taken from [https://research.fs.usda.gov/products/dataandtools/fia-datamart](https://research.fs.usda.gov/products/dataandtools/fia-datamart) (Accessed 09-April-2026)

> **Background**
The Forest Inventory and Analysis (FIA) research program has been in existence since mandated by Congress in 1928. FIA reports on status and trends in forest area and location; in the species, size, and health of trees; in total tree growth, mortality, and removals by harvest; in wood production and utilization rates by various products; and in forest land ownership. Before 1999, all inventories were conducted on a periodic basis. The passage of the 1998 Farm Bill requires FIA to collect data annually on plots within each State. This kind of up-to-date information is essential to frame realistic forest policies and programs. USDA Forest Service regional research stations are responsible for conducting these inventories and publishing summary reports for individual States.
In addition to published reports, the Forest Service provides data collected in each inventory to those interested in further analysis. This report describes a standard format in which data can be obtained. This standard format, referred to as the Forest Inventory and Analysis Database (FIADB) structure, was developed to provide users with as much data as possible in a consistent manner among States. A number of inventories conducted prior to the implementation of the annual inventory are available in the FIADB. However, various data attributes may be empty or the items may have been collected or computed differently. Annual inventories use a common plot design and common data collection procedures nationwide, resulting in greater consistency among FIA work units than earlier inventories. Data field definitions note inconsistencies caused by different sampling designs and processing methods.

> **The FIA Database**
The Forest Inventory and Analysis Database (FIADB) was developed to provide users with data in a consistent format, spanning all States and inventories. The first version of FIADB replaced two FIA regional databases; the Eastern States (Eastwide database) documented by Hansen and others (1992), and Western States (Westwide database) documented by Woudenberg and Farrenkopf (1995). A new national plot design provided the impetus for replacing these two databases, and FIA work units adopted the new design in all State inventories initiated after 1998. The FIADB table structure is currently derived from the National Information Management System (NIMS), which was designed to process and store annual inventory data. A number of changes in the FIADB structure have been made to accommodate the data processing and storage requirements of NIMS. As a result, data from periodic inventories are stored in a format consistent with annual inventory data.

Documentation for FIA database is concentrated in the following documents below (The order being the version history of the user manual documentation):

Currently documented in @burrill2024d, with past documentation in @burrill2024a, @burrill2024b, @burrill2024c, @burrill2023, @burrill2021, @woudenberg2010, @woudenberg1995westwide, @hansen1992eastwide.

Annotations for analysis relevant tables were manually transcribed (@burrill2024d) and cross checked.

  + Burrill, Elizabeth A.; DiTommaso, Andrea M.; Turner, Jeffery A.; Pugh, Scott A.; Christensen, Glenn; Kralicek, Karin M.; Perry, Carol J.; Lepine, Lucie C.; Walker, David M.; Conkling, Barbara L. 2024. The Forest Inventory and Analysis Database, FIADB user guides, volume: database description (version 9.4), nationwide forest inventory (NFI). U.S. Department of Agriculture, Forest Service. 1016 p. [Online]. Available at web address: https://research.fs.usda.gov/understory/forest-inventory-and-analysis-database-user-guide-nfi
  + Burrill, Elizabeth A.; DiTommaso, Andrea M.; Turner, Jeffery A.; Pugh, Scott A.; Christensen, Glenn; Kralicek, Karin M.; Perry, Carol J.; Lepine, Lucie C.; Walker, David M.; Conkling, Barbara L. 2024. The Forest Inventory and Analysis Database, FIADB user guides, volume: database description (version 9.3), nationwide forest inventory (NFI). U.S. Department of Agriculture, Forest Service. 1026 p. [Online]. Available at web address: https://research.fs.usda.gov/understory/forest-inventory-and-analysis-database-user-guide-nfi
  + Burrill, Elizabeth A.; DiTommaso, Andrea M.; Turner, Jeffery A.; Pugh, Scott A.; Christensen, Glenn; Perry, Carol J.; Lepine, Lucie C.; Walker, David M.; Conkling, Barbara L. 2024. The Forest Inventory and Analysis Database, FIADB user guides, volume database description (version 9.2), nationwide forest inventory (NFI). U.S. Department of Agriculture, Forest Service. 1042 p. [Online]. Available at web address: https://www.fs.usda.gov/research/products/dataandtools/datasets/fia-datamart
  + Burrill, Elizabeth A.; DiTommaso, Andrea M.; Turner, Jeffery A.; Pugh, Scott A.; Christensen, Glenn; Perry, Carol J.; Lepine, Lucie C.; Walker, David M.; Williams, Mary A.; Conkling, Barbara L. 2024. The Forest Inventory and Analysis Database: database description and user guide version 9.1 for Phase 2. U.S. Department of Agriculture, Forest Service. 1066 p. [Online]. Available at web address: https://www.fia.fs.usda.gov/library/database-documentation/index.
  + Burrill, Elizabeth A.; DiTommaso, Andrea M.; Turner, Jeffery A.; Pugh, Scott A.; Menlove, James; Christensen, Glenn; Perry, Carol J.; Conkling, Barbara L. 2023. The Forest Inventory and Analysis Database: database description and user guide version 9.1 for Phase 2. U.S. Department of Agriculture, Forest Service. 1066 p. [Online]. Available at web address: https://www.fia.fs.usda.gov/library/database-documentation/index.
  + Burrill, Elizabeth A.; DiTommaso, Andrea M.; Turner, Jeffery A.; Pugh, Scott A.; Menlove, James; Christiansen, Glenn; Perry, Carol J.; Conkling, Barbara L. 2021. The Forest Inventory and Analysis Database: database description and user guide version 9.0.1 for Phase 2. U.S. Department of Agriculture, Forest Service. 1026 p. [Online]. Available at web address: http://www.fia.fs.fed.us/library/database-documentation/.
  + Woudenberg, Sharon W.; Conkling, Barbara L.; O’Connell, Barbara M.; LaPoint, Elizabeth B.; Turner, Jeffery A.; Waddell, Karen L. 2010. The Forest Inventory and Analysis Database: Database description and users manual version 4.0 for Phase 2. Gen. Tech. Rep. RMRS-GTR-245. Fort Collins, CO: U.S. Department of Agriculture, Forest Service, Rocky Mountain Research Station. 339 p.
  + Woudenberg, S.W.; Farrenkopf, T.O. 1995. The Westwide forest inventory data base: user's manual. Gen. Tech. Rep. INT-GTR-317. Ogden, UT: U.S. Department of Agriculture, Forest Service, Intermountain Research Station. 67 p.
  + Hansen, M.H.; Frieswyk, T.; Glover, J.F.; Kelly, J.F. 1992. The Eastwide forest inventory data base: users manual. Gen. Tech. Rep. NC-151. St. Paul, MN: U.S. Department of Agriculture, Forest Service, North Central Forest Experiment Station. 48 p.

The annotations table draws heavily on these above documents with augmentation from private communication with Forest Service scientist.

## Files

This data is identified as a data source for ISCN4 data harmonization and curation project.
There are six tables currently annotated in the FIADB data collection.
This represents the data directly relevant for layer-level: 

  + soil carbon measurements: only provided in table *FIA2024_table_subplot_soil_sample_layer* 
    - bulk density [BULK_DENSITY]
    - bulk density fine fraction [BULK_DENSITY_FINES]
    - organic carbon fraction [C_ORG_PCT]
    - total carbon [C_TOTAL_PCT]
    - inorganic carbon [C_INORG_PCT]
    - coarse fraction [COARSE_FRACTION_PCT]
  + depth of sample and layer information:
    - layer thickness [LAYER_THICKNESS] *FIA2024_table_subplot_soil_sample_layer*
    - layer type [LAYER_TYPE] *FIA2024_table_subplot_soil_sample_layer*
    - core size [CORE_SIZE] *FIA2024_TableSubPlot_Soil_Sample_Location.csv*
    - core length [CORE_LENGTH] *FIA2024_TableSubPlot_Soil_Sample_Location.csv*
    - depth of refuse [RESTRICTION_DEPTH_X] *FIA2024_TableSubPlot_Soil_Sample_Location.csv*
  + sampling time: 
    - key variables are: inventory year [INVYR] (this information is present in all the tables extracted)
    - measurement day [MEASDAY]
    - measurement month [MEASMON]
    - measurement year [MEASYEAR]
  + sample location: 
    - Latitude and Longitude coordinates [*FIA2024_TableEntirePLOT.csv*] (only table providing for the sites)
    - state code [STATECD] (provided in all tables)
    - county code [COUNTYCD] (provided in all tables except Survey table)

This represents the data directly relevant for layer-level soil carbon measurements including bulk density, organic carbon fraction, coarse fraction, depth of sample, sampling time, and sample location.

Please note that there is quite a bit more data in the FIADB and we welcome interested parties extending these annotations.

These files are in the FIADB 2024 data rescue.

-   [Readme](README_FIA2024.qmd)
    -   This is the primary file that documents the transcriptions and decision made during data rescue.
-   [FIA2024.bib](FIA2024.bib)
    -   citation for database transcribed from user manuals and other documentation
-   Annotations
    -   this includes the annotations constructed using user manual (latest version 9.4; https://research.fs.usda.gov/understory/forest-inventory-and-analysis-database-user-guide-nfi) pulled from fia data mart portal with individual tables documented in the manual (plot, subplot, subplot layer, subplot location, survey, tree tables)
        + [FIA2024_TableEntirePLOT.csv](FIA2024_TableEntirePLOT.csv)
        + [FIA2024_TableSurvey.csv](FIA2024_TableSurvey.csv)
        + [FIA2024_TableSubPlot.csv](FIA2024_TableSubPlot.csv)
        + [FIA2024_TableSubPlot_Soil_Sample_Location.csv](FIA2024_TableSubPlot_Soil_Sample_Location.csv)
        + [FIA2024_TableSubplot_Soil_Sample_Layer.csv](FIA2024_TableSubplot_Soil_Sample_Layer.csv)
        + [FIA2024_TableTREE.csv](FIA2024_TableTREE.csv)
    - Note: the soil erosion and visit table is not exclusively annotated in the documented user manual; hence, we pulled annotations for these tables using column header search in the manual. 
- [temp/](temp/)
        + scratch folder that will not be archived on GitHub, it should include local copies of the data resource (likely a PDF article) and any spreadsheets or other tools you used for the transcriptions.

# Data Rescue working notes 

## Data Rescue Plan

+ Lead
  - [x] Read the article or data resource and the discription of the target project
  - [x] Create a fresh fork of the primary repository https://github.com/ktoddbrown/SoilDRaH/tree/main, it might be useful to rename it 'SoilDRaH_AuthorYYYY'
  - [x] Create a copy of this template folder and rename it to AuthorYYYY
  - [x] Fill out Data summary section (except for the Files), the Data Rescue Plan (here!), and update the author section of the yml to reflect who is working on the rescue
  - [x] Create a GitHub issue from the Data Rescue template and get feedback from a reviewer on the project (Dr Todd-Brown or a grad student typically).
  - [x] Add the reviewer to the author list in the yml
+ Do
  - [x] Code up data download from FIADB Data Mart
  - [x] Transcribe the Data tables in FIADB manual v.9.4 for the selected tables
  - [x] Create a bib file with the recommended citations
  - [x] Read in primary tables and variable description information from the manual
  - [ ] read function match `02_DataHarmonization/readFIA.R` function inputs and output formats
  - [ ] ask one of the undergrads to provide a secondary transcription for the TemplateCV and TemplateSubmit pdfs
  - [ ] push everything to the lead fork
+ Measure
  - [ ] cross check with second transcriber for tables and methods, and resolve any differences
  - [ ] compile or render this ReadMe with all chunks up through the Level 0 data rescue section with `eval: true`
  - [ ] push to your local fork and create a pull request to the main branch of the primary repository
  - [ ] address any revisions in your pull request review
  - [ ] celebrate!

## Annotations notes:
-   The soils visit, soil erosion tables do not exist as explicit annotation tables in the pdf manual, hence, we do not have separate annotations for these tables

## Download FIA 2024

# Data processing

```{r}
#| label: downloadData
#| echo: false
#| eval: false # i already have data downloaded to my local
# This data pull may have a high memory demand
# Ensure verbose is defined
if (!exists("verbose")) verbose <- T
if(verbose) warning('This function may have a high memory demand.')

# # Download entire FIA database
# downloadUrl <- "https://apps.fs.usda.gov/fia/datamart/CSV/CSV_FIADB_ENTIRE.zip"
# # What name is the downloaded file
# filename <- "CSV_FIADB_ENTIRE.zip"
if(!file.exists(file.path(dataDir, filename))){
  if(verbose) message('Database not found, downloading...')
  if(verbose) warning('This is a large file and may require a manual download https://www.fs.usda.gov/research/products/dataandtools/tools/fia-datamart, select the `Entire FIADB CSV Archive` download under `Additional Download Options`.' )
  utils::download.file(url = downloadUrl, 
                       destfile = file.path(dataDir, 
                                            filename))
}
  
  # Unzip zip file
if(!file.exists(file.path(dataDir, 'CSV_FIADB_ENTIRE'))){
  if(verbose) message('Unzipping file...')
  if(verbose) warning('This large file may require a manual unzip. Navigate to the data directory and unzip the file to a folder called `CSV_FIADB_ENTIRE`.')
  utils::unzip(file.path(dataDir, filename), exdir = dataDir, overwrite = FALSE)
  if(verbose) message('Unzipping done.')
}

```

Reading in FIA ENTIRE PLOT csv table

```{r}
#| label: table1Check
#| eval: false #remove when you have the file transcribed

read_csv(file = file.path(dataDir, 'ENTIRE_PLOT.csv'), #read in the specific file
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) # 1,988,296 × 63 tibble

```

Reading in FIA ENTIRE SURVEY csv table

```{r}
#| label: table2Check
#| eval: false #remove when you have the file transcribed

read_csv(file = file.path(dataDir, 'ENTIRE_SURVEY.csv'), #read in the specific file
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) # 1,775 × 14 tibble

```

Reading in FIA ENTIRE SUBPLOT csv table

```{r}
#| label: table3Check
#| eval: false #remove when you have the file transcribed

read_csv(file = file.path(dataDir, 'ENTIRE_SUBPLOT.csv'), #read in the specific file
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) # 7,640,714 × 48 tibble

```

Reading in FIA ENTIRE SUBPLOT SOIL SAMPLE LOCATION csv table

```{r}
#| label: table4Check
#| eval: false #remove when you have the file transcribed

read_csv(file = file.path(dataDir, 'ENTIRE_SUBP_SOIL_SAMPLE_LOC.csv'), #read in the specific file
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) # 3,716 × 32 tibble

```

Reading in FIA ENTIRE SUBPLOT SOIL SAMPLE LAYER csv table

```{r}
#| label: table5Check
#| eval: false #remove when you have the file transcribed

read_csv(file = file.path(dataDir, 'ENTIRE_SUBP_SOIL_SAMPLE_LAYER.csv'), #read in the specific file
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) # 14,243 × 57 tibble

```

Reading in FIA ENTIRE TREE csv table

```{r}
#| label: table6Check
#| eval: false #remove when you have the file transcribed

read_csv(file = file.path(dataDir, 'ENTIRE_TREE.csv'), #read in the specific file
         n_max = 1000, # just 1000 rows 
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) 
# this table is high in memory (13.32 GB data) with 198 columns

```

Reading in FIA SOILS EROSION csv table

```{r}
#| label: table7Check
#| eval: false #remove when you have the file transcribed

read_csv(file = file.path(dataDir, 'ENTIRE_SOILS_EROSION.csv'), #read in the specific file
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) # 44,874 × 16 tibble

```

Reading in FIA SOILS VISIT csv table

```{r}
#| label: table8Check
#| eval: false #remove when you have the file transcribed

read_csv(file = file.path(dataDir, 'ENTIRE_SOILS_VISIT.csv'), #read in the specific file
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) # 12,683 × 12 tibble
```

Using the FIABD pdf manual v.9.4(2024), we manually transcribe the annotations into different tables; plot, subplot, subplot layer, subplot location, survey, tree tables for level0. These annotations will further be extended to a standard data model for our level01 data.

```{r}
annotations <- list(
  FIA2024_annotations = lapply(annotationFilenames, function(xx) {
    readr::read_delim(
      file = xx,
      delim = ",",
      col_types = readr::cols(.default = readr::col_character())
    )
  }))

### Read in control vocabulary###
#Read in the results as tables or figure transcriptions. This includes
#...the caption as well as the tables themselves
controlVocabulary.ls <- list(
  primary_CV = readr::read_delim(file = controlVocabulary,
    delim = ',', 
    skip = 5, # we want to only keep the headers without the preamble text.
    col_types = readr::cols(.default = readr::col_character())))

```

# Level 0 data

```{r}
#This chunk has two purposes. 
#...1) Check the formatting by reading in everything 
#...2) Create a list of everything to process later in level 1

original_tables <- function(dataDir,
                            data_files = c('ENTIRE_PLOT.csv',
                                           'ENTIRE_SURVEY.csv',
                                           'ENTIRE_SUBPLOT.csv',
                                           'ENTIRE_SUBP_SOIL_SAMPLE_LOC.csv',
                                           'ENTIRE_SUBP_SOIL_SAMPLE_LAYER.csv',
                                          # 'ENTIRE_TREE.csv',
                                           'ENTIRE_SOILS_EROSION.csv',
                                           'ENTIRE_SOILS_VISIT.csv')) {
  
  # create an empty list to hold the data
  tables <- vector("list", length(data_files))
  
  # read files
  for (i in seq_along(data_files)) {
    tables[[i]] <- readr::read_csv(
      file = file.path(dataDir, data_files[i]),
      col_types = readr::cols(.default = readr::col_character())
    )
  }

  table_names <- tools::file_path_sans_ext(basename(data_files)) # strip .csv
  names(tables) <- table_names 
  
  return(tables)
}

primary_data <- original_tables(dataDir)

data.lvl0.ls <- list(
  #Read in a list of all the bib files
  citation.ls = read.bib(file = primaryCitation.file),
  primaryData.ls = primary_data,
  primaryAnnotations.ls = annotations,
  controlVocabulary.ls = controlVocabulary.ls)

```
