---
title: "Data Rescue: Austin 2010"
format:
  html:
    toc: true
    eval: false #default to not evaluating the chunks, override this the template is completed
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - Austin2010_Methods.bib #contains the citations in the methods section
  - Austin2010.bib #contains citation for the data source, generally a single citation of the primary paper
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  - id: ashbonner310 #use the github name as an id
    name:
      given: Ashley
      family: Bonner
    orcid: 0000-0001-7882-6353
    affiliation:
      - ref: uf-ees
    role:
      - transcription: lead 
      
  - id: ktoddbrown 
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role: 
      - transcription: review


affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
    
timelog:
  - activity: 
      description: Start moving rescue from old template to new template, particularly the README.
      who: ashbonner310 
      start: 2026-01-29 T16:45
      end: 2026-01-29 T17:15
      duration: P30M
  - activity: 
      description: Continue moving rescue from old template to new template, particularly moving files other than README.
      who: ashbonner310 
      start: 2026-01-30 T15:02
      end: 2026-01-30 T15:34
      duration: P32M
  - activity: 
      description: Finish moving rescue from old template to new template.
      who: ashbonner310 
      start: 2026-01-30 T17:51
      end: 2026-01-30 T18:45
      duration: P54M
  - activity: 
      description: Fixing some old errors, starting the cross-check.
      who: ashbonner310 
      start: 2026-04-10 T15:45
      end: 2026-04-10 T17:00
      duration: P1H15M
  - activity: 
      description: Finishing cross-check, rebuilding plot from paper.
      who: ashbonner310 
      start: 2026-04-10 T15:45
      end: 2026-04-10 T17:00
      duration: P1H15M
version: 'vs202512' 
---


```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false


library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# data files for level 0
methods.file <- "Austin2010_Methods.md"
figure1.file <- 'Austin2010_Figure1.csv'

# Bibliograph files
primaryCitation.file <- 'Austin2010.bib'
methodsCitation.file <- 'Austin2010_Methods.bib'

# read function
# readsource.file <- '../../../R/readAuthorYYYY.R'
# source('readsource.file')

#semantic files for level 1
# of_variable.file <- '../../semantics/SoilDRaHVocabulary_of_variable.csv'
# is_type.file <- '../../semantics/SoilDRaHVocabulary_is_type.csv'

```


# Data Summary

> A.T. Austin, & C.L. Ballaré, Dual role of lignin in plant litter decomposition in terrestrial ecosystems, Proceedings of the National Academy of Sciences. U.S.A. 107 (10) 4618-4622, https://doi.org/10.1073/pnas.0909396107 (2010).

For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/119.

@Austin2010 was interested in photodegredation rates of litter decomposition based on litter chemistry (specifically based on the levels of lignin and cellulose). 
They also used wavelength-specific light attenuation filters to determine what wavelengths of light contribute to abiotic mass loss.

## Tables and figures

- Table
This paper contains no tables.

- Figure 
  1) Mass loss over experimental attenuation of specific regions of the solar spectrum; panel A covers mass loss of litter and panel B covers mass loss of a synthetic cellulose-lignin substrate.
  2) Radiation absorbance spectra of synthetic cellulose-lignin substrates enriched with different quantities of lignin.
  3) Degradation of synthetic cellulose-lignin substrates under biotic conditions and abiotic conditions with full solar radiation.
  4) Panel A shows initial to final lignin concentrations of the synthetic cellulose-lignin substrate in both abiotic and biotic experiments, while panels B and C compare observed lignin loss compared to predictions based on the assumption that all degradation was due to biotic or abiotic conditions.

Since we are only interested in the litter decay, we are rescuing only Figure 1.

## Fit for purpose: DeAD

This paper was identified for data rescue with the Decomposition Across Drylands (DeAD) project. 
The DeAD data rescue project is concerned with documenting and preserving data which documents litter decay rates in arid and semi-arid ecosystems.

The paper contains data from two seperate experiments, one using a synthetic cellulose-lignin substrate which we are not including in the data rescue, and an experiment using collected litter which we are rescuing.

 - Location: a perennial grassland in the province of Córdoba, Argentina (31° 04 S, 64° 31 W)
 - Duration: 120 days
 - Mass loss: Mass loss or decay rate or mass remaining of litter sample over duration
 - Solar exposure: Litter treated with filters (see treatments below)
 - Litter species: Senescent litter was gathered on site, and were perennial tussock grasses. Identified species were *Festuca hieronymi* and *Stipa trichotoma*
 - Vegetation Coverage: Open Woodland
 - Litter Chemistry: Found via standard acid detergent method:
    + cellulose and sugars, 49.1 ± 0.4%
    + hemicellulose, 37.3 ± 0.2%
    + lignin,7.3 ± 0.2%.
 - Litter bag materials and method: 0.750 g of litter in an open-topped plastic polystyrene box measuring 8 cm wide × 13 cm long × 2.5 cm.
 - Treatments: 
    + Plastic box placed in chicken wire cage to allow for free flow of air.
    + Rainout shelter
    + No contact with ground; wire cage suspended 30cm above ground.
    + Four solar filter treatments: control (transparent filter), attenuation of UV-B (280–315 nm), attenuation of all UV radiation (280–400 nm), attenuation of UV and blue wavelengths (280–450 nm)


## Files

These files are in the Austin2010 data rescue.

- [Readme](README_Austin2010.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Austin2010.bib](Austin2010.bib)
  + citation for article transcribed
- [Austin2010_Methods.bib](Austin2010_Methods.bib)
  + citations for the methods section of the article
- [Austin2010_Methods.md](Austin2010_Methods.md)
  + methods transcribed from primary article
- [Austin2010_Figure1.csv](Austin2010_TableN.csv)
  + Plot-digitized values of Figure 1 from primary article, up to one significant figure more than the y-axis of the bar graph in the figure. 
- [temp/](temp/)
    + scratch folder that will not be archived on GitHub, it should include local copies of the data resource (likely a PDF article) and any spreadsheets or other tools you used for the transcriptions.

# Data Rescue working notes 

## Data Rescue Plan

+ Lead
  - [X] Read the article or data resource and the discription of the target project
  - [X] Create a fresh fork of the primary repository https://github.com/ktoddbrown/SoilDRaH/tree/main, it might be useful to rename it 'SoilDRaH_Austin2010'
  - [X] Create a copy of this template folder and rename it to Austin2010
  - [X] Fill out Data summary section (except for the Files), the Data Rescue Plan (here!), and update the author section of the yml to reflect who is working on the rescue
  - [X] Create a GitHub issue from the Data Rescue template and get feedback from a reviewer on the project (Dr Todd-Brown or a grad student typically).
  - [X] Add the reviewer to the author list in the yml
+ Do
  - [X] Extract points from Figure1 and export it to csv
  - [X] Transcribe or copy the method section
  - [X] Insert references to bib entries (note: none included in this paper's methods section.)
  - [X] reformat equations and mathematical symbols as [LaTex](https://en.wikibooks.org/wiki/LaTeX/Mathematics)
  - [X] reformat text using [Markdown](https://www.markdownguide.org/cheat-sheet/)
  - [X] Create a bib file with the citations from the methods section (note: none included in this paper's methods section.)
  - [X] Push to your local fork and get a second transcription pushed to your local fork by someone else on the project
+ Measure
  - [X] cross check with second transcriber for tables and methods, and resolve any differences
  - [ ] compile ReadMe with all chunks up through the Level 0 data rescue section with `eval: true`
  - [ ] push to your local fork and create a pull request to the main branch of the primary repository
  - [ ] address any revisions suggested in your pull request review
  - [ ] celebrate!

## Figure 1

Figure 1 is the only data retrieved from this paper as the rest of the data is using a synthetic substrate, and we are only interested in data on nonsynthetic litter decay.
Initial digitization for Figure 1 was performed with the [Plot Digitization app](https://plotdigitizer.com/app).
Values were rounded to one additional decimal place than the y-axis on the original graph.
For this figure, the original bar graph had a y-axis labeled to the hundredths place, so the outputs from the digitizer were rounded to the thousandths place.
The measurement in (% d$^{-1}$) was converted to (% per day) for ease of notation.

The Figure .csv has five columns: substrate, transmitted wavelengths in nm, Mass Loss in $\%$ per day, height of the error bar (SEM), and letters indicating significant differences for Tukey HSD post-hoc comparisons.

```{r figure1}

read_csv(file = figure1.file,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = figure1.file, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)

Figure1.df<- read_csv(file = figure1.file,
                   skip = 1)

ggplot(data = Figure1.df) + 
  geom_col(aes(x = `Transmitted Wavelengths (nm)`,
               y = `Mass loss (% per day)`)) + 
  geom_errorbar(aes(x = `Transmitted Wavelengths (nm)`,
                    ymin = `Mass loss (% per day)`,
                    ymax = `Mass loss (% per day) SEM`),
                width = 0.2) + 
  geom_text(
    aes(label = Letter, 
        x = `Transmitted Wavelengths (nm)`,
        y = `Mass loss (% per day) SEM` + 0.002),
    position = position_dodge(0.9),
    vjust = 0
  ) +
  facet_wrap(~Substrate,
             ncol = 1,
             scales = 'free_x')

```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareFigure1
#| eval: false

diff --strip-trailing-cr --suppress-common-lines -y temp/Austin2010_Figure1_LK.csv Austin2010_Figure1.csv
```

## Methods 

Any different to the cross check for the methods section will be output after this line.

```{bash eval=FALSE}
#| label: compareMethods
#| eval: false
diff --strip-trailing-cr --suppress-common-lines -y temp/Austin2010_Methods_LK.md Austin2010_Methods.md
```

### Methods  - Text

The final text transcription from the Methods section is rendered below.

```{r}
#| label: readMethods
#| output: asis
#| eval: true

# This revised section will append two ### to any existing headers here so that it renders nicely with the other headers in the document.

methodsLines <- readLines(methods.file) |>
  str_replace_all(pattern = '^#', replacement = '####')

cat(paste0(methodsLines, collapse = '\n'))

```

<!----
## Citation notes from Methods

Below are the citation notes from the methods section.
This generally doesn't read well in the final report so commenting it out here but it is very useful for tracking citations when you are doing the transcriptions.
Some citations are missing, those are noted here and how best guess were generated.
Think of this as a lab notebook documenting how you found each citation.

- [x] (Van Soest 1963) @VanSoest1963
  + Van Soest PJ (1963) Use of detergents in analysis of fibrous feeds II: A rapid method for the determination of fiber and lignin. J Assoc Of c Agr Chem 46:829–835.
  + direct export
---->

# Level 0 Data rescue

This level 0 data read parses the transcribed files with minimal cleaning or transformations.

```{r}
#| label: readLevel0
#This chunk has two purposes. 
#...1) Check the formatting by reading in everything 
#...2) Create a list of everything to process later in level 1

data.lvl0.ls <- list(
  #Read in a list of all the bib files
  citation = list(
    #Citation for the article transcriptions are pulled from
    primary = read.bib(file = primaryCitation.file), 
    #Citations for all referenced articles
    methods = read.bib(file = methodsCitation.file)
  ),
  #Read in the text transcription of the article's methods section
  method = read_lines(file = methods.file),
  #Read in the results as tables or figure transcriptions. This includes
  #...the caption as well as the tables themselves
  data = list(
    Table1 = list(
      #Read the caption as a text string. Captions are the first cell on 
      #...the first row.
      caption = read_csv(file = figure1.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      #Read in all the data, skipping the first row with the caption and read
      #...in the table as character. This element is a tibble (data.frame).
      primary = read_csv(file = figure1.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    ) #End 'Table1' element
    #Repeat for all tables and figures...
   ) #end 'data' element
)

```

The code is also in `readsource.file` and returned with the `dataLevel = 'level0'` argument.
Below verifies that the data from the function matches what is coded above.

```{r}
#| label: checkReadFunction_lvl0

if(file.exists(readsource.file)){
  datafunction.lvl0.ls <- readAuthorYYYY(dataDir = '.',
                                         dataLevel = 'level0')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function matches code here.')
  }else{
    print('There is a mismatch in the data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}
```

# Level 1 data - CURATION PROJECT

This data is identified as a data source for the ...
Broad classes of information that we are looking for include:

<!--- Include description of where the data is located in the source.--->
- Location: 
- Soil carbon stock: 
- Soil type: 
- Land use: 

Information from the methods and other sections are pulled into tables with line by line source citations.
This information can be values, units, methods, or other useful notes and is manually extracted from the transcriptions loaded into the level 0 data above.

The final format of the data frame here will include the following 

  + grouping ids
  + `of_variable`
  + `is_type`
  + `with_entry`
  + `from_source`

The `of_variable` being used here are:

 + citation
 
 <!--- include other variables being used, this should match vocabulary in the SoilDRaHVocabulary resources --->

```{r}
#| label: createLvl1
#| eval: false

#### Method-section only data ####
#Create a table for the data in the method-section that is not directly 
#...associated with any table. Often this is study level information.

studyMeta <- tribble(~of_variable, ~is_type, ~with_entry, ~from_source,
                   'region', 'value', 'Honaunau Forest on the southwestern slopes of Mauna Loa, island of Hawaii', paste('Method ln5:', paste(data.lvl0.ls$method[5], collapse = ' ')))

#### Table/Figure N ####
# Process the primary data from Table/Figure M by maintaining data groups and 
#...standardizing the variable name so that we can link this to the study 
#...metadata and data in the other tables when needed. Link in the methods 
#...and other information from the method-section as needed.

Table1Primary <- data.lvl0.ls$data$Table1$primary |>
  #Add a row id preserving row grouping -- not unique
  mutate(row_id = paste0('R', 1:n())) |>
  #Transform the table to long data by row group and column names.
  #...This allows us to then create the cross links between the tables.
  pivot_longer(cols = -row_id,
               names_to = 'column_name', values_to = 'with_entry',
               values_drop_na = TRUE) |> #drop missing values
  #Link the variable names to the column names using a switch statement
  mutate(of_variable = case_when(
           #Soil PH is given as an example below
           column_name == 'SOC' ~ 'soil_organic_carbon')) |>
  #Flag the source of this data as Table 1, this table is merged later with
  #...other data so this keeps track of where the data came from
  mutate(from_source = 'Table 1')

## Create the metadata for table one from the header information and cited methods
Table1Meta <- Table1Primary |>
  select(column_name, of_variable) |>
  unique() |>
  #Grab everything between the parentheses as units and attribute the source as the column names.
  mutate(unit = str_extract(column_name, pattern = '(?<=\\().*(?=\\))'),
         from_source = 'Table 1 column names.') |>
  #If there aren't units then drop the row
  filter(!is.na(unit)) |>
  #Don't keep the column name now that you have the units, cross link via the variable
  select(-column_name) |>
  #Stack into the table the methods for each variable that we can find
  bind_rows(
  tribble(~of_variable, ~method, ~from_source,
            'soil_organic_carbon', paste0(data.lvl0.ls$method[72:73], collapse = ' '), 'Methods ln72-73') ) |>
  #Stack onto the tables the controlled vocabulary used
  bind_rows(
    tribble(~of_variable, ~control_vocabulary, ~from_source,
            'soil_class', 'Histosol:USDA classification for histosol soil type|Andisols:USDA classification for andisol soil type', 'expert informed')
  ) |>
  # Push the dimensions of the variable into a single column. Making this a long table.
  pivot_longer(cols = c(unit, method, control_vocabulary),
               names_to = 'is_type',
               values_drop_na = TRUE,
               values_to = 'with_entry')

#### Create level 1 #####
## Pull everything together into by stacking the meta and primary data tables
data.lvl1.ls <- list(
  meta = bind_rows(studyMeta, Table1Meta),
  primary = bind_rows(Table1Primary)|>
    arrange(row_id, column_name,
            of_variable, is_type, with_entry, from_source)
  )

```

## Check for new variables

`of_variable` in the data files should be defined in the SoilDRaH Vocabulary files.

```{r}
#| label: check_of_variable

#Read in the list of current 'of_variable' entries
SoilDRaH_variable_list <- read_csv(file =of_variable.file,
                                   show_col_types = FALSE)

#compare the variables in the level 1 tables to the SoilDRaH variables
new_variables <- lapply(data.lvl1.ls, function(xx){
  setdiff(unique(xx$of_variable), 
          unique(SoilDRaH_variable_list$of_variable))
}) |>
  unlist()

#Check that there are no new variables
if(length(new_variables) == 0){
  print('All of_variables are in the SoilDRaH variable list. No additions required')
}else{
  cat(paste('Please start a issue for new variables [',
               new_variables, ']', sep = '', collapse = '\n'))
}
```

## Check for is_type

`is_type` in the data files should be defined in the SoilDRaH Vocabulary files.

```{r}
#| label: check_is_type

#read in the current types and what variable they are associated with
SoilDRaH_type_list <- read_csv(file = is_type.file,
                                   show_col_types = FALSE) |>
  #remove the project definition and only keep the variable and type
  select(of_variable, is_type) |>
  #variables are pushed together in the type defined table. Separate the
  #...entries here
  separate_longer_delim(cols = 'of_variable', delim = ',')

#Pull all the of_variable-is_type pairs from the source data
source_pairs <- tibble()

#If there are a large number of entries, consider replacing this for-loop.
#...However this avoids having to install the plyr or other libraries, and 
#...there are generally no more then 3-5 tables
for(dataTable in data.lvl1.ls){
  source_pairs <- dataTable |>
    select(of_variable, is_type) |> 
    unique() |>
    bind_rows(source_pairs)
}

undefined_data_type <- source_pairs |>
  #subset for variables that are not 'incomplete'
  semi_join(SoilDRaH_variable_list |>
              filter(project_status != 'incomplete'),
            by = join_by(of_variable)) |>
  #check for of_variable-is_type tuples that are not in the SoilDRaH type definitions
  anti_join(SoilDRaH_type_list,
            by = join_by(of_variable, is_type))

#Tell the user what you found.
if(nrow(undefined_data_type) == 0){
  print('All of_variables-is_types are in the SoilDRaH type list. No additions required')
}else{
  print(paste0('Please start a issue for new of_variable-is_type [', paste(paste(undefined_data_type$of_variable, undefined_data_type$is_type, sep = '-'), collapse = ','), ']'))
}
```

## Check function match

Check for an exact match with the read function.

```{r}
#| label: checkReadFunction_lvl1

if(file.exists(readsource.file)){
  datafunction.lvl1.ls <- readAuthorYYYY(dataDir = '.',
                                         dataLevel = 'level1')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function output matches ReadMe code for level 1.')
  }else{
    print('There is a mismatch in the level 1 data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}

```


# References

<!---- This space will be filled with the citations when rendered. Leave this blank. ---->

