---
title: "Data Rescue: Wang 2017"
format:
  html:
    toc: true
    eval: true #default to not evaluating the chunks, override this the template is completed
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - Wang2017_Methods.bib 
  - Wang2017.bib 
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  - id: JennavieveK #use the github name as an id
    name:
      given: Jennavieve
      family: Keefer
    affiliation:
      - ref: uf
    role:
      - "Transcription"
  - id: AmandaLopez14 #use the github name as an id
    name:
      given: Amanda
      family: Lopez
    affiliation:
      - ref: uf
    role:
      - "Transcription"

affiliations:
  - id: uf
    name: University of Florida
    city: Gainesville
    state: Florida
    country: USA
    postal-code: 32611
    url: https://www.ufl.edu/
    
timelog:
  - activity: 
      description: add general UF affilation and filled in portion of template
      who: JennavieveK 
      start: 2026-07-14 T1130
      duration: P60M
  - activity: 
      description: data summary and fit for purpose
      who: JennavieveK #github username
      #when you started and ended, can be just a year or down to the minute
      start: 2026-07-15 T1050
      #start: YYYY-MM-DD THH:MM
      #alternatively duration can be combined with a being/end
      duration: P45M
  - activity: 
      description: added figures 2 and 3, took out tables section
      who: AmandaLopez14 #github username
      #when you started and ended, can be just a year or down to the minute
      start: 2026-07-20 T1013
      #start: YYYY-MM-DD THH:MM
      #alternatively duration can be combined with a being/end
      duration: P50M
  - activity: 
      description: cross-checked figure 2
      who: JennavieveK 
      #when you started and ended, can be just a year or down to the minute
      start: 2026-08-06 T1115
      #start: YYYY-MM-DD THH:MM
      #alternatively duration can be combined with a being/end
      duration: P60M
  - activity: 
      description: cross-checked figure 3
      who: JennavieveK 
      #when you started and ended, can be just a year or down to the minute
      start: 2026-08-07 T1440
      #start: YYYY-MM-DD THH:MM
      #alternatively duration can be combined with a being/end
      duration: P70M
  - activity: 
      description: cross-checked methods section
      who: JennavieveK 
      #when you started and ended, can be just a year or down to the minute
      start: 2026-08-08 T1620
      #start: YYYY-MM-DD THH:MM
      #alternatively duration can be combined with a being/end
      duration: P120M
  - activity: 
      description: fixed pull-request changes
      who: JennavieveK 
      #when you started and ended, can be just a year or down to the minute
      start: 2026-09-06 T1200
      #start: YYYY-MM-DD THH:MM
      #alternatively duration can be combined with a being/end
      duration: P90M
version: 'vs202512' #template version, remove before use
---


```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false
#| eval: true


library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# data files for level 0
methods.file <- "Wang2017_Methods.md"
figure2.file <- 'Wang2017_Figure2.csv'
figure3.file <- 'Wang2017_Figure3.csv'

# Bibliography files
primaryCitation.file <- 'Wang2017.bib'
methodsCitation.file <- 'Wang2017_Methods.bib'

# read function; you will need to create this function after you have your data
#... rescue reviewed and before you finish your data harmonization.
readsource.file <- '../../02_DataHarmonization/readWang2017.R'
if(file.exists(readsource.file)){
  source(readsource.file)
}

#semantic files for level 1
#of_variable.file <- '../../semantics/SoilDRaHVocabulary_of_variable.csv'
#is_type.file <- '../../semantics/SoilDRaHVocabulary_is_type.csv'

```


# Data Summary

Wang, J., Liu, L., Wang, X., Yang, S., Zhang, B., Li, P., ... & Liu, W. (2017). High night‐time humidity and dissolved organic carbon content support rapid decomposition of standing litter in a semi‐arid landscape. Functional Ecology, 31(8), 1659-1668.https://doi.org/10.1111/1365-2435.12854

For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/251.

@Wang2017 primarily measures how varying positioning treatments, abiotic factors, and microbial activities affect the mass loss of C. squarrosa litter in a Chinese semi-arid temperate steppe. 
It demonstrates that standing litter decomposes significantly faster than ground-level litter due to higher night-time relative humidity, a trend that persists even after the litter falls to the soil surface.
They conducted a litterbag field experiment, and then some of that litter was freeze-dried for microbial community composition analysis.
Another portion was refrigerated for laboratory incubation of how the standing phase affects further decomposition with regards to standing vs surface litter.

## Tables and figures

- Table 
  1) Effects of position treatments on temperature and relative humidity (RH) in the semi-arid ecosystem.
  2) Soil organic carbon (SOC) formation during laboratory incubation for litter pre-treated under different position treatments.

- Figure 
  1) Temperature and precipitation at the study site and temperature and relative humidity (RH) on the surface of litter.
  2) Effects of position treatments on litter decomposition, litter mass remaining over time under different position treatments.
  3)  Litter dissolved organic carbon over time, litter position treatment effects on litter DOC.
  4) The relationships between the highest litter moisture, microbial respiration rate, night-time average relative humidity (RH), litter total phospholipid fatty acids (PLFAs) and litter decomposition rate.
  5) Litter decomposition during laboratory incubation for litter pre-treated under different position treatments.
  6) The potential mechanisms regulating decomposition and soil organic carbon (SOC) formation efficiency for soil surface litter (b) and standing litter (c) in semi-arid ecosystems.


## Fit for purpose: DeAD

This data is identified as a data source for DeAD.
The Decomposition Across Drylands (DeAD) data collection is a new data effort targeting dryland decomposition studies.
The purpose of this data collection is to develop better litter decay models for dryland ecosystems.

## Main Variables of Interest

  - Location: The study was conducted in a temperate steppe located at the Duolun Restoration Ecology Research Station, Inner Mongolia, China (42°02ʹ N, 116°17ʹ E, 1324 m a.s.l.).
  - Duration: Field experiment ran from 14 July to 15 September in 2014 and 11 May to 25 July in 2015. 
One litterbag was collected from each frame on days 31, 57, 88, 117, 318, 346, 384, 409, 474, 711, and 787.
Laboratory incubation ran for 192 days, with measurements taken at days 1, 2, 3, 4, 5, 7, 9, 11, 16, 25, 43, and 60.
  - Mass loss: Mass loss identified in Figure 2 and Figure 5 under the Statistical Analysis section of Materials and Methods section.
  - Solar exposure: Solar exposure identified in Materials and methods, under Field Experiment Design (290-390 nm).
  
### Other Relevant Variables

  - Litter species: Primary litter species was identified as C. squarrosa (C4 grass).
  - Microsite data: Identified in Field Experiment Design section of Materials and methods. 
The litterbags were placed horizontally on the surface of soil or on the supporting mesh, and above-ground vegetation was periodically clipped during growing seasons to avoid shading.
  - Litter chemical composition: The initial dissolved organic carbon (DOC) concentration, N concentration and C/N ratio of C. squarrosa litter were 12·44 mg g−1, 0·64% and 65·80, respectively.
  - Microclimate relative humidity: Identified in Table 1.
  - Litter bag materials and method: 1-mm × 1·5-mm nylon mesh litterbags (10 cm × 15 cm).
  
## Files

<!--- Include a brief explaination to all the files in the folder--->

These files are in the @Wang2017 data rescue.

- [Readme](README_Wang2017.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Wang2017.bib](Wang2017.bib)
  + citation for article transcribed
- [Wang2017_Methods.bib](Wang2017_Methods.bib)
  + citations for the methods section of the article
- [Wang2017_Methods.md](Wang2017_Methods.md)
  + methods transcribed from primary article
- [Wang2017_Figure2.csv](Wang2017_Figure2.csv)
  + Figure 2 from primary article
- [Wang2017_Figure3.csv](Wang2017_Figure3.csv)
  + Figure 3 from primary article
- [temp/](temp/)
    + scratch folder that will not be archived on GitHub, it should include local copies of the data resource (likely a PDF article) and any spreadsheets or other tools you used for the transcriptions.

# Data Rescue working notes 

## Data Rescue Plan

+ Lead
  - [x] Read the article or data resource and the discription of the target project
  - [x] Create a fresh fork of the primary repository https://github.com/ktoddbrown/SoilDRaH/tree/main, it might be useful to rename it 'SoilDRaH_Wang2017'
  - [x] Create a copy of this template folder and rename it to Wang2017
  - [x] Fill out Data summary section (except for the Files), the Data Rescue Plan (here!), and update the author section of the yml to reflect who is working on the rescue
  - [x] Create a GitHub issue from the Data Rescue template and get feedback from a reviewer on the project (Dr Todd-Brown or a grad student typically).
  - [ ] Add the reviewer to the author list in the yml
+ Do
  - [x] Extract points from Figure2 and export it to csv
  - [x] Extract points from Figure3 and export it to csv
  - [x] Transcribe or copy the method section
  - [x] Insert references to bib entries
  - [x] reformat equations and mathematical symbols as [LaTex](https://en.wikibooks.org/wiki/LaTeX/Mathematics)
  - [x] reformat text using [Markdown](https://www.markdownguide.org/cheat-sheet/)
  - [x] Create a bib file with the citations from the methods section
  - [x] Push to your local fork and get a second transcription pushed to your local fork by someone else on the project
+ Measure
  - [x] cross check with second transcriber for tables and methods, and resolve any differences
  - [x] compile or render this ReadMe with all chunks up through the Level 0 data rescue section with `eval: true`
  - [x] push to your local fork and create a pull request to the main branch of the primary repository
  - [ ] address any revisions in your pull request review
  - [ ] celebrate!

## Figure 2

<!--- Create a plot to make the figure easy to compare to the table. You don't have to be exactly the same but it should be obvious, at a glance, if the transcript matches.---->

Figure 2 was modified by removing the error bars, statistical significance marker, and inset plot. 

```{r}
#| label: Figure2Check

figure2_raw <- read_csv(file = figure2.file, col_types = cols(.default = col_character())
)

caption_text <- colnames(figure2_raw)[1]

figure2.df <- read_csv(file = figure2.file, #read in the specific file
         skip = 1, #skip the first line with the caption
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) |>
  mutate(across(c(`Time (days)`, `Litter mass remaining (%)`), as.numeric))

ggplot(data = figure2.df, 
       mapping = aes(x = `Time (days)`, y = `Litter mass remaining (%)`, group = `Litter type`, color = `Litter type`)) +
  geom_line() +
  geom_point() +
    labs(caption = caption_text)

```

And is crosschecked with the secondary transcription (expected to be empty)

```{bash}
#| label: compareFigure2

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.
#... Note that the template should not have a conflict here.

diff --strip-trailing-cr --suppress-common-lines -y temp/Wang2017_Figure2_JK.csv Wang2017_Figure2.csv
```

## Figure 3

<!--- Create a plot to make the figure easy to compare to the table. You don't have to be exactly the same but it should be obvious, at a glance, if the transcript matches.---->

Figure 3 was modified in graph a where the times period 400-500 days is highlighted as 2014 and 700-800 days is highlighted as 2015. 

```{r}
#| label: figure3Check
#| fig-height: 9
#| fig-width: 6.5

library(ggplot2)
library(dplyr)
library(grid)

# 1. Load Data & Clean Labels
figure3.df <- read_csv(
  file = figure3.file, 
  skip = 1, 
  col_types = cols(.default = col_character())
) |>
  mutate(
    `litter type` = trimws(`litter type`),
    `litter type` = case_when(
      grepl("surf", `litter type`, ignore.case = TRUE) ~ "Surface",
      grepl("stand", `litter type`, ignore.case = TRUE) ~ "Standing",
      TRUE ~ `litter type`
    ),
    panel_letter = trimws(tolower(gsub("[^a-zA-Z]", "", `graph label`))),
    panel_letter = substr(panel_letter, 1, 1),
    `x value` = as.numeric(`x value`),
    `DOC (mg g^-1)` = as.numeric(`DOC (mg g^-1)`)
  )

color_palette <- c("Surface" = "black", "Standing" = "grey50")

# 2. Panel (a): Top Plot
plot_a <- ggplot(
  data = figure3.df |> filter(panel_letter == "a"),
  mapping = aes(x = `x value`, y = `DOC (mg g^-1)`, color = `litter type`, group = `litter type`)
) +
  geom_line() +
  geom_point(size = 2.5) +
  scale_color_manual(values = color_palette) +
  scale_y_continuous(limits = c(0, 22)) + 
  labs(x = "Time (days)", y = expression(DOC~(mg~g^-1)), title = "(a)") +
  theme_classic() +
  theme(
    legend.position = "right",
    legend.title = element_blank()
  )

# 3. Middle Row: Panels (b) and (c) with Shared Y-Axis (0 to 8)
plot_b <- ggplot(
  data = figure3.df |> filter(panel_letter == "b"),
  mapping = aes(x = `x value`, y = `DOC (mg g^-1)`, color = `litter type`)
) +
  geom_point(size = 2.5) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  scale_color_manual(values = color_palette) +
  scale_y_continuous(limits = c(0, 8)) +
  labs(x = NULL, y = expression(DOC~(mg~g^-1)), title = "(b) 2014") +
  theme_classic() +
  theme(legend.position = "none")

plot_c <- ggplot(
  data = figure3.df |> filter(panel_letter == "c"),
  mapping = aes(x = `x value`, y = `DOC (mg g^-1)`, color = `litter type`)
) +
  geom_point(size = 2.5) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  scale_color_manual(values = color_palette) +
  scale_y_continuous(limits = c(0, 8)) +
  labs(x = NULL, y = NULL, title = "(c) 2014") +
  theme_classic() +
  theme(legend.position = "none")

# 4. Bottom Row: Panels (d) and (e) with Custom X-Axis Labels & Shared Y-Axis (0.5 to 2.5)
plot_d <- ggplot(
  data = figure3.df |> filter(panel_letter == "d"),
  mapping = aes(x = `x value`, y = `DOC (mg g^-1)`, color = `litter type`)
) +
  geom_point(size = 2.5) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  scale_color_manual(values = color_palette) +
  scale_y_continuous(limits = c(0.5, 2.5)) +
  labs(x = "CV of temperature", y = expression(DOC~(mg~g^-1)), title = "(d) 2015") +
  theme_classic() +
  theme(legend.position = "none")

plot_e <- ggplot(
  data = figure3.df |> filter(panel_letter == "e"),
  mapping = aes(x = `x value`, y = `DOC (mg g^-1)`, color = `litter type`)
) +
  geom_point(size = 2.5) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  scale_color_manual(values = color_palette) +
  scale_y_continuous(limits = c(0.5, 2.5)) +
  labs(x = "CV of RH", y = NULL, title = "(e) 2015") +
  theme_classic() +
  theme(legend.position = "none")

# 5. Render exact 3-row layout grid
grid::grid.newpage()
grid::pushViewport(grid::viewport(layout = grid::grid.layout(3, 2, heights = grid::unit(c(1, 0.8, 0.8), "null"))))

# Row 1: Plot (a) spanning both columns
print(plot_a, vp = grid::viewport(layout.pos.row = 1, layout.pos.col = 1:2))

# Row 2: Plots (b) and (c)
print(plot_b, vp = grid::viewport(layout.pos.row = 2, layout.pos.col = 1))
print(plot_c, vp = grid::viewport(layout.pos.row = 2, layout.pos.col = 2))

# Row 3: Plots (d) and (e) with distinct CV titles
print(plot_d, vp = grid::viewport(layout.pos.row = 3, layout.pos.col = 1))
print(plot_e, vp = grid::viewport(layout.pos.row = 3, layout.pos.col = 2))

```

And is crosschecked with the secondary transcription (expected to be empty)

```{bash}
#| label: compareFigure3

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.
#... Note that the template should not have a conflict here.

diff --strip-trailing-cr --suppress-common-lines -y temp/Wang2017_Figure3_JK.csv Wang2017_Figure3.csv
```

## Methods

Any different to the cross check for the methods section will be output after this line.

```{bash}
#| label: compareMethods
#| eval: false
# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.
diff --strip-trailing-cr --suppress-common-lines -y temp/Wang2017_Methods_JK.md Wang2017_Methods.md
```

### Methods  - Text

The final text transcription from the Methods section is rendered below.

```{r}
#| label: readMethods
#| output: asis
#| eval: true

# This revised section will append two ### to any existing headers here so that it renders nicely with the other headers in the document.

methodsLines <- readLines(methods.file) |>
  str_replace_all(pattern = '^#', replacement = '####')

cat(paste0(methodsLines, collapse = '\n'))

```

<!----
## Citation notes from Methods

Below are the citation notes from the methods section.
This generally doesn't read well in the final report so commenting it out here but it is very useful for tracking citations when you are doing the transcriptions.
Some citations are missing, those are noted here and how best guess were generated.
Think of this as a lab notebook documenting how you found each citation.

- [ ] short citation
  + copy from bib
  + manually entered or direct export
---->

# Level 0 Data rescue

This level 0 data read parses the transcribed files with minimal cleaning or transformations.

```{r}
#| label: readLevel0
#This chunk has two purposes. 
#...1) Check the formatting by reading in everything 
#...2) Create a list of everything to process later in level 1

data.lvl0.ls <- list(
  #Read in a list of all the bib files
  citation = list(
    #Citation for the article transcriptions are pulled from
    primary = read.bib(file = primaryCitation.file), 
    #Citations for all referenced articles
    methods = read.bib(file = methodsCitation.file)
  ),
  #Read in the text transcription of the article's methods section
  method = read_lines(file = methods.file),
  #Read in the results as tables or figure transcriptions. This includes
  #...the caption as well as the tables themselves
  data = list(
    Figure2 = list(
      #Read the caption as a text string. Captions are the first cell on 
      #...the first row.
      caption = read_csv(file = figure2.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      #Read in all the data, skipping the first row with the caption and read
      #...in the table as character. This element is a tibble (data.frame).
      primary = read_csv(file = figure2.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    ), #End 'Table1' element
    Figure3 = list(
      #Read the caption as a text string. Captions are the first cell on 
      #...the first row.
      caption = read_csv(file = figure3.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      #Read in all the data, skipping the first row with the caption and read
      #...in the table as character. This element is a tibble (data.frame).
      primary = read_csv(file = figure3.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    ) #End 'Table1' element
    #Repeat for all tables and figures...
   ) #end 'data' element
)

```

The code is also in `readsource.file` and returned with the `dataLevel = 'level0'` argument.
Below verifies that the data from the function matches what is coded above.

```{r}
#| label: checkReadFunction_lvl0

if(file.exists(readsource.file)){
  datafunction.lvl0.ls <- readWang2017(dataDir = '.',
                                         dataLevel = 'level0')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function matches code here.')
  }else{
    print('There is a mismatch in the data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}
```

# Level 1 data - CURATION PROJECT

This data is identified as a data source for the ...
Broad classes of information that we are looking for include:

<!--- Include description of where the data is located in the source.--->
- Location: 
- Soil carbon stock: 
- Soil type: 
- Land use: 

Information from the methods and other sections are pulled into tables with line by line source citations.
This information can be values, units, methods, or other useful notes and is manually extracted from the transcriptions loaded into the level 0 data above.

The final format of the data frame here will include the following 

  + grouping ids
  + `of_variable`
  + `is_type`
  + `with_entry`
  + `from_source`

The `of_variable` being used here are:

 + citation
 
 <!--- include other variables being used, this should match vocabulary in the SoilDRaHVocabulary resources --->

```{r}
#| label: createLvl1
#| eval: false

#### Method-section only data ####
#Create a table for the data in the method-section that is not directly 
#...associated with any table. Often this is study level information.

studyMeta <- tribble(~of_variable, ~is_type, ~with_entry, ~from_source,
                   'region', 'value', 'Honaunau Forest on the southwestern slopes of Mauna Loa, island of Hawaii', paste('Method ln5:', paste(data.lvl0.ls$method[5], collapse = ' ')))

#### Table/Figure N ####
# Process the primary data from Table/Figure 2 by maintaining data groups and 
#...standardizing the variable name so that we can link this to the study 
#...metadata and data in the other tables when needed. Link in the methods 
#...and other information from the method-section as needed.

Table1Primary <- data.lvl0.ls$data$Table1$primary |>
  #Add a row id preserving row grouping -- not unique
  mutate(row_id = paste0('R', 1:n())) |>
  #Transform the table to long data by row group and column names.
  #...This allows us to then create the cross links between the tables.
  pivot_longer(cols = -row_id,
               names_to = 'column_name', values_to = 'with_entry',
               values_drop_na = TRUE) |> #drop missing values
  #Link the variable names to the column names using a switch statement
  mutate(of_variable = case_when(
           #Soil PH is given as an example below
           column_name == 'SOC' ~ 'soil_organic_carbon')) |>
  #Flag the source of this data as Table 1, this table is merged later with
  #...other data so this keeps track of where the data came from
  mutate(from_source = 'Table 1')

## Create the metadata for table one from the header information and cited methods
Table1Meta <- Table1Primary |>
  select(column_name, of_variable) |>
  unique() |>
  #Grab everything between the parentheses as units and attribute the source as the column names.
  mutate(unit = str_extract(column_name, pattern = '(?<=\\().*(?=\\))'),
         from_source = 'Table 1 column names.') |>
  #If there aren't units then drop the row
  filter(!is.na(unit)) |>
  #Don't keep the column name now that you have the units, cross link via the variable
  select(-column_name) |>
  #Stack into the table the methods for each variable that we can find
  bind_rows(
  tribble(~of_variable, ~method, ~from_source,
            'soil_organic_carbon', paste0(data.lvl0.ls$method[72:73], collapse = ' '), 'Methods ln72-73') ) |>
  #Stack onto the tables the controlled vocabulary used
  bind_rows(
    tribble(~of_variable, ~control_vocabulary, ~from_source,
            'soil_class', 'Histosol:USDA classification for histosol soil type|Andisols:USDA classification for andisol soil type', 'expert informed')
  ) |>
  # Push the dimensions of the variable into a single column. Making this a long table.
  pivot_longer(cols = c(unit, method, control_vocabulary),
               names_to = 'is_type',
               values_drop_na = TRUE,
               values_to = 'with_entry')

#### Create level 1 #####
## Pull everything together into by stacking the meta and primary data tables
data.lvl1.ls <- list(
  meta = bind_rows(studyMeta, Table1Meta),
  primary = bind_rows(Table1Primary)|>
    arrange(row_id, column_name,
            of_variable, is_type, with_entry, from_source)
  )

```

## Check for new variables

`of_variable` in the data files should be defined in the SoilDRaH Vocabulary files.

```{r}
#| label: check_of_variable
#| eval: false

#Read in the list of current 'of_variable' entries
SoilDRaH_variable_list <- read_csv(file =of_variable.file,
                                   show_col_types = FALSE)

#compare the variables in the level 1 tables to the SoilDRaH variables
new_variables <- lapply(data.lvl1.ls, function(xx){
  setdiff(unique(xx$of_variable), 
          unique(SoilDRaH_variable_list$of_variable))
}) |>
  unlist()

#Check that there are no new variables
if(length(new_variables) == 0){
  print('All of_variables are in the SoilDRaH variable list. No additions required')
}else{
  cat(paste('Please start a issue for new variables [',
               new_variables, ']', sep = '', collapse = '\n'))
}
```

## Check for is_type

`is_type` in the data files should be defined in the SoilDRaH Vocabulary files.

```{r}
#| label: check_is_type
#| eval: false

#read in the current types and what variable they are associated with
SoilDRaH_type_list <- read_csv(file = is_type.file,
                                   show_col_types = FALSE) |>
  #remove the project definition and only keep the variable and type
  select(of_variable, is_type) |>
  #variables are pushed together in the type defined table. Separate the
  #...entries here
  separate_longer_delim(cols = 'of_variable', delim = ',')

#Pull all the of_variable-is_type pairs from the source data
source_pairs <- tibble()

#If there are a large number of entries, consider replacing this for-loop.
#...However this avoids having to install the plyr or other libraries, and 
#...there are generally no more then 3-5 tables
for(dataTable in data.lvl1.ls){
  source_pairs <- dataTable |>
    select(of_variable, is_type) |> 
    unique() |>
    bind_rows(source_pairs)
}

undefined_data_type <- source_pairs |>
  #subset for variables that are not 'incomplete'
  semi_join(SoilDRaH_variable_list |>
              filter(project_status != 'incomplete'),
            by = join_by(of_variable)) |>
  #check for of_variable-is_type tuples that are not in the SoilDRaH type definitions
  anti_join(SoilDRaH_type_list,
            by = join_by(of_variable, is_type))

#Tell the user what you found.
if(nrow(undefined_data_type) == 0){
  print('All of_variables-is_types are in the SoilDRaH type list. No additions required')
}else{
  print(paste0('Please start a issue for new of_variable-is_type [', paste(paste(undefined_data_type$of_variable, undefined_data_type$is_type, sep = '-'), collapse = ','), ']'))
}
```

## Check function match

Check for an exact match with the read function.

```{r}
#| label: checkReadFunction_lvl1

if(file.exists(readsource.file)){
  datafunction.lvl1.ls <- readWang2017(dataDir = '.',
                                         dataLevel = 'level1')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function output matches ReadMe code for level 1.')
  }else{
    print('There is a mismatch in the level 1 data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}

```


# References

<!---- This space will be filled with the citations when rendered. Leave this blank. ---->

