---
title: "Data Curation: DeAD 2025"
format:
  html:
    toc: true
    eval: false #default to not evaluating the chunks, override this the template is completed
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - DeAD2025.bib #contains citation that provide background on the project
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  - id: ktoddbrown #use the github name as an id
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role: 
      - curation: lead #review
  - id: AshBonner310 #use the github name as an id
    name:
      given: Ashley
      family: Bonner
    orcid: 0000-0001-7882-6353
    affiliation:
      - ref: uf-ees
    role: 
      - curation: lead #review

affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
timelog:
  - activity: 
      description: Drafted initial template
      who: ktoddbrown #github username
      #when you started and ended, can be just a year or down to the minute
      end: 2025-12-07T1637
      #start: YYYY-MM-DD THH:MM
      #alternatively duration can be combined with a being/end
      #duration: P1H15M
---


```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false


library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# Bibliograph files
refernceCitation.file <- 'ProjectYYYY.bib'

#semantic files for level 1
of_variable.file <- '../../semantics/SoilDRaHVocabulary_of_variable.csv'
is_type.file <- '../../semantics/SoilDRaHVocabulary_is_type.csv'

```

# Purpose

The Decomposition Across Drylands (DeAD) data collection is a new data effort targeting dryland decomposition studies. 
The purpose of this data collection is to develop better litter decay models for dryland ecosystems.
We are primarily interested in litter bag studies currently, but please contact us in the discussions if you have other datasets of interest.

# Variables of interest

<!--- Briefly describe what you want the data rescuers to look for when digitizing a data resource.---->

  * **Location**: latitude/longitude, associated datum, geospatial method, precision
  * **Duration**: start and end time of exposure to the day resolution if possible, month is acceptable
  * **Mass loss**: Mass loss or decay rate or mass remaining of litter sample over duration
  * **Solar exposure**: Degree of shading of litter bags relative to full sun

We anticipate the following information to also be relevant for reanalysis.

  - *Litter species*
  - Microsite data including *vegetation coverage*
  - *Litter chemical composition* including lignin content and C:N ration
  - *Extracellular enzyme* and microbial *'omics*
  - Microclimate *relative humidity*
  - Litter bag *materials* and method including litter bag area and *mesh size*
  - *Treatments* including rain out shelters or canopy shading or filters
  - *Age* of ecosystem
  - *Land management*/use or disturbances

# Data sources

For this project, we have identified the following data sources through personal communications with H Throop associated with @Throop2025.
We are focusing on datasets that are not in [Aridec](https://essd.copernicus.org/articles/14/3471/2022/).

<!---- Move here to final table when data curation in process. ---->

## Data Harmonized - tupled

<!---- Move to this table when the data tuple harmonization is complete. --->

| AuthorYYYY | citation | Data Rescue - GitHub Issue | Folder link | Data Harmonization - GitHub Issue |
|------------|-----------|----------|----------|--------|

## Data Rescue - digitized

<!---- Move to this table when the data rescue is complete and it is ready for tuple harmonization. --->

| AuthorYYYY | citation | Data Rescue - GitHub Issue | Folder link |
|------------|-----------|----------|----------|


## In process

<!---- move files into this table when someone takes lead on the data rescue, update to add in git hub issue when it is create --->

| AuthorYYYY | citation | Data Rescue - GitHub Issue | lead point of contact|
|------------|----------|-----------|----------|
| Ares2001 | \@Ares2001 | [77](https://github.com/ktoddbrown/SoilDRaH/issues/77) | ktoddbrown |
<!---- remove and replace the example above ---->

## Imported to bib

<!---- Move here when data source bib entry has been made ---->

| AuthorYYYY | citation | 
|------------|------------|
| Austin2010 | @Austin2010 |
| Wang2017 | @Wang2017 |
| Santos1984 | @Santos1984 |
| Barba 2016 | @Barba2016 |
| Martínez 2010 | @Martinez2010 |
| Blanco 2011 | @Blanco2011 |
| Rozema 1997 | @Rozema1997 |
| LadoMonserrat 2016 | @LadoMonserrat2016 |
| Kemp 2003 | @Kemp2003 |
|Lyons 2010 | @Lyons2010 |
| Almagro 2012 | @Almagro2012 |
| Duguay 2000 | @Duguay2000 |
| Dirks 2010 | @Dirks2010 |
| BravoOvideo 2017 | @BravoOvideo2017 |
| Day 2007 | @Day2007 |
| Dimitrakopoulos 2010 | @Dimitrakopoulos2010 |
| Gaxiola 2015 | @Gaxiola2015 |
| Kavvadias 2001 | @Kavvadias2001 |
| Kumada 2008 | @Kumada2008 |
| Gliksman 2016 | @Gliksman2016 |
| Lin 2015 | @Lin2015 |
| Qasemian 2014 | @Qasemian2014 |
| Sagi 2019 | @Sagi2019 |
| Bates 2007 | @Bates2007 |
| dosSantos 2019 | @dosSantos2019 |
| Bing 2018 | @Bing2018 |
| Fraser 2013 | @Fraser2013 | 
| Huang 2017 | @Huang2017 |
| Lin 2018 | @Lin2018 |
| Mao 2016 | @Mao2016 |
| HerreraArreola 2007 | @HerreraArreola2007 |
| Gallo 2009 | @Gallo2009 |
| Kotilainen 2009 | @Kotilainen2009 |
| Smith 2010 | @Smith2010 |
| DíazGuerra 2018 | @DiazGuerra2018 |
| Lee 2014 | @Lee2014 |
| Barnes 2012 | @Barnes2012 |
| Gallo 2006 | @Gallo2006 |
| Kirschbaum 2011 | @Kirschbaum2011 |
| delCid 2019 | @delCid2019 |
| Foereid 2010 | @Foereid2010 |
| LaporteFauret 2021 | @LaporteFauret2021 |
| Arroita 2013 | @Arroita2013 |
| Brandt 2009 | @Brandt2009 |

## Identified

# Data Level 1 

```{r}
#Source and append all the level 1 tuples

```

# Data Level 2

```{r}
#Map the comparable methods and align any needed vocabulary
```

```{r}
#Data viz checks
```

# References

<!---- This will be populated when you render the document.---->
