---
title: "Data Rescue: ISCN3 Feb 2026"
format:
  html:
    toc: true
    code-fold: true
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - ISCN3_Contributors.bib  #contains the citations for contributed datasets by authors
  - ISCN3_2015.bib #contains citation for the data source, generally a single citation of the primary paper
nocite: |
  @* #add all citations
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  - id: ktoddbrown #use the github name as an id
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role: #TODO define these roles and add reference
      - transcription: lead #secondary, review
      - standardization: lead #review
      - curation: lead #review
  - id: vaasukimarupaka
    name:
      given: Vaasuki
      family: Marupaka
    orcid: 0000-0002-3854-1764
    affiliation:
      - ref: uf-ees
    role:
      - transcription: lead #secondary, review
      - standardization: lead
      - curation: lead #review
affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
    
timelog:
  - activity: 
      description: Lead Transcription and ReadMe Update
      who: Vaasuki Marupaka
      end: 2026-03-25
      start: 2026-02-06
      duration: PT30H
      timelog:
---

```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false
#| eval: true


# libraries
library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R
library(stringr) # regular expressions

# data directory, primary annotations & control vocabularies
dataDownload.dir <- 'temp'
controlVocabulary <- 'ISCN3_TemplateCVs.csv'
rescuedContributedCitations <- 'ISCN3_ISCN3_citation.csv'
readME_templateSubmit <- 'README_TemplateSubmit.md'

# The below csv tables are used to construct annotations for ISCN4
AnnotationsDir <- '.'
ISCN3_layerAnnotations <- 'ISCN3_TemplateSubmit_TableLayer.csv'
ISCN3_profileAnnotations <- 'ISCN3_TemplateSubmit_TableProfile.csv'
ISCN3_siteAnnotations <- 'ISCN3_TemplateSubmit_TableSite.csv'
ISCN3_disturbanceAnnotations <- 'ISCN3_TemplateSubmit_TableDisturbance.csv'
ISCN3_fractionAnnotations <- 'ISCN3_TemplateSubmit_TableFraction.csv'

annotationFilenames <- c(layerAnnotations = ISCN3_layerAnnotations,
                          profileAnnotations = ISCN3_profileAnnotations,
                          siteAnnotations = ISCN3_siteAnnotations,
                          disturbanceAnnotations = ISCN3_disturbanceAnnotations,
                          fractionAnnotations = ISCN3_fractionAnnotations)

# Bibliograph files
primaryCitation.file <- 'ISCN3_2015.bib'
contributionsCitation.file <- 'ISCN3_Contributors.bib'

# read function
readsource.file <- 'readISCN3.R'
if(file.exists(readsource.file)){
  source(readsource.file)
}

#semantic files for level 1
of_variable.file <- '../../02_DataHarmonization/SoilDRaHVocabulary_of_variable.csv'
is_type.file <- '../../02_DataHarmonization/SoilDRaHVocabulary_is_type.csv'

```

# Data Summary

> The ISCN is an international scientific community devoted to the advancement of soil carbon research. The ISCN manages an open-access, community-driven soil carbon database. This is version 3-1 of the ISCN Database, released in December 2015. It gathers 38 separate data set contributions, totaling 67,112 sites with data from 71,198 soil profiles and 431,324 soil layers.

*Data source citation:* Nave, L., K. Johnson, C. van Ingen, D. Agarwal, M. Humphrey, and N. Beekwilder. 2022. International Soil Carbon Network version 3 Database (ISCN3) ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/cc751923c5576b95a6d6a227d5afe8ba (Accessed 2026-03-01).

## Fit for purpose: ISCN4

This data is identified as a data source for ISCN2025.
The purpose ISCN 2025 (aka ISCN4) is to create a data product aggregating data on geo-located soil organic carbon profiles.
With this data we will examine the assumption of steady state on field surveyed organic carbon.

ISCN3 contains four primary data tables:

  * `profile` containing the profile information including geolocation and site information, 
  * `layer` containing layer level soil characterizations, 
  * `citation` containing the citation and additional references information for the contributed datasets,
  * `dataset` with study information that extends the basic citation

## Files

<!--- Include a brief explanation to all the files in the folder--->

These files are in the ISCN3 data rescue.

-   [Readme](README_ISCN3.qmd)
    -   This is the primary file that documents the transcriptions and decision made during data rescue.
-   [ISCN3_2015.bib](ISCN3_2015.bib)
    -   citation for project website and primary journal article transcribed
-   [ISCN3_Contributors.bib](ISCN3_Contributors.bib)
    -   citations for the contributed datasets to ISCN
-   [AnnotationsDir](AnnotationsDir)
    -   this includes the annotations constructed using template.pdfs tables pulled from edi with individual tables (layer, profile, site, fraction, and disturbance)
        + [ISCN3_TemplateSubmit_TableSite.csv](ISCN3_TemplateSubmit_TableSite.csv)
        + [ISCN3_TemplateSubmit_TableProfile.csv](ISCN3_TemplateSubmit_TableProfile.csv)
        + [ISCN3_TemplateSubmit_TableLayer.csv](ISCN3_TemplateSubmit_TableLayer.csv)
        + [ISCN3_TemplateSubmit_TableDisturbance.csv](ISCN3_TemplateSubmit_TableDisturbance.csv)
        + [ISCN3_TemplateSubmit_TableFraction.csv](ISCN3_TemplateSubmit_TableFraction.csv)
        + [ISCN3_TemplateCVs.csv](ISCN3_TemplateCVs.csv)
        + [README_TemplateCV.md](README_TemplateCV.md)
        + [README_TemplateSubmit.md](README_TemplateSubmit.md)
-   [temp/](temp/)
    -   the download location for ISCN3 files archived on EDI web portal
    -   this folder includes the following files:
        + [ISCN3_citation.csv](ISCN3_citation.csv)
        + [ISCN3_dataset.csv](ISCN3_dataset.csv)
        + [ISCN3_layer.csv](ISCN3_layer.csv)
        + [ISCN3_profile.csv](ISCN3_profile.csv)
        + [ISCNtemplate.xlsx](ISCNtemplate.xlsx)
        + [ISCNTranscribed_TemplateCVs.xlsx](ISCNTranscribed_TemplateCVs.xlsx)
        + [TemplateCVs.pdf](TemplateCVs.pdf)
        + [TemplateSubmit.pdf](TemplateSubmit.pdf)

# Data Rescue working notes 

## Data Rescue Plan

+ Lead
  - [x] Read the article or data resource and the discription of the target project
  - [x] Create a fresh fork of the primary repository https://github.com/ktoddbrown/SoilDRaH/tree/main, it might be useful to rename it 'SoilDRaH_AuthorYYYY'
  - [x] Create a copy of this template folder and rename it to AuthorYYYY
  - [x] Fill out Data summary section (except for the Files), the Data Rescue Plan (here!), and update the author section of the yml to reflect who is working on the rescue
  - [x] Create a GitHub issue from the Data Rescue template and get feedback from a reviewer on the project (Dr Todd-Brown or a grad student typically).
  - [x] Add the reviewer to the author list in the yml
+ Do
  - [x] Code up data download from EDI
  - [x] Transcribe the EDI tables in `TemplateSubmit.pdf` into `ISCN3_TemplateSubmit_TableX.csv` 
  - [ ] Transcribe `TemplateSubmit.pdf` markdown for preamble text
  - [X] Transcribe the EDI tables in `TemplateCV.pdf` into `ISCN_TemplateCV.csv`
  - [ ] Transcribe `TemplateCV.pdf` markdown for preamble text
  - [X] Create a bib file with the recommended citations
  - [ ] Replace the citations with bib references and treat `ISCN3_ISCN3_citation.csv` as a rescue object 
  - [ ] Create a recommended citations and acknowledgements table based on 'citations' as a markdown document that inserts the bib entries
  - [ ] Read in primary tables and template information 
  - [ ] read function match `02_DataHarmonization/readAres2001.R` function inputs and output formats
  - [x] push annotations into an `old` folder as a place holder
  - [x] ask one of the undergrads to provide a secondary transcription for the TemplateCV and TemplateSubmit pdfs
  - [ ] push everything to the lead fork
+ Measure
  - [ ] cross check with second transcriber for tables and methods, and resolve any differences
  - [ ] compile or render this ReadMe with all chunks up through the Level 0 data rescue section with `eval: true`
  - [ ] push to your local fork and create a pull request to the main branch of the primary repository
  - [ ] address any revisions in your pull request review
  - [ ] celebrate!

## Citation notes

  - `AK DSC Project SOC stock computation` had a bad url for the doi that linked to a spam site. Updating doi with the EDI pasta link.

## Download ISCN3

The ISCN3 is downloaded from EDI (@ISCN3). Below we loop through the files that are expected and download them.

```{r}
#| label: dataDownloadEDI
#| code-summary: "Download and identify each file in the EDI package."

### Download the data ####
# link the table name to the download url and file path(s) for data package
pasta_package <- 'https://pasta.lternet.edu/package/data/eml/edi/1160/1/'

download_table <- tibble::tribble(
  ~table, ~entity_id, ~file_name, 
  'layer', "4af719a84f8981fcc63f1f92760cb253", 'ISCN3_layer.csv',
  'profile', "40527580cc045d33d9a5aaf728bf204e", 'ISCN3_profile.csv',
  'citation', "320e31ca911f187550ca2143c31fd408", 'ISCN3_citation.csv',
  'dataset', "cdd0c7a4cac3f28d6d788c91f506775f", 'ISCN3_dataset.csv',
  'control_vocabulary', "a8eef6e94b669b365e443c15d9402a03", 'ISCNTranscribed_TemplateCVs.xlsx',
  'template', "cb5c1cd86e9ab17f699f7a05044325b7", 'ISCNtemplate.xlsx',
  'other_pdfs', "eb320ae7b57296765f543cbb370b0f24", 'TemplateCVs.pdf',
  'other_pdfs', "b81a7f4214d176280a5ff4e0f0d52d8b", 'TemplateSubmit.pdf')

#One by one, download the data in each row from the download_table
#... and store that in the appropriate file path location on your computer
for(row_index in 1:nrow(download_table)){
  if(!file.exists(file.path(dataDownload.dir,download_table$file_name[row_index]))){
    print(paste('Download file:', 
                download_table$file_name[row_index]))
    utils::download.file(
      paste0(pasta_package, download_table$entity_id[row_index]), 
      file.path(dataDownload.dir,download_table$file_name[row_index]), 
      extra  = "--max-time 300")
  }
}

originalTables <- plyr::dlply(
    .data = download_table |>
      dplyr::filter(table %in% c('citation', 'dataset', 'profile', 'layer')), 
    .variables = c('table'),
    .fun = function(xx){
      return(
        readr::read_delim(file = file.path(dataDir = dataDownload.dir, 
                                           xx$file_name), 
                          delim = ';', 
                          col_types = readr::cols(.default = readr::col_character())))
    }
  )

```

Using the [TemplateCVs.pdf](TemplateCVs.pdf) and [TemplateSubmit.pdf](TemplateSubmit.pdf), we manually transcribe the annotations into different tables; layer, profile, site, disturbance, and fraction for level0. These annotations will further be extended to a standard data model for our level01 data.

```{r}
annotations <- list(
  caption = readr::read_lines(file = readME_templateSubmit),
  ISCN3_annotations = lapply(annotationFilenames, function(xx) {
    readr::read_delim(
      file = file.path(AnnotationsDir, xx),
      delim = ",",
      col_types = readr::cols(.default = readr::col_character())
    )
  }))

### Read in control vocabulary###
#Read in the results as tables or figure transcriptions. This includes
#...the caption as well as the tables themselves
controlVocabulary.ls <- list(
  caption = readr::read_delim(file = controlVocabulary,
                       col_types = cols(.default = col_character()),
                       n_max = 5, col_names = FALSE)$X1[1:5],
  primary_CV = readr::read_delim(file = controlVocabulary,
    delim = ',', 
    skip = 5, # we want to only keep the headers without the preamble text.
    col_types = readr::cols(.default = readr::col_character())))

```

# Level 0 data

```{r}
#This chunk has two purposes. 
#...1) Check the formatting by reading in everything 
#...2) Create a list of everything to process later in level 1

data.lvl0.ls <- list(
  #Read in a list of all the bib files
  citation.ls = list(
    #Citation for the article transcriptions are pulled from
    primary = read.bib(file = primaryCitation.file), 
    #Citations for all referenced articles
    contributed = read.bib(file = contributionsCitation.file),
    #Read in the bib transcription for the contributed bib rescued object
    rescuedCitations = read_delim(file = rescuedContributedCitations,
                                skip = 1)
  ),
  primaryData.ls = originalTables,
  primaryAnnotations.ls = annotations,
  controlVocabulary.ls = controlVocabulary.ls)

```

## Primary data: layer table

```{r}
knitr::kable(data.lvl0.ls$primaryData.ls$layer |> 
               head()) |>
  kable_paper() |>
  scroll_box(width = "100%", height = "300px")

```

## Primary annotations: layer table

```{r}
knitr::kable(data.lvl0.ls$primaryAnnotations.ls$ISCN3_annotations$layerAnnotations |> 
               head()) |>
  kable_paper() |>
  scroll_box(width = "100%", height = "300px")

```

## Primary rescued citations

```{r}
knitr::kable(data.lvl0.ls$citation.ls$rescuedCitations |> 
               head()) |>
  kable_paper() |>
  scroll_box(width = "100%", height = "300px")

```
