---
title: "Data Rescue: ISCN3"
format:
  html:
    toc: true
    code-fold: true
    eval: true
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - ISCN3_Contributors.bib  #contains the citations for contributed datasets by authors
  - ISCN3_2015.bib #contains citation for the data source, generally a single citation of the primary paper
nocite: |
  @* #add all citations
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  - id: ktoddbrown #use the github name as an id
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role: #TODO define these roles and add reference
      - transcription: lead #secondary, review
      - standardization: lead #review
      - curation: lead #review
  - id: vaasukimarupaka
    name:
      given: Vaasuki
      family: Marupaka
    orcid: 0000-0002-3854-1764
    affiliation:
      - ref: uf-ees
    role:
      - transcription: lead #secondary, review
      - standardization: lead
      - curation: lead #review
affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
    
timelog:
  - activity: 
      description: Lead Transcription and ReadMe Update
      who: Vaasuki Marupaka
      end: 2026-03-25
      start: 2026-02-06
      duration: PT30H
  - activity: 
      description: Revise readme
      who: ktoddbrown
      end: 2026-06-02T16:28
      start: 2026-06-02
---

```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false
#| eval: true


# libraries
library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R
library(stringr) # regular expressions

# Data download directory from EDI
dataDownload.dir <- 'temp'

### Level 0 rescued files ####
# These files have some level of manual rescue/transcription to them

# Control vocabulary transcribed from EDI file `TemplateCVs.pdf`
controlVocabulary <- 'ISCN3_TemplateCVs.csv'

## Files from the `TemplateSubmit.pdf` document include a formatted header
##... and the table discriptions

# Transcription of the header from the EDI file `TemplateSubmit.pdf`
readME_templateSubmit <- 'README_TemplateSubmit.md'

# Rescued transcriptions of individual table meta data from the 
#... EDI file `TemplateSubmit.pdf` for the layer, profile, site
#... disturbance, and fraction tables. These do not match exactly the 
#... EDI archived tables but do provide the metadata for those tables
ISCN3_layerAnnotations <- 'ISCN3_TemplateSubmit_TableLayer.csv'
ISCN3_profileAnnotations <- 'ISCN3_TemplateSubmit_TableProfile.csv'
ISCN3_siteAnnotations <- 'ISCN3_TemplateSubmit_TableSite.csv'
ISCN3_disturbanceAnnotations <- 'ISCN3_TemplateSubmit_TableDisturbance.csv'
ISCN3_fractionAnnotations <- 'ISCN3_TemplateSubmit_TableFraction.csv'

#### Files that have to do with the fixed citations

# Transcribed citations from the EDI file `ISCN3_citation.csv`
#...table and then extended manually when files were mis-cited
rescuedContributedCitations <- 'ISCN3_ISCN3_citation.csv'

# Bibliography files
primaryCitation.file <- 'ISCN3_2015.bib'
contributionsCitation.file <- 'ISCN3_Contributors.bib'

### Reference to the read function
# read function
readsource.file <- '../../02_DataHarmonization/readISCN3.R'
if(file.exists(readsource.file)){
  source(readsource.file)
}

#semantic files for level 1
of_variable.file <- '../../02_DataHarmonization/SoilDRaHVocabulary_of_variable.csv'
is_type.file <- '../../02_DataHarmonization/SoilDRaHVocabulary_is_type.csv'

level1_controlVocabulary <- list(soc = 'carbon',
                                 bd = 'bulk_density',
                                 cf = 'coarse_fraction',
                                 layer = 'layer',
                                 geolocation = 'geolocation',
                                 om = 'organic_fraction')

```

# Data Summary

> The ISCN is an international scientific community devoted to the advancement of soil carbon research. The ISCN manages an open-access, community-driven soil carbon database. This is version 3-1 of the ISCN Database, released in December 2015. It gathers 38 separate data set contributions, totaling 67,112 sites with data from 71,198 soil profiles and 431,324 soil layers.

*Data source citation:* Nave, L., K. Johnson, C. van Ingen, D. Agarwal, M. Humphrey, and N. Beekwilder. 2022. International Soil Carbon Network version 3 Database (ISCN3) ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/cc751923c5576b95a6d6a227d5afe8ba (Accessed 2026-03-01).

## Fit for purpose: ISCN4

This data is identified as a data source for ISCN2025.
The purpose ISCN 2025 (aka ISCN4) is to create a data product aggregating data on geo-located soil organic carbon profiles.
With this data we will examine the assumption of steady state on field surveyed organic carbon.

ISCN3 contains four primary data tables:

  * `profile` containing the profile information including geolocation and site information, 
  * `layer` containing layer level soil characterizations, 
  * `citation` containing the citation and additional references information for the contributed datasets,
  * `dataset` with study information that extends the basic citation

The EDI data archive also includes metadata in submission templates.
This information is in pdfs files and required a manual transcription.

## Files

<!--- Include a brief explanation to all the files in the folder--->

These files are in the ISCN3 data rescue.

- [Readme_ISCN3.qmd](README_ISCN3.qmd) This is the primary file that documents the transcriptions and decision made during data rescue.
- [ISCN3_2015.bib](ISCN3_2015.bib) citation for project website and primary journal article transcribed
- [ISCN3_Contributors.bib](ISCN3_Contributors.bib) citations for the contributed datasets to ISCN
- [ISCN3_ISCN3_citation.csv](ISCN3_ISCN3_citation.csv) Rescued citations where the citations were replaced with bib references
- [README_TemplateCV.md](README_TemplateCV.md) control vocabulary text information
- [ISCN3_TemplateCVs.csv](ISCN3_TemplateCVs.csv) control vocabulary tables
- [README_TemplateSubmit.md](README_TemplateSubmit.md)  submission guidance header
- [ISCN3_TemplateSubmit_TableSite.csv](ISCN3_TemplateSubmit_TableSite.csv) submission guidence that has meta data for the site table
- [ISCN3_TemplateSubmit_TableProfile.csv](ISCN3_TemplateSubmit_TableProfile.csv) submission guidance that has meta data for the profile table
- [ISCN3_TemplateSubmit_TableLayer.csv](ISCN3_TemplateSubmit_TableLayer.csv) submission guidance that has meta data for the layer table
- [ISCN3_TemplateSubmit_TableDisturbance.csv](ISCN3_TemplateSubmit_TableDisturbance.csv) submission guidance that has meta data for the disturbance table (not in ISCN3 tables)
- [ISCN3_TemplateSubmit_TableFraction.csv](ISCN3_TemplateSubmit_TableFraction.csv) submission guidance that has meta data for the fraction table (not in ISCN3 tables)
- [temp/README.md](temp/README.md) a placeholder file for the download folder that contains the EDI citation.

The ISCN3 files are archived on EDI web portal and will be downloaded to the `temp` folder in the chunk below:

  + [ISCN3_citation.csv](ISCN3_citation.csv)
  + [ISCN3_dataset.csv](ISCN3_dataset.csv)
  + [ISCN3_layer.csv](ISCN3_layer.csv)
  + [ISCN3_profile.csv](ISCN3_profile.csv)
  + [ISCNtemplate.xlsx](ISCNtemplate.xlsx)
  + [ISCNTranscribed_TemplateCVs.xlsx](ISCNTranscribed_TemplateCVs.xlsx)
  + [TemplateCVs.pdf](TemplateCVs.pdf) transcribed source pdf
  + [TemplateSubmit.pdf](TemplateSubmit.pdf) transcribed source pdf

<!---- # Data Rescue working notes 

## Data Rescue Plan

+ Lead
  - [x] Read the article or data resource and the discription of the target project
  - [x] Create a fresh fork of the primary repository https://github.com/ktoddbrown/SoilDRaH/tree/main, it might be useful to rename it 'SoilDRaH_AuthorYYYY'
  - [x] Create a copy of this template folder and rename it to AuthorYYYY
  - [x] Fill out Data summary section (except for the Files), the Data Rescue Plan (here!), and update the author section of the yml to reflect who is working on the rescue
  - [x] Create a GitHub issue from the Data Rescue template and get feedback from a reviewer on the project (Dr Todd-Brown or a grad student typically).
  - [x] Add the reviewer to the author list in the yml
+ Do
  - [x] Code up data download from EDI
  - [x] Transcribe the EDI tables in `TemplateSubmit.pdf` into `ISCN3_TemplateSubmit_TableX.csv` 
  - [x] Transcribe `TemplateSubmit.pdf` markdown for preamble text
  - [x] Transcribe the EDI tables in `TemplateCV.pdf` into `ISCN_TemplateCV.csv`
  - [x] Transcribe `TemplateCV.pdf` markdown for preamble text
  - [x] Create a bib file with the recommended citations
  - [x] Replace the citations with bib references and treat `ISCN3_ISCN3_citation.csv` as a rescue object 
  - [ ] Create a recommended citations and acknowledgements table based on 'citations' as a markdown document that inserts the bib entries
  - [ ] Read in primary tables and template information 
  - [ ] read function match `02_DataHarmonization/readAres2001.R` function inputs and output formats
  - [x] push annotations into an `old` folder as a place holder
  - [x] ask one of the undergrads to provide a secondary transcription for the TemplateCV and TemplateSubmit pdfs
  - [ ] push everything to the lead fork
+ Measure
  - [x] cross check with second transcriber for tables and methods, and resolve any differences
  - [ ] compile or render this ReadMe with all chunks up through the Level 0 data rescue section with `eval: true`
  - [ ] push to your local fork and create a pull request to the main branch of the primary repository
  - [ ] address any revisions in your pull request review
  - [ ] celebrate!

## Citation notes

  - `AK DSC Project SOC stock computation` had a bad url for the doi that linked to a spam site. Updating doi with the EDI pasta link.
  - `Jorgensen_ARCN` has this reference citation: Jorgenson, M.T., Swanson, D.K., Macander, M. 2001. Ecological subsections of Noatak National Preserve. 73. But when looked up, it does not match the destination file.
  - `Kane` dataset has this citation reference: Valentine, David. 2004. Soil profiles along productivity gradients in interior Alaska: Summer 2003, Bonanza Creek LTER - University of Alaska Fairbanks. BNZ:133, http://www.lter.uaf.edu/data_detail.cfm?datafile_pkey=133. Fixed this citation in rescued citations file with the correct reference citation.
  - `NASIS-2014` link doesn't work anymore.
  - `Worldwide soil carbon and nitrogen data` dataset has invalid doi.
--->
## Download ISCN3

The ISCN3 is downloaded from EDI (@ISCN3). 
Below we loop through the files that are expected and download them.

```{r}
#| label: dataDownloadEDI
#| code-summary: "Download and identify each file in the EDI package."

### Download the data ####
# link the table name to the download url and file path(s) for data package
pasta_package <- 'https://pasta.lternet.edu/package/data/eml/edi/1160/1/'

#transcribe from the EDI site, should be stable and not need future updates
download_table <- tibble::tribble(
  ~table, ~entity_id, ~file_name, 
  'layer', "4af719a84f8981fcc63f1f92760cb253", 'ISCN3_layer.csv',
  'profile', "40527580cc045d33d9a5aaf728bf204e", 'ISCN3_profile.csv',
  'citation', "320e31ca911f187550ca2143c31fd408", 'ISCN3_citation.csv',
  'dataset', "cdd0c7a4cac3f28d6d788c91f506775f", 'ISCN3_dataset.csv',
  'control_vocabulary', "a8eef6e94b669b365e443c15d9402a03", 'ISCNTranscribed_TemplateCVs.xlsx',
  'template', "cb5c1cd86e9ab17f699f7a05044325b7", 'ISCNtemplate.xlsx',
  'other_pdfs', "eb320ae7b57296765f543cbb370b0f24", 'TemplateCVs.pdf',
  'other_pdfs', "b81a7f4214d176280a5ff4e0f0d52d8b", 'TemplateSubmit.pdf')

#One by one, download the data in each row from the download_table
#... and store that in the appropriate file path location on your computer
for(row_index in 1:nrow(download_table)){
  #check to see if the file exists before trying to download it
  if(!file.exists(file.path(dataDownload.dir,
                            download_table$file_name[row_index]))){
    #tell the user you are downloading the file
    print(paste('Download file:', 
                download_table$file_name[row_index]))
    #use the utility download function to pull the file based on the 
    #... pasta url for ISCN3 and the entity ID, match that up with the file
    #... name on that row
    utils::download.file(
      paste0(pasta_package, download_table$entity_id[row_index]), 
      file.path(dataDownload.dir, download_table$file_name[row_index]), 
      extra  = "--max-time 300")
  } #end file exists check
} #end for loop through the download table

```

# Level 0 data

```{r}
#| label: readLevel0
#This chunk has two purposes. 
#...1) Check the formatting by reading in everything 
#...2) Create a list of everything to process later in level 1

data.lvl0.ls <- list(
  ####Read in a list of all the bib files####
  citation = list(
    #Citation for the article transcriptions are pulled from
    primary = read.bib(file = primaryCitation.file), 
    #Citations for all referenced articles
    contributed = read.bib(file = contributionsCitation.file)
  ),
  ####Read in the tables and their meta data###
  data = list(
    ####Citation table####
    TableCitation = list(
      #manual correction to out of date citations, originally
      #...this was the `temp/ISCN_citation.csv`
      primary = readr::read_delim(file = 'ISCN3_ISCN3_citation.csv', 
                                  skip = 1,
                                  delim = ';', 
                                  col_types = 
                                    readr::cols(.default = 
                                                  readr::col_character())),
      caption = readr::read_delim(file = 'ISCN3_ISCN3_citation.csv', 
                                  n_max = 1,
                                  delim = ';',
                                  col_names = FALSE,
                                  col_types = 
                                    readr::cols(.default = 
                                                  readr::col_character()))[[1]]
      #look in the metadata in the profile table for some variables
    ), #end citation list
    ####Dataset table####
    TableDataset = list(
      primary = readr::read_delim(file = 'temp/ISCN3_dataset.csv', 
                                  delim = ';', 
                                  col_types = 
                                    readr::cols(.default = 
                                                  readr::col_character()))
      #look in the metadata in the profile table for some variables
    ), #end dataset list
    ###Site table####
    TableSite = list(
      #no primary data here for sites, look in the profile and layer tables
      meta = readr::read_delim(file = 'ISCN3_TemplateSubmit_TableSite.csv',
                                delim = ",",
        col_types = readr::cols(.default = readr::col_character())
      )
    ), #end site
    TableFraction = list(
      #no primary data here
      meta = readr::read_delim(file = 'ISCN3_TemplateSubmit_TableFraction.csv',
                                delim = ",",
        col_types = readr::cols(.default = readr::col_character())
      )
    ), #end fraction
    TableDisturbance = list(
      #no primary data here
      meta = readr::read_delim(file = 'ISCN3_TemplateSubmit_TableDisturbance.csv',
                                delim = ",",
        col_types = readr::cols(.default = readr::col_character())
      )
    ), #end disturbance
    ###Profile table####
    TableProfile = list(
      primary = readr::read_delim(file = 'temp/ISCN3_profile.csv', 
                                  delim = ';', 
                                  col_types = 
                                    readr::cols(.default = 
                                                  readr::col_character())),
      meta = readr::read_delim( file = 'ISCN3_TemplateSubmit_TableProfile.csv',
                                delim = ",",
                                col_types = 
                                  readr::cols(.default =
                                                readr::col_character()))
    ), #end profile
    #### Layer table ####
    TableLayer = list(
      primary = readr::read_delim(file = 'temp/ISCN3_layer.csv', 
                                  delim = ';', 
                                  col_types = 
                                    readr::cols(.default = 
                                                  readr::col_character())),
      meta = readr::read_delim( file = 'ISCN3_TemplateSubmit_TableLayer.csv', 
                                delim = ",",
                                col_types = readr::cols(
                                  .default = readr::col_character()))
    ) #end layer table
    ), # end data list
  #### General control vocabulary that is not broke down by table ####
  control_vocabulary = list(
    readme = read_lines(file = 'README_TemplateCV.md'),
    caption = readr::read_delim( file = 'ISCN3_TemplateCVs.csv', 
                       n_max = 1,
                       col_names = FALSE,
                       delim = ",",
                       col_types = readr::cols(
                         .default = readr::col_character()))[[1]],
    primary = readr::read_delim( file = 'ISCN3_TemplateCVs.csv', 
                       skip = 1,
                       delim = ",",
                       col_types = readr::cols(
                         .default = readr::col_character())))
  )
  
```

<!---- consider making a diagram or nested list to identify the data structures here--->

## Check function match

Check for an exact match with the read function.

```{r}
#| label: checkReadFunction_lvl0

if(file.exists(readsource.file)){
  datafunction.lvl0.ls <- readISCN3(dataDir = '.',
                                         dataLevel = 'level0')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function matches code here.')
  }else{
    print('There is a mismatch in the data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}
```

# Level 1 data - ISCN 2025

This data is identified as a data source for the ISCN2025
Broad classes of information that we are looking for include:

For this analysis we are interested in:

  * **Location**: latitude/longitude, associated datum, geo-spatial method, precision
  * **Observation time**: observation year
  * **Soil organic carbon stocks**: organic carbon fraction, bulk density, coarse fraction, soil organic carbon by area, soil organic carbon by volume
  * **Layer depth**: Depth of surveyed soil layer

Note that level 1 data processing below removes the NRCS data contribution from ISCN3.
See the SoilDRaH entry for the NCSS data as a updated replacement for this or work directly with the level 0 data above.

This code takes a two step approach. First it isolates each of the variables of interest with the `is_type` observations as columns. Any known corrections or filters are applied in this first step.

```{r}
#| label: level1_constants

precisions.ls <- list(latLong_digits = 4, #assume lat/lon precision of 0.0001
                      #cluster resolution in lat/lon degrees
                      geolocationCluster_resolution = 0.001, 
                      #cluster resolution in years
                      observationYear_resolution = 2, 
                      #assume precision of 1cm
                      depth_digits = 0) 
```

## Space-time locations: resolution and clusters

ISCN3 provides latitude/longitude and observation date.
We have identified three main issues with the data collection:

  1) Locations appear overly precise with several extending to 8 or more reported degrees of precision that appear to be numerical errors.
    - *To resolve* round the lat/lon values to `{r} precisions.ls$latLong_digits` digits.
  2) Several locations appear to be repeated with both the lat/long reported extremely closely and shared values of bulk density or soil carbon elements reported.
    - *To resolve* this complete the degree distance between all lat/lon and applied a `{r} precisions.ls$geolocationCluster_resolution` cluster height to define geolocation groups.
    - *Note* that this ignores datum corrections, if exact location matters (<100m) you may need to reprocess the data from level 0.
  3) Observation dates appear to have standard export issues often seen in Access/Excel csv exports. Observation dates do not adhere to the reported YYYY-MM-DD format and are instead numerical values. Given the distribution of points its likely that some of these values were orginally entered as years and others reported as days-since-origin. Excel has historically exported their dates as days since 1900-01-01.
    - *To resolve* any values below 2015 were assumed to be CE-year and assigned the first of that year. Values above 2015 were assumed to be days since 1900-01-01.
  4) Observation dates also had minor disagreements between likely duplicate entries.
    - *To resolve*, similar to geolocations, a distance matrix was calculated and clusters within `{r} precisions.ls$observationYear_resolution` were assigned unique identifies.

```{r}
#| label: ISCN3_geo
#| code-summary: 'ISCN3 repeats geolocation, identify clusters by location and observation time. ISCN3 also has inconsistant formating in the observation date; if number below 2015 assume year, otherwise days since origin.'

#Format numerical values and impose precision
site_org.df <- data.lvl0.ls$data$TableProfile$primary |>
  select("dataset_name_sub", "dataset_name_soc", #ID
         "site_name", "profile_name", #ID
         "lat (dec. deg)", "long (dec. deg)", "datum (datum)", #geolocation
         "state (state_province)", "country (country)", #region
         "observation_date (YYYY-MM-DD)", #observation_date
         "profile_zero_ref (profile_zero_ref)")  |>
  #remove 75 651 profiles. Work with the current NRCS data download instead
  filter(!str_detect(dataset_name_sub, 'NRCS')) |>
  #correct the observation, some are entered in as years an others 
  #...as days since origin (excel convention)
  mutate(date_number = as.numeric(`observation_date (YYYY-MM-DD)`)) |>
  mutate(observation_date_formula = case_when(is.na(date_number) ~ 'missing',
                                              date_number > 2015 ~ 'days since origin',
                                              date_number <= 2015 ~ 'year CE',
                                              .default = 'default') |>
           as.factor(),
    observation_date = 
           case_when(is.na(date_number) ~ NA_Date_,
                     #assign origin based on Excel/Access historical convention
                     date_number > 2015 ~ as_date(date_number, origin = ymd('1900-01-01')),
                     #process things with just the year specified using the truncated flag
                     date_number <= 2015 ~ ymd(`observation_date (YYYY-MM-DD)`, 
                                               quiet = TRUE,
                                               truncated = 2),
         .default = NA_Date_)) |>
  #set the geolocations to numerics
  mutate(across(c(`lat (dec. deg)`, `long (dec. deg)`), as.numeric)) |>
  #round to 4 digits of precision or about 10 m https://wiki.openstreetmap.org/wiki/Precision_of_coordinates
  #... many numbers look overly precise and there was likely numerical error
  #....introduced during initial processing (excel/access exports?)
  mutate(across(c(`lat (dec. deg)`, `long (dec. deg)`), 
                .fns = round, digits = precisions.ls$latLong_digits))

#Define the location clusters if there are finite latitude/longitude reported using clusters on eucludian distances
site_geocluster <- site_org.df |>
  select("dataset_name_sub", #ID
         "site_name",  #ID
         "lat (dec. deg)", "long (dec. deg)") |>
  #there is one site in the 'World soil carbon and nitrogen data' that has two
  #... lat-long locations specified. Kick it out.
  unique() |> filter(n() == 1, .by = c(dataset_name_sub, site_name)) |>
  #remove any location that is missing lat/lon, don't assign a geocluster
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`)) |>
  #Create a distance matrix based on the latitude and longitude,
  #...cluster the sites, and cut out similar groups. 
  #...Looking at the tree heights and the spread of the lat/lon, 
  #...it looks like the height corresponds to the degree separation of the 
  #...points. Degrees 0.001 will correspond to 100m and is one order of 
  #...magnitude greater then the enforced precision.
  mutate(geolocation_id = sprintf('geocluster_%05d',
           dist(matrix(c(`lat (dec. deg)`, `long (dec. deg)`), ncol = 2)) |>
           hclust() |>
           cutree(h = precisions.ls$geolocationCluster_resolution))) #|> #Using euclidean distance
  #filter(n() > 1, .by = geolocation_id) #take a look at the clustered locations

#What happens if we change the tolerance
# temp <-  dist(matrix(c(site_geocluster$`lat (dec. deg)`,
#              site_geocluster$`long (dec. deg)`), ncol = 2)) |>
#           hclust()
#summary(temp$height > 0.0001) #5443 unique clusters
#summary(temp$height > 0.001) #5117 unique clusters
#summary(temp$height > 0.01) #4205 unique clusters

#Define time clusters similarlly to the spatial clusters
site_temporalCluster <- site_org.df |>
  select("dataset_name_sub", #ID
         "site_name",  #ID
         observation_date) |>
  #only define a cluster if we have an observation date
  filter(is.finite(observation_date)) |> unique() |>
  #create the temporal cluster
  mutate(date_year = year(observation_date),
         temporal_id = sprintf(
           'timecluster_%02d',
           dist(matrix(decimal_date(observation_date), ncol = 1)) |> 
             hclust() |>
             cutree(h = precisions.ls$observationYear_resolution))) #max height here corresponds to year span, consider two years to be effectively the same observation time

#temp <-  dist(matrix(decimal_date(site_temporalCluster$observation_date), ncol = 1)) |>
#           hclust()
#range(temp$height) # 0.00000 94.60383
#summary(temp$height > 1) #75 clusters
#summary(temp$height > 2) #43 clusters
#summary(temp$height > 5) #21 clusters
#summary(temp$height > 10) #11 clusters

#reconstruct the site information with the clusters developed above
#...note that we keep the exact lat/long and observate date provided 
#...and they are now associated with neighborhood clusters
site.df <- site_geocluster |>
  full_join(site_temporalCluster,
            by = join_by(dataset_name_sub, site_name)) |>
  full_join(site_org.df,
            by = join_by(dataset_name_sub, site_name,
                         `lat (dec. deg)`, `long (dec. deg)`,
                         observation_date))

uniqueSpaceTime <- site.df |>
  filter(!is.na(temporal_id) & ! is.na(geolocation_id)) |>
  select(temporal_id, geolocation_id) |> unique() |> nrow()
```

There are `{r} uniqueSpaceTime` unique geolocation-time pairs reported.

## Layer processing: constructing identifiers

Processing the layer table 

  1) Layer observations appear to be duplicated for different SOC calculations and entered in under multiple datasets.
    - *To resolve* add in the clusters developed in the site processing above so that we can better identify duplicate entries of variables when processing each variable type later in the code.
  2) Identifiers provided under the `.*_name` pattern are not unique leading to issues pivoting the data during processing.
    - *To resolve* construct a `layer_id` from available zero reference and the upper/lower bounds. Ensure that the upper/lower bounds have a realistic resolution of 1 cm. Also construct a `sample_id` from the `dataset_name`, `site_name`, `profile_name`, and `layer_name`. The sample id is slightly redundent but allows us to consolodate the identifiers.

Things to also note here:
  - Removing the NRCS layers reduced row counts from 445 829 to 72 503.
  - Location information is given in both the site and layer tables. We assume the site is correct and ignore the layer information.

```{r}
#| label: ISCN3preprocess_layer

layer.df <- data.lvl0.ls$data$TableLayer$primary |>
  select(#"ISCN 1-1 (2015-12-10)", 
         "dataset_name_sub", "dataset_name_soc", 
         "site_name", "profile_name", 
         #"lat (dec. deg)", "long (dec. deg)", "state (state_province)", "country (country)", "observation_date (YYYY-MM-DD)", #use location information from the site not layer
         "layer_name", "layer_top (cm)", "layer_bot (cm)", "layer_note", 
         #"hzn_desgn_other", "hzn", "hzn_desgn", "color",
         #"vegclass_local", 
         #"soil_taxon", "soil_series", 
         "bd_method", "bd_samp (g cm-3)", "bd_tot (g cm-3)", "bd_whole (g cm-3)", "bd_other (g cm-3)", "bdNRCS_prep_code", 
         "cNRCS_prep_code", "c_method", "c_tot (percent)", "oc (percent)", "loi (percent)", 
         #"n_tot (percent)", "c_to_n (mass ratio)", 
         "soc (g cm-2)", "soc_carbon_flag", "soc_method", 
         #"ph_method", "ph_cacl", "ph_h2o", "ph_other", 
         "caco3 (percent)", 
         #"sand_tot_psa (percent)", "silt_tot_psa (percent)", "clay_tot_psa (percent)", 
         "wpg2_method", "wpg2 (percent)", 
         #"cat_exch (cmol H+ kg-1)", 
         #"al_dith (specified by al_fe_units)", "al_ox (specified by al_fe_units)", "al_other (specified by al_fe_units)", "fe_dith (specified by al_fe_units)", "fe_ox (specified by al_fe_units)", "fe_other (specified by al_fe_units)", "mn_dith (specified by al_fe_units)", "mn_ox (specified by al_fe_units)", "mn_other (specified by al_fe_units)", "al_fe_units (extract_units)", "al_fe_method",
         #"ca_al (specified by bc_units)", "ca_ext (specified by bc_units)", "k_ext (specified by bc_units)", "mg_ext (specified by bc_units)", "na_ext (specified by bc_units)", "bc_units (extract_units)", "bc_method", "base_sum (specified by cec_h_units)", "cec_sum (specified by cec_h_units)", "ecec (specified by cec_h_units)", "cec_h_units (extract_units)", 
         #"bs (percent)", "bs_sum (percent)", 
         #"h_ext (specified by metal_ext_units)", "zn_ext (specified by metal_ext_units)", "metal_ext_units (extract_units)", "metal_ext_method", 
         #"p_bray (specified by p_units)", "p_ox (specified by p_units)", "p_meh (specified by p_units)", "p_other (specified by p_units)", "p_units (extract_units)", "p_method", 
         #"root_quant_size", "root_weight (g)",
         #"15n (‰)", "13c (‰)", "14c (‰)", "14c_sigma (‰)", "14c_age (BP)", "14c_age_sigma (BP)", "fraction_modern", "fraction_modern_sigma", 
         #"textureClass", "locator_parent_alias"
         ) |>
  #remove 373 326 layers. Work with the current NRCS data download instead
  filter(!str_detect(dataset_name_sub, 'NRCS')) |> #72503 obs of 25 variables
  #remove columns that are empty because some of these are NRCS only observations
  filter_out(if_all(everything(), is.na)) |> #does not remove columns right now but might if the variables were expanded
  #add in the site information so that we can get the cluster ids
  left_join(site.df,
            by = join_by(dataset_name_sub, dataset_name_soc, site_name, profile_name)) |> 
  mutate(across(c(`layer_top (cm)`, `layer_bot (cm)`), as.numeric)) |>
  #Construct the layer ID from the rounded depths and 
  #...add in the zero reference if it exists
  mutate(layer_id = case_when( 
    !is.na(`profile_zero_ref (profile_zero_ref)`) &
      !is.na(`layer_top (cm)` + `layer_bot (cm)`) ~
      sprintf('layer (zero-ref:%s) [%d,%d]',
              `profile_zero_ref (profile_zero_ref)`,
              round(`layer_top (cm)`, digits = precisions.ls$depth_digits),
              round(`layer_bot (cm)`, digits = precisions.ls$depth_digits)),
    is.na(`profile_zero_ref (profile_zero_ref)`) &
      !is.na(`layer_top (cm)` + `layer_bot (cm)`) ~ 
      sprintf('layer [%d,%d]',
              round(`layer_top (cm)`, digits = precisions.ls$depth_digits),
              round(`layer_bot (cm)`, digits = precisions.ls$depth_digits)),
    !is.na(`profile_zero_ref (profile_zero_ref)`) &
      is.na(`layer_top (cm)` + `layer_bot (cm)`) ~
      sprintf('layer (zero-ref:%s)',
              `profile_zero_ref (profile_zero_ref)`),
    is.na(`profile_zero_ref (profile_zero_ref)`) &
      is.na(`layer_top (cm)` + `layer_bot (cm)`) ~ NA_character_,
    .default = 'missing'
    )) |>
  arrange(geolocation_id, temporal_id, layer_id) |> #full address
  #create a sample_id for simplification later from key:value pairs of the names
  mutate(sample_id = sprintf('dataset:%s,site:%s,profile:%s,layer:%s',
                               dataset_name_sub, site_name, profile_name, layer_name))
```

```{r}
#| label: ISCN3_bulkDensity
bulkdensity.df <- layer.df  |>
  select(c(ends_with('_id'), starts_with('bd')), dataset_name = dataset_name_sub) |>
  mutate(across(.cols = ends_with('(g cm-3)'), .fns = as.numeric)) |>
  mutate(across(.cols = ends_with('(g cm-3)'), .fns = round, digits = 2)) |> #72 503
  pivot_longer(cols = where(is.numeric), names_to = 'column_name_bd',
               values_to = 'value_bd', values_drop_na = TRUE) |>
  filter(value_bd >= 0) |> #ensure there are positive values
  mutate(method_bd = case_when(
    !is.na(bd_method) & !is.na(bdNRCS_prep_code) ~ 
      sprintf('bd_method:%s;bdNRCS_prep_code:%s', bd_method, bdNRCS_prep_code),
    is.na(bd_method) & !is.na(bdNRCS_prep_code) ~ 
      sprintf('bdNRCS_prep_code:%s', bdNRCS_prep_code),
    !is.na(bd_method) & is.na(bdNRCS_prep_code) ~ 
      sprintf('bd_method:%s', bd_method),
    .default = NA_character_)) |>
  select(geolocation_id, temporal_id, layer_id, sample_id,
         column_name_bd, value_bd, method_bd, dataset_name) |> #59 459
  #Deal with duplicate values
  mutate(sample_id = ifelse(n() == 1 | length(unique(sample_id)) == 1,
                            unique(sample_id), 
                            paste0('[MultiSourced]', paste(unique(sample_id), collapse = ';'))),
         column_name_bd = ifelse(n() == 1 & length(unique(column_name_bd)) > 1,
                            column_name_bd, 
                            paste(sort(unique(column_name_bd)), collapse = ';')),
         method_bd = paste(unique(method_bd[!is.na(method_bd)]), collapse = ';'),
         .by = c(geolocation_id, temporal_id, layer_id, value_bd)) |> 
  unique() |> #49 976, roughly 15% removed as duplicate values
  #Assign the variable names
  mutate(method_bd = paste(case_when(
    #Bulk density types taken from metadata states that whole soil bulk density with coarse fraction removed
    #Nave et al. New Forests. 2019 implies that this is a fine earth bulk density in it's methods description of how ISCN3 was used.
    column_name_bd %in% c( "bd_samp (g cm-3)",
                           'bd_other (g cm-3);bd_samp (g cm-3)') ~ 'bulk_density:fine_earth',
    #yes this bd_whole and bd_tot are correct according to the meta data
    column_name_bd %in% c("bd_whole (g cm-3)") ~ 'bulk_density:fine_earth:estimated',
    column_name_bd %in% c( "bd_tot (g cm-3)") ~ 'bulk_density:whole_soil',
    #contradictory values match, label it as unknown type
    column_name_bd %in% c("bd_samp (g cm-3);bd_tot (g cm-3)", 
                        'bd_other (g cm-3)') ~ 'bulk_density:unspecified',
    column_name_bd == c('bd_other (g cm-3);bd_tot (g cm-3)') ~ 'bulk_density:whole_soil',
    column_name_bd == c('bd_samp (g cm-3);bd_tot (g cm-3)') ~ 'bulk_density:fine_earth',
    .default = 'bad code'), method_bd, sep = ';'))|>
  arrange(geolocation_id, temporal_id, layer_id) |>
  unique() |> #nrow #49 976
  # filter(n() == 1,
  #        .by = c(geolocation_id, temporal_id, layer_id)) #nrow 16 148
  # remove observations which are poorly specified, roughly 22%
  filter(n() == 1,
         #length(unique(value_bd)) > 1,
         .by = c(geolocation_id, temporal_id, layer_id, method_bd)) |> #nrow 39 679
  mutate(of_variable = 'bulk_density',
         unit = 'g cm-3') |>
  select(ends_with('_id'), variable = of_variable, value = value_bd, 
         unit = unit, method = method_bd, dataset_name)
```

```{r}
#| label: ISCN3_soc
soc.df <- layer.df  |>
  select(c(ends_with('_id'), starts_with('soc'), dataset_name_soc))  |>
  filter(!is.na(`soc (g cm-2)`)) |> #46 343
  #pick digits based on hist(str_length(soc.df$`soc (g cm-2)`))
  mutate(value = round(as.numeric(`soc (g cm-2)`), digits = 5),
         method_soc = paste0('dataset_name_soc:', dataset_name_soc, ';', case_when(
           !is.na(soc_method) & !is.na(soc_carbon_flag) ~ 
             sprintf('soc_method:%s;soc_carbon_flag:%s',
                     soc_method, soc_carbon_flag),
           !is.na(soc_method) & is.na(soc_carbon_flag) ~ 
             sprintf('soc_method:%s',
                     soc_method),
           is.na(soc_method) & !is.na(soc_carbon_flag) ~ 
             sprintf('soc_carbon_flag:%s',
                     soc_carbon_flag),
           .default = 'unspecified'))) |>
  filter(value >= 0) |> #46 337
  #Deal with duplicate values
  mutate(sample_id = ifelse(n() == 1 | length(unique(sample_id)) == 1,
                            unique(sample_id), 
                            paste0('[MultiSourced]', paste(unique(sample_id), collapse = ';'))),
         method = paste(unique(method_soc[!is.na(method_soc)]), collapse = ';'),
         dataset_name_soc = paste(unique(dataset_name_soc[!is.na(dataset_name_soc)]), collapse = ';'), #keep seperate for citations later
         .by = c(geolocation_id, temporal_id, layer_id, value)) |> 
  unique() |> #41 750
  select(geolocation_id, temporal_id, layer_id, sample_id, value, method, 
         dataset_name = dataset_name_soc) |>
  mutate(variable = 'organic_carbon_stock',
         unit = 'g cm-2')
```

```{r}
#|label: ISCN3_carbonfraction
carbon_fraction.df <- layer.df  |>
  select(c(ends_with('_id'), dataset_name = dataset_name_sub,
           `oc (percent)`, `c_tot (percent)`, `caco3 (percent)`,
           cNRCS_prep_code, c_method))  |>
  pivot_longer(cols = contains('percent'),
               names_to = 'column_name', values_to = 'value', 
               values_drop_na = TRUE) |>
  mutate(value = as.numeric(value), 
         method = case_when(
           !is.na(c_method) & !is.na(cNRCS_prep_code) ~ 
             sprintf('c_method:%s;cNRCS_prep_code:%s',
                     c_method, cNRCS_prep_code),
           !is.na(c_method) & is.na(cNRCS_prep_code) ~ 
             sprintf('c_method:%s',
                     c_method),
           is.na(c_method) & !is.na(cNRCS_prep_code) ~ 
             sprintf('cNRCS_prep_code:%s',
                     cNRCS_prep_code),
           .default = 'unspecified')) |> #51 977 between 0 and 87
  mutate(method = paste(case_when(column_name == 'oc (percent)' ~ 'carbon_organic',
                                  column_name == 'c_tot (percent)' ~ 'carbon_total',
                                  column_name == 'caco3 (percent)' ~ 'carbon_inorganic_CaCO3'),
                        method, sep = ';')) |>
  #Deal with duplicate values
  mutate(sample_id = ifelse(n() == 1 | length(unique(sample_id)) == 1,
                            unique(sample_id), 
                            paste0('[MultiSourced]', paste(unique(sample_id), collapse = ';'))),
         method = paste(unique(method[!is.na(method)]), collapse = ';'),
         #there are a small ammount of layers that share OC, C and/or CaCO3 values
         #... which is possible, so group on this as well.
         .by = c(geolocation_id, temporal_id, layer_id, column_name, value)) |> 
  unique() |> #45 278
  select(geolocation_id, temporal_id, layer_id, sample_id, value, method, dataset_name) |>
  mutate(variable = 'carbon_fraction',
         unit = 'mass-percent (carbon to fine earth)')
```

```{r}
#| label: ISCN3_organicFraction
#| 
organic_fraction.df <- layer.df  |>
  select(ends_with('_id'), dataset_name = dataset_name_sub,
         `loi (percent)`)  |>
  filter(!is.na(`loi (percent)`)) |> #938
  #pick digits based on hist(str_length(soc.df$`loi (percent)`))
  mutate(value = round(as.numeric(`loi (percent)`), digits = 5)) |>
  filter(value >= 0) |>
  #Deal with duplicate values
  mutate(sample_id = ifelse(n() == 1 | length(unique(sample_id)) == 1,
                            unique(sample_id), 
                            paste0('[MultiSourced]', paste(unique(sample_id), collapse = ';'))),
         .by = c(geolocation_id, temporal_id, layer_id, value)) |> 
  unique() |> #936
  select(geolocation_id, temporal_id, layer_id, sample_id, value, dataset_name) |>
  mutate(variable = 'organic_matter_fraction',
         unit = 'mass-percent (organic to fine earth)',
         method = 'loss on ignition')
```

```{r}
#| label: ISCN3_depth
depth.df <- layer.df  |>
  select(site_name, #cross with the site dataframe
         ends_with('_id'), starts_with('layer'), 
         `profile_zero_ref (profile_zero_ref)`) |> #72 503
  filter(!is.na(`layer_top (cm)` + `layer_bot (cm)`)) |> #72 503
  #pick digits based on hist(str_length(soc.df$`loi (percent)`))
  mutate(upper = round(as.numeric(`layer_top (cm)`), digits = 0),
         lower = round(as.numeric(`layer_bot (cm)`), digits = 0),
         zero_ref = `profile_zero_ref (profile_zero_ref)`) |>
  #Deal with duplicate values
  mutate(sample_id = ifelse(n() == 1 | length(unique(sample_id)) == 1,
                            unique(sample_id), 
                            paste0('[MultiSourced]', paste(unique(sample_id), collapse = ';'))),
         .by = c(geolocation_id, temporal_id, layer_id, upper, lower, zero_ref)) |> 
  unique() |> #54 770
  select(geolocation_id, temporal_id, layer_id, sample_id, upper, lower, zero_ref) |>
  mutate(variable = 'depth_interval',
         unit = 'cm')
```

```{r}
#| label: ISCN3_spaceTime
geolocation.df <- layer.df |>
  select(geolocation_id, `lat (dec. deg)`, `long (dec. deg)`, datum = `datum (datum)`) |>
  unique() |> #5408
  #Deal with duplicate values
  reframe(latitude = mean(`lat (dec. deg)`),
          longitude = mean(`long (dec. deg)`),
          resolution = mean(c(diff(range(`lat (dec. deg)`)),
                              diff(range(`long (dec. deg)`)))),
          count = n(),
          .by = c(geolocation_id, datum)) |> #5133
  mutate(unit = 'decimal degrees',
         variable = 'geolocation')

region.df <- layer.df |>
  select(geolocation_id, country = `country (country)`, state = `state (state_province)`) |>
  pivot_longer(c(country, state),
               names_to = 'region', values_to = 'value',
                values_drop_na = TRUE) |>
  filter(value != 'Unknown') |>
  unique() |> #5383
  #Deal with duplicate values
  reframe(value = paste(sort(unique(value)), collapse = ';'),
          .by = c(geolocation_id, region)) |>
  pivot_wider(names_from = region, values_from = value) |>
  mutate(variable = 'region')

temporal.df <- layer.df |> 
  select(temporal_id, observation_date, `observation_date (YYYY-MM-DD)`, observation_date_formula) |>
  unique() |> #667
  #Deal with duplicate values
  reframe(date = mean(observation_date),
          resolution = as.numeric(difftime(max(observation_date), min(observation_date), 
                                           units = 'days'))/365,
          count = n(),
         .by = c(temporal_id)) |>
  mutate(unit = 'decimal year (CE)',
         variable = 'observation_date')
  
```

```{r}
#| label: ISCN3_corrections_citations
#| code-summary: 'ISCN3 needs expanded citations for the permafrost RCN.'


#collection table
collection.df <- 
  #### Create cross key for the citation dataset naming to include the double
  #### citations needed for the Permafrost RCN data sets
  data.lvl0.ls$data$TableDataset$primary |>
  select("dataset_name", #matches dataset_name_sub or dataset_name_soc
         #"dataset_type (dataset_type)",
         "curator_name", "curator_organization", "curator_email", 
         "contact_name", "contact_email", 
         "dataset_description"
  ) |>
  #expand the Permafrost RCN datasets so we can pull in that citation
  reframe(dataset_name_citation = c('Permafrost_RCN', dataset_name)[c(str_detect(dataset_name, '^Permafrost RCN_'), TRUE)],
          .by = everything()) |>
  full_join(
    data.lvl0.ls$data$TableCitation$primary |>
      select(dataset_name_citation = "dataset_name", #matches dataset_name_sub or dataset_name_soc
             "reference", 
             "citation", "citation_usage", 
             "acknowledgement", "acknowledgement_usage") |>
      mutate(source_from = 'Table Citation'),
    relationship = "many-to-many",
    by = join_by(dataset_name_citation)) |>
  #### Trim NRCS, remove 2 collection. Work with the current NRCS data download instead
  filter(!str_detect(dataset_name, 'NRCS')) 
  
```


```{r}
#| label: build_lvl1
data.lvl1.ls <- list(
  sample = soc.df |>
    bind_rows(bulkdensity.df) |>
    bind_rows(organic_fraction.df) |>
    bind_rows(carbon_fraction.df) |>
    rename(citation_id = dataset_name),
  depth = depth.df,
  geolocation = geolocation.df |>
    bind_rows(region.df),
  temporal = temporal.df,
  study_bibKey = 'ISCN3'
)

data.lvl1.ls$citation_key <-   data.lvl1.ls$sample |>
    select(citation_id) |>
    unique() |>
    reframe(dataset_name = str_split(citation_id, patter = ';') |>
              unlist(),
            .by = citation_id) |>
    left_join(
      collection.df |>
        pivot_longer(cols = c(citation, reference),
                     names_to = 'variable',
                     values_to = 'text', 
                     values_drop_na = TRUE) |>
        reframe(bib_key = 
                  str_extract_all(text,'(?<=@)\\w+') |> unlist(),
                .by = c(dataset_name, variable)),
      by = join_by(dataset_name),
      relationship = "many-to-many")

allKeys <- data.lvl1.ls$citation_key |>
  filter(is.finite(bib_key))
data.lvl1.ls$citation <- c(data.lvl0.ls$citation$primary,
                           data.lvl0.ls$citation$contributed[unique(allKeys$bib_key)])
```

##DONT RUN BELOW

```{r}
soc.df |>
  select(geolocation_id, temporal_id, layer_id) |>
  unique() |> #nrow() #46 354 layers total
  mutate(has_location = !is.na(geolocation_id),
         has_time = !is.na(temporal_id),
         has_depth = str_detect(layer_id, '\\[\\d+,\\d+\\]')) |>
  reframe(count = n(),
          .by = starts_with('has_')) |>
  mutate(fraction = round(count / sum(count), digits = 2)) |>
  arrange(desc(count))
```

```{r}

  
layer.df <- data.lvl0.ls$data$TableLayer$primary |>
  select(#"ISCN 1-1 (2015-12-10)", 
         "dataset_name_sub", "dataset_name_soc", 
         "site_name", "profile_name", 
         #"lat (dec. deg)", "long (dec. deg)", "datum (datum)",
         #"state (state_province)", "country (country)", 
         #"observation_date (YYYY-MM-DD)",
         "layer_name", "layer_top (cm)", "layer_bot (cm)", "layer_note", 
         #"hzn_desgn_other", "hzn", "hzn_desgn", "color",
         #"vegclass_local", 
         #"soil_taxon", "soil_series", 
         "bd_method", "bd_samp (g cm-3)", "bd_tot (g cm-3)", "bd_whole (g cm-3)", "bd_other (g cm-3)", "bdNRCS_prep_code", 
         "cNRCS_prep_code", "c_method", "c_tot (percent)", "oc (percent)", "loi (percent)", 
         #"n_tot (percent)", 
         #"c_to_n (mass ratio)", 
         "soc (g cm-2)", "soc_carbon_flag", "soc_method", 
         #"ph_method", "ph_cacl", "ph_h2o", "ph_other", 
         "caco3 (percent)", 
         #"sand_tot_psa (percent)", "silt_tot_psa (percent)", "clay_tot_psa (percent)", 
         "wpg2_method", "wpg2 (percent)", 
         #"cat_exch (cmol H+ kg-1)", 
         #"al_dith (specified by al_fe_units)", "al_ox (specified by al_fe_units)", "al_other (specified by al_fe_units)", "fe_dith (specified by al_fe_units)", "fe_ox (specified by al_fe_units)", "fe_other (specified by al_fe_units)", "mn_dith (specified by al_fe_units)", "mn_ox (specified by al_fe_units)", "mn_other (specified by al_fe_units)", "al_fe_units (extract_units)", "al_fe_method",
         #"ca_al (specified by bc_units)", "ca_ext (specified by bc_units)", "k_ext (specified by bc_units)", "mg_ext (specified by bc_units)", "na_ext (specified by bc_units)", "bc_units (extract_units)", "bc_method", "base_sum (specified by cec_h_units)", "cec_sum (specified by cec_h_units)", "ecec (specified by cec_h_units)", "cec_h_units (extract_units)", 
         #"bs (percent)", "bs_sum (percent)", 
         #"h_ext (specified by metal_ext_units)", "zn_ext (specified by metal_ext_units)", "metal_ext_units (extract_units)", "metal_ext_method", 
         #"p_bray (specified by p_units)", "p_ox (specified by p_units)", "p_meh (specified by p_units)", "p_other (specified by p_units)", "p_units (extract_units)", "p_method", 
         #"root_quant_size", "root_weight (g)",
         #"15n (‰)", "13c (‰)", "14c (‰)", "14c_sigma (‰)", "14c_age (BP)", "14c_age_sigma (BP)", "fraction_modern", "fraction_modern_sigma", 
         #"textureClass", "locator_parent_alias"
         ) |>
  #remove 373 326. Work with the current NRCS data download instead
  filter(str_detect(dataset_name_sub, 'NRCS')) |>
  #There are 2434 layers that are not uniquely defined
  mutate(row_name = paste0('R', 1:n()))
  
  #make layer names unique, some were numeric based and got truncated 
  mutate(layer_name = sprintf('%s (%s to %s)', layer_name, `layer_top (cm)`, `layer_bot (cm)`))
```

```{r}
#| label: createLvl1
#| code-summary: 'Create a level 1 data set with SoilDRaH standard vocabulary and table structures.'

#Site table
site.df <- data.lvl0.ls$data$TableProfile$primary |>
  select("dataset_name_sub", "dataset_name_soc", #ID
         "site_name", "profile_name", #ID
         "lat (dec. deg)", "long (dec. deg)", "datum (datum)", #geolocation
         "state (state_province)", "country (country)", #region
         "observation_date (YYYY-MM-DD)", #observation_date
         "profile_zero_ref (profile_zero_ref)") |>  #depth
  pivot_longer(cols = all_of(c(
    "lat (dec. deg)", "long (dec. deg)", "datum (datum)", #geolocation
    "state (state_province)", "country (country)", #region
    "observation_date (YYYY-MM-DD)", #observation_date
    "profile_zero_ref (profile_zero_ref)"
    )),
    names_to = 'column_header', values_to = 'with_entry',
    values_drop_na = TRUE) |>
  mutate(of_variable = case_when(
    column_header %in% c("lat (dec. deg)", "long (dec. deg)", "datum (datum)") ~ 'geolocation',
    column_header %in% c("state (state_province)", "country (country)") ~ 'region', #region
    column_header %in% c("observation_date (YYYY-MM-DD)") ~ 'observation_date', #observation_date
    column_header %in% c("profile_zero_ref (profile_zero_ref)") ~ 'depth',
      .default = NA_character_),
  is_type = case_when(
    column_header %in% c("lat (dec. deg)") ~ 'latitude',
    column_header %in% c("long (dec. deg)") ~ 'longitude',
    column_header %in% c("datum (datum)") ~ 'datum',
    column_header %in% c("state (state_province)") ~ 'state',
    column_header %in% c("country (country)") ~ 'country',
    column_header %in% c("observation_date (YYYY-MM-DD)") ~ 'date', #observation_date
    column_header %in% c("profile_zero_ref (profile_zero_ref)") ~ 'zero_reference',
      .default = NA_character_)
    ) |>
  mutate(from_source = 'Table Profile')

#collection table
collection.df <- data.lvl0.ls$data$TableDataset$primary |>
  select("dataset_name", #matches dataset_name_sub or dataset_name_soc
         #"dataset_type (dataset_type)",
         "curator_name", "curator_organization", "curator_email", 
         "contact_name", "contact_email", 
         "dataset_description") |>
  mutate(source_from = 'Table Dataset') |>
  full_join(
    data.lvl0.ls$data$TableCitation$primary |>
      select("dataset_name", #matches dataset_name_sub or dataset_name_soc
             "reference", 
             "citation", "citation_usage", 
             "acknowledgement", "acknowledgement_usage") |>
      mutate(source_from = 'Table Citation'),
    by = join_by(dataset_name, source_from)) |>
  pivot_longer(cols = -c(dataset_name, source_from),
               names_to = 'column_header', values_to = 'with_entry',
               values_drop_na = TRUE) |>
  mutate(of_variable = str_extract(column_header, '^[^_]+'),
         is_type = str_extract(column_header, '[^_]+$')) |>
  mutate(is_type = if_else(of_variable == is_type, 'value', is_type))

layer.df <- data.lvl0.ls$data$TableLayer$primary |>
  #make layer names unique, some were numeric based and got truncated 
  mutate(layer_name = sprintf('%s (%s to %s)', layer_name, `layer_top (cm)`, `layer_bot (cm)`)) |>
  #There are 2136 layers that are not uniquely defined, looks like rows doubled form NRCS prep codes
  mutate(row_name = paste0('R', 1:n())) |>
  select('row_name', 
         #"ISCN 1-1 (2015-12-10)", 
         "dataset_name_sub", "dataset_name_soc", 
         "site_name", "profile_name", 
         #"lat (dec. deg)", "long (dec. deg)", "datum (datum)",
         #"state (state_province)", "country (country)", 
         #"observation_date (YYYY-MM-DD)",
         "layer_name", "layer_top (cm)", "layer_bot (cm)", "layer_note", 
         #"hzn_desgn_other", "hzn", "hzn_desgn", "color",
         #"vegclass_local", 
         #"soil_taxon", "soil_series", 
         "bd_method", "bd_samp (g cm-3)", "bd_tot (g cm-3)", "bd_whole (g cm-3)", "bd_other (g cm-3)", "bdNRCS_prep_code", 
         "cNRCS_prep_code", "c_method", "c_tot (percent)", "oc (percent)", "loi (percent)", 
         #"n_tot (percent)", 
         #"c_to_n (mass ratio)", 
         "soc (g cm-2)", "soc_carbon_flag", "soc_method", 
         #"ph_method", "ph_cacl", "ph_h2o", "ph_other", 
         "caco3 (percent)", 
         #"sand_tot_psa (percent)", "silt_tot_psa (percent)", "clay_tot_psa (percent)", 
         "wpg2_method", "wpg2 (percent)", 
         #"cat_exch (cmol H+ kg-1)", 
         #"al_dith (specified by al_fe_units)", "al_ox (specified by al_fe_units)", "al_other (specified by al_fe_units)", "fe_dith (specified by al_fe_units)", "fe_ox (specified by al_fe_units)", "fe_other (specified by al_fe_units)", "mn_dith (specified by al_fe_units)", "mn_ox (specified by al_fe_units)", "mn_other (specified by al_fe_units)", "al_fe_units (extract_units)", "al_fe_method",
         #"ca_al (specified by bc_units)", "ca_ext (specified by bc_units)", "k_ext (specified by bc_units)", "mg_ext (specified by bc_units)", "na_ext (specified by bc_units)", "bc_units (extract_units)", "bc_method", "base_sum (specified by cec_h_units)", "cec_sum (specified by cec_h_units)", "ecec (specified by cec_h_units)", "cec_h_units (extract_units)", 
         #"bs (percent)", "bs_sum (percent)", 
         #"h_ext (specified by metal_ext_units)", "zn_ext (specified by metal_ext_units)", "metal_ext_units (extract_units)", "metal_ext_method", 
         #"p_bray (specified by p_units)", "p_ox (specified by p_units)", "p_meh (specified by p_units)", "p_other (specified by p_units)", "p_units (extract_units)", "p_method", 
         #"root_quant_size", "root_weight (g)",
         #"15n (‰)", "13c (‰)", "14c (‰)", "14c_sigma (‰)", "14c_age (BP)", "14c_age_sigma (BP)", "fraction_modern", "fraction_modern_sigma", 
         #"textureClass", "locator_parent_alias"
         ) |>
  pivot_longer(cols = -contains('_name'),
               names_to = 'column_header', values_to = 'with_entry',
               values_drop_na = TRUE) |>
  mutate(of_variable = 
           case_when(
             column_header %in% c("layer_top (cm)", "layer_bot (cm)", "layer_note") ~ 'layer',
             #column_header %in% c("color") ~ 'soil_color',
             #column_header %in% c("hzn_desgn_other", "hzn", "hzn_desgn") ~ 'soil_horizon',
             #column_header %in% c("soil_taxon", "soil_series") ~ 'soil_classificaction',
             #Bulk density types taken from metadata states that whole soil bulk density with coarse fraction removed
             #Nave et al. New Forests. 2019 implies that this is a fine earth bulk density in it's methods description of how ISCN3 was used.
             column_header %in% c( "bd_samp (g cm-3)") ~ 'bulk_density_fine_earth',
             #yes this bd_whole and bd_tot are correct according to the meta data
             column_header %in% c("bd_whole (g cm-3)") ~ 'bulk_density_estimated_fine_earth',
             column_header %in% c( "bd_tot (g cm-3)") ~ 'bulk_density_whole_soil',
             column_header %in% c("bd_method", "bd_other (g cm-3)", "bdNRCS_prep_code") ~ 'bulk_density',
             #And the splitting strategy with soil carbon
             column_header %in% c("oc (percent)") ~ 'carbon_organic_fraction',
             column_header %in% c("soc (g cm-2)", "soc_carbon_flag", "soc_method") ~ 'carbon_organic_area_density',
             column_header %in% c("loi (percent)") ~'loss_on_ignition',
             #need to associated with other carbon variables later
             column_header %in% c("cNRCS_prep_code", "c_method") ~'carbon',
             column_header %in% c("c_tot (percent)") ~ 'carbon_total',
             column_header %in% c("caco3 (percent)") ~ 'carbon_inorganic',
             column_header %in% c("wpg2_method", "wpg2 (percent)") ~ 'coarse_fraction',
             .default = NA_character_
           ),
         is_type = 
           case_when(
             str_detect(column_header, 'note') ~ 'note',
             str_detect(column_header, 'method') ~ 'method',
             str_detect(column_header, 'NRCS_prep_code') ~ 'NRCS_prep_code',
             str_detect(column_header, 'flag') ~ 'flag',
             column_header == "layer_top (cm)" ~ 'upper',
             column_header == "layer_bot (cm)" ~ 'lower',
             .default = 'value')) |>
  #Check the classifications of the columns
#   select(column_header, of_variable, is_type) |> unique() |> view()
#   Check for uniquenesss
#  filter(n() > 1, .by =c(contains('_name'), of_variable, is_type))
  mutate(source_from = 'Table Layer' )

#this is nto exactly clean but I think it's workable
  layer.meta <- layer.df |>
    select(column_header, of_variable) |>
    unique() |>
    mutate(ISCN_variable = str_extract(column_header, '^\\S+')) |>
        mutate(across(everything(), str_trim)) |>
    left_join(data.lvl0.ls$data$TableLayer$meta |>
        mutate(across(everything(), str_trim)),
              by = c('ISCN_variable' = 'Variable')) |>
    select(column_header, of_variable, 
           ISCN3_vocabulary = ISCN_variable, name = Name, unit = Units, description = Description) |>
    pivot_longer(cols = c(ISCN3_vocabulary, name, unit, description),
                 names_to = 'is_type', values_to = 'with_entry',
                 values_drop_na = TRUE) |>
    mutate(from_source = 'Meta for Layer (ISCN3_TemplateSubmit_TableLayer.csv)')

  

# What is true across all of ISCN3
study.df <- tribble(~of_variable, ~is_type, ~with_entry, ~from_source,
                   'citation', 'value', '@ISCN3', 'EDI download page',
                   'version', 'value', "ISCN 1-1 (2015-12-10)", 'orginal table header number 1',
                   'geolocation', 'unit', 'decimal degrees', 'profile and layer table column names',
                   'observation_date', 'format', 'days since 1900-01-01', 'infered given past history in Excel.') |> 
  bind_rows(layer.meta)

#### Create level 1 #####
## Pull everything together
data.lvl1.ls <- list(
  study = study.df,
  collection = collection.df,
  site = site.df,
  layer = layer.df,
  citations = c(data.lvl0.ls$citation$primary, data.lvl0.ls$citation$contributed)
  )

rm(list = setdiff(ls(), c('data.lvl0.ls','data.lvl1.ls', 
                          'readsource.file', 'readISCN3')))

```

## Check semantics

```{r}
#| label: buildSemanticYAML

library(yaml)

variable_type.df <- plyr::ldply(data.lvl1.ls, function(xx){
  if(!is.data.frame(xx)){
    return(NULL)
  }else{
  ans <- xx |>
    select(of_variable, is_type) |>
      unique()
  
  return(ans)
  }
}) 

variable_type.ls <- variable_type.df |>
  select(of_variable, is_type) |>
  #filter(n() > 1, .by = everything()) #citation-value is repeated and ok
  unique() |>
  arrange(of_variable, is_type) |>
  plyr::dlply(.variables = 'of_variable', .fun = function(xx){
    return(list(is_type = xx$is_type))
  }) 


#use this to construct the opening list at the top of the level 1 section
# as.yaml(variable_type.ls) |> cat()

#TODO change this when the variable/type is moved over to a yaml
ref.ls <- variable_type.ls

check.df <- variable_type.df |>
  mutate(has_variable = of_variable %in% names(ref.ls)) |>
  mutate(has_type = is_type %in% ref.ls[[unique(of_variable)]]$is_type,
         .by = of_variable)

if(!all(check.df$has_variable)){
  stop('There are variables that needed to be added to the semantics')
}

if(!all(check.df$has_type)){
  stop('There are types that need to be added to the semantics')
}
```

## Check function match

Check for an exact match with the read function.

```{r}
#| label: checkReadFunction_lvl1

if(file.exists(readsource.file)){
  datafunction.lvl1.ls <- readISCN3(dataDir = '.',
                                         dataLevel = 'level1')
  
  if(isTRUE(all.equal(datafunction.lvl1.ls, data.lvl1.ls))){
    print('Read function output matches ReadMe code for level 1.')
  }else{
    print('There is a mismatch in the level 1 data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}

```


# References

<!---- This space will be filled with the citations when rendered. Leave this blank. ---->
