---
title: "Data Rescue: ISCN3"
format:
  html:
    toc: true
    code-fold: true
    eval: true
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - ISCN3_Contributors.bib  #contains the citations for contributed datasets by authors
  - ISCN3_2015.bib #contains citation for the data source, generally a single citation of the primary paper
nocite: |
  @* #add all citations
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  - id: ktoddbrown #use the github name as an id
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role: #TODO define these roles and add reference
      - transcription: lead #secondary, review
      - standardization: lead #review
      - curation: lead #review
  - id: vaasukimarupaka
    name:
      given: Vaasuki
      family: Marupaka
    orcid: 0000-0002-3854-1764
    affiliation:
      - ref: uf-ees
    role:
      - transcription: lead #secondary, review
      - standardization: lead
      - curation: lead #review
affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
    
timelog:
  - activity: 
      description: Lead Transcription and ReadMe Update
      who: Vaasuki Marupaka
      end: 2026-03-25
      start: 2026-02-06
      duration: PT30H
  - activity: 
      description: Revise readme
      who: ktoddbrown
      end: 2026-06-02T16:28
      start: 2026-06-02
---

```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false
#| eval: true


# libraries
library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R
library(stringr) # regular expressions

# Data download directory from EDI
dataDownload.dir <- 'temp'

### Level 0 rescued files ####
# These files have some level of manual rescue/transcription to them

# Control vocabulary transcribed from EDI file `TemplateCVs.pdf`
controlVocabulary <- 'ISCN3_TemplateCVs.csv'
controlVocabulary_header <- 'ISCN3_TemplateCV_preamble.md'

## Files from the `TemplateSubmit.pdf` document include a formatted header
##... and the table discriptions

# Transcription of the header from the EDI file `TemplateSubmit.pdf`
readME_templateSubmit <- 'ISCN3_TemplateSubmit_preamble.md'

# Rescued transcriptions of individual table meta data from the 
#... EDI file `TemplateSubmit.pdf` for the layer, profile, site
#... disturbance, and fraction tables. These do not match exactly the 
#... EDI archived tables but do provide the metadata for those tables
ISCN3_layerAnnotations <- 'ISCN3_TemplateSubmit_TableLayer.csv'
ISCN3_profileAnnotations <- 'ISCN3_TemplateSubmit_TableProfile.csv'
ISCN3_siteAnnotations <- 'ISCN3_TemplateSubmit_TableSite.csv'
ISCN3_disturbanceAnnotations <- 'ISCN3_TemplateSubmit_TableDisturbance.csv'
ISCN3_fractionAnnotations <- 'ISCN3_TemplateSubmit_TableFraction.csv'

#### Files that have to do with the fixed citations

# Transcribed citations from the EDI file `ISCN3_citation.csv`
#...table and then extended manually when files were mis-cited
rescuedContributedCitations <- 'ISCN3_ISCN3_citation.csv'

# Bibliography files
primaryCitation.file <- 'ISCN3_2015.bib'
contributionsCitation.file <- 'ISCN3_Contributors.bib'

### Reference to the read function
# read function
readsource.file <- '../../02_DataHarmonization/readISCN3.R'
if(file.exists(readsource.file)){
  source(readsource.file)
}

#semantic files for level 1
of_variable.file <- '../../02_DataHarmonization/SoilDRaHVocabulary_of_variable.csv'
is_type.file <- '../../02_DataHarmonization/SoilDRaHVocabulary_is_type.csv'

level1_controlVocabulary <- list(soc = 'carbon',
                                 bd = 'bulk_density',
                                 cf = 'coarse_fraction',
                                 layer = 'layer',
                                 geolocation = 'geolocation',
                                 om = 'organic_fraction')

```

# Data Summary

> The ISCN is an international scientific community devoted to the advancement of soil carbon research. The ISCN manages an open-access, community-driven soil carbon database. This is version 3-1 of the ISCN Database, released in December 2015. It gathers 38 separate data set contributions, totaling 67,112 sites with data from 71,198 soil profiles and 431,324 soil layers.

*Data source citation:* Nave, L., K. Johnson, C. van Ingen, D. Agarwal, M. Humphrey, and N. Beekwilder. 2022. International Soil Carbon Network version 3 Database (ISCN3) ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/cc751923c5576b95a6d6a227d5afe8ba (Accessed 2026-03-01).

## Fit for purpose: ISCN4

This data is identified as a data source for ISCN2025.
The purpose ISCN 2025 (aka ISCN4) is to create a data product aggregating data on geo-located soil organic carbon profiles.
With this data we will examine the assumption of steady state on field surveyed organic carbon.

ISCN3 contains four primary data tables:

  * `profile` containing the profile information including geolocation and site information, 
  * `layer` containing layer level soil characterizations, 
  * `citation` containing the citation and additional references information for the contributed datasets,
  * `dataset` with study information that extends the basic citation

The EDI data archive also includes metadata in submission templates.
This information is in pdfs files and required a manual transcription.

## Files

<!--- Include a brief explanation to all the files in the folder--->

These files are in the ISCN3 data rescue.

- [Readme_ISCN3.qmd](README_ISCN3.qmd) This is the primary file that documents the transcriptions and decision made during data rescue.
- [ISCN3_2015.bib](ISCN3_2015.bib) citation for project website and primary journal article transcribed
- [ISCN3_Contributors.bib](ISCN3_Contributors.bib) citations for the contributed datasets to ISCN
- [ISCN3_ISCN3_citation.csv](ISCN3_ISCN3_citation.csv) Rescued citations where the citations were replaced with bib references
- [ISCN3_TemplateCV_preamble.md](ISCN3_TemplateCV_preamble.md) control vocabulary text information
- [ISCN3_TemplateCVs.csv](ISCN3_TemplateCVs.csv) control vocabulary tables
- [ISCN3_TemplateSubmit_preamble.md](ISCN3_TemplateSubmit_preamble.md)  submission guidance header
- [ISCN3_TemplateSubmit_TableSite.csv](ISCN3_TemplateSubmit_TableSite.csv) submission guidence that has meta data for the site table
- [ISCN3_TemplateSubmit_TableProfile.csv](ISCN3_TemplateSubmit_TableProfile.csv) submission guidance that has meta data for the profile table
- [ISCN3_TemplateSubmit_TableLayer.csv](ISCN3_TemplateSubmit_TableLayer.csv) submission guidance that has meta data for the layer table
- [ISCN3_TemplateSubmit_TableDisturbance.csv](ISCN3_TemplateSubmit_TableDisturbance.csv) submission guidance that has meta data for the disturbance table (not in ISCN3 tables)
- [ISCN3_TemplateSubmit_TableFraction.csv](ISCN3_TemplateSubmit_TableFraction.csv) submission guidance that has meta data for the fraction table (not in ISCN3 tables)
- [temp/README.md](temp/README.md) a placeholder file for the download folder that contains the EDI citation.

The ISCN3 files are archived on EDI web portal and will be downloaded to the `temp` folder in the chunk below:

  + [ISCN3_citation.csv](ISCN3_citation.csv)
  + [ISCN3_dataset.csv](ISCN3_dataset.csv)
  + [ISCN3_layer.csv](ISCN3_layer.csv)
  + [ISCN3_profile.csv](ISCN3_profile.csv)
  + [ISCNtemplate.xlsx](ISCNtemplate.xlsx)
  + [ISCNTranscribed_TemplateCVs.xlsx](ISCNTranscribed_TemplateCVs.xlsx)
  + [TemplateCVs.pdf](TemplateCVs.pdf) transcribed source pdf
  + [TemplateSubmit.pdf](TemplateSubmit.pdf) transcribed source pdf

<!---- # Data Rescue working notes 

## Data Rescue Plan

+ Lead
  - [x] Read the article or data resource and the discription of the target project
  - [x] Create a fresh fork of the primary repository https://github.com/ktoddbrown/SoilDRaH/tree/main, it might be useful to rename it 'SoilDRaH_AuthorYYYY'
  - [x] Create a copy of this template folder and rename it to AuthorYYYY
  - [x] Fill out Data summary section (except for the Files), the Data Rescue Plan (here!), and update the author section of the yml to reflect who is working on the rescue
  - [x] Create a GitHub issue from the Data Rescue template and get feedback from a reviewer on the project (Dr Todd-Brown or a grad student typically).
  - [x] Add the reviewer to the author list in the yml
+ Do
  - [x] Code up data download from EDI
  - [x] Transcribe the EDI tables in `TemplateSubmit.pdf` into `ISCN3_TemplateSubmit_TableX.csv` 
  - [x] Transcribe `TemplateSubmit.pdf` markdown for preamble text
  - [x] Transcribe the EDI tables in `TemplateCV.pdf` into `ISCN_TemplateCV.csv`
  - [x] Transcribe `TemplateCV.pdf` markdown for preamble text
  - [x] Create a bib file with the recommended citations
  - [x] Replace the citations with bib references and treat `ISCN3_ISCN3_citation.csv` as a rescue object 
  - [ ] Create a recommended citations and acknowledgements table based on 'citations' as a markdown document that inserts the bib entries
  - [ ] Read in primary tables and template information 
  - [ ] read function match `02_DataHarmonization/readAres2001.R` function inputs and output formats
  - [x] push annotations into an `old` folder as a place holder
  - [x] ask one of the undergrads to provide a secondary transcription for the TemplateCV and TemplateSubmit pdfs
  - [ ] push everything to the lead fork
+ Measure
  - [x] cross check with second transcriber for tables and methods, and resolve any differences
  - [ ] compile or render this ReadMe with all chunks up through the Level 0 data rescue section with `eval: true`
  - [ ] push to your local fork and create a pull request to the main branch of the primary repository
  - [ ] address any revisions in your pull request review
  - [ ] celebrate!

## Citation notes

  - `AK DSC Project SOC stock computation` had a bad url for the doi that linked to a spam site. Updating doi with the EDI pasta link.
  - `Jorgensen_ARCN` has this reference citation: Jorgenson, M.T., Swanson, D.K., Macander, M. 2001. Ecological subsections of Noatak National Preserve. 73. But when looked up, it does not match the destination file.
  - `Kane` dataset has this citation reference: Valentine, David. 2004. Soil profiles along productivity gradients in interior Alaska: Summer 2003, Bonanza Creek LTER - University of Alaska Fairbanks. BNZ:133, http://www.lter.uaf.edu/data_detail.cfm?datafile_pkey=133. Fixed this citation in rescued citations file with the correct reference citation.
  - `NASIS-2014` link doesn't work anymore.
  - `Worldwide soil carbon and nitrogen data` dataset has invalid doi.
--->
## Download ISCN3

The ISCN3 is downloaded from EDI (@ISCN3). 
Below we loop through the files that are expected and download them.

```{r}
#| label: dataDownloadEDI
#| code-summary: "Download and identify each file in the EDI package."

### Download the data ####
# link the table name to the download url and file path(s) for data package
pasta_package <- 'https://pasta.lternet.edu/package/data/eml/edi/1160/1/'

#transcribe from the EDI site, should be stable and not need future updates
download_table <- tibble::tribble(
  ~table, ~entity_id, ~file_name, 
  'layer', "4af719a84f8981fcc63f1f92760cb253", 'ISCN3_layer.csv',
  'profile', "40527580cc045d33d9a5aaf728bf204e", 'ISCN3_profile.csv',
  'citation', "320e31ca911f187550ca2143c31fd408", 'ISCN3_citation.csv',
  'dataset', "cdd0c7a4cac3f28d6d788c91f506775f", 'ISCN3_dataset.csv',
  'control_vocabulary', "a8eef6e94b669b365e443c15d9402a03", 'ISCNTranscribed_TemplateCVs.xlsx',
  'template', "cb5c1cd86e9ab17f699f7a05044325b7", 'ISCNtemplate.xlsx',
  'other_pdfs', "eb320ae7b57296765f543cbb370b0f24", 'TemplateCVs.pdf',
  'other_pdfs', "b81a7f4214d176280a5ff4e0f0d52d8b", 'TemplateSubmit.pdf')

#One by one, download the data in each row from the download_table
#... and store that in the appropriate file path location on your computer
for(row_index in 1:nrow(download_table)){
  #check to see if the file exists before trying to download it
  if(!file.exists(file.path(dataDownload.dir,
                            download_table$file_name[row_index]))){
    #tell the user you are downloading the file
    print(paste('Download file:', 
                download_table$file_name[row_index]))
    #use the utility download function to pull the file based on the 
    #... pasta url for ISCN3 and the entity ID, match that up with the file
    #... name on that row
    utils::download.file(
      paste0(pasta_package, download_table$entity_id[row_index]), 
      file.path(dataDownload.dir, download_table$file_name[row_index]), 
      extra  = "--max-time 300")
  } #end file exists check
} #end for loop through the download table

```

# Level 0 data

```{r}
#| label: readLevel0
#This chunk has two purposes. 
#...1) Check the formatting by reading in everything 
#...2) Create a list of everything to process later in level 1

data.lvl0.ls <- list(
  ####Read in a list of all the bib files####
  citation = list(
    #Citation for the article transcriptions are pulled from
    primary = read.bib(file = primaryCitation.file), 
    #Citations for all referenced articles
    contributed = read.bib(file = contributionsCitation.file)
  ),
  ####Read in the tables and their meta data###
  data = list(
    ####Citation table####
    TableCitation = list(
      #manual correction to out of date citations, originally
      #...this was the `temp/ISCN_citation.csv`
      primary = readr::read_delim(file = 'ISCN3_ISCN3_citation.csv', 
                                  skip = 1,
                                  delim = ';', 
                                  col_types = 
                                    readr::cols(.default = 
                                                  readr::col_character())),
      caption = readr::read_delim(file = 'ISCN3_ISCN3_citation.csv', 
                                  n_max = 1,
                                  delim = ';',
                                  col_names = FALSE,
                                  col_types = 
                                    readr::cols(.default = 
                                                  readr::col_character()))[[1]]
      #look in the metadata in the profile table for some variables
    ), #end citation list
    ####Dataset table####
    TableDataset = list(
      primary = readr::read_delim(file = 'temp/ISCN3_dataset.csv', 
                                  delim = ';', 
                                  col_types = 
                                    readr::cols(.default = 
                                                  readr::col_character()))
      #look in the metadata in the profile table for some variables
    ), #end dataset list
    ###Site table####
    TableSite = list(
      #no primary data here for sites, look in the profile and layer tables
      meta = readr::read_delim(file = 'ISCN3_TemplateSubmit_TableSite.csv',
                                delim = ",",
        col_types = readr::cols(.default = readr::col_character())
      )
    ), #end site
    TableFraction = list(
      #no primary data here
      meta = readr::read_delim(file = 'ISCN3_TemplateSubmit_TableFraction.csv',
                                delim = ",",
        col_types = readr::cols(.default = readr::col_character())
      )
    ), #end fraction
    TableDisturbance = list(
      #no primary data here
      meta = readr::read_delim(file = 'ISCN3_TemplateSubmit_TableDisturbance.csv',
                                delim = ",",
        col_types = readr::cols(.default = readr::col_character())
      )
    ), #end disturbance
    ###Profile table####
    TableProfile = list(
      primary = readr::read_delim(file = 'temp/ISCN3_profile.csv', 
                                  delim = ';', 
                                  col_types = 
                                    readr::cols(.default = 
                                                  readr::col_character())),
      meta = readr::read_delim( file = 'ISCN3_TemplateSubmit_TableProfile.csv',
                                delim = ",",
                                col_types = 
                                  readr::cols(.default =
                                                readr::col_character()))
    ), #end profile
    #### Layer table ####
    TableLayer = list(
      primary = readr::read_delim(file = 'temp/ISCN3_layer.csv', 
                                  delim = ';', 
                                  col_types = 
                                    readr::cols(.default = 
                                                  readr::col_character())),
      meta = readr::read_delim( file = 'ISCN3_TemplateSubmit_TableLayer.csv', 
                                delim = ",",
                                col_types = readr::cols(
                                  .default = readr::col_character()))
    ) #end layer table
    ), # end data list
  #### General control vocabulary that is not broke down by table ####
  control_vocabulary = list(
    readme = read_lines(file = 'ISCN3_TemplateCV_preamble.md'),
    caption = readr::read_delim( file = 'ISCN3_TemplateCVs.csv', 
                       n_max = 1,
                       col_names = FALSE,
                       delim = ",",
                       col_types = readr::cols(
                         .default = readr::col_character()))[[1]],
    primary = readr::read_delim( file = 'ISCN3_TemplateCVs.csv', 
                       skip = 1,
                       delim = ",",
                       col_types = readr::cols(
                         .default = readr::col_character())))
  )
  
```

<!---- consider making a diagram or nested list to identify the data structures here--->

## Check function match

Check for an exact match with the read function.

```{r}
#| label: checkReadFunction_lvl0

if(file.exists(readsource.file)){
  datafunction.lvl0.ls <- readISCN3(dataDir = '.',
                                         dataLevel = 'level0')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function matches code here.')
  }else{
    print('There is a mismatch in the data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}
```

# Level 1 data - ISCN 2025

This data is identified as a data source for the ISCN2025
Broad classes of information that we are looking for include:

For this analysis we are interested in:

  * **Location**: latitude/longitude, associated datum, geo-spatial method, precision
  * **Observation time**: observation year
  * **Soil organic carbon stocks**: organic carbon fraction, bulk density, coarse fraction, soil organic carbon by area, soil organic carbon by volume
  * **Layer depth**: Depth of surveyed soil layer

Note that level 1 data processing below removes the NRCS data contribution from ISCN3.
See the SoilDRaH entry for the NCSS data as a updated replacement for this or work directly with the level 0 data above.

This code takes a two step approach. First it isolates each of the variables of interest with the `is_type` observations as columns. Any known corrections or filters are applied in this first step.

```{r}
#| label: level1_constants

precisions.ls <- list(latLong_digits = 4, #assume lat/lon precision of 0.0001
                      #cluster resolution in lat/lon degrees
                      geolocationCluster_resolution = 0.001, 
                      #cluster resolution in years
                      observationYear_resolution = 2, 
                      #assume precision of 1cm
                      depth_digits = 0,
                      #assume precision of bulk density at 0.01 g cm-3
                      bulkDensity_digits = 2,
                      #assume precision of 0.1% for loss on ignition
                      organicFraction_digits = 1,
                      #assume precision of 0.01% for carbon fraction that is often measured using an elemental analyizer, this could be method specific in the future.
                      carbonFraction_digits = 2,
                      #since this is normalized to the layer thickness, use a very high precision here
                      carbonStock_digits = 10) 
```

## Citations pre-processing: expanding keys

Citations are keyed to datasets however the Permafrost RCN has seperate citations for the layer and profile level data. We use the site information from the profile and observations from the layer data, and need to expand the citations keyed to those datasets.

```{r}
#| label: ISCN3_corrections_citations
#| code-summary: 'ISCN3 needs expanded citations for the permafrost RCN.'


#collection table
collection.df <- 
  #### Create cross key for the citation dataset naming to include the double
  #### citations needed for the Permafrost RCN data sets
  data.lvl0.ls$data$TableDataset$primary |>
  select("dataset_name", #matches dataset_name_sub or dataset_name_soc
         #"dataset_type (dataset_type)",
         "curator_name", "curator_organization", "curator_email", 
         "contact_name", "contact_email", 
         "dataset_description"
  ) |>
  mutate(source_from = 'Table Dataset') |>
  #expand the Permafrost RCN datasets so we can pull in that citation
  reframe(dataset_name_citation = c('Permafrost_RCN', dataset_name)[c(str_detect(dataset_name, '^Permafrost RCN_'), TRUE)],
          .by = everything()) |>
  full_join(
    data.lvl0.ls$data$TableCitation$primary |>
      select(dataset_name_citation = "dataset_name", #matches dataset_name_sub or dataset_name_soc
             "reference", 
             "citation", "citation_usage", 
             "acknowledgement", "acknowledgement_usage") |>
      mutate(source_from = 'Table Citation'),
    relationship = "many-to-many",
    by = join_by(dataset_name_citation)) |>
  #### Trim NRCS, remove 2 collection. Work with the current NRCS data download instead
  filter(!str_detect(dataset_name, 'NRCS')) |>
  reframe(value = paste(sort(unique(citation)), collapse = ';'),
          from_source = paste(sort(unique(c(source_from.x, source_from.y))), collapse = ';'),
          .by = dataset_name) |>
  pivot_longer(cols = c('value', 'from_source'),
               names_to = 'is_type', values_to = 'with_entry',
               values_drop_na = TRUE) |>
  mutate(of_variable = 'bibkey')
  
```


## Space-time pre-processing: resolution and clusters

ISCN3 provides latitude/longitude and observation date.
We have identified three main issues with the data collection:

  1) Locations appear overly precise with several extending to 8 or more reported degrees of precision that appear to be numerical errors.
    - *To resolve* round the lat/lon values to `{r} precisions.ls$latLong_digits` digits.
  2) Several locations appear to be repeated with both the lat/long reported extremely closely and shared values of bulk density or soil carbon elements reported.
    - *To resolve* this complete the degree distance between all lat/lon and applied a `{r} precisions.ls$geolocationCluster_resolution` cluster height to define geolocation groups.
    - *Assume*: this ignores datum corrections, if exact location matters (<100m) you may need to reprocess the data from level 0.
  3) Observation dates appear to have standard export issues often seen in Access/Excel csv exports. Observation dates do not adhere to the reported YYYY-MM-DD format and are instead numerical values. Given the distribution of points its likely that some of these values were orginally entered as years and others reported as days-since-origin. Excel has historically exported their dates as days since 1900-01-01.
    - *To resolve* any values below 2015 were assumed to be CE-year and assigned the first of that year. Values above 2015 were assumed to be days since 1900-01-01.
  4) Observation dates with minor disagreements are likely duplicate entries.
    - *To resolve*, similar to geolocations, a distance matrix was calculated and clusters within `{r} precisions.ls$observationYear_resolution` were assigned unique identifies.

```{r}
#| label: ISCN3_geo
#| code-summary: 'ISCN3 repeats geolocation, identify clusters by location and observation time. ISCN3 also has inconsistant formating in the observation date; if number below 2015 assume year, otherwise days since origin.'

#> data.lvl0.ls$data$TableProfile$primary$`lat (dec. deg)` |> unique() |> as.numeric() |> sort() |> diff() |> quantile(c(0.01, 0.05, 0.1, 0.2, 0.5))
#     1%      5%     10%     20%     50% 
#0.00001 0.00002 0.00004 0.00008 0.00028

#Format numerical values and impose precision
site_org.df <- data.lvl0.ls$data$TableProfile$primary |>
  select("dataset_name_sub", "dataset_name_soc", #ID
         "site_name", "profile_name", #ID
         "lat (dec. deg)", "long (dec. deg)", "datum (datum)", #geolocation
         "state (state_province)", "country (country)", #region
         "observation_date (YYYY-MM-DD)", #observation_date
         "profile_zero_ref (profile_zero_ref)")  |>
  #remove 75 651 profiles. Work with the current NRCS data download instead
  filter(!str_detect(dataset_name_sub, 'NRCS')) |>
  #correct the observation, some are entered in as years an others 
  #...as days since origin (excel convention)
  mutate(date_number = as.numeric(`observation_date (YYYY-MM-DD)`)) |>
  mutate(observation_date_formula = case_when(is.na(date_number) ~ 'missing',
                                              date_number > 2015 ~ 'days since origin',
                                              date_number <= 2015 ~ 'year CE',
                                              .default = 'default') |>
           as.factor(),
    observation_date = 
           case_when(is.na(date_number) ~ NA_Date_,
                     #assign origin based on Excel/Access historical convention
                     date_number > 2015 ~ as_date(date_number, origin = ymd('1900-01-01')),
                     #process things with just the year specified using the truncated flag
                     date_number <= 2015 ~ ymd(`observation_date (YYYY-MM-DD)`, 
                                               quiet = TRUE, truncated = 2),
         .default = NA_Date_)) |>
  #set the geolocations to numerics
  mutate(across(c(`lat (dec. deg)`, `long (dec. deg)`), .fns = as.numeric)) |>
  #round to 4 digits of precision or about 10 m https://wiki.openstreetmap.org/wiki/Precision_of_coordinates
  #... many numbers look overly precise and there was likely numerical error
  #....introduced during initial processing (excel/access exports?)
  mutate(across(c(`lat (dec. deg)`, `long (dec. deg)`), 
                .fns = ~ round(.x, digits = precisions.ls$latLong_digits)))

#Define the location clusters if there are finite latitude/longitude reported using clusters on eucludian distances
site_geocluster <- site_org.df |>
  select("dataset_name_sub", #ID
         "site_name",  #ID
         "lat (dec. deg)", "long (dec. deg)", 'datum (datum)') |>
  #there is one site in the 'World soil carbon and nitrogen data' that has two
  #... lat-long locations specified. Kick it out.
  unique() |> filter(n() == 1, .by = c(dataset_name_sub, site_name)) |>
  #remove any location that is missing lat/lon, don't assign a geocluster
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`)) |>
  #Create a distance matrix based on the latitude and longitude,
  #...cluster the sites, and cut out similar groups. 
  #...Looking at the tree heights and the spread of the lat/lon, 
  #...it looks like the height corresponds to the degree separation of the 
  #...points. Degrees 0.001 will correspond to 100m and is one order of 
  #...magnitude greater then the enforced precision.
  mutate(geolocation_id = sprintf('geocluster_%05d_%s',
           dist(matrix(c(`lat (dec. deg)`, `long (dec. deg)`), ncol = 2)) |>
           hclust() |>
           cutree(h = precisions.ls$geolocationCluster_resolution),
         `datum (datum)`)) #|> #Using euclidean distance
  #filter(n() > 1, .by = geolocation_id) #take a look at the clustered locations

#What happens if we change the tolerance
# temp <-  dist(matrix(c(site_geocluster$`lat (dec. deg)`,
#              site_geocluster$`long (dec. deg)`), ncol = 2)) |>
#           hclust()
#summary(temp$height > 0.0001) #5443 unique clusters
#summary(temp$height > 0.001) #5117 unique clusters
#summary(temp$height > 0.01) #4205 unique clusters

#Define time clusters similarlly to the spatial clusters
site_temporalCluster <- site_org.df |>
  select("dataset_name_sub", #ID
         "site_name",  #ID
         observation_date) |>
  #only define a cluster if we have an observation date
  filter(is.finite(observation_date)) |> unique() |>
  #create the temporal cluster
  mutate(date_year = year(observation_date),
         temporal_id = sprintf(
           'timecluster_%02d',
           dist(matrix(decimal_date(observation_date), ncol = 1)) |> 
             hclust() |>
             cutree(h = precisions.ls$observationYear_resolution))) #max height here corresponds to year span, consider two years to be effectively the same observation time

#temp <-  dist(matrix(decimal_date(site_temporalCluster$observation_date), ncol = 1)) |>
#           hclust()
#range(temp$height) # 0.00000 94.60383
#summary(temp$height > 1) #75 clusters
#summary(temp$height > 2) #43 clusters
#summary(temp$height > 5) #21 clusters
#summary(temp$height > 10) #11 clusters

#reconstruct the site information with the clusters developed above
#...note that we keep the exact lat/long and observate date provided 
#...and they are now associated with neighborhood clusters
site.df <- site_geocluster |>
  full_join(site_temporalCluster,
            by = join_by(dataset_name_sub, site_name)) |>
  full_join(site_org.df,
            by = join_by(dataset_name_sub, site_name,
                         `lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`,
                         observation_date)) 

uniqueSpaceTime <- site.df |>
  filter(!is.na(temporal_id) & ! is.na(geolocation_id)) |>
  select(temporal_id, geolocation_id) |> unique() |> nrow()
```

There are `{r} uniqueSpaceTime` unique geolocation-time pairs reported.

## Layer pre-processing: constructing identifiers

Processing the layer table considerations

  1) Identifiers provided under the `.*_name` pattern are not unique leading to issues pivoting the data during processing.
    - *To resolve* construct a `layer_id` from available zero reference and the upper/lower bounds. Ensure that the upper/lower bounds have a realistic resolution of `{r} precisions.ls$depth_digits` digits. Also construct a `sample_id` from the `dataset_name`, `site_name`, `profile_name`, and `layer_name`. The sample id is slightly redundent but allows us to consolodate the identifiers.
  2) Layer observations appear to be duplicated for different SOC calculations and duplications listed under multiple datasets.
    - *To resolve* add in the clusters developed in the site processing above so that we can better identify duplicate entries of variables when processing each variable type later in the code. Collapse the entry if the values match at the same geolocation-time-depth.
  

Things to also note here:
  - Removing the NRCS layers reduced row counts from 445 829 to 72 503.
  - Location information is given in both the site and layer tables. We assume the site is correct and ignore the layer information.

```{r}
#| label: ISCN3preprocess_layer

layer.df <- data.lvl0.ls$data$TableLayer$primary |>
  select(#"ISCN 1-1 (2015-12-10)", 
         "dataset_name_sub", "dataset_name_soc", 
         "site_name", "profile_name", 
         #"lat (dec. deg)", "long (dec. deg)", "state (state_province)", "country (country)", "observation_date (YYYY-MM-DD)", #use location information from the site not layer
         "layer_name", "layer_top (cm)", "layer_bot (cm)", "layer_note", 
         #"hzn_desgn_other", "hzn", "hzn_desgn", "color",
         #"vegclass_local", 
         #"soil_taxon", "soil_series", 
         "bd_method", "bd_samp (g cm-3)", "bd_tot (g cm-3)", "bd_whole (g cm-3)", "bd_other (g cm-3)", "bdNRCS_prep_code", 
         "cNRCS_prep_code", "c_method", "c_tot (percent)", "oc (percent)", "loi (percent)", 
         #"n_tot (percent)", "c_to_n (mass ratio)", 
         "soc (g cm-2)", "soc_carbon_flag", "soc_method", 
         #"ph_method", "ph_cacl", "ph_h2o", "ph_other", 
         "caco3 (percent)", 
         #"sand_tot_psa (percent)", "silt_tot_psa (percent)", "clay_tot_psa (percent)", 
         "wpg2_method", "wpg2 (percent)", 
         #"cat_exch (cmol H+ kg-1)", 
         #"al_dith (specified by al_fe_units)", "al_ox (specified by al_fe_units)", "al_other (specified by al_fe_units)", "fe_dith (specified by al_fe_units)", "fe_ox (specified by al_fe_units)", "fe_other (specified by al_fe_units)", "mn_dith (specified by al_fe_units)", "mn_ox (specified by al_fe_units)", "mn_other (specified by al_fe_units)", "al_fe_units (extract_units)", "al_fe_method",
         #"ca_al (specified by bc_units)", "ca_ext (specified by bc_units)", "k_ext (specified by bc_units)", "mg_ext (specified by bc_units)", "na_ext (specified by bc_units)", "bc_units (extract_units)", "bc_method", "base_sum (specified by cec_h_units)", "cec_sum (specified by cec_h_units)", "ecec (specified by cec_h_units)", "cec_h_units (extract_units)", 
         #"bs (percent)", "bs_sum (percent)", 
         #"h_ext (specified by metal_ext_units)", "zn_ext (specified by metal_ext_units)", "metal_ext_units (extract_units)", "metal_ext_method", 
         #"p_bray (specified by p_units)", "p_ox (specified by p_units)", "p_meh (specified by p_units)", "p_other (specified by p_units)", "p_units (extract_units)", "p_method", 
         #"root_quant_size", "root_weight (g)",
         #"15n (‰)", "13c (‰)", "14c (‰)", "14c_sigma (‰)", "14c_age (BP)", "14c_age_sigma (BP)", "fraction_modern", "fraction_modern_sigma", 
         #"textureClass", "locator_parent_alias"
         ) |>
  #remove 373 326 layers. Work with the current NRCS data download instead
  filter(!str_detect(dataset_name_sub, 'NRCS')) |> #72503 obs of 25 variables
  #remove columns that are empty because some of these are NRCS only observations
  filter_out(if_all(everything(), is.na)) |> #does not remove columns right now but might if the variables were expanded
  #add in the site information so that we can get the cluster ids
  left_join(site.df,
            by = join_by(dataset_name_sub, dataset_name_soc, site_name, profile_name)) |> 
  #enforce precision
  mutate(across(c(`layer_top (cm)`, `layer_bot (cm)`), 
                .fns= ~ as.numeric(.x, digits = precisions.ls$depth_digits) )) |>
  #Construct the layer ID from the rounded depths and 
  #...add in the zero reference if it exists
  mutate(layer_id = case_when( 
    !is.na(`profile_zero_ref (profile_zero_ref)`) &
      !is.na(`layer_top (cm)` + `layer_bot (cm)`) ~
      sprintf('layer (zero-ref:%s) [%d,%d]',
              `profile_zero_ref (profile_zero_ref)`,
              `layer_top (cm)`,`layer_bot (cm)`),
    is.na(`profile_zero_ref (profile_zero_ref)`) &
      !is.na(`layer_top (cm)` + `layer_bot (cm)`) ~ 
      sprintf('layer [%d,%d]', `layer_top (cm)`, `layer_bot (cm)`),
    !is.na(`profile_zero_ref (profile_zero_ref)`) &
      is.na(`layer_top (cm)` + `layer_bot (cm)`) ~
      sprintf('layer (zero-ref:%s)',
              `profile_zero_ref (profile_zero_ref)`),
    is.na(`profile_zero_ref (profile_zero_ref)`) &
      is.na(`layer_top (cm)` + `layer_bot (cm)`) ~ NA_character_,
    .default = 'missing'
    )) |> 
  #create a sample_id for simplification later from key:value pairs of the names
  #and the row number
  mutate(sample_id = sprintf('dataset:%s,site:%s,profile:%s,layer:%s, row:%d',
                             dataset_name_sub, site_name, 
                             profile_name, layer_name,
                             1:n()))
```

## Space-time processing: resolution and replication

Given the geo-location and temporal clusters:

1) Calculate the associated lat/lon, and observation times with each cluster as a `mean` and record the range of values as a `resolution.` 
2) Collapse the state/country for each geolocation cluster.

```{r}
#| label: ISCN3_spaceTime

geolocation.df <- layer.df |>
  select(geolocation_id, `lat (dec. deg)`, `long (dec. deg)`, datum = `datum (datum)`) |>
  filter(!is.na(geolocation_id)) |>
  unique() |> #5405
  #Deal with duplicate values
  reframe(latitude = mean(`lat (dec. deg)`),
          longitude = mean(`long (dec. deg)`),
          resolution = mean(c(diff(range(`lat (dec. deg)`)),
                              diff(range(`long (dec. deg)`)))),
          observation_count = n(),
          .by = c(geolocation_id, datum)) |> #5131
  mutate(unit = 'decimal degrees',
         of_variable = 'geolocation',
         #TODO set precision
         across(c(latitude, longitude, resolution, observation_count, datum), as.character)) |>
  pivot_longer(cols = c(latitude, longitude, unit, resolution, observation_count, datum),
               names_to = 'is_type', values_to = 'with_entry',
               values_drop_na = TRUE)

region.df <- layer.df |>
  select(geolocation_id, country = `country (country)`, state = `state (state_province)`) |>
  pivot_longer(c(country, state),
               names_to = 'is_type', values_to = 'with_entry',
                values_drop_na = TRUE) |>
  filter(with_entry != 'Unknown', !is.na(geolocation_id)) |>
  unique() |> #7767
  #Deal with duplicate values
  reframe(with_entry = paste(sort(unique(with_entry)), collapse = ';'),
          .by = c(geolocation_id, is_type)) |>
  #pivot_wider(names_from = is_type, values_from = with_entry) |> #5131
  mutate(of_variable = 'region') 

temporal.df <- layer.df |> 
  select(temporal_id, observation_date, `observation_date (YYYY-MM-DD)`, observation_date_formula) |>
  filter(!is.na(temporal_id)) |>
  unique() |> #664 obs
  #Deal with duplicate values
  reframe(date = mean(decimal_date(observation_date)) |>
            round(digits = 2),
          resolution = round(as.numeric(difftime(max(observation_date), min(observation_date), 
                                           units = 'days'))/365, digits = 2),
          observation_count = n(),
         .by = c(temporal_id)) |>
  mutate(unit = 'decimal year (CE)',
         of_variable = 'observation_year') |> #44 obs
  mutate(across(.cols = c(date, resolution, observation_count, unit), as.character)) |>
  pivot_longer(cols = c(date, resolution, observation_count, unit),
               names_to = 'is_type', values_to = 'with_entry',
                 values_drop_na = TRUE)

```

## Depth processing: trim

Most of the layer identifiers for upper and lower bounds were dealt with in the layer pre-processing. Create a depth interval data frame by selecting the layer related columns. 

```{r}
#| label: ISCN3_depth
depth.df <- layer.df  |>
  select(layer_id, 
         upper = `layer_top (cm)`, lower = `layer_bot (cm)`,
         zero_ref = `profile_zero_ref (profile_zero_ref)`) |> #72 503 obs
  filter(!is.na(upper + lower)) |> #59 895 obs
  mutate(of_variable = 'depth_interval',
         unit = 'cm',
         across(.cols = c(upper, lower), .fns = as.character)) |>
    pivot_longer(cols = c(upper, lower, zero_ref, unit),
                 names_to = 'is_type', values_to = 'with_entry',
                 values_drop_na = TRUE)
```


## Organic fraction: precision and replication

The organic fraction corresponds to a soil fraction that is made up of organic material. Fraction, in this context, is a physiochemical sepearation rather then a proportion.

1) Organic matter fraction values in the data frame generally have one digit of precision however several values carried out to over 6 digits with repeated 0's or 9's. These are likely Excel export artifacts
  - *To resolve* round the values to `{r} precisions.ls$organicFraction_digits` digits of precision.
2) There are likely duplicate observations (see geolocation).
  - *To resolve* collapse observations that are numerically the same and share location-time-layer cluster ids.

```{r}
#| label: ISCN3_organicFraction
#| 
organic_fraction.df <- layer.df  |>
  select(ends_with('_id'), dataset_name = dataset_name_sub,
         `loi (percent)`)  |>
  filter(!is.na(`loi (percent)`),  #938 obs
         !is.na(geolocation_id), !is.na(temporal_id)) |> #555 obs
  #remove excel export errors
  mutate(value = sprintf(paste0('%.', precisions.ls$carbonFraction_digits, 'f'),
                         as.numeric(`loi (percent)`))) |>
  #Deal with duplicate values
  reframe(sample_id = ifelse(n() == 1,
                            unique(sample_id), 
                            paste0('[MultiSourced]',
                                   paste(sort(unique(sample_id)), 
                                         collapse = ';'))),
         .by = c(geolocation_id, temporal_id, layer_id, 
                 dataset_name, #track for citation
                 value)) |> #553 obs
  select(geolocation_id, temporal_id, layer_id, sample_id,
         value, dataset_name) |>
  mutate(of_variable = 'organic_fraction',
         unit = 'mass-percent',
         method = 'loss on ignition from fine earth') |>
  pivot_longer(cols = c(value, unit, method),
               names_to = 'is_type', values_to = 'with_entry',
               values_drop_na = TRUE) #|>
#check that rows are uniquely identified
#  pivot_wider(names_from = 'is_type', values_from = 'with_entry')
```

## Carbon fraction: precision, method harmonization, and duplicates

Processing carbon fraction considerations:

1) Carbon fraction values generally showed two decimals of precision however several values carried out to over 6 digits with repeated 0's or 9's. These are likely Excel export artifacts
  - *To resolve* format the carbon fraction values to `{r} precisions.ls$carbonFraction_digits` digits of precision.
2) Three types of organci carbon measurements are identified by ISCN3 `oc (percent)`, `c_tot (percent)`, and `caco3 (percent)`. 
  - *To resolve* manually code a table of all unique column name and method notes mapping these to a standardized vocabulary. This is then mapped in turn with a `case_when` statement to the SoilDRaH vocabulary. Hopefully, this two step process will make it easier to maintain.
  - Note that some of the collapsed values have conflicting methods. Keep this conflict noted in the methods section for now and deal with in the curation step.
3) There are likely duplicate observations (see geolocation).
  - *To resolve* collapse observations that are numerically the same and share location-time cluster ids.

Note that in this case fraction does not refer to a ratio but rather a physical or chemical subset of the soil. 

```{r}
#|label: ISCN3_carbonfraction

#### Carbon fraction map ####
#### With the removal of the NRCS data the cNRCS_prep_code is all missing
#### thus not included here
carbonFraction_map <- tribble(~column_name,~c_method, ~what, ~how,
"oc (percent)","MIR", 'carbon_fraction:organic', 'formula:MIR',
"c_tot (percent)","dry combustion, LECO analyzer", 'carbon_fraction:total', 'dry combustion:elemental analyzer',
"c_tot (percent)",NA, 'carbon_fraction:total', NA,
"c_tot (percent)","dry combustion",'carbon_fraction:total', 'dry combustion',
"c_tot (percent)","Perkin Elmer CHN (model 2400).", 'carbon_fraction:total', 'elemental analyzer',
"c_tot (percent)","M", 'carbon_fraction:total', NA,
"c_tot (percent)","G", 'carbon_fraction:total', NA,
"oc (percent)","L", 'carbon_fraction:organic', NA,
"oc (percent)","A", 'carbon_fraction:organic', NA,
"oc (percent)","T", 'carbon_fraction:organic', NA,
"oc (percent)","wet oxidation", 'carbon_fraction:organic', 'wet oxidation',
"caco3 (percent)","dry combustion", 'carbon_fraction:inorganic:CaCO3', 'dry combustion',
"caco3 (percent)",NA, 'carbon_fraction:inorganic:CaCO3', NA,
"oc (percent)","Determined total %C from high temp combustion + elemental analyzer (on <2mm fine earth fraction)", 'carbon_fraction:organic', 'dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2399",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2400",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2417",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2418",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2401",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2402",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2419",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2420",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2405",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2406",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2423",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2424",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2407",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2408",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2425",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2426",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2411",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2412",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2429",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2430",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2415",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2416",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2433",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2434",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2403",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2404",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2421",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2422",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2409",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2410",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2427",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2428",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2413",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2414",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2431",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"c_tot (percent)","mean of ≥3 samples, dry combustion, Perkin-Elmer 2432",'carbon_fraction:total', 'mean:dry combustion:elemental analyzer',
"caco3 (percent)","A", 'carbon_fraction:inorganic:CaCO3', NA,
"c_tot (percent)","Leco TruSpec CHN (model 630-100-400).", 'carbon_fraction:total', 'elemental analyzer',
"oc (percent)","manometric", 'carbon_fraction:organic', 'manometric',
"c_tot (percent)","dry combustion (CHN) of oven-dried, root-free sample, <2mm fraction", 'carbon_fraction:total', 'dry combustion:elemental analyzer',
"c_tot (percent)","Element Analyzer.  Loss of Ignition", 'carbon_fraction:total', 'formula:loss on ignition:elemental analyzer',
"c_tot (percent)","Carlo Erba CN Elemental Analyzer.", 'carbon_fraction:total', 'elemental analyzer',
"caco3 (percent)","wet oxidation",'carbon_fraction:inorganic:CaCO3', 'wet oxidation',
"oc (percent)",NA, 'carbon_fraction:organic', NA,
"oc (percent)","assumed to have no inorganic carbonate ; Combustion analyser 1200 deg C  (Manies et al, 2006)", 'carbon_fraction:organic', 'dry combustion:elemental analyzer:assumed no inorganics',
"oc (percent)","Fumigated in 12 M HCl (48 h) to remove carbonates (Hedges and Stern, 1984);.freeze-dried and kept in a desiccator; elemental analyzer (Carlo Erba 2500) used to determine C% and N% content (all samples were we have N% values).", 'carbon_fraction:organic', 'elemental analyzer:acid removal of inorganics',
"c_tot (percent)","assumed to have no inorganic carbonate ; Combustion analyser 1200 deg C  (Manies et al, 2006)", 'carbon_fraction:total', 'dry combustion:elemental analyzer',
"c_tot (percent)","all mineral samples fumigated in 12M HCl (48h) to remove carbonates (Hedges and Stern 1984) then run by combustion analyzer re Manies et al, 2006", 'carbon_fraction:total', 'elemental analyzer',
"oc (percent)","all mineral samples fumigated in 12M HCl (48h) to remove carbonates (Hedges and Stern 1984) then run by combustion analyzer re Manies et al, 2006", 'carbon_fraction:organic', 'elemental analyzer:acid removal of inorganics',
"oc (percent)","Site vegetation reflects pH; acidic vegetation sites do not have carbonates, but sites with non-tussock vegetation could also contain carbonates within the %C measurement reported here", 'carbon_fraction:total', NA,
"oc (percent)","8",'carbon_fraction:organic', NA,
"oc (percent)","Loss on Ignition (LOI, weight %) at 550° C (for 6 h) was used to determine organic content, and at 950° C (for 2h) to determine carbonate content (Dean, 1974; Heiri et al., 2001).  A 3rd order polynomial regression model based on individual soil samples where both LOI and C% (from elemental analyser) were available (n= 171, R2= 0.98) was used to predict C% (y) for the soil samples where only LOI% was available: y= (-0.00005x3)+(0.0059x2)+(0.362x)", 'carbon_fraction:organic', 'formula:loss on ignition')

#### Process carbon fraction data####

carbon_fraction.df <- layer.df  |>
  select(c(ends_with('_id'), dataset_name = dataset_name_sub,
           `oc (percent)`, `c_tot (percent)`, `caco3 (percent)`,
           #cNRCS_prep_code, #empty after the removal of NRCS data
           c_method))  |>
  pivot_longer(cols = contains('percent'),
               names_to = 'column_name', values_to = 'value', 
               values_drop_na = TRUE) |>
  #merge in the column names to sample IDs
  mutate(sample_id = paste0(sample_id, ',column_name:', column_name)) |>
  # apply precision
  mutate(value = sprintf(paste0('%.', precisions.ls$carbonFraction_digits, 'f'),
                         as.numeric(value))
  ) |>
  # standardize methods
  left_join(carbonFraction_map,
            by = join_by(column_name, c_method)) |>
  mutate(of_variable = str_extract(what, pattern = '\\w+'),
         method = if_else(is.na(how),
                          str_remove(what, pattern = '\\w+:'),
                          paste(str_remove(what, pattern = '\\w+:'),
                                how,sep = '-'))) |> #51 059 obs
  #Deal with duplicate values
  reframe(sample_id = ifelse(n() == 1 | length(unique(sample_id)) == 1,
                            unique(sample_id), 
                            paste0('[MultiSourced]', paste(unique(sample_id), collapse = ';'))),
         method = paste(unique(method[!is.na(method)]), collapse = ';'),
         #there are a small ammount of layers that share OC, C and/or CaCO3 values
         #... which is possible, so group on this as well.
         #observation_count = n(),
         dataset_name = paste(sort(unique(dataset_name)), sep = ';'),
         .by = c(geolocation_id, temporal_id, layer_id, of_variable, value)) |> #44 902 obs
  mutate(unit = 'mass-percent') |>
  pivot_longer(cols = c(unit, value, method),
               names_to = 'is_type', values_to = 'with_entry',
               values_drop_na = TRUE) #|>
  #pivot_wider(names_from = 'is_type', values_from = 'with_entry')
```

## Bulk Density: map methods, enforce precision, and ensure unique id

Processing bulk density considerations:

1) Bulk density values generally showed two decimals of precision however several values carried out to over 6 digits with repeated 0's or 9's. Theres are likely Excel export artifacts
  - *To resolve* round the bulk density values to `{r} precisions.ls$bulkDensity_digits` digits of precision.
2) Three types of bulk density measurements are identified by ISCN3 `bd_samp`, `bd_tot`, `bd_whole` and a catch all `bd_other`. 
  - *To resolve* manually code a table of all unique column name and method notes mapping these to a standardized vocabulary. This is then mapped in turn with a `case_when` statement to the SoilDRaH vocabulary. Hopefully, this two step process will make it easier to maintain.
  - Note that some of the collapsed values have conflicting methods. Keep this conflict noted in the methods section for now
3) Some layers have multiple bulk density observations making their layer/sample identifiers non-unique for these observations.
  - *To resolve* create a unique observation identifier.
4) There are likely duplicate observations (see geolocation).
  - *To resolve* collapse observations that are numerically the same and share location-time cluster ids.


```{r}
#| label: ISCN3_bulkDensity

### Create bulk density method map####
# Map the methods to column names and bulk density 
bulkDensity_methodMap <- tribble(~column_name, ~bd_method, ~bdNRCS_prep_code, ~what, ~how,
"bd_samp (g cm-3)","bd_samp values are means of multiple observations",NA, 'bulk_density-fine_earth', 'measured:mean',
"bd_other (g cm-3)","bd = dry weight/volume",NA, 'bulk_density-unspecified',NA,
#TODO likely NRCS prep codes, G. and M. seem to be leaders for air dried prep
"bd_tot (g cm-3)","G",NA, 'bulk_density-whole_soil', NA,
"bd_tot (g cm-3)","M",NA, 'bulk_density-whole_soil', NA,
"bd_other (g cm-3)","G",NA, 'bulk_density-unspecified', NA,
"bd_other (g cm-3)","M",NA, 'bulk_density-unspecified', NA,
"bd_samp (g cm-3)","measured as weight per unit volume",NA, 'bulk_density-fine_earth', NA,
"bd_other (g cm-3)","predicted from %C",NA, 'bulk_density-unspecified', 'formula:carbon_fraction',
"bd_other (g cm-3)","estimated from weight per unit area",NA, 'bulk_density-unspecified', 'estimated:weight per area',
"bd_other (g cm-3)","gap-filled with value from nearby sample",NA, 'bulk_density-unspecified', 'gap_filled:neighboring_samples',
"bd_samp (g cm-3)","calculated based on soil texture, rock fragment, and organic matter content",NA, 'bulk_density-fine_earth', 'formula:soil texture, rock fragment, organic matter',
"bd_tot (g cm-3)","gap-filled value from re-measurement of same plot in 2000",NA, 'bulk_density-whole_soil', 'gap_filled:resampled',
"bd_tot (g cm-3)","core volume/mass",NA, 'bulk_density-whole_soil', NA,
"bd_samp (g cm-3)","Calculated as oven-dry mass of core, <2mm fraction, divided by core volume.",NA, 'bulk_density-stock', 'measured:fine dry mass over core volume',
"bd_other (g cm-3)","bd = (1.45*0.111)/((loi/100)*1.45+(1-(loi/100))*0.111)",NA, 'bulk_density-unspecified', 'formula:organic_matter_fraction',
"bd_other (g cm-3)","interpolated from adjacent horizons",NA, 'bulk_density-unspecified', 'gap_filled:neighboring horizons',
"bd_tot (g cm-3)",NA,NA, 'bulk_density-whole_soil', NA,
"bd_samp (g cm-3)",NA,NA, 'bulk_density-fine_earth', NA,
"bd_other (g cm-3)","g of <2mm fraction divided by total volume of material (often includes roots)",NA, 'bulk_density-stock', 'measured:fine dry mass over core volume',
"bd_other (g cm-3)","gap-filled from horizon type",NA, 'bulk_density-unspecified', 'gap_filled:neighboring_horizons',
"bd_other (g cm-3)","BD was estimated based on C % using a power-based regression model based on individual soil samples where both C % (x) and DBD (y) were available (n= 831, R2= 0.73): y= 1.2559x-0.6831. In samples with permafrsot, the regression-derived BD was corrected for volumetric ice-content",NA, 'bulk_density-unspecified', 'formula:carbon_fraction',
"bd_other (g cm-3)","took average of similar horizons types",NA, 'bulk_density-unspecified', 'gap_filled:similar horizons')

### Process bulk density ####
# Process the data frame
bulkdensity.df <- layer.df  |>  #72 503 obs
  select(ends_with('_id'),
         "bd_samp (g cm-3)", "bd_tot (g cm-3)", "bd_whole (g cm-3)",
         "bd_other (g cm-3)",
         "bdNRCS_prep_code", "bd_method", 
         dataset_name = dataset_name_sub) |> #dataset_name keep for citation development
  #Cast the bulk density values, ensure they are positive (-999 flags present), 
  #... and impose a precision to clean up export issues
  mutate(across(.cols = ends_with('(g cm-3)'), .fns = as.numeric)) |>
  pivot_longer(cols = ends_with('(g cm-3)'), names_to = 'column_name',
               values_to = 'value', values_drop_na = TRUE) |>
  filter(value >= 0) |> #ensure there are positive values
  #impose precision
  mutate(value = sprintf(paste0('%.', precisions.ls$carbonFraction_digits, 'f'),
                         value)) |>
  mutate(sample_id = paste0(sample_id, ',column_name:', column_name)) |>
  #Join the manual methods map
  left_join(bulkDensity_methodMap,
            by = join_by(bd_method, bdNRCS_prep_code, column_name)) |>
  mutate(of_variable = str_extract(what, pattern = '^\\w+'),
         method = if_else(is.na(how),
                          str_remove(what, pattern = '^\\w+-'),
                          paste(str_remove(what, pattern = '^\\w+-'), 
                                how, sep = '-'))) |> #59 455 obs
  #For each set of IDs and unique values, if there are multiple samples
  #... collapse the other columns
  reframe(sample_id = paste0('[MultiSourced]', paste(unique(sample_id), collapse = ';')),
         dataset_name = paste(sort(unique(dataset_name)), collapse = ';'),
         method = paste(sort(unique(method)), collapse = ';'),
         .by = c(geolocation_id, temporal_id, layer_id, #IDs
                 of_variable, value)) |> #50 037 obs, roughly 15% removed as duplicate values
  mutate(unit = 'g cm-3') |>
  pivot_longer(cols = c(value, method, unit), 
               names_to = 'is_type', values_to = 'with_entry',
               values_drop_na = TRUE) #|>
  #pivot_wider(names_from = 'is_type', values_from = 'with_entry')
```

## Coarse fraction

Processing coarse fraction considerations:

1) Coarse fraction values are integers in the ISCN3 table. No precision needed. 
3) Some layers have multiple coarse fraction observations making their layer/sample identifiers non-unique for these observations.
  - *To resolve* create a unique observation identifier.
4) There are likely duplicate observations (see geolocation).
  - *To resolve* collapse observations that are numerically the same and share location-time cluster ids.


```{r}
#| label: ISCN3_coarseFraction

coarse_fraction.df <- layer.df  |>
  select(c(ends_with('_id'), dataset_name = dataset_name_sub,
           value = `wpg2 (percent)`, `wpg2_method`)) |>
  filter(!is.na(value)) |> #16 509 obs
  # # apply precision
  # mutate(value = sprintf(paste0('%.', precisions.ls$carbonFraction_digits, 'f'),
  #                        as.numeric(value))
  # ) |>
  # standardize methods
  mutate(of_variable = 'coarse_fraction',
         method = case_when(
           is.na(wpg2_method) ~ NA,
           wpg2_method == 'as observed in profile/field' ~ 'estimated:in field',
           wpg2_method == 'assumed' ~ 'estimated:assumed',
           wpg2_method == 'estimated' ~ 'estimated',
           wpg2_method == 'measured' ~ 'measured',
           wpg2_method == 'weight of >2mm fraction/weight of total sample' ~ NA,
           .default = '[bad coding STOP]'
         )) |> #16 509 obs
  #Deal with duplicate values
  reframe(sample_id = ifelse(n() == 1 | length(unique(sample_id)) == 1,
                            unique(sample_id), 
                            paste0('[MultiSourced]', paste(unique(sample_id), collapse = ';'))),
         method = paste(unique(method[!is.na(method)]), collapse = ';'),
         #observation_count = n(),
         dataset_name = paste(sort(unique(dataset_name)), sep = ';'),
         .by = c(geolocation_id, temporal_id, layer_id, of_variable, value)) |> #13 954 obs, roughly 15% removed
  #make sure the sample_id is unique for multiple observations of the same site
  #filter(n() > 1, .by = c(geolocation_id, temporal_id, layer_id, of_variable))
  mutate(unit = 'mass-percent') |>
  pivot_longer(cols = c(unit, value, method),
               names_to = 'is_type', values_to = 'with_entry',
               values_drop_na = TRUE) #|>
#make sure we have unique identifers
  #pivot_wider(names_from = 'is_type', values_from = 'with_entry')

```

## Create level 1

```{r}

id_key <- bulkdensity.df |>
    bind_rows(coarse_fraction.df) |>
    bind_rows(carbon_fraction.df) |>
  select(geolocation_id, temporal_id, layer_id) |>
  unique()

site.df <- id_key |>
  right_join(geolocation.df |> 
              bind_rows(region.df),
            by = join_by(geolocation_id),
            relationship = "many-to-many") |>
  bind_rows(
    id_key |>
      right_join(temporal.df,
                by = join_by(temporal_id),
                relationship = "many-to-many"))

data.lvl1.ls <- list(
  site = site.df,
  layer = bulkdensity.df |>
    bind_rows(coarse_fraction.df) |>
    bind_rows(carbon_fraction.df),
  collection = collection.df,
  citations = c(data.lvl0.ls$citation$primary,
                data.lvl0.ls$citation$contributed)
)
```

## Check semantics

```{r}
#| label: ISCN4SemanticProposal

proposedSemantics.filename <- '../../03_DataCuration/ISCN2025/proposedSemantics.csv'

semantics.df <- data.lvl1.ls$site |>
  select(of_variable, is_type) |>
  unique() |>
  bind_rows(
    data.lvl1.ls$layer |>
      select(of_variable, is_type) |>
      unique()
  ) |>
  bind_rows(
    data.lvl1.ls$collection |>
      select(of_variable, is_type) |>
      unique()
  ) |>
  mutate(site_id = 'ISCN3_2015') |>
  arrange(of_variable) 

#Write things the first time and then just check
#write_csv(semantics.df, file = proposedSemantics.filename)

ISCN4Semantics <- read_csv(proposedSemantics.filename)

if(anti_join(semantics.df, ISCN4Semantics,
             by = join_by(of_variable, is_type, site_id)) |> nrow() > 0){
  stop('new variables')
}else{
  print('Proposed semantics are included.')
}

```

```{r}
#| label: buildSemanticYAML

library(yaml)

variable_type.df <- plyr::ldply(data.lvl1.ls, function(xx){
  if(!is.data.frame(xx)){
    return(NULL)
  }else{
  ans <- xx |>
    select(of_variable, is_type) |>
      unique()
  
  return(ans)
  }
}) 

variable_type.ls <- variable_type.df |>
  select(of_variable, is_type) |>
  #filter(n() > 1, .by = everything()) #citation-value is repeated and ok
  unique() |>
  arrange(of_variable, is_type) |>
  plyr::dlply(.variables = 'of_variable', .fun = function(xx){
    return(list(is_type = xx$is_type))
  }) 


#use this to construct the opening list at the top of the level 1 section
# as.yaml(variable_type.ls) |> cat()

#TODO change this when the variable/type is moved over to a yaml
ref.ls <- variable_type.ls

check.df <- variable_type.df |>
  mutate(has_variable = of_variable %in% names(ref.ls)) |>
  mutate(has_type = is_type %in% ref.ls[[unique(of_variable)]]$is_type,
         .by = of_variable)

if(!all(check.df$has_variable)){
  stop('There are variables that needed to be added to the semantics')
}

if(!all(check.df$has_type)){
  stop('There are types that need to be added to the semantics')
}
```

## Check function match

Check for an exact match with the read function.

```{r}
#| label: checkReadFunction_lvl1

if(file.exists(readsource.file)){
  datafunction.lvl1.ls <- readISCN3(dataDir = '.',
                                         dataLevel = 'level1')
  
  if(isTRUE(all.equal(datafunction.lvl1.ls, data.lvl1.ls))){
    print('Read function output matches ReadMe code for level 1.')
  }else{
    print('There is a mismatch in the level 1 data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}

```


# References

<!---- This space will be filled with the citations when rendered. Leave this blank. ---->
