---
title: "Data Rescue: ISCN3 Feb 2026"
format:
  html:
    toc: true
    code-fold: true
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - ISCN3_Contributors.bib  #contains the citations for contributed datasets by authors
  - ISCN3_2015.bib #contains citation for the data source, generally a single citation of the primary paper
nocite: |
  @* #add all citations
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  - id: ktoddbrown #use the github name as an id
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role: #TODO define these roles and add reference
      - transcription: lead #secondary, review
      - standardization: lead #review
      - curation: lead #review
  - id: vaasukimarupaka
    name:
      given: Vaasuki
      family: Marupaka
    orcid: 0000-0002-3854-1764
    affiliation:
      - ref: uf-ees
    role:
      - transcription: lead #secondary, review
      - standardization: lead
      - curation: lead #review
affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
---

```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false
#| eval: true


# libraries
library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R
library(stringr) # regular expressions

# data directory, primary annotations & control vocabularies
dataDownload.dir <- '~/Documents/GitHub/SoilDRaH_forked/01_DataRescue/ISCN3_2015/temp/ISCN3'
dataAnnotations <- '~/Documents/GitHub/SoilDRaH_forked/01_DataRescue/ISCN3_2015/ISCN_dataAnnotations2026.csv'
controlVocabulary <- '~/Documents/GitHub/SoilDRaH_forked/01_DataRescue/ISCN3_2015/ISCN3_2015_Control_vocab.csv'

# The below csv tables are used to conctruct annotations for ISCN4
Annotations <- '~/Documents/GitHub/SoilDRaH_forked/01_DataRescue/ISCN3_2015/ISCN3_NewAnnotations1March2026/'
ISCN3_layerAnnotations <- file.path(Annotations, 'ISCN3_layerAnnotations.csv')
ISCN3_profileAnnotations <- file.path(Annotations, 'ISCN3_profileAnnotations.csv')
ISCN3_siteAnnotations <- file.path(Annotations, 'ISCN3_siteAnnotations.csv')
ISCN3_citationAnnotations <- file.path(Annotations, 'ISCN3_citationAnnotations.csv')
ISCN3_disturbance <- file.path(Annotations, 'ISCN3_disturbanceAnnotations.csv')
ISCN3_fractionAnnotations <- file.path(Annotations, 'ISCN3_fractionAnnotations.csv')

# Bibliograph files
primaryCitation.file <- '~/Documents/GitHub/SoilDRaH_forked/01_DataRescue/ISCN3_2015/ISCN3_2015.bib'
contributionsCitation.file <- '~/Documents/GitHub/SoilDRaH_forked/01_DataRescue/ISCN3_2015/ISCN3_Contributors.bib'

# read function
readsource.file <- '~/Documents/GitHub/SoilDRaH_forked/01_DataRescue/ISCN3_2015/readISCN3.R'
if(file.exists(readsource.file)){
  source(readsource.file)
}

#semantic files for level 1
of_variable.file <- '../../02_DataHarmonization/SoilDRaHVocabulary_of_variable.csv'
is_type.file <- '../../02_DataHarmonization/SoilDRaHVocabulary_is_type.csv'

```

# Data Summary

The purpose of this document is to summarize the International Soil Carbon Network Database and demonstrate how to use the `readISCN3` function.

> The ISCN is an international scientific community devoted to the advancement of soil carbon research. The ISCN manages an open-access, community-driven soil carbon database. This is version 3-1 of the ISCN Database, released in December 2015. It gathers 38 separate data set contributions, totaling 67,112 sites with data from 71,198 soil profiles and 431,324 soil layers. For more information about the ISCN, its scientific community and resources, data policies and partner networks visit: http://iscn.fluxdata.org/. For information about processes used to construct the DB: https://iscn.fluxdata.org/data/data-information/.

*Data source citation:* Nave, L., K. Johnson, C. van Ingen, D. Agarwal, M. Humphrey, and N. Beekwilder. 2022. International Soil Carbon Network version 3 Database (ISCN3) ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/cc751923c5576b95a6d6a227d5afe8ba (Accessed 2026-03-01).

## Fit for purpose: ISCN4

This data is identified as a data source for ISCN2025. The purpose ISCN 2025 (aka ISCN4) is to create a data product aggregating data on geo-located soil organic carbon profiles. With this data we will examine the assumption of steady state on field surveyed organic carbon.

ISCN3 contains four primary data tables:: a `profile` table containing the profile information, a `layer` table containing layer level data for the different soil profiles, a `citation` table containing the citation and additional references information for the contributed datasets and finally a `dataset` table with study information. 
For the above These data annotation table is provided in a `<id>-of_variable-is_type-with_entry` tuple data model.

## Files

<!--- Include a brief explanation to all the files in the folder--->

These files are in the ISCN3 data rescue.

-   [Readme](README_ISCN3.qmd)
    -   This is the primary file that documents the transcriptions and decision made during data rescue.
-   [ISCN3_2015.bib](ISCN3_2015.bib)
    -   citation for project website and primary journal article transcribed
-   [ISCN3_Contributors.bib](ISCN3_Contributors.bib)
    -   citations for the contributed datasets to ISCN
    [ISCN3_dataAnnotations2026](ISCN3_dataAnnotations2026)
    -   this includes the annotations constructed using csv tables pulled from edi
-   [temp/](temp/)
    -   the download location for ISCN3 files archived on EDI web portal
    -   this folder includes the following files:
        + [ISCN3_citation.csv](ISCN3_citation.csv)
        + [ISCN3_dataset.csv](ISCN3_dataset.csv)
        + [ISCN3_layer.csv](ISCN3_layer.csv)
        + [ISCN3_profile.csv](ISCN3_profile.csv)
        + [ISCNtemplate.xlsx](ISCNtemplate.xlsx)
        + [ISCNTranscribed_TemplateCVs.xlsx](ISCNTranscribed_TemplateCVs.xlsx)
        + [TemplateCVs.pdf](TemplateCVs.pdf)
        + [TemplateSubmit.pdf](TemplateSubmit.pdf)

# Data Rescue working notes 

## Data Rescue Plan

+ Lead
  - [ ] Read the article or data resource and the discription of the target project
  - [ ] Create a fresh fork of the primary repository https://github.com/ktoddbrown/SoilDRaH/tree/main, it might be useful to rename it 'SoilDRaH_AuthorYYYY'
  - [ ] Create a copy of this template folder and rename it to AuthorYYYY
  - [ ] Fill out Data summary section (except for the Files), the Data Rescue Plan (here!), and update the author section of the yml to reflect who is working on the rescue
  - [ ] Create a GitHub issue from the Data Rescue template and get feedback from a reviewer on the project (Dr Todd-Brown or a grad student typically).
  - [ ] Add the reviewer to the author list in the yml
+ Do
  - [ ] Create a spreadsheet with TableN and export it to csv
  - [ ] Extract points from FigureM and export it to csv
  - [ ] Transcribe or copy the method section
  - [ ] Insert references to bib entries
  - [ ] reformat equations and mathematical symbols as [LaTex](https://en.wikibooks.org/wiki/LaTeX/Mathematics)
  - [ ] reformat text using [Markdown](https://www.markdownguide.org/cheat-sheet/)
  - [ ] Create a bib file with the citations from the methods section
  - [ ] Push to your local fork and get a second transcription pushed to your local fork by someone else on the project
+ Measure
  - [ ] cross check with second transcriber for tables and methods, and resolve any differences
  - [ ] compile or render this ReadMe with all chunks up through the Level 0 data rescue section with `eval: true`
  - [ ] push to your local fork and create a pull request to the main branch of the primary repository
  - [ ] address any revisions in your pull request review
  - [ ] celebrate!

## Download ISCN3

The ISCN3 is downloaded from EDI (@ISCN3). Below we loop through the files that are expected and download them.

```{r}
#| label: dataDownloadEDI
#| code-summary: "Download and identify each file in the EDI package."

### Download the data ####
# link the table name to the download url and file path(s) for data package
pasta_package <- 'https://pasta.lternet.edu/package/data/eml/edi/1160/1/'

download_table <- tibble::tribble(
  ~table, ~entity_id, ~file_name, 
  'layer', "4af719a84f8981fcc63f1f92760cb253", 'ISCN3_layer.csv',
  'profile', "40527580cc045d33d9a5aaf728bf204e", 'ISCN3_profile.csv',
  'citation', "320e31ca911f187550ca2143c31fd408", 'ISCN3_citation.csv',
  'dataset', "cdd0c7a4cac3f28d6d788c91f506775f", 'ISCN3_dataset.csv',
  'control_vocabulary', "a8eef6e94b669b365e443c15d9402a03", 'ISCNTranscribed_TemplateCVs.xlsx',
  'template', "cb5c1cd86e9ab17f699f7a05044325b7", 'ISCNtemplate.xlsx',
  'other_pdfs', "eb320ae7b57296765f543cbb370b0f24", 'TemplateCVs.pdf',
  'other_pdfs', "b81a7f4214d176280a5ff4e0f0d52d8b", 'TemplateSubmit.pdf')

#One by one, download the data in each row from the download_table
#... and store that in the appropriate file path location on your computer
for(row_index in 1:nrow(download_table)){
  if(!file.exists(file.path(dataDownload.dir,download_table$file_name[row_index]))){
    print(paste('Download file:', 
                download_table$file_name[row_index]))
    utils::download.file(
      paste0(pasta_package, download_table$entity_id[row_index]), 
      file.path(dataDownload.dir,download_table$file_name[row_index]), 
      extra  = "--max-time 300")
  }
}

```

## Constructing Annotations

The metadata along with control vocabulary for the data variables are provided on EDI (@ISCN3). From the [ISCNTranscribed_TemplateCVs.xlsx](ISCNTranscribed_TemplateCVs.xlsx) file, we will extract the annotations for our profile, layer, citation tables. For dataset table, there are no additional annotations available on the edi portal.

```{r}
#| label: constructAnnotations
#| code-summary: "Identify and extract each annotation csv in the EDI package"

# read in the csv here for the different annotation table
ISCN_baseAnnotations <- read_csv(file = ISCN3_layerAnnotations,
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) |> 
  mutate(table_id = 'layer') |> 
  bind_rows(read_csv(file = ISCN3_profileAnnotations, 
                     col_types = cols(.default = col_character())) |> 
              mutate(table_id = 'profile')) |> 
  bind_rows(read_csv(file = ISCN3_siteAnnotations,
                     col_types = cols(.default = col_character())) |> 
              mutate(table_id = 'site')) |> 
  bind_rows(read_csv(file = ISCN3_citationAnnotations, 
                     col_types = cols(.default = col_character())) |> 
              mutate(table_id = 'citation')) |> 
  pivot_longer(cols = c(UNITS, DESCRIPTION),
               names_to = "source",
               values_to = "with_entry",
               values_drop_na = T) |> 
  mutate(is_type = case_when(
    source == 'UNITS' & with_entry == 'CV' ~ 'control vocabulary',
    source == 'UNITS' ~ 'units',
    source == 'DESCRIPTION' ~ 'description')) |> 
  select(-source) |> 
  select(table_id, 
         column_id = VARIABLE,
         of_variable = NAME,
         is_type,
         with_entry) 

ISCN_methodAnnotations <- ISCN_baseAnnotations |> 
  dplyr::filter(stringr::str_detect(of_variable,
                                    regex('\\bmethod\\b', 
                                          ignore_case = T))) |> 
  dplyr::transmute(column_id, 
            of_variable, 
            table_id,
            is_type = 'method',
            with_entry = of_variable)

ISCN_dataAnnotations <- bind_rows(ISCN_baseAnnotations, ISCN_methodAnnotations) |> 
  mutate(is_type = factor(is_type, 
                          levels = c('description', 'units',
                                     'method', 'control vocabulary'))) |> 
 # arrange(table_id, VARIABLE, is_type) |> 
  unique() # 337 obs of 5 vars

#write_csv(ISCN_dataAnnotations, "ISCN_dataAnnotations2026.csv")

```

# Level 0 data

Below is an example of how to use the `readISCN3` function and basic reporting on the soil organic carbon related measurements.

```{r}
#| label: pullDatalevel0
#| code-summary: "pull level 0 data using readISCN3 function"

iscn.ls <- readISCN3(dataDir = dataDownload.dir,
                     annotationFilename = dataAnnotations,
                     controlVocabularyFile = controlVocabulary,
                     primaryBibFile = primaryCitation.file,
                     contributedBibfile = contributionsCitation.file,
                     format = 'level0')

```

## Primary data: layer table

```{r}
knitr::kable(iscn.ls$level0_data$layer |> 
               head()) |>
  kable_paper() |>
  scroll_box(width = "100%", height = "300px")

```

## Table: `citation`

```{r}
knitr::kable(iscn.ls$annotation |>
               dplyr::filter(is_type == 'description',
                      table_id == 'citation') |>
               dplyr::select(column_id, of_variable, description = with_entry)) |>
  kable_paper() |>
  scroll_box(width = "100%", height = "300px")
```

## Table: `site`

```{r}
knitr::kable(iscn.ls$annotation |>
               dplyr::filter(is_type == 'description',
                      table_id == 'site') |>
               dplyr::select(column_id, of_variable, description = with_entry)) |>
  kable_paper() |>
  scroll_box(width = "100%", height = "300px")
```

## Table: `profile`

```{r}
knitr::kable(iscn.ls$annotation |>
               dplyr::filter(is_type == 'description',
                      table_id == 'profile') |>
               dplyr::select(column_id, of_variable, description = with_entry)) |>
  kable_paper() |>
  scroll_box(width = "100%", height = "300px")
```

## Table: `layer`

```{r}
knitr::kable(iscn.ls$annotation |>
               dplyr::filter(is_type == 'description',
                      table_id == 'layer') |>
               dplyr::select(column_id, of_variable, description = with_entry)) |>
  kable_paper() |>
  scroll_box(width = "100%", height = "300px")
```
