---
title: "Data Rescue: Crow 2016"
format:
  html: https://doi.org/10.1080/17583004.2016.1213140
    toc: true
date: 2026-04-16
date-format: YYYY-MM-DD
bibliography:
  - Crow2016_Methods.bib #contains citations in methods section
  - Crow2016.bib #contains citations for data source
author:
  - name:
      given: Katherine
      family: Todd-Brown
      orcid: 0000-0002-3109-8130
      note: https://github.com/ktoddbrown
      affiliations:
        - id: ESS UF
          name: University of Florida
          department: Environmental Engineering Sciences
          city: Gainesville
          state: FL
          country: USA
      role: 
        - "Transcription"
        - "Review"
   - id: waldera
    name:
      given: Ashlynn
      family: Walder
    affiliation:
      - uf-ise
    role:
      - transcription: lead
      #- standardization:
      #- curation:
  - id: savascott
    name:
      given: Savannah
      family: Scott
    affiliation:
      - uf-ees
    role: 
      - transcription: secondary
      #- standardization: 
      #- curation: 
affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
  - id: uf-mae
    name: University of Florida
    department: Industrial and Systems Eningeering
    city: Gainesville
    state: Florida
    county: USA
    address: "303 Weil Hall \nP.O. Box 116595"
    postal-code: 32603
    url: https://ise.ufl.edu
    
timelog:
  - activity: 
      description: Revise based on feedback
      who: waldera #github username
      start: 2025-09-30
      end: 2026-4-02
      duration: 20H
---
---
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(kableExtra)
library(bibtex)
```

This documents the data transcription for  .
For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/.

> Crow, S. E., Reeves, M., Turn, S., Taniguchi, S., Schubert, O. S., & Koch, N. (2016). Carbon balance implications of land use change from pasture to managed eucalyptus forest in Hawaii. Carbon Management, 7(3-4), 327–341. https://doi.org/10.1080/17583004.2016.1213140

For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/123

# Data Summary

@Crow2016 evaluates carbon stocks in Hawaiian eucalyptus forests planted on former pasture land. It compares methods for measuring soil carbon to assess the impact of land-use change. The study considers both carbon sequestration and emissions from forest management to estimate the system’s net climate benefit.

# Tables and figures

- Tables
  Table 1: Elevation, climate, soil series, and stand age for paired pasture and eucalyptus plots
  Table 2: Soil carbon stocks by bulk density and equivalent mass methods, biomass carbon, and percent change from pasture
  Table 3: Pasture productivity and carbon input rates
  Table 4: Soil carbon concentration and bulk density by depth with statistical comparisons
  Table 5: CO₂-equivalent emissions for a 7-year eucalyptus production cycle

- Figures
  Figure 1: Location of eucalyptus study plots and forest boundary over soil taxonomic layers
  Figure 2: Life cycle inventory for greenhouse gas impact calculations
  Figure 3: Soil carbon stocks by depth and mass for pasture vs. eucalyptus plots

## Fit for purpose: HiCSC

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the HiCSC.

- Location: Island of Hawaii, Hamakua Coast region; pasture land planted with managed Eucalyptus grandis forest.
- Soil carbon stock: Approximately 533 Mg C ha⁻¹ in the top 1 m.
- Soil type: Volcanic ash–derived soils with very strong organic matter and clay mineral interaction.
- Land use: Conversion of former pasture to monoculture eucalyptus plantation.

# Files

These files are in the AuthorYYYY data rescue.

- [Readme](Crow2016_README.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Crow2016.bib](Crow2016.bib)
  + citation for article transcribed
- [Crow2016_Methods.bib](Crow2016_Methods.bib)
  + citations for the methods section of the article
- [Crow2016_Methods.md](Crow2016_Methods.md)
  + methods transcribed from primary article
- [Crow2016_Table1.csv](Crow2016_Table1.csv)
  + table 1 from primary article with 
- [Crow2016_Table2.csv](Crow2016_Table2.csv)
  + table 1 from primary article with 
- [temp/](temp/)
- [Crow2016_Table4.csv](Crow2016_Table4.csv)
  + table 1 from primary article with 
- [temp/](temp/)
- [Crow2016_Figure3.csv](Crow2016_Figure3.csv)
  + table 1 from primary article with 
- [temp/](temp/)
    + scratch folder that will not be archived includes excel and article pdf document

# Working notes 

## Data Rescue Plan

- [ ] describe files in human readable form inside ReadMe
- [ ] description with Git issue with link
- [ ] add contribution yaml
- [ ] create excel sheet with Tables 1-5
- [ ] export excel table to csv
- [ ] copy over methods section into markdown
- [ ] pull down citations in methods section to bib file
- [ ] add in citation links to methods
- [ ] cross check with second transcriber for tables and methods
- [ ] submit to independent review
- [ ] archive on SoilDRaH

### Table N

Table N was modified ....

```{r table1}
#| code-fold: true
#| message: false

filename <- 'Table.csv'
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```


{{< include AuthorYYYY_Methods.Rmd >}}

# Transcription comparison

```{bash eval=FALSE}
diff --strip-trailing-cr --suppress-common-lines -y temp/AuthorYYYY_Methods_ABC.md AuthorYYYY_Methods_BCD.Rmd
```


## Table 1

<!--- Common table modifications include: inserting an extra column to pivot sub-headers, filling in NA to identify blanks, and adding Latex or Markdown to denote formatting.---->

Table 1 was modified ....

```{r}
#| label: table1Check
#| eval: false #remove when you have the file transcribed


read_csv(file = "Crow2016_Table1.csv", #read in the specific file
         skip = 1, #skip the first line with the caption
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = "Crow2016_Table1.csv", #read in that file again but...
                           n_max = 1, #only read the first line
                           col_names = 'caption', #call it the caption
                           #read everything as a character
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE #do not return messages
                           )$caption #and take it out of the table
  )
```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareTableN
#| eval: false

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y Crow2016_Table1_AW.csv Crow2016_Table1_SRS.csv

```

## Table 2

<!--- Common table modifications include: inserting an extra column to pivot sub-headers, filling in NA to identify blanks, and adding Latex or Markdown to denote formatting.---->

Table 2 was modified ....

```{r}
#| label: table2Check
#| eval: false #remove when you have the file transcribed


read_csv(file = "Crow2016_Table2.csv", #read in the specific file
         skip = 1, #skip the first line with the caption
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = "Crow2016_Table2.csv", #read in that file again but...
                           n_max = 1, #only read the first line
                           col_names = 'caption', #call it the caption
                           #read everything as a character
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE #do not return messages
                           )$caption #and take it out of the table
  )
```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareTableN
#| eval: false

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y Crow2016_Table2_AW.csv Crow2016_Table2_SRS.csv



```

## Table 4

<!--- Common table modifications include: inserting an extra column to pivot sub-headers, filling in NA to identify blanks, and adding Latex or Markdown to denote formatting.---->

Table 4 was modified ....

```{r}
#| label: table4Check
#| eval: false #remove when you have the file transcribed


read_csv(file = "Crow2016_Table4.csv", #read in the specific file
         skip = 1, #skip the first line with the caption
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = "Crow2016_Table4.csv", #read in that file again but...
                           n_max = 1, #only read the first line
                           col_names = 'caption', #call it the caption
                           #read everything as a character
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE #do not return messages
                           )$caption #and take it out of the table
  )
```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareTableN
#| eval: false

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y Crow2016_Table4_AW.csv Crow2016_Table4_SRS.csv

```

## Figure 3

<!--- Common table modifications include: inserting an extra column to pivot sub-headers, filling in NA to identify blanks, and adding Latex or Markdown to denote formatting.---->
Figure 3 was modified ....

```{r}
#| label: Figure3Check
#| eval: false #remove when you have the file transcribed


read_csv(file = "Crow2016_Figure3.csv", #read in the specific file
         skip = 1, #skip the first line with the caption
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = "Crow2016_Figure3.csv", #read in that file again but...
                           n_max = 1, #only read the first line
                           col_names = 'caption', #call it the caption
                           #read everything as a character
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE #do not return messages
                           )$caption #and take it out of the table
  )
```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareTableN
#| eval: false

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y Crow2016_Figure3_AW.csv Crow2016_Figure3_SRS.csv

```

## Methods

Any different to the cross check for the methods section will be output after this line.

```{bash}
#| label: compareMethods
#| eval: false
# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y Crow2016_Methods_AW.md  Crow2016_Methods_SRS.md


```


```

### Methods  - Text

The final text transcription from the Methods section is rendered below.

```{r}
#| label: readMethods
#| output: asis
#| eval: true

# This revised section will append two ### to any existing headers here so that it renders nicely with the other headers in the document.

methodsLines <- readLines("Crow2016_Methods.md") |>
  str_replace_all(pattern = '^#', replacement = '####')

cat(paste0(methodsLines, collapse = '\n'))

```

<!----
## Citation notes from Methods

Below are the citation notes from the methods section.
This generally doesn't read well in the final report so commenting it out here but it is very useful for tracking citations when you are doing the transcriptions.
Some citations are missing, those are noted here and how best guess were generated.
Think of this as a lab notebook documenting how you found each citation.

- [x] Anon. (2008). SimaPro Craft Version 7.1.8. PRé Consultants. http://www.pre-sustainability.com/simapro-lca-software  
  + copy from bib

- [x] Anon. (2011). U.S. Life Cycle Inventory Database. National Renewable Energy Laboratory. http://www.nrel.gov/lci/  
  + copy from bib

- [x] Bare, J. (2011). TRACI 2.0: The tool for the reduction and assessment of chemical and other environmental impacts 2.0. Clean Technologies and Environmental Policy, 13, 687–696. https://doi.org/10.1007/s10098-010-0338-9  
  + copy from bib

- [x] Blackmore, M., & Vitousek, P. M. (2000). Cattle grazing, forest loss, and fuel loading in a dry forest ecosystem at Pu’u Wa’aWa’a Ranch, Hawai‘i. Biotropica, 32, 625–632. https://doi.org/10.1111/j.1744-7429.2000.tb00509.x  
  + copy from bib

- [x] Chapin, F. S., Bloom, A. J., Field, C. B., & Waring, R. H. (1987). Plant responses to multiple environmental factors. BioScience, 37, 49–57. https://doi.org/10.2307/1310177  
  + copy from bib

- [x] Ecoinvent. (2007). Ecoinvent Data v2.0 Ecoinvent Reports No. 1–25.  
  + copy from bib

- [x] Ellsworth, L. M., Litton, C. M., Taylor, A. D., & Kauffman, J. B. (2013). Spatial and temporal variability of Guinea grass (Megathyrsus maximus) fuel loads and moisture on Oahu, Hawaii. International Journal of Wildland Fire, 22, 1083–1092. https://doi.org/10.1071/WF12051  
  + copy from bib

- [x] Giambelluca, T. W., Chen, Q., Frazier, A. G., et al. (2013). Online rainfall atlas of Hawai‘i. Bulletin of the American Meteorological Society, 94, 313–316. https://doi.org/10.1175/BAMS-D-11-00228.1  
  + copy from bib

- [x] Giardina, C. P., & Ryan, M. G. (2002). Total belowground carbon allocation in a fast-growing Eucalyptus plantation estimated using a carbon balance approach. Ecosystems, 5, 487–499. https://doi.org/10.1007/s10021-002-0130-8  
  + copy from bib

- [x] Gifford, R. M., & Roderick, M. L. (2003). Soil carbon stocks and bulk density: Spatial or cumulative mass coordinates as a basis of expression. Global Change Biology, 9, 1507–1514. https://doi.org/10.1046/j.1365-2486.2003.00677.x  
  + copy from bib

- [x] Houghton, J. T., Ding, Y., Griggs, D. J., et al. (2001). Climate change 2001: The scientific basis. Cambridge University Press.  
  + copy from bib

- [x] Misra, R. K., Turnbull, C. R. A., Cromer, R. N., Gibbons, A. K., & LaSala, A. V. (1998). Below- and above-ground growth of Eucalyptus nitens in a young plantation. Forest Ecology and Management, 106, 283–293. https://doi.org/10.1016/S0378-1127(97)00339-3  
  + copy from bib

- [x] Motooka, P. S., Nelson, D., Nagai, G., & Ching, L. (2003). Weeds of Hawaiʻi’s pastures and natural areas: An identification and management guide. College of Tropical Agriculture. https://www.ctahr.hawaii.edu/invweed/weedshi.html  
  + copy from bib

- [x] Reeves, M. I. (2012). The potential carbon sequestration of Eucalyptus grandis in conjunction with its use as a biofuel feedstock. University of Hawaii at Manoa. https://scholarspace.manoa.hawaii.edu/server/api/core/bitstreams/c0d4b15a-e96e-4122-9c37-c4627aa771b93/content  
  + copy from bib

- [x] Saint-André, L., Thongo M’Bou, A., Mabiala, A., et al. (2005). Age-related equations for above- and below-ground biomass of a Eucalyptus hybrid in Congo. Forest Ecology and Management, 205, 199–214. https://doi.org/10.1016/j.foreco.2004.10.006  
  + copy from bib

- [x] Schukrt, T. H., Strand, R. F., Cole, T. G., & McDuffie, K. E. (1988). Equations for predicting biomass of six introduced tree species, island of Hawaii. U.S. Department of Agriculture Forest Service. https://doi.org/10.2737/PSW-RN-401  
  + copy from bib

- [x] Thorne, M. S., & Hewlett, J. P. (2013). Hawaii forage production estimator tool. https://globalrangelands.org/hawaii/hawaii-forage-production-estimator-tool  
  + copy from bib

- [x] Turn, S., Keffer, V., Dalla, S., Evans, D., & Wang, W. (2005). Preliminary gasification testing of Eucalyptus fuels. http://hawaii.gov/dbedt/info/energy/publications/biomass-gasification-2005.pdf  
  + copy from bib

- [x] Turn, S., Smith, L., Koch, N., & Taniguchi, S. (2014). Energy analysis of Eucalyptus grandis production in Hawaii. Journal of Renewable and Sustainable Energy, 6, 043104. https://doi.org/10.1063/1.4885119  
  + copy from bib

- [x] Wendt, J. W., & Hauser, S. (2013). An equivalent soil mass procedure for monitoring soil organic carbon in multiple soil layers. European Journal of Soil Science, 64, 58–65. https://doi.org/10.1111/ejss.12002  
  + copy from bib

# Level 0 Data rescue

This level 0 data read parses the transcribed files with minimal cleaning or transformations.

```{r}
#| label: readLevel0
#This chunk has two purposes. 
#...1) Check the formatting by reading in everything 
#...2) Create a list of everything to process later in level 1

data.lvl0.ls <- list(
  #Read in a list of all the bib files
  citation = list(
    #Citation for the article transcriptions are pulled from
    primary = read.bib(file = "Crow2016.bib"), 
    #Citations for all referenced articles
    methods = read.bib(file = "Crow2016_Methods.bib")
  ),
  #Read in the text transcription of the article's methods section
  method = read_lines(file = "Crow2016_Methods."),
  #Read in the results as tables or figure transcriptions. This includes
  #...the caption as well as the tables themselves
  data = list(
    Table1 = list(
      #Read the caption as a text string. Captions are the first cell on 
      #...the first row.
      caption = read_csv(file = "Crow2016_Table1.csv",
                       col_types = cols(.default = col_character()),
                       n_max = 1, col_names = FALSE)$X1[1],

    #Read in all the data, skipping the first row with the caption and read
    #...in the table as character. This element is a tibble (data.frame).
    primary = read_csv(file = "Crow2016_Table1.csv",
                       col_types = cols(.default = col_character()),
                       skip = 1)
  ), #End 'Table1' element

    #Repeat for all tables and figures...
   ) #end 'data' element
)

```

The code is also in `readsource.file` and returned with the `dataLevel = 'level0'` argument.
Below verifies that the data from the function matches what is coded above.

```{r}
#| label: checkReadFunction_lvl0

if(file.exists(readsource.file)){
  datafunction.lvl0.ls <- readAuthorYYYY(dataDir = '.',
                                         dataLevel = 'level0')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function matches code here.')
  }else{
    print('There is a mismatch in the data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}
```

# Level 1 data - CURATION PROJECT

This data is identified as a data source for the ...
Broad classes of information that we are looking for include:

<!--- Include description of where the data is located in the source.--->
- Location: 
- Soil carbon stock: 
- Soil type: 
- Land use: 

Information from the methods and other sections are pulled into tables with line by line source citations.
This information can be values, units, methods, or other useful notes and is manually extracted from the transcriptions loaded into the level 0 data above.

The final format of the data frame here will include the following 

  + grouping ids
  + `of_variable`
  + `is_type`
  + `with_entry`
  + `from_source`

The `of_variable` being used here are:

 + citation
 
 <!--- include other variables being used, this should match vocabulary in the SoilDRaHVocabulary resources --->

```{r}
#| label: createLvl1
#| eval: false

#### Method-section only data ####
#Create a table for the data in the method-section that is not directly 
#...associated with any table. Often this is study level information.

studyMeta <- tribble(~of_variable, ~is_type, ~with_entry, ~from_source,
                   'region', 'value', 'Honaunau Forest on the southwestern slopes of Mauna Loa, island of Hawaii', paste('Method ln5:', paste(data.lvl0.ls$method[5], collapse = ' ')))

#### Table/Figure N ####
# Process the primary data from Table/Figure M by maintaining data groups and 
#...standardizing the variable name so that we can link this to the study 
#...metadata and data in the other tables when needed. Link in the methods 
#...and other information from the method-section as needed.

Table1Primary <- data.lvl0.ls$data$Table1$primary |>
  #Add a row id preserving row grouping -- not unique
  mutate(row_id = paste0('R', 1:n())) |>
  #Transform the table to long data by row group and column names.
  #...This allows us to then create the cross links between the tables.
  pivot_longer(cols = -row_id,
               names_to = 'column_name', values_to = 'with_entry',
               values_drop_na = TRUE) |> #drop missing values
  #Link the variable names to the column names using a switch statement
  mutate(of_variable = case_when(
           #Soil PH is given as an example below
           column_name == 'SOC' ~ 'soil_organic_carbon')) |>
  #Flag the source of this data as Table 1, this table is merged later with
  #...other data so this keeps track of where the data came from
  mutate(from_source = 'Table 1')

## Create the metadata for table one from the header information and cited methods
Table1Meta <- Table1Primary |>
  select(column_name, of_variable) |>
  unique() |>
  #Grab everything between the parentheses as units and attribute the source as the column names.
  mutate(unit = str_extract(column_name, pattern = '(?<=\\().*(?=\\))'),
         from_source = 'Table 1 column names.') |>
  #If there aren't units then drop the row
  filter(!is.na(unit)) |>
  #Don't keep the column name now that you have the units, cross link via the variable
  select(-column_name) |>
  #Stack into the table the methods for each variable that we can find
  bind_rows(
  tribble(~of_variable, ~method, ~from_source,
            'soil_organic_carbon', paste0(data.lvl0.ls$method[72:73], collapse = ' '), 'Methods ln72-73') ) |>
  #Stack onto the tables the controlled vocabulary used
  bind_rows(
    tribble(~of_variable, ~control_vocabulary, ~from_source,
            'soil_class', 'Histosol:USDA classification for histosol soil type|Andisols:USDA classification for andisol soil type', 'expert informed')
  ) |>
  # Push the dimensions of the variable into a single column. Making this a long table.
  pivot_longer(cols = c(unit, method, control_vocabulary),
               names_to = 'is_type',
               values_drop_na = TRUE,
               values_to = 'with_entry')

#### Create level 1 #####
## Pull everything together into by stacking the meta and primary data tables
data.lvl1.ls <- list(
  meta = bind_rows(studyMeta, Table1Meta),
  primary = bind_rows(Table1Primary)|>
    arrange(row_id, column_name,
            of_variable, is_type, with_entry, from_source)
  )

```

## Check for new variables

`of_variable` in the data files should be defined in the SoilDRaH Vocabulary files.

```{r}
#| label: check_of_variable

#Read in the list of current 'of_variable' entries
SoilDRaH_variable_list <- read_csv(file =of_variable.file,
                                   show_col_types = FALSE)

#compare the variables in the level 1 tables to the SoilDRaH variables
new_variables <- lapply(data.lvl1.ls, function(xx){
  setdiff(unique(xx$of_variable), 
          unique(SoilDRaH_variable_list$of_variable))
}) |>
  unlist()

#Check that there are no new variables
if(length(new_variables) == 0){
  print('All of_variables are in the SoilDRaH variable list. No additions required')
}else{
  cat(paste('Please start a issue for new variables [',
               new_variables, ']', sep = '', collapse = '\n'))
}
```

## Check for is_type

`is_type` in the data files should be defined in the SoilDRaH Vocabulary files.

```{r}
#| label: check_is_type

#read in the current types and what variable they are associated with
SoilDRaH_type_list <- read_csv(file = is_type.file,
                                   show_col_types = FALSE) |>
  #remove the project definition and only keep the variable and type
  select(of_variable, is_type) |>
  #variables are pushed together in the type defined table. Separate the
  #...entries here
  separate_longer_delim(cols = 'of_variable', delim = ',')

#Pull all the of_variable-is_type pairs from the source data
source_pairs <- tibble()

#If there are a large number of entries, consider replacing this for-loop.
#...However this avoids having to install the plyr or other libraries, and 
#...there are generally no more then 3-5 tables
for(dataTable in data.lvl1.ls){
  source_pairs <- dataTable |>
    select(of_variable, is_type) |> 
    unique() |>
    bind_rows(source_pairs)
}

undefined_data_type <- source_pairs |>
  #subset for variables that are not 'incomplete'
  semi_join(SoilDRaH_variable_list |>
              filter(project_status != 'incomplete'),
            by = join_by(of_variable)) |>
  #check for of_variable-is_type tuples that are not in the SoilDRaH type definitions
  anti_join(SoilDRaH_type_list,
            by = join_by(of_variable, is_type))

#Tell the user what you found.
if(nrow(undefined_data_type) == 0){
  print('All of_variables-is_types are in the SoilDRaH type list. No additions required')
}else{
  print(paste0('Please start a issue for new of_variable-is_type [', paste(paste(undefined_data_type$of_variable, undefined_data_type$is_type, sep = '-'), collapse = ','), ']'))
}
```

## Check function match

Check for an exact match with the read function.

```{r}
#| label: checkReadFunction_lvl1

if(file.exists(readsource.file)){
  datafunction.lvl1.ls <- readAuthorYYYY(dataDir = '.',
                                         dataLevel = 'level1')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function output matches ReadMe code for level 1.')
  }else{
    print('There is a mismatch in the level 1 data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}

```


# References

<!---- This space will be filled with the citations when rendered. Leave this blank. ---->
      - uf-ise
    role:
      - transcription: lead
      #- standardization:
      #- curation:
  - id: savascott
    name:
      given: Savannah
      family: Scott
    affiliation:
      - uf-ees
    role: 
      - transcription: secondary
      #- standardization: 
      #- curation: 
affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
  - id: uf-mae
    name: University of Florida
    department: Industrial and Systems Eningeering
    city: Gainesville
    state: Florida
    county: USA
    address: "303 Weil Hall \nP.O. Box 116595"
    postal-code: 32603
    url: https://ise.ufl.edu
    
timelog:
  - activity: 
      description: Revise based on feedback
      who: waldera #github username
      start: 2025-09-30
      end: 2026-4-02
      duration: 20H
---
---
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(kableExtra)
library(bibtex)
```

This documents the data transcription for  .
For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/.

> Crow, S. E., Reeves, M., Turn, S., Taniguchi, S., Schubert, O. S., & Koch, N. (2016). Carbon balance implications of land use change from pasture to managed eucalyptus forest in Hawaii. Carbon Management, 7(3-4), 327–341. https://doi.org/10.1080/17583004.2016.1213140

For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/123

# Data Summary

@Crow2016 evaluates carbon stocks in Hawaiian eucalyptus forests planted on former pasture land. It compares methods for measuring soil carbon to assess the impact of land-use change. The study considers both carbon sequestration and emissions from forest management to estimate the system’s net climate benefit.

# Tables and figures

- Tables
  Table 1: Elevation, climate, soil series, and stand age for paired pasture and eucalyptus plots
  Table 2: Soil carbon stocks by bulk density and equivalent mass methods, biomass carbon, and percent change from pasture
  Table 3: Pasture productivity and carbon input rates
  Table 4: Soil carbon concentration and bulk density by depth with statistical comparisons
  Table 5: CO₂-equivalent emissions for a 7-year eucalyptus production cycle

- Figures
  Figure 1: Location of eucalyptus study plots and forest boundary over soil taxonomic layers
  Figure 2: Life cycle inventory for greenhouse gas impact calculations
  Figure 3: Soil carbon stocks by depth and mass for pasture vs. eucalyptus plots

## Fit for purpose: HiCSC

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the HiCSC.

- Location: Island of Hawaii, Hamakua Coast region; pasture land planted with managed Eucalyptus grandis forest.
- Soil carbon stock: Approximately 533 Mg C ha⁻¹ in the top 1 m.
- Soil type: Volcanic ash–derived soils with very strong organic matter and clay mineral interaction.
- Land use: Conversion of former pasture to monoculture eucalyptus plantation.

# Files

These files are in the AuthorYYYY data rescue.

- [Readme](Crow2016_README.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Crow2016.bib](Crow2016.bib)
  + citation for article transcribed
- [Crow2016_Methods.bib](Crow2016_Methods.bib)
  + citations for the methods section of the article
- [Crow2016_Methods.md](Crow2016_Methods.md)
  + methods transcribed from primary article
- [Crow2016_Table1.csv](Crow2016_Table1.csv)
  + table 1 from primary article with 
- [Crow2016_Table2.csv](Crow2016_Table2.csv)
  + table 1 from primary article with 
- [temp/](temp/)
- [Crow2016_Table4.csv](Crow2016_Table4.csv)
  + table 1 from primary article with 
- [temp/](temp/)
- [Crow2016_Figure3.csv](Crow2016_Figure3.csv)
  + table 1 from primary article with 
- [temp/](temp/)
    + scratch folder that will not be archived includes excel and article pdf document

# Working notes 

## Data Rescue Plan

- [ ] describe files in human readable form inside ReadMe
- [ ] description with Git issue with link
- [ ] add contribution yaml
- [ ] create excel sheet with Tables 1-5
- [ ] export excel table to csv
- [ ] copy over methods section into markdown
- [ ] pull down citations in methods section to bib file
- [ ] add in citation links to methods
- [ ] cross check with second transcriber for tables and methods
- [ ] submit to independent review
- [ ] archive on SoilDRaH

### Table N

Table N was modified ....

```{r table1}
#| code-fold: true
#| message: false

filename <- 'Table.csv'
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```


{{< include AuthorYYYY_Methods.Rmd >}}

# Transcription comparison

```{bash eval=FALSE}
diff --strip-trailing-cr --suppress-common-lines -y temp/AuthorYYYY_Methods_ABC.md AuthorYYYY_Methods_BCD.Rmd
```


## Table 1

<!--- Common table modifications include: inserting an extra column to pivot sub-headers, filling in NA to identify blanks, and adding Latex or Markdown to denote formatting.---->

Table 1 was modified ....

```{r}
#| label: table1Check
#| eval: false #remove when you have the file transcribed


read_csv(file = TableN.file, #read in the specific file
         skip = 1, #skip the first line with the caption
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = tableN.file, #read in that file again but...
                           n_max = 1, #only read the first line
                           col_names = 'caption', #call it the caption
                           #read everything as a character
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE #do not return messages
                           )$caption #and take it out of the table
  )
```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareTableN
#| eval: false

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y Crow2016_Table1_AW.csv Crow2016_Table1_SRS.csv

```

## Table 2

<!--- Common table modifications include: inserting an extra column to pivot sub-headers, filling in NA to identify blanks, and adding Latex or Markdown to denote formatting.---->

Table 2 was modified ....

```{r}
#| label: table2Check
#| eval: false #remove when you have the file transcribed


read_csv(file = TableN.file, #read in the specific file
         skip = 1, #skip the first line with the caption
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = tableN.file, #read in that file again but...
                           n_max = 1, #only read the first line
                           col_names = 'caption', #call it the caption
                           #read everything as a character
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE #do not return messages
                           )$caption #and take it out of the table
  )
```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareTableN
#| eval: false

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y Crow2016_Table2_AW.csv Crow2016_Table2_SRS.csv

```

## Table 4

<!--- Common table modifications include: inserting an extra column to pivot sub-headers, filling in NA to identify blanks, and adding Latex or Markdown to denote formatting.---->

Table 4 was modified ....

```{r}
#| label: table4Check
#| eval: false #remove when you have the file transcribed


read_csv(file = TableN.file, #read in the specific file
         skip = 1, #skip the first line with the caption
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = tableN.file, #read in that file again but...
                           n_max = 1, #only read the first line
                           col_names = 'caption', #call it the caption
                           #read everything as a character
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE #do not return messages
                           )$caption #and take it out of the table
  )
```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareTableN
#| eval: false

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y Crow2016_Table4_AW.csv Crow2016_Table4_SRS.csv

```

## Figure 3

<!--- Common table modifications include: inserting an extra column to pivot sub-headers, filling in NA to identify blanks, and adding Latex or Markdown to denote formatting.---->
Figure 3 was modified ....

```{r}
#| label: Figure3Check
#| eval: false #remove when you have the file transcribed


read_csv(file = TableN.file, #read in the specific file
         skip = 1, #skip the first line with the caption
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = tableN.file, #read in that file again but...
                           n_max = 1, #only read the first line
                           col_names = 'caption', #call it the caption
                           #read everything as a character
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE #do not return messages
                           )$caption #and take it out of the table
  )
```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareTableN
#| eval: false

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y Crow2016_Figure3_AW.csv Crow2016_Figure3_SRS.csv

```

## Methods

Any different to the cross check for the methods section will be output after this line.

```{bash}
#| label: compareMethods
#| eval: false
# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y Crow2016_Methods_AW.md  Crow2016_Methods_SRS.md


```

```



## Methods

Any different to the cross check for the methods section will be output after this line.

```{bash}
#| label: compareMethods
#| eval: false
# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y Crow2016_Methods_AW.md  Crow2016_Methods_SRS.md


```

### Methods  - Text

The final text transcription from the Methods section is rendered below.

```{r}
#| label: readMethods
#| output: asis
#| eval: true

# This revised section will append two ### to any existing headers here so that it renders nicely with the other headers in the document.

methodsLines <- readLines(methods.file) |>
  str_replace_all(pattern = '^#', replacement = '####')

cat(paste0(methodsLines, collapse = '\n'))

```

<!----
## Citation notes from Methods

Below are the citation notes from the methods section.
This generally doesn't read well in the final report so commenting it out here but it is very useful for tracking citations when you are doing the transcriptions.
Some citations are missing, those are noted here and how best guess were generated.
Think of this as a lab notebook documenting how you found each citation.

- [x] Anon. (2008). SimaPro Craft Version 7.1.8. PRé Consultants. http://www.pre-sustainability.com/simapro-lca-software  
  + copy from bib

- [x] Anon. (2011). U.S. Life Cycle Inventory Database. National Renewable Energy Laboratory. http://www.nrel.gov/lci/  
  + copy from bib

- [x] Bare, J. (2011). TRACI 2.0: The tool for the reduction and assessment of chemical and other environmental impacts 2.0. Clean Technologies and Environmental Policy, 13, 687–696. https://doi.org/10.1007/s10098-010-0338-9  
  + copy from bib

- [x] Blackmore, M., & Vitousek, P. M. (2000). Cattle grazing, forest loss, and fuel loading in a dry forest ecosystem at Pu’u Wa’aWa’a Ranch, Hawai‘i. Biotropica, 32, 625–632. https://doi.org/10.1111/j.1744-7429.2000.tb00509.x  
  + copy from bib

- [x] Chapin, F. S., Bloom, A. J., Field, C. B., & Waring, R. H. (1987). Plant responses to multiple environmental factors. BioScience, 37, 49–57. https://doi.org/10.2307/1310177  
  + copy from bib

- [x] Ecoinvent. (2007). Ecoinvent Data v2.0 Ecoinvent Reports No. 1–25.  
  + copy from bib

- [x] Ellsworth, L. M., Litton, C. M., Taylor, A. D., & Kauffman, J. B. (2013). Spatial and temporal variability of Guinea grass (Megathyrsus maximus) fuel loads and moisture on Oahu, Hawaii. International Journal of Wildland Fire, 22, 1083–1092. https://doi.org/10.1071/WF12051  
  + copy from bib

- [x] Giambelluca, T. W., Chen, Q., Frazier, A. G., et al. (2013). Online rainfall atlas of Hawai‘i. Bulletin of the American Meteorological Society, 94, 313–316. https://doi.org/10.1175/BAMS-D-11-00228.1  
  + copy from bib

- [x] Giardina, C. P., & Ryan, M. G. (2002). Total belowground carbon allocation in a fast-growing Eucalyptus plantation estimated using a carbon balance approach. Ecosystems, 5, 487–499. https://doi.org/10.1007/s10021-002-0130-8  
  + copy from bib

- [x] Gifford, R. M., & Roderick, M. L. (2003). Soil carbon stocks and bulk density: Spatial or cumulative mass coordinates as a basis of expression. Global Change Biology, 9, 1507–1514. https://doi.org/10.1046/j.1365-2486.2003.00677.x  
  + copy from bib

- [x] Houghton, J. T., Ding, Y., Griggs, D. J., et al. (2001). Climate change 2001: The scientific basis. Cambridge University Press.  
  + copy from bib

- [x] Misra, R. K., Turnbull, C. R. A., Cromer, R. N., Gibbons, A. K., & LaSala, A. V. (1998). Below- and above-ground growth of Eucalyptus nitens in a young plantation. Forest Ecology and Management, 106, 283–293. https://doi.org/10.1016/S0378-1127(97)00339-3  
  + copy from bib

- [x] Motooka, P. S., Nelson, D., Nagai, G., & Ching, L. (2003). Weeds of Hawaiʻi’s pastures and natural areas: An identification and management guide. College of Tropical Agriculture. https://www.ctahr.hawaii.edu/invweed/weedshi.html  
  + copy from bib

- [x] Reeves, M. I. (2012). The potential carbon sequestration of Eucalyptus grandis in conjunction with its use as a biofuel feedstock. University of Hawaii at Manoa. https://scholarspace.manoa.hawaii.edu/server/api/core/bitstreams/c0d4b15a-e96e-4122-9c37-c4627aa771b93/content  
  + copy from bib

- [x] Saint-André, L., Thongo M’Bou, A., Mabiala, A., et al. (2005). Age-related equations for above- and below-ground biomass of a Eucalyptus hybrid in Congo. Forest Ecology and Management, 205, 199–214. https://doi.org/10.1016/j.foreco.2004.10.006  
  + copy from bib

- [x] Schukrt, T. H., Strand, R. F., Cole, T. G., & McDuffie, K. E. (1988). Equations for predicting biomass of six introduced tree species, island of Hawaii. U.S. Department of Agriculture Forest Service. https://doi.org/10.2737/PSW-RN-401  
  + copy from bib

- [x] Thorne, M. S., & Hewlett, J. P. (2013). Hawaii forage production estimator tool. https://globalrangelands.org/hawaii/hawaii-forage-production-estimator-tool  
  + copy from bib

- [x] Turn, S., Keffer, V., Dalla, S., Evans, D., & Wang, W. (2005). Preliminary gasification testing of Eucalyptus fuels. http://hawaii.gov/dbedt/info/energy/publications/biomass-gasification-2005.pdf  
  + copy from bib

- [x] Turn, S., Smith, L., Koch, N., & Taniguchi, S. (2014). Energy analysis of Eucalyptus grandis production in Hawaii. Journal of Renewable and Sustainable Energy, 6, 043104. https://doi.org/10.1063/1.4885119  
  + copy from bib

- [x] Wendt, J. W., & Hauser, S. (2013). An equivalent soil mass procedure for monitoring soil organic carbon in multiple soil layers. European Journal of Soil Science, 64, 58–65. https://doi.org/10.1111/ejss.12002  
  + copy from bib

# Level 0 Data rescue

This level 0 data read parses the transcribed files with minimal cleaning or transformations.

```{r}
#| label: readLevel0
#This chunk has two purposes. 
#...1) Check the formatting by reading in everything 
#...2) Create a list of everything to process later in level 1

data.lvl0.ls <- list(
  #Read in a list of all the bib files
  citation = list(
    #Citation for the article transcriptions are pulled from
    primary = read.bib(file = primaryCitation.file), 
    #Citations for all referenced articles
    methods = read.bib(file = methodsCitation.file)
  ),
  #Read in the text transcription of the article's methods section
  method = read_lines(file = methods.file),
  #Read in the results as tables or figure transcriptions. This includes
  #...the caption as well as the tables themselves
  data = list(
    Table1 = list(
      #Read the caption as a text string. Captions are the first cell on 
      #...the first row.
      caption = read_csv(file = tableN.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      #Read in all the data, skipping the first row with the caption and read
      #...in the table as character. This element is a tibble (data.frame).
      primary = read_csv(file = tableN.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    ) #End 'Table1' element
    #Repeat for all tables and figures...
   ) #end 'data' element
)

```

The code is also in `readsource.file` and returned with the `dataLevel = 'level0'` argument.
Below verifies that the data from the function matches what is coded above.

```{r}
#| label: checkReadFunction_lvl0

if(file.exists(readsource.file)){
  datafunction.lvl0.ls <- readAuthorYYYY(dataDir = '.',
                                         dataLevel = 'level0')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function matches code here.')
  }else{
    print('There is a mismatch in the data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}
```

# Level 1 data - CURATION PROJECT

This data is identified as a data source for the ...
Broad classes of information that we are looking for include:

<!--- Include description of where the data is located in the source.--->
- Location: 
- Soil carbon stock: 
- Soil type: 
- Land use: 

Information from the methods and other sections are pulled into tables with line by line source citations.
This information can be values, units, methods, or other useful notes and is manually extracted from the transcriptions loaded into the level 0 data above.

The final format of the data frame here will include the following 

  + grouping ids
  + `of_variable`
  + `is_type`
  + `with_entry`
  + `from_source`

The `of_variable` being used here are:

 + citation
 
 <!--- include other variables being used, this should match vocabulary in the SoilDRaHVocabulary resources --->

```{r}
#| label: createLvl1
#| eval: false

#### Method-section only data ####
#Create a table for the data in the method-section that is not directly 
#...associated with any table. Often this is study level information.

studyMeta <- tribble(~of_variable, ~is_type, ~with_entry, ~from_source,
                   'region', 'value', 'Honaunau Forest on the southwestern slopes of Mauna Loa, island of Hawaii', paste('Method ln5:', paste(data.lvl0.ls$method[5], collapse = ' ')))

#### Table/Figure N ####
# Process the primary data from Table/Figure M by maintaining data groups and 
#...standardizing the variable name so that we can link this to the study 
#...metadata and data in the other tables when needed. Link in the methods 
#...and other information from the method-section as needed.

Table1Primary <- data.lvl0.ls$data$Table1$primary |>
  #Add a row id preserving row grouping -- not unique
  mutate(row_id = paste0('R', 1:n())) |>
  #Transform the table to long data by row group and column names.
  #...This allows us to then create the cross links between the tables.
  pivot_longer(cols = -row_id,
               names_to = 'column_name', values_to = 'with_entry',
               values_drop_na = TRUE) |> #drop missing values
  #Link the variable names to the column names using a switch statement
  mutate(of_variable = case_when(
           #Soil PH is given as an example below
           column_name == 'SOC' ~ 'soil_organic_carbon')) |>
  #Flag the source of this data as Table 1, this table is merged later with
  #...other data so this keeps track of where the data came from
  mutate(from_source = 'Table 1')

## Create the metadata for table one from the header information and cited methods
Table1Meta <- Table1Primary |>
  select(column_name, of_variable) |>
  unique() |>
  #Grab everything between the parentheses as units and attribute the source as the column names.
  mutate(unit = str_extract(column_name, pattern = '(?<=\\().*(?=\\))'),
         from_source = 'Table 1 column names.') |>
  #If there aren't units then drop the row
  filter(!is.na(unit)) |>
  #Don't keep the column name now that you have the units, cross link via the variable
  select(-column_name) |>
  #Stack into the table the methods for each variable that we can find
  bind_rows(
  tribble(~of_variable, ~method, ~from_source,
            'soil_organic_carbon', paste0(data.lvl0.ls$method[72:73], collapse = ' '), 'Methods ln72-73') ) |>
  #Stack onto the tables the controlled vocabulary used
  bind_rows(
    tribble(~of_variable, ~control_vocabulary, ~from_source,
            'soil_class', 'Histosol:USDA classification for histosol soil type|Andisols:USDA classification for andisol soil type', 'expert informed')
  ) |>
  # Push the dimensions of the variable into a single column. Making this a long table.
  pivot_longer(cols = c(unit, method, control_vocabulary),
               names_to = 'is_type',
               values_drop_na = TRUE,
               values_to = 'with_entry')

#### Create level 1 #####
## Pull everything together into by stacking the meta and primary data tables
data.lvl1.ls <- list(
  meta = bind_rows(studyMeta, Table1Meta),
  primary = bind_rows(Table1Primary)|>
    arrange(row_id, column_name,
            of_variable, is_type, with_entry, from_source)
  )

```

## Check for new variables

`of_variable` in the data files should be defined in the SoilDRaH Vocabulary files.

```{r}
#| label: check_of_variable

#Read in the list of current 'of_variable' entries
SoilDRaH_variable_list <- read_csv(file =of_variable.file,
                                   show_col_types = FALSE)

#compare the variables in the level 1 tables to the SoilDRaH variables
new_variables <- lapply(data.lvl1.ls, function(xx){
  setdiff(unique(xx$of_variable), 
          unique(SoilDRaH_variable_list$of_variable))
}) |>
  unlist()

#Check that there are no new variables
if(length(new_variables) == 0){
  print('All of_variables are in the SoilDRaH variable list. No additions required')
}else{
  cat(paste('Please start a issue for new variables [',
               new_variables, ']', sep = '', collapse = '\n'))
}
```

## Check for is_type

`is_type` in the data files should be defined in the SoilDRaH Vocabulary files.

```{r}
#| label: check_is_type

#read in the current types and what variable they are associated with
SoilDRaH_type_list <- read_csv(file = is_type.file,
                                   show_col_types = FALSE) |>
  #remove the project definition and only keep the variable and type
  select(of_variable, is_type) |>
  #variables are pushed together in the type defined table. Separate the
  #...entries here
  separate_longer_delim(cols = 'of_variable', delim = ',')

#Pull all the of_variable-is_type pairs from the source data
source_pairs <- tibble()

#If there are a large number of entries, consider replacing this for-loop.
#...However this avoids having to install the plyr or other libraries, and 
#...there are generally no more then 3-5 tables
for(dataTable in data.lvl1.ls){
  source_pairs <- dataTable |>
    select(of_variable, is_type) |> 
    unique() |>
    bind_rows(source_pairs)
}

undefined_data_type <- source_pairs |>
  #subset for variables that are not 'incomplete'
  semi_join(SoilDRaH_variable_list |>
              filter(project_status != 'incomplete'),
            by = join_by(of_variable)) |>
  #check for of_variable-is_type tuples that are not in the SoilDRaH type definitions
  anti_join(SoilDRaH_type_list,
            by = join_by(of_variable, is_type))

#Tell the user what you found.
if(nrow(undefined_data_type) == 0){
  print('All of_variables-is_types are in the SoilDRaH type list. No additions required')
}else{
  print(paste0('Please start a issue for new of_variable-is_type [', paste(paste(undefined_data_type$of_variable, undefined_data_type$is_type, sep = '-'), collapse = ','), ']'))
}
```

## Check function match

Check for an exact match with the read function.

```{r}
#| label: checkReadFunction_lvl1

if(file.exists(readsource.file)){
  datafunction.lvl1.ls <- readAuthorYYYY(dataDir = '.',
                                         dataLevel = 'level1')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function output matches ReadMe code for level 1.')
  }else{
    print('There is a mismatch in the level 1 data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}

```


# References

<!---- This space will be filled with the citations when rendered. Leave this blank. ---->

