---
title: "Data Curation: PROJECT YYYY"
format:
  html:
    toc: true
    eval: false #default to not evaluating the chunks, override this the template is completed
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - ProjectYYYY.bib #contains citation that provide background on the project
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  - id: ktoddbrown #use the github name as an id
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role: #TODO define these roles and add reference
      - transcription: lead #secondary, review
      - standardization: lead #review
      - curation: lead #review

affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
timelog:
  - activity: 
      description: Drafted initial template
      who: @ktoddbrown #github username
      #when you started and ended, can be just a year or down to the minute
      end: 2025-11-25T18:45
      #start: YYYY-MM-DD THH:MM
      #alternatively duration can be combined with a being/end
      duration: P1H15M
version: 'vs202512' #template version, remove before use
---


```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false


library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# Bibliograph files
refernceCitation.file <- 'ProjectYYYY.bib'

#semantic files for level 1
of_variable.file <- '../../semantics/SoilDRaHVocabulary_of_variable.csv'
is_type.file <- '../../semantics/SoilDRaHVocabulary_is_type.csv'

```

# Purpose

The purpose PROJECTYYYY is to create a data product aggregating data on ...
With this data we will ...
If you are interested in contributing to this project please contact .... OR We have closed contributions for this project, if you would like to restart or extend the current dataset please contact ....

# Variables of interest

<!--- Briefly describe what you want the data rescuers to look for when digitizing a data resource.---->

For this analysis we are interested in:

  * **Location**: latitude/longitude, associated datum, geo-spatial method, precision
  * **Soil organic carbon stocks**: organic carbon fraction, bulk density, coarse fraction, soil organic carbon by area, soil organic carbon by volume
  * **Layer depth**: Depth of surveyed soil layer

If available we will also pull in:

  * **Land history**: land management , duration of practice, definition of land management
  * **Mean annual temperature**: mean annual temperature, unit of temperature
  
# Data sources

For this project, we have identified the following data sources by ...

<!---- Data stage corresponds to: Identified, In progress, level 0: digitized data rescue, level 1: harmonized into standard data tuple, level 2: integrated into a project ---->

| AuthorYYYY | citation | GitHub Issue | Stage |
|--------|--|------------|-----|
| Ares2001 | @Ares2001 | [77](https://github.com/ktoddbrown/SoilDRaH/issues/77) | 2 |
<!---- remove and replace the example above ---->

# Data Level 1 

```{r}
#Source and append all the level 1 tuples

```

# Data Level 2

```{r}
#Map the comparable methods and align any needed vocabulary
```

```{r}
#Data viz checks
```

# References

<!---- This will be populated when you render the document.---->
