---
title: "Data Rescue: Author YYYY"
format:
  html:
    toc: true
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - AuthorYYYY_Methods_KTB.bib
  - AuthorYYYY.bib
author:
  - name:
      given: Katherine
      family: Todd-Brown
      orcid: 0000-0002-3109-8130
      note: https://github.com/ktoddbrown
      affiliations:
        - id: ESS UF
          name: University of Florida
          department: Environmental Engineering Sciences
          city: Gainesville
          state: FL
          country: USA
      role: 
        - "Transcription"
        - "Review"
  - name:
      given: Jane
      family: Gator
      affiliation:
        - id: UF
          name: University of Florida
          city: Gainesville
          state: FL
          country: USA
      role: "Transcription"
      note: https://github.com/
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(kableExtra)
```

This documents the data transcription for  .
For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/.

> Full citation here

# Data Summary

AuthorYYYY primarily measures ... They use this data to look at ...

## Fit for purpose: HiCSC

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the HiCSC.

- Location: 
- Soil carbon stock: 
- Soil type: 
- Land use: 

# Tables and figures

- Table 
  1) 

- Figure 
  1) 
  
### Table N

Table N was modified ....

```{r table1}
#| code-fold: true
#| message: false

filename <- 'Table.csv'
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```


{{< include AuthorYYYY_Methods.Rmd >}}

# Transcription comparison

```{bash eval=FALSE}
diff --strip-trailing-cr --suppress-common-lines -y temp/AuthorYYYY_Methods_ABC.md AuthorYYYY_Methods_BCD.Rmd
```

# Files

These files are in the AuthorYYYY data rescue.

- [Readme](README_KTB.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [AuthorYYYY.bib](AuthorYYYY.bib)
  + citation for article transcribed
- [AuthorYYYY_Methods.bib](AuthorYYYY_Methods.bib)
  + citations for the methods section of the article
- [AuthorYYYY_Methods.Rmd](AuthorYYYY_Methods.Rmd)
  + methods transcribed from primary article
- [AuthorYYYY_TableN.csv](AuthorYYYY_TableN.csv)
  + table N from primary article with ...
- [temp/](temp/)
    + scratch folder that will not be archived includes excel and article pdf document

# Working notes 

## Data Rescue Plan

- [ ] describe files in human readable form inside ReadMe
- [ ] description with Git issue with link
- [ ] add contribution yaml
- [ ] create excel sheet with Table 1 and 3
- [ ] export excel table to csv
- [ ] copy over methods section into markdown
- [ ] pull down citations in methods section to bib file
- [ ] add in citation links to methods
- [ ] cross check with second transcriber for tables and methods
- [ ] submit to independent review
- [ ] archive on SoilDRaH


## Citation notes from Methods

Below are the citation notes from the methods section.
Some citations are missing, those are noted here and how best guess were generated.

- [ ] short citation
  + copy from bib
  + manually entered or direct export



