---
title: "Data Rescue: Litton 2008"
format:
  html:
    toc: true
date: 2025-07
date-format: YYYY-MMMM
bibliography:
  - Litton2008_Methods.bib
  - Litton2008.bib
author:
  - name:
      given: Katherine
      family: Todd-Brown
      orcid: 0000-0002-3109-8130
      note: https://github.com/ktoddbrown
      affiliations:
        - id: ESS UF
          name: University of Florida
          department: Environmental Engineering Sciences
          city: Gainesville
          state: FL
          country: USA
      role: 
        - "Review"
  - name:
      given: Layla
      family: Kalandjian
      affiliation:
        - id: UF
          name: University of Florida
          city: Gainesville
          state: FL
          country: USA
      role: "Transcription"
      note: https://github.com/laylakalandjian
  - name:
      given: Savannah
      family: Scott
      affiliation:
        - id: UF
          name: University of Florida
          city: Gainesville
          state: FL
          country: USA
      role: "Transcription"
      note: https://github.com/SavaScott
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(kableExtra)

methods.file <- 'Litton2008_Methods.md'
table2.file <- 'Litton2008_Table2.csv'
figure6.file <- 'Litton2008_Figure6.csv'
primaryCitation.file <- 'Litton2008.bib'
methodsCitation.file <- 'Litton2008_Methods.bib'
```

This documents the data transcription for Litton 2008.
For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/91.

LITTON, C.M., SANDQUIST, D.R. and CORDELL, S. (2008), A non-native invasive grass increases soil carbon flux in a Hawaiian tropical dry forest. Global Change Biology, 14: 726-739. https://doi.org/10.1111/j.1365-2486.2008.01546.x

# Data Summary

This paper focuses on the invasion of non-native plants in terrestrial ecosystems, particularly by looking at the presence of *P. setaceum* in the Kaupulehu Dry Forest Preserve and its impact on underground C cycling. Litton examines both C pools and fluxes in invaded and non-invaded plots over a 2-year time period.

## Fit for purpose: HiCSC

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the HiCSC.

- Location: Latitude/Longitude as well as time of observation identified in methods section under *Study area* subsection.
- Soil carbon stock: Depth of soil mentioned in methods section under *Soil C pools* subsection.
- Soil type: Soil type identified in methods section. 
- Land use: Identified in methods section under *Study area* subsection.

# Tables and figures

Figure 1: Two-year time course of treatment means for monthly soil-surface CO2 efflux, weekly soil temperature, and weekly soil water potential with weekly cumulative precipitation
Figure 2: Soil surface CO2 efflux vs soil water potential
Figure 3: Daily soil-surface CO2 efflux and precipitation
Figure 4: Soil-surface CO2 efflux vs total leaf area index
Figure 5: Two-year time course of monthly tree litterfall and total annual aboveground litterfall
Figure 6: *In situ* foliage decomposition of dominant tree with non-native invasive grass vs without
Figure 7: C4 grass contribution (%) to soil C pools and fluxes

Table 1: Stand characteristics of study site with and without *Pennisetum setaceum*
Table 2: Soil carbon pools in grass and grass removal plots
Table 3: Cumulative annual flux of C from soils in grass and grass removal plots
  
### Table 2

Table 2 was modified to combine row subheadings, separated with a hyphen. Column names were combined with sublabels. Additionally, notes at bottom of table combined with table title in the caption.

```{r table2}
#| code-fold: true
#| message: false

filename <- table2.file
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```

### Figure 6

Data from Figure 6 pulled using Plot Digitizer and altered into a csv. SE stands for Standard Error, Standard error high and lows from bar graph included as separate columns. Letters indicating statistically significant differences were listed in a new column titled "letter".

```{r figure6}
#| label: figure6Check
#| code-fold: true
#| message: false

fig6Caption <- read_csv(file = figure6.file, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption
fig6Data <- read_csv(file = figure6.file,
                   skip = 1,
                                      col_types = cols(.default = col_character()))

fig6Data$`Litter Type` <- factor(fig6Data$`Litter Type`,
                                  levels = unique(fig6Data$`Litter Type`))
fig6Data$`Mass remaining (%)` <- as.numeric(fig6Data$`Mass remaining (%)`)
fig6Data$`SE High`<- as.numeric(fig6Data$`SE High`)
fig6Data$`SE Low`<- as.numeric(fig6Data$`SE Low`)
fig6Data$`Type Label` <- factor(fig6Data$`Type Label`, levels = c("Removal", "Grass"))

ggplot(fig6Data, aes(x = `Litter Type`, y = `Mass remaining (%)`, fill = interaction(`Litter Type`, `Type Label`))) +
  geom_bar(stat = "identity", color = "black", width = 0.5, position = position_dodge(width = 1)) +
  scale_y_continuous(breaks = c(0, 15, 30, 45, 60, 75), limits = c(0, 75)) +
  scale_fill_manual(values = c("white", "lightgray", "darkgray")) +
  facet_grid(~ `Type Label`, scales = "free_x", space = "free_x") +
  geom_errorbar(aes(ymin = `SE Low`, ymax = `SE High`, group = `Type Label`),
                position = position_dodge(width = 1),
                width = 0.25) +
  theme_classic() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 14, face = "bold"),
    axis.line = element_line(color = "black"),
    panel.spacing = unit(0, "lines")
  ) +
  scale_x_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  geom_text(aes(
    label = `Letter`, y = `SE High` + 2),
    position = position_dodge(width = 1),
    vjust = 0,
    size = 5
  ) +
    labs(y = "Mass Remaining (%)",
         x = "Species")

```

{{< include Litton2008_Methods.md >}}

# Transcription comparison

```{bash eval=FALSE}
diff --strip-trailing-cr --suppress-common-lines -y Litton2008_Table2.csv Litton2008_Table2_SRS.csv
```

# Files

These files are in the Litton2008 data rescue.

- [Readme](README_Litton2008.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Litton2008.bib](Litton2008.bib)
  + citation for article transcribed.
- [Litton2008_Methods.bib](Litton2008_Methods.bib)
  + citations for the methods section of the article.
- [Litton2008_Methods.md](Litton2008_Methods.md)
  + methods transcribed from primary article.
- [Litton2008_Table2.csv](Litton2008_Table2.csv)
  + table 2 from primary article with Soil C values.
- [Litton2008_Figure6.csv](Litton2008_Figure6.csv)
  + figure 6 from primary article with litter decomposition.
- [temp/](temp/)
    + scratch folder that will not be archived includes excel and article pdf documents.

# Working notes 

## Data Rescue Plan

- [x] describe files in human readable form inside ReadMe
- [x] description with Git issue with link
- [x] add contribution yaml
- [x] create excel sheet with Table 2
- [x] export excel table to csv
- [x] copy over methods section into markdown
- [x] use plot digitizer to pull figure 6
- [x] pull down citations in methods section to bib file
- [x] add in citation links to methods
- [x] cross check with second transcriber for tables and methods
- [x] submit to independent review
- [x] archive on SoilDRaH

# Level 0 Data rescue

This level 0 data read parses the transcribed files with minimal cleaning or transformations.

```{r readLevel0}
#This chunk has two purposes. 
#...1) Check the formatting by reading in everything 
#...2) Create a list of everything to process later in level 1

data.lvl0.ls <- list(
  #Read in a list of all the bib files
  citation = list(
    #Citation for the article transcriptions are pulled from
    primary = read.bib(file = primaryCitation.file), 
    #Citations for all referenced articles in methods section
    methods = read.bib(file = methodsCitation.file)
  ),
  #Read in the text transcription of the article's methods section
  method = read_lines(file = methods.file),
  #Read in the results as tables or figure transcriptions. This includes
  #...the caption as well as the tables themselves
  data = list(
    Table2 = list(
      #Read the caption as a text string. Captions are the first cell on 
      #...the first row.
      caption = read_csv(file = table2.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      #Read in all the data, skipping the first row with the caption and read
      #...in the table as character. This element is a tibble (data.frame).
      primary = read_csv(file = table2.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    ), #End 'Table1' element
    Figure6 = list(
      caption = read_csv(file = figure6.file,
                         col_types = cols(.default = col_character()), n_max = 1, col_names = FALSE)$X1[1],
      primary = read_csv(file = figure6.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    )
    #Repeat for all tables and figures...
   ) #end 'data' element
)

```

```{r}
#| label: checkReadFunction_lvl0

datafunction.lvl0.ls <- readAuthorYYYY(dataDir = '.',
                                    dataLevel = 'level0')

if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
  print('Read function matches code here.')
}else{
  print('There is a mismatch in the data objects between the code here and the read function.')
}

```
