---
title: "Data Rescue: Litton 2008"
format:
  html:
    toc: true
date: 2025-07
date-format: YYYY-MMMM
bibliography:
  - Litton2008_Methods.bib
  - Litton2008.bib
author:
  - name:
      given: Katherine
      family: Todd-Brown
      orcid: 0000-0002-3109-8130
      note: https://github.com/ktoddbrown
      affiliations:
        - id: ESS UF
          name: University of Florida
          department: Environmental Engineering Sciences
          city: Gainesville
          state: FL
          country: USA
      role: 
        - "Review"
  - name:
      given: Layla
      family: Kalandjian
      affiliation:
        - id: UF
          name: University of Florida
          city: Gainesville
          state: FL
          country: USA
      role: "Transcription"
      note: https://github.com/laylakalandjian
  - name:
      given: Savannah
      family: Scott
      affiliation:
        - id: UF
          name: University of Florida
          city: Gainesville
          state: FL
          country: USA
      role: "Transcription"
      note: https://github.com/SavaScott
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(kableExtra)

methods.file <- 'Litton2008_Methods.md'
table2.file <- 'Litton2008_Table2.csv'
figure6.file <- 'Litton2008_Figure6.csv'
primaryCitation.file <- 'Litton2008.bib'
methodsCitation.file <- 'Litton2008_Methods.bib'
```

This documents the data transcription for Litton 2008.
For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/91.

LITTON, C.M., SANDQUIST, D.R. and CORDELL, S. (2008), A non-native invasive grass increases soil carbon flux in a Hawaiian tropical dry forest. Global Change Biology, 14: 726-739. https://doi.org/10.1111/j.1365-2486.2008.01546.x

# Data Summary

This paper focuses on the invasion of non-native plants in terrestrial ecosystems, particularly by looking at the presence of *P. setaceum* in the Kaupulehu Dry Forest Preserve and its impact on underground C cycling. Litton examines both C pools and fluxes in invaded and non-invaded plots over a 2-year time period.

## Fit for purpose: HiCSC

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the HiCSC.

- Location: Latitude/Longitude as well as time of observation identified in methods section under *Study area* subsection.
- Soil carbon stock: Depth of soil mentioned in methods section under *Soil C pools* subsection.
- Soil type: Soil type identified in methods section. 
- Land use: Identified in methods section under *Study area* subsection.

# Tables and figures

Figure 1: Two-year time course of treatment means for monthly soil-surface CO2 efflux, weekly soil temperature, and weekly soil water potential with weekly cumulative precipitation
Figure 2: Soil surface CO2 efflux vs soil water potential
Figure 3: Daily soil-surface CO2 efflux and precipitation
Figure 4: Soil-surface CO2 efflux vs total leaf area index
Figure 5: Two-year time course of monthly tree litterfall and total annual aboveground litterfall
Figure 6: *In situ* foliage decomposition of dominant tree with non-native invasive grass vs without
Figure 7: C4 grass contribution (%) to soil C pools and fluxes

Table 1: Stand characteristics of study site with and without *Pennisetum setaceum*
Table 2: Soil carbon pools in grass and grass removal plots
Table 3: Cumulative annual flux of C from soils in grass and grass removal plots
  
### Table 2

Table 2 was modified to combine row subheadings, separated with a hyphen. Column names were combined with sublabels. Additionally, notes at bottom of table combined with table title in the caption.

```{r table2}
#| code-fold: true
#| message: false

filename <- table2.file
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```

### Figure 6

Data from Figure 6 pulled using Plot Digitizer and altered into a csv. SD stands for Standard Deviation, Standard deviation high and lows from bar graph included as separate columns.

```{r figure6}
#| code-fold: true
#| message: false

filename <- figure6.file
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```

{{< include Litton2008_Methods.md >}}

# Transcription comparison

```{bash eval=FALSE}
diff --strip-trailing-cr --suppress-common-lines -y Litton2008_Table2.csv Litton2008_Table2_SRS.csv
```

# Files

These files are in the Litton2008 data rescue.

- [Readme](README_Litton2008.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Litton2008.bib](Litton2008.bib)
  + citation for article transcribed.
- [Litton2008_Methods.bib](Litton2008_Methods.bib)
  + citations for the methods section of the article.
- [Litton2008_Methods.md](Litton2008_Methods.md)
  + methods transcribed from primary article.
- [Litton2008_Table2.csv](Litton2008_Table2.csv)
  + table 2 from primary article with Soil C values.
- [Litton2008_Figure6.csv](Litton2008_Figure6.csv)
  + figure 6 from primary article with litter decomposition.
- [temp/](temp/)
    + scratch folder that will not be archived includes excel and article pdf documents.

# Working notes 

## Data Rescue Plan

- [x] describe files in human readable form inside ReadMe
- [x] description with Git issue with link
- [x] add contribution yaml
- [x] create excel sheet with Table 2
- [x] export excel table to csv
- [x] copy over methods section into markdown
- [x] use plot digitizer to pull figure 6
- [x] pull down citations in methods section to bib file
- [x] add in citation links to methods
- [x] cross check with second transcriber for tables and methods
- [x] submit to independent review
- [ ] archive on SoilDRaH


## Citation notes from Methods

Below are the citation notes from the methods section.
Some citations are missing, those are noted here and how best guess were generated.

- [ ] short citation
  + copy from bib
  + manually entered or direct export