---
title: "Data Rescue: Rozema 1997"
format:
  html:
    toc: true
    eval: true 
date: last-modified
date-format: YYYY-MM
bibliography:
  - Rozema1997_Methods.bib #contains the citations in the methods section
  - Rozema1997.bib #contains citation for the data source, generally a single citation of the primary paper
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  - id: ashbonner310 #use the github name as an id
    name:
      given: Ashley
      family: Bonner
    orcid: 0000-0001-7882-6353
    affiliation:
      - ref: uf-ees
    role: 
      - transcription: lead (methods bib, Figure 1)
      - transcription: secondary (Table 1, Table 2)

  - id: anegondia #use the github name as an id
    name:
      given: Arunima
      family: Anegondi
    affiliation:
      - ref: uf
    role: 
      - transcription: lead (Table 1, methods)
      
  - id: Misheph #use the github name as an id
    name:
      given: Mirah
      family: Blanchard
    affiliation:
      - ref: uf
    role: 
      - transcription: lead (Table 1)
      - transcription: secondary (Figure 1, methods)
      
affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
  - id: uf
    name: University of Florida
    city: Gainesville
    state: Florida
    country: USA
    postal-code: 32611
    url: https://www.ufl.edu/
    
timelog:
  - activity: 
      description: updated methods citation notes and author contributions
      who: ashbonner310 
      start: 2026-04-13 T1515
      duration: P30M
  - activity:
      description: transcriptions (Table 1, Figure 1, methods)
      who: Misheph
      start: 2026-05-01
      duration: P1H30M    
  - activity:
      description: transcriptions (Table 1, methods)
      who: anegondia
      start: 2026-04-01
      duration: P1H
  - activity: 
      description: coding in file names, updating rescue plan and rescue info, cross checking tables and figure 
      who: ashbonner310
      start: 2026-05-14 T18:45
      duration: P2H45M
  - activity: 
      description: finish methods cross check and final edits 
      who: ashbonner310
      start: 2026-05-15 T17:00
      duration: P2H30min
---


```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false
#| eval: true


library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# data files for level 0
methods.file <- "Rozema1997_Methods.md"
table1.file <- 'Rozema1997_Table1.csv'
table2.file <- 'Rozema1997_Table2.csv'
figure1.file <- 'Rozema1997_Figure1.csv'

# Bibliography files
primaryCitation.file <- 'Rozema1997.bib'
methodsCitation.file <- 'Rozema1997_Methods.bib'

# read function; you will need to create this function after you have your data
#... rescue reviewed and before you finish your data harmonization.
readsource.file <- '../../02_DataHarmonization/readAuthorYYYY.R'
if(file.exists(readsource.file)){
  source(readsource.file)
}

#semantic files for level 1
of_variable.file <- '../../semantics/SoilDRaHVocabulary_of_variable.csv'
is_type.file <- '../../semantics/SoilDRaHVocabulary_is_type.csv'

```


# Data Summary

> Rozema, J., Tosserams, M., Nelissen, H.J.M. et al. *Stratospheric ozone reduction and ecosystem processes: enhanced UV-B radiation affects chemical quality and decomposition of leaves of the dune grassland species Calamagrostis epigeios*. Plant Ecology 128, 285–294 (1997). https://doi.org/10.1023/A:1009723210062

For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/#221.

In this study conducted in the Netherlands, the researchers wanted to know whether enhanced UV-B radiation would affect the chemical composition of dune grassland plants, as well as the decomposition rate of plant material. 
They studied this by growing *Calamagrostis epigeios* grass under either ambient or enhanced UV-B in a controlled, experimental field and analyzing its chemical composition. 
They examined decomposition using mesh litter bags over a 50-60 day period in either the experimental field or a dune grassland site. 
The study found that UV-B causes plant tissues to grow significantly more lignin, which slows decomposition, but also directly speeds up the breakdown of plant matter at the same time.

## Tables and figures

- Table 
  1) Percentage of of various organic compounds (\% Dry Weight) in litter grown under ambient vs enhanced UV-B, before and after decomposition under each lighting condition.
  2) Total carbon and nitrogen (\% Dry Weight) and C/N ratio of material in leaf grown under ambient vs enhanced UV-B, before and after decomposing under each lighting condition

- Figure 
  1)  Mass loss (\% Dry Weight) in leaf litter grown under ambient vs enhanced UV-B, before and after decomposition under each lighting condition, after: a) over 60 days in dune grassland in 1994 b) over 50 days in dune grassland in 1995 c) over 60 days on the experimental field.

## Fit for purpose: DeAD

This data was identified as a data source for the the **De**composition **A**cross **D**rylands (DeAD) data collection.
This is a new data effort targeting dryland decomposition studies. 
The purpose of this data collection is to develop better litter decay models for dryland ecosystems..

###Main Variables of Interest:

**Location**: Latitude/longitude explicitly stated for dune grassland site (52°30'N, 4°40'E) and Vrije Universiteit Amsterdam campus named in the Methods section.

**Duration**: Start and end time of exposure to the day resolution if possible, month is acceptable Exact dates and number of days for both decomposition experiment periods stated in Methods section under Decomposition Experiments: August 18, 1994 - October 17, 1994 (60 days) and September 9, 1995 - October 25, 1995 (50 days).

**Mass loss**: Results reported in Figure 1 as percentage dry weight loss.

**Solar exposure**: Litter was exposed to “ambient” UV-B radiation at around 5 kJ per m^2 per day or “enhanced” UV-B radiation at 7.5 kJ per m^2 per day.
 Lamps provided increased radiation, the “enhanced” radiation was filtered such that radiation less than 290 nm was blocked, “ambient” had foil filters to block radiation below 313 nm (to control for UV-A radiation)
 
###Other Variables of Interest:

*Litter species*: Calamagrostis epigejos/epigeios, a dune grass (common name bushgrass).
*Microsite data*: canopy vs bare soil distinction noted in Methods under Decomposition Experiments, but no quantitative data given.
*Litter chemical composition*: reported in table 1, C:N reported in table 2
*Litterbags*: 8.5 x 8.5 cm bags with 1 mm on the lower side, 3 mm on the upper side, 2g of dried material
*Treatments*: UV-B lamps, ambient control lamps, and cellulose acetate filters
  
## Files

These files are in the Rozema1997 data rescue.

- [Readme](README_Rozema1997.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Rozema1997.bib](Rozema1997.bib)
  + citation for article transcribed
- [Rozema1997_Methods.bib](Rozema1997_Methods.bib)
  + citations for the methods section of the article
- [Rozema1997_Methods.md](Rozema1997_Methods.md)
  + methods transcribed from primary article
- [Rozema1997_Table1.csv](Rozema1997_Table1.csv)
  + table 1 from primary article, with statistical significance values included as columns.
- [Rozema1997_Table2.csv](Rozema1997_Table2.csv)
  + table 2 from primary article,  with statistical significance values included as columns.
- [Rozema1997_Figure1.csv](Rozema1997_Figure1.csv)
  + a plot digitization of Figure 1 from primary article, with values resported to the tenths place (one degree of precision more than axis labels)
- [temp/](temp/)
    + scratch folder including second transcirptions that will be archived on GitHub and any files that will not be archived on GitHub, including local copies of the data resource (likely a PDF article) and any spreadsheets or other tools you used for the transcriptions.

# Data Rescue working notes 

## Data Rescue Plan

+ Lead
  - [X] Read the article or data resource and the discription of the target project
  - [X] Create a fresh fork of the primary repository https://github.com/ktoddbrown/SoilDRaH/tree/main
  - [X] Create a copy of this template folder and rename it to Rozema1997
  - [X] Fill out Data summary section (except for the Files), the Data Rescue Plan (here!), and update the author section of the yml to reflect who is working on the rescue
  - [X] Create a GitHub issue from the Data Rescue template and get feedback from a reviewer on the project (Dr Todd-Brown or a grad student typically).
  - [X] Add the reviewer to the author list in the yml
+ Do
  - [X] Create a spreadsheet with Table1 and export it to csv
  - [X] Create a spreadsheet with Table2 and export it to csv
  - [X] Extract points from Figure1 and export it to csv
  - [X] Transcribe or copy the method section
  - [X] Insert references to bib entries
  - [X] Create a bib file with the citations from the methods section
  - [X] Push to your local fork and get a second transcription pushed to your local fork by someone else on the project
+ Measure
  - [X] cross check with second transcriber for tables and methods, and resolve any differences
  - [X] compile or render this ReadMe with all chunks up through the Level 0 data rescue section with `eval: true`
  - [X] push to your local fork and create a pull request to the main branch of the primary repository
  - [X] address any revisions in your pull request review
  - [X] celebrate!

## Table 1

The second half of Table 1 uses a number of asterisks (*) to relay information about the statistical significance of different qualities (Litter Quality, Decomposition, Photodegradation) on the results.
The translation of the asterisks is included in the table caption.
These are saved in three extra columns on the CSV, corresponding to the organic compounds which are listed as rows (Soluble Carbohydrates, Starch, α-Cellulose, Hemicellulose, Tannins, Lignin).

```{r}
#| label: table1Check
#| eval: true

read_csv(file = table1.file, #read in the specific file
         skip = 1, #skip the first line with the caption
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = table1.file, #read in that file again but...
                           n_max = 1, #only read the first line
                           col_names = 'caption', #call it the caption
                           #read everything as a character
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE #do not return messages
                           )$caption #and take it out of the table
  )
```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareTable1
#| eval: true

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y temp/Rozema1997_Table1_AB.csv Rozema1997_Table1.csv
```

## Table 2

Table 2 was recorded as-is.
The caption includes details for interpreting the rows on statistical significance of litter quality, decomposition, and UV-B radiation.

```{r}
#| label: table2Check
#| eval: true


read_csv(file = table2.file, #read in the specific file
         skip = 1, #skip the first line with the caption
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = table2.file, #read in that file again but...
                           n_max = 1, #only read the first line
                           col_names = 'caption', #call it the caption
                           #read everything as a character
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE #do not return messages
                           )$caption #and take it out of the table
  )
```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareTable2
#| eval: true

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y temp/Rozema1997_Table2_AB.csv Rozema1997_Table2.csv
```

## Figure 1

Figure 1 was captured with plot digitizer and formatted into a single csv.

```{r}
#| label: Figure1Check
#| eval: true

read_csv(file = figure1.file, #read in the specific file
         skip = 1, #skip the first line with the caption
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = figure1.file, #read in that file again but...
                           n_max = 1, #only read the first line
                           col_names = 'caption', #call it the caption
                           #read everything as a character
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE #do not return messages
                           )$caption #and take it out of the table
  )

Figure1.df <- read_csv(file = figure1.file,
                   skip = 1,
                   show_col_types = FALSE) %>%
  mutate(Duration = str_c(`Duration (days)`, " days")) %>%
  unite(col = "Subgraphs",
        c("Site", "Year", "Duration"),
        sep = ", ") %>%
  select(!`Duration (days)`) %>%
  unite(col = "Treatments",
        c("Litter Quality", "Radiation"),
        sep = " ")

ggplot(data = Figure1.df) + 
  geom_col(aes(x = `Treatments`,
               y = `Mass loss (% DW)`),
           width = 0.5) + 
  geom_errorbar(aes(x = `Treatments`,
                    ymin = `Mass loss (% DW)`,
                    ymax = `Error Bar Height`),
                width = 0.2) + 
  facet_wrap(~Subgraphs,
             ncol = 1,
             scales = 'free_x')

```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareFigure1
#| eval: true

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y temp/Rozema1997_Figure1_MB.csv Rozema1997_Figure1.csv
```


## Methods

Any different to the cross check for the methods section will be output after this line.

```{bash}
#| label: compareMethods
#| eval: true
#| 
diff --strip-trailing-cr --suppress-common-lines -y temp/Rozema1997_Methods_MB.md Rozema1997_Methods.md
```

### Methods  - Text

The final text transcription from the Methods section is rendered below.

```{r}
#| label: readMethods
#| output: asis
#| eval: true

# This revised section will append two ### to any existing headers here so that it renders nicely with the other headers in the document.

methodsLines <- readLines(methods.file) |>
  str_replace_all(pattern = '^#', replacement = '####')

cat(paste0(methodsLines, collapse = '\n'))

```

<!----
## Citation notes from Methods

Below are the citation notes from the methods section.
This generally doesn't read well in the final report so commenting it out here but it is very useful for tracking citations when you are doing the transcriptions.
Some citations are missing, those are noted here and how best guess were generated.
Think of this as a lab notebook documenting how you found each citation.

- [X] Tosserams et al. 1995
  + copy from bib
  + direct export
  + Source mentions "Tosserams et al. 1997" but that citation doesn't exist in references
  + Investigation into this citation matches the reference made in the methods section, the methods markdown file changes the erroneous reference to this one
  
- [X] Tosserams et al. 1996
  + direct export
  
- [X] Allen 1989
  + book, manually entered 
  
- [X] Green et al. 1980
  +  manually entered 

- [X] Kuiters 1990
  +  direct export

- [X] Rozema et al. 1995
  +  manually entered 
  +  chapter/article within a book within a series. 
  +  added details about series book appears in

- [X] Sokal & Rohlf 1981
  + book, manually entered 
  
- [X] Tolsma 1989
  + doctorate thesis
  + manually entered

- [X] Van Beckhoven 1995
  + doctorate thesis
  + manually entered
---->

# Level 0 Data rescue

This level 0 data read parses the transcribed files with minimal cleaning or transformations.

```{r}
#| label: readLevel0
#| eval: true
#This chunk has two purposes. 
#...1) Check the formatting by reading in everything 
#...2) Create a list of everything to process later in level 1

data.lvl0.ls <- list(
  #Read in a list of all the bib files
  citation = list(
    #Citation for the article transcriptions are pulled from
    primary = read.bib(file = primaryCitation.file), 
    #Citations for all referenced articles
    methods = read.bib(file = methodsCitation.file)
  ),
  #Read in the text transcription of the article's methods section
  method = read_lines(file = methods.file),
  #Read in the results as tables or figure transcriptions. This includes
  #...the caption as well as the tables themselves
  data = list(
    Table1 = list(
      #Read the caption as a text string. Captions are the first cell on 
      #...the first row.
      caption = read_csv(file = table1.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      #Read in all the data, skipping the first row with the caption and read
      #...in the table as character. This element is a tibble (data.frame).
      primary = read_csv(file = table1.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    ), #End 'Table1' element
  Table2 = list(
      #Read the caption as a text string. Captions are the first cell on 
      #...the first row.
      caption = read_csv(file = table2.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      #Read in all the data, skipping the first row with the caption and read
      #...in the table as character. This element is a tibble (data.frame).
      primary = read_csv(file = table2.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    ), #End 'Table2' element  
  Figure1 = list(
      #Read the caption as a text string. Captions are the first cell on 
      #...the first row.
      caption = read_csv(file = figure1.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      #Read in all the data, skipping the first row with the caption and read
      #...in the table as character. This element is a tibble (data.frame).
      primary = read_csv(file = figure1.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    ) #End 'Figure1' element  
   ) #end 'data' element
)

```

The code is also in `readsource.file` and returned with the `dataLevel = 'level0'` argument.
Below verifies that the data from the function matches what is coded above.

```{r}
#| label: checkReadFunction_lvl0
#| eval: false

if(file.exists(readsource.file)){
  datafunction.lvl0.ls <- readAuthorYYYY(dataDir = '.',
                                         dataLevel = 'level0')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function matches code here.')
  }else{
    print('There is a mismatch in the data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}
```

# Level 1 data - CURATION PROJECT

This data is identified as a data source for the ...
Broad classes of information that we are looking for include:

<!--- Include description of where the data is located in the source.--->
- Location: 
- Soil carbon stock: 
- Soil type: 
- Land use: 

Information from the methods and other sections are pulled into tables with line by line source citations.
This information can be values, units, methods, or other useful notes and is manually extracted from the transcriptions loaded into the level 0 data above.

The final format of the data frame here will include the following 

  + grouping ids
  + `of_variable`
  + `is_type`
  + `with_entry`
  + `from_source`

The `of_variable` being used here are:

 + citation
 
 <!--- include other variables being used, this should match vocabulary in the SoilDRaHVocabulary resources --->

```{r}
#| label: createLvl1
#| eval: false

#### Method-section only data ####
#Create a table for the data in the method-section that is not directly 
#...associated with any table. Often this is study level information.

studyMeta <- tribble(~of_variable, ~is_type, ~with_entry, ~from_source,
                   'region', 'value', 'Honaunau Forest on the southwestern slopes of Mauna Loa, island of Hawaii', paste('Method ln5:', paste(data.lvl0.ls$method[5], collapse = ' ')))

#### Table/Figure N ####
# Process the primary data from Table/Figure M by maintaining data groups and 
#...standardizing the variable name so that we can link this to the study 
#...metadata and data in the other tables when needed. Link in the methods 
#...and other information from the method-section as needed.

Table1Primary <- data.lvl0.ls$data$Table1$primary |>
  #Add a row id preserving row grouping -- not unique
  mutate(row_id = paste0('R', 1:n())) |>
  #Transform the table to long data by row group and column names.
  #...This allows us to then create the cross links between the tables.
  pivot_longer(cols = -row_id,
               names_to = 'column_name', values_to = 'with_entry',
               values_drop_na = TRUE) |> #drop missing values
  #Link the variable names to the column names using a switch statement
  mutate(of_variable = case_when(
           #Soil PH is given as an example below
           column_name == 'SOC' ~ 'soil_organic_carbon')) |>
  #Flag the source of this data as Table 1, this table is merged later with
  #...other data so this keeps track of where the data came from
  mutate(from_source = 'Table 1')

## Create the metadata for table one from the header information and cited methods
Table1Meta <- Table1Primary |>
  select(column_name, of_variable) |>
  unique() |>
  #Grab everything between the parentheses as units and attribute the source as the column names.
  mutate(unit = str_extract(column_name, pattern = '(?<=\\().*(?=\\))'),
         from_source = 'Table 1 column names.') |>
  #If there aren't units then drop the row
  filter(!is.na(unit)) |>
  #Don't keep the column name now that you have the units, cross link via the variable
  select(-column_name) |>
  #Stack into the table the methods for each variable that we can find
  bind_rows(
  tribble(~of_variable, ~method, ~from_source,
            'soil_organic_carbon', paste0(data.lvl0.ls$method[72:73], collapse = ' '), 'Methods ln72-73') ) |>
  #Stack onto the tables the controlled vocabulary used
  bind_rows(
    tribble(~of_variable, ~control_vocabulary, ~from_source,
            'soil_class', 'Histosol:USDA classification for histosol soil type|Andisols:USDA classification for andisol soil type', 'expert informed')
  ) |>
  # Push the dimensions of the variable into a single column. Making this a long table.
  pivot_longer(cols = c(unit, method, control_vocabulary),
               names_to = 'is_type',
               values_drop_na = TRUE,
               values_to = 'with_entry')

#### Create level 1 #####
## Pull everything together into by stacking the meta and primary data tables
data.lvl1.ls <- list(
  meta = bind_rows(studyMeta, Table1Meta),
  primary = bind_rows(Table1Primary)|>
    arrange(row_id, column_name,
            of_variable, is_type, with_entry, from_source)
  )

```

## Check for new variables

`of_variable` in the data files should be defined in the SoilDRaH Vocabulary files.

```{r}
#| label: check_of_variable
#| eval: false

#Read in the list of current 'of_variable' entries
SoilDRaH_variable_list <- read_csv(file =of_variable.file,
                                   show_col_types = FALSE)

#compare the variables in the level 1 tables to the SoilDRaH variables
new_variables <- lapply(data.lvl1.ls, function(xx){
  setdiff(unique(xx$of_variable), 
          unique(SoilDRaH_variable_list$of_variable))
}) |>
  unlist()

#Check that there are no new variables
if(length(new_variables) == 0){
  print('All of_variables are in the SoilDRaH variable list. No additions required')
}else{
  cat(paste('Please start a issue for new variables [',
               new_variables, ']', sep = '', collapse = '\n'))
}
```

## Check for is_type

`is_type` in the data files should be defined in the SoilDRaH Vocabulary files.

```{r}
#| label: check_is_type
#| eval: false

#read in the current types and what variable they are associated with
SoilDRaH_type_list <- read_csv(file = is_type.file,
                                   show_col_types = FALSE) |>
  #remove the project definition and only keep the variable and type
  select(of_variable, is_type) |>
  #variables are pushed together in the type defined table. Separate the
  #...entries here
  separate_longer_delim(cols = 'of_variable', delim = ',')

#Pull all the of_variable-is_type pairs from the source data
source_pairs <- tibble()

#If there are a large number of entries, consider replacing this for-loop.
#...However this avoids having to install the plyr or other libraries, and 
#...there are generally no more then 3-5 tables
for(dataTable in data.lvl1.ls){
  source_pairs <- dataTable |>
    select(of_variable, is_type) |> 
    unique() |>
    bind_rows(source_pairs)
}

undefined_data_type <- source_pairs |>
  #subset for variables that are not 'incomplete'
  semi_join(SoilDRaH_variable_list |>
              filter(project_status != 'incomplete'),
            by = join_by(of_variable)) |>
  #check for of_variable-is_type tuples that are not in the SoilDRaH type definitions
  anti_join(SoilDRaH_type_list,
            by = join_by(of_variable, is_type))

#Tell the user what you found.
if(nrow(undefined_data_type) == 0){
  print('All of_variables-is_types are in the SoilDRaH type list. No additions required')
}else{
  print(paste0('Please start a issue for new of_variable-is_type [', paste(paste(undefined_data_type$of_variable, undefined_data_type$is_type, sep = '-'), collapse = ','), ']'))
}
```

## Check function match

Check for an exact match with the read function.

```{r}
#| label: checkReadFunction_lvl1
#| eval: false

if(file.exists(readsource.file)){
  datafunction.lvl1.ls <- readAuthorYYYY(dataDir = '.',
                                         dataLevel = 'level1')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function output matches ReadMe code for level 1.')
  }else{
    print('There is a mismatch in the level 1 data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}

```


# References

<!---- This space will be filled with the citations when rendered. Leave this blank. ---->

