---
title: "Data Rescue: Youkhana2009"
format:
  html:
    toc: true
date: 2025-06
date-format: YYYY-MMMM
bibliography:
  - Youkhana2009_Methods.bib
  - Youkhana2009.bib
author:
  - name:
      given: Katherine
      family: Todd-Brown
      orcid: 0000-0002-3109-8130
      note: https://github.com/ktoddbrown
      affiliations:
        - id: ESS UF
          name: University of Florida
          department: Environmental Engineering Sciences
          city: Gainesville
          state: FL
          country: USA
      role: 
        - "Transcription"
        - "Review"
  - name:
      given: Savannah
      family: Scott
      affiliation:
        - id: ESS UF
          name: University of Florida
          city: Gainesville
          state: FL
          country: USA
      role: "Transcription"
      note: https://github.com/SavaScott
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(kableExtra)
library(bibtex)

methods.file <- 'Youkhana2009_Methods.md'
table1.file <- 'Youkhana2009_Table1.csv'
table2.file <- 'Youkhana2009_Table2.csv'
table4.file <- 'Youkhana2009_Table4.csv'
table5.file <- 'Youkhana2009_Table5.csv'
primaryCitation.file <- 'Youkhana2009.bib'
methodsCitation.file <- 'Youkhana2009_Methods.bib'
```

This documents the data transcription for @Youkhana2009.
For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/84

> Youkhana, Adel, and Travis Idol. “Tree pruning mulch increases soil C and N in a shaded coffee agroecosystem in Hawaii.” Soil Biology and Biochemistry, vol. 41, no. 12, Dec. 2009, pp. 2527–2534, https://doi.org/10.1016/j.soilbio.2009.09.011.

# Data Summary

Youkhana2009 primarily measures the relationship between tree pruning mulch application and the retention of soil carbon (C) and nitrogen (N) in a shaded coffee agroecosystem. 
The study involved applying chipped mulch from the *Leucaena* hybrid *KX2* to coffee plots over three years.
Mulch decomposition and nitrogen loss were measured during the first year, and changes in soil C and N were monitored over the subsequent two years.
They use this data to test if there were any significant changes in the soil composition due to the addition of mulch to the agroecosystem. 

## Fit for purpose: HiCSC

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the HiCSC.

- Location: Locational data was found in the material and methods section, subsection 2.1 Study Site.
- Soil carbon stock: The depth of the soil was mentioned in the materials and methods section.
- Soil type: Soil type is mentioned in the materials and methods section, subsection 2.1 Study Site
- Land use: The land use of the soil is found within the materials and methods section, subsection 2.1 Study Site

# Tables and figures

- Table 
  1) *Leucaena* mulch additions from tree pollarding
  2) Initial soil characteristics (2005) in a shaded coffee agroecosystem
  3) Significance of mass loss and changes in proportion of mulch biochemical components over time
  4) Changes in carbon, nitrogen, and lignin concentrations and the C:N and lignin:N ratios of tree mulch in shade coffee systems
  5) Changes in soil C and N due to mulch additions in a shade coffee agroecosystem

- Figure 
  1) Mean monthly air temperature and relative humidity and cumulative rainfall at the study site during 2006
  2) Mulch decomposition in shade coffee system
  3) Nitrogen loss from liter in shade coffee system
  4) Mass loss from mulch biochemical components in shade coffee system
  5) Changes in proportion of mulch biochemical components in shade coffee system
  
### Table 1

Table 1 was modified to include the footer in the caption.

```{r table1}
#| code-fold: true
#| message: false

filename <- table1.file
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```
