---
title: "Data Rescue: Youkhana2009"
format:
  html:
    toc: true
date: 2025-06
date-format: YYYY-MMMM
bibliography:
  - Youkhana2009_Methods.bib
  - Youkhana2009.bib
author:
  - name:
      given: Katherine
      family: Todd-Brown
      orcid: 0000-0002-3109-8130
      note: https://github.com/ktoddbrown
      affiliations:
        - id: ESS UF
          name: University of Florida
          department: Environmental Engineering Sciences
          city: Gainesville
          state: FL
          country: USA
      role:
        - "Review"
  - name:
      given: Savannah
      family: Scott
      affiliation:
        - id: ESS UF
          name: University of Florida
          city: Gainesville
          state: FL
          country: USA
      role: "Transcription"
      note: https://github.com/SavaScott
 - name:
      given: Layla
      family: Kalandjian
      affiliation:
        - id: ESS UF
          name: University of Florida
          city: Gainesville
          state: FL
          country: USA
      role: "Transcription"
      note: https://github.com/laylakalandjian
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(kableExtra)
library(bibtex)

methods.file <- 'Youkhana2009_Methods.md'
table1.file <- 'Youkhana2009_Table1.csv'
table2.file <- 'Youkhana2009_Table2.csv'
table4.file <- 'Youkhana2009_Table4.csv'
table5.file <- 'Youkhana2009_Table5.csv'
primaryCitation.file <- 'Youkhana2009.bib'
methodsCitation.file <- 'Youkhana2009_Methods.bib'
```

This documents the data transcription for @Youkhana2009.
For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/84

> Youkhana, Adel, and Travis Idol. “Tree pruning mulch increases soil C and N in a shaded coffee agroecosystem in Hawaii.” Soil Biology and Biochemistry, vol. 41, no. 12, Dec. 2009, pp. 2527–2534, https://doi.org/10.1016/j.soilbio.2009.09.011.

# Data Summary

Youkhana2009 primarily measures the relationship between tree pruning mulch application and the retention of soil carbon (C) and nitrogen (N) in a shaded coffee agroecosystem. 
The study involved applying chipped mulch from the *Leucaena* hybrid *KX2* to coffee plots over three years.
Mulch decomposition and nitrogen loss were measured during the first year, and changes in soil C and N were monitored over the subsequent two years.
They use this data to test if there were any significant changes in the soil composition due to the addition of mulch to the agroecosystem. 

## Fit for purpose: HiCSC

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the HiCSC.

- Location: Locational data was found in the material and methods section, subsection 2.1 Study Site.
- Soil carbon stock: The depth of the soil was mentioned in the materials and methods section.
- Soil type: Soil type is mentioned in the materials and methods section, subsection 2.1 Study Site
- Land use: The land use of the soil is found within the materials and methods section, subsection 2.1 Study Site

# Tables and figures

- Table 
  1) *Leucaena* mulch additions from tree pollarding
  2) Initial soil characteristics (2005) in a shaded coffee agroecosystem
  3) Significance of mass loss and changes in proportion of mulch biochemical components over time
  4) Changes in carbon, nitrogen, and lignin concentrations and the C:N and lignin:N ratios of tree mulch in shade coffee systems
  5) Changes in soil C and N due to mulch additions in a shade coffee agroecosystem

- Figure 
  1) Mean monthly air temperature and relative humidity and cumulative rainfall at the study site during 2006
  2) Mulch decomposition in shade coffee system
  3) Nitrogen loss from liter in shade coffee system
  4) Mass loss from mulch biochemical components in shade coffee system
  5) Changes in proportion of mulch biochemical components in shade coffee system
  
### Table 1

Table 1 was modified to include the footer in the caption.

```{r table1}
#| code-fold: true
#| message: false

filename <- table1.file
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```

### Table 2

Table 2 was modified to include the footer in the caption.

```{r table2}
#| code-fold: true
#| message: false

filename <- table2.file
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```

### Table 4

Table 4 was modified to include the footer in the caption.

```{r table4}
#| code-fold: true
#| message: false

filename <- table4.file
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```

### Table 5

Table 5 was modified to include the footer in the caption.

```{r table5}
#| code-fold: true
#| message: false

filename <- table5.file
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```

{{< include Youkhana2009_Methods.md >}}

# Transcription comparison

```{bash eval=FALSE, echo=FALSE}
diff --strip-trailing-cr --suppress-common-lines -y temp/AuthorYYYY_Methods_SRS.md AuthorYYYY_Methods_LK.Rmd
```

# Files

These files are in the @Youkhana2009 data rescue.

- [Readme](README_Youkhana2009.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Youkhana2009.bib](Youkhana2009.bib)
  + citation for article transcribed
- [Youkhana2009_Methods.bib](Youkhana2009_Methods.bib)
  + citations for the methods section of the article
- [Youkhana2009_Methods.md](Youkhana2009_Methods.md)
  + methods transcribed from primary article
- [Youkhana2009_Table1.csv](Youkhana2009_Table1.csv)
  + table 1 from primary article with *Leucaena* mulch additons
- [Youkhana2009_Table2.csv](Youkhana2009_Table2.csv)
  + table 2 from primary article with inital soil characteristics (2005)
- [Youkhana2009_Table4.csv](Youkhana2009_Table4.csv)
  + table 4 from primary article with changes in carbon concentrations
- [Youkhana2009_Table5.csv](Youkhana2009_Table5.csv)
  + table 5 from primary article with changes in soil C due to mulch addition

# Working notes 

## Data Rescue Plan

- [x] describe files in human readable form inside ReadMe
- [x] description with Git issue with link
- [x] add contribution yaml
- [x] create excel sheet with Table 1, 2, 4 and 5
- [x] export excel table to csv
- [x] copy over methods section into markdown
- [x] pull down citations in methods section to bib file
- [x] add in citation links to methods
- [x] cross check with second transcriber for tables and methods
- [x] submit to independent review
- [ ] archive on SoilDRaH


## Citation notes from Methods

Below are the citation notes from the methods section.
Some citations are missing, those are noted here and how best guess were generated.

- [ ] short citation
  + copy from bib
  + manually entered or direct export

- [x] Giambelluca, T. W., Chen, Q., Frazier, A. G., Price, J. P., Chen, Y. L., Chu, P. S., ... & Delparte, D. M. (2013). Online rainfall atlas of Hawai ‘i. Bulletin of the American Meteorological Society, 94(3), 313-316.
  + direct export
  
- [x] National Climatic Data Center 2002
   + no citation

- [x] Brewbaker, J. L. (2008). Registration of KX2‐Hawaii, Interspecific‐Hybrid Leucaena. Journal of Plant Registrations, 2(3), 190-193.
  + direct export
  
- [x] Herrick, J. E. (1995). Simple method for determination of mass loss rates for soil-contaminated samples in decomposition studies.
  + direct export
  
- [x] Idol, T. W., Holzbaur, K. A., Pope, P. E., & Ponder, F. (2002). Control‐bag correction for forest floor litterbag contamination. Soil Science Society of America Journal, 66(2), 620-623.
  + direct export

- [x] Holechek, J. L., & Vavra, M. (1982). Comparison of micro-and macro-digestion methods for fiber analysis in forages and ruminant diets. Rangeland Ecology & Management/Journal of Range Management Archives, 35(6), 799-801.
  + direct export
  
- [x] Anderson, J. M., & Ingram, J. S. (1994). Tropical soil biology and fertility: a handbook of methods. Soil Science, 157(4), 265.
  + direct export
  
- [x] Chapin III, F. S., Matson, P. A., & Vitousek, P. (2011). Principles of terrestrial ecosystem ecology. Springer Science & Business Media.
  + direct export
  
- [x] SAS/STAT User's Guide Version 9 1990
  + no citation
  
- [x] von Ende, C. N. (2020). Repeated-measures analysis: growth and other time-dependent measures. In Design and analysis of ecological experiments (pp. 113-137). Chapman and Hall/CRC.
  + direct export
  
- [x] Maxwell, S. E., Delaney, H. D., & Kelley, K. (2017). Designing experiments and analyzing data: A model comparison perspective. Routledge.
  + direct export
   
- [x] Scheiner, S. M. (2020). MANOVA: multiple response variables and multispecies interactions. In Design and analysis of ecological experiments (pp. 94-112). Chapman and Hall/CRC.
  + direct export

# Level 0: raw data reads
```{r readLevel0}

data.lvl0.df <- list(citation = 
                    list(primary = 
                           read.bib(file = primaryCitation.file), 
                         methods = 
                           read.bib(file = methodsCitation.file)
                    ),
                  method = read_lines(file = methods.file),
                  data = list(Table1 = 
                                list(caption = 
                                       read_csv(file = table1.file,
                                                col_types = cols(.default = col_character()),
                                                n_max = 1, col_names = FALSE)$X1[1],
                                     primary = 
                                       read_csv(file = table1.file,
                                                col_types = cols(.default = col_character()),
                                                skip = 1)
                                ),
                              Table2 = list(
                                caption = read_csv(file = table2.file,
                                                col_types = cols(.default = col_character()),
                                                n_max = 1, col_names = FALSE)$X1[1],
                                     primary = 
                                       read_csv(file = table2.file,
                                                col_types = cols(.default = col_character()),
                                                skip = 1)
                                ),
                               Table4 = 
                                list(caption = 
                                       read_csv(file = table4.file,
                                                col_types = cols(.default = col_character()),
                                                n_max = 1, col_names = FALSE)$X1[1],
                                     primary = 
                                       read_csv(file = table4.file,
                                                col_types = cols(.default = col_character()),
                                                skip = 1)
                                ),
                                Table5 = 
                                list(caption = 
                                       read_csv(file = table5.file,
                                                col_types = cols(.default = col_character()),
                                                n_max = 1, col_names = FALSE)$X1[1],
                                     primary = 
                                       read_csv(file = table5.file,
                                                col_types = cols(.default = col_character()),
                                                skip = 1)
                                )))
```
