---
title: "Data Rescue: Crews 1995"
format:
  html:
    toc: true
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - Crews1995_Methods.bib
  - Crews1995.bib
author:
  - name:
      given: Katherine
      family: Todd-Brown
      orcid: 0000-0002-3109-8130
      note: https://github.com/ktoddbrown
      affiliations:
        - id: ESS UF
          name: University of Florida
          department: Environmental Engineering Sciences
          city: Gainesville
          state: FL
          country: USA
      role: 
        - "Review"
  - name:
      given: Layla
      family: Kalandjian
      affiliation:
        - id: UF
          name: University of Florida
          city: Gainesville
          state: FL
          country: USA
      role: "Transcription"
      note: https://github.com/laylakalandjian
  - name:
      given: Savannah
      family: Scott
      affiliation:
        - id: UF
          name: University of Florida
          city: Gainesville
          state: FL
          country: USA
      role: "Transcription"
      note: https://github.com/SavaScott
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(kableExtra)

methods.file <- 'Crews1995_Methods.md'
primaryCitation.file <- 'Crews1995.bib'
methodsCitation.file <- 'Crews1995_Methods.bib'
```

This documents the data transcription for  .
For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/87.

Crews, T.E., Kitayama, K., Fownes, J.H., Riley, R.H., Herbert, D.A., Mueller-Dombois, D. and Vitousek, P.M. (1995), Changes in Soil Phosphorus Fractions and Ecosystem Dynamics across a Long Chronosequence in Hawaii. Ecology, 76: 1407-1424. https://doi.org/10.2307/1938144

# Data Summary

Crews (1995) looks at changes in soil phosphorous fractions at six chronosequence sites across Hawaii, with total carbon also recorded for additional soil analyses. While the paper primarily focuses on changes in soil P fractions in comparison to the Walker and Syers model, carbon data was used to create 50-cm soil profiles from each site and element ratios with nitrogen and organic phosphorus. Litter of dominant _Metrosideros polymorpha_ tree species is also analyzed for contributions to soil composition.

## Fit for purpose: HiCSC

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the [HiCSC](https://github.com/ktoddbrown/SoilDRaH/wiki/Hawaii-SOC).

Location: The chronosequence section in this paper does not list specific latitude/longitude of the 6 sites but does include a figure which could be used to approximate specific latitude and longitude. The paper also gives detailed regional descriptions and elevations which could be used for inference. Year of observation is not given in the text.
Soil carbon stock: Depth of the sample of soil organic carbon is given in the methods section, under subsection _Other soil analyses_.
Soil type: Detailed classification provided in the chronosequence section.
Land use: Detailed land use and land use history in the chronosequence section.

# Tables and figures

Figure 1: Model of phosphorus transformations with time
Figure 2: Locations and ages of the six chronosequence field sites in Hawaii
Figure 3: Changes in phosphorus fractions with time at six chronosequence sites presented as mass per volume and fractions of total P
Figure 4: Percentage of leaf litter mass remaining versus decomposition time up to 2 years
Figure 5: Exponential decomposition constants for litter decomposed over 2 years
Figure 6: Changes in percentage cover of six dominant tree and shrub genera at six chronosequence sites
Figure 7: Detrended correspondence analysis (DCA) scores of chronosequence sites
Figure 8: Cluster analysis and similarity comparison of species composition of chronosequence sites
Figure 9: Changes in percentage of P and N in foliage of _Metrosideros polymorpha_ across chronosequence sites

Table 1: Hawaiian chronosequence sites including age, elevation, mean temperature and precipitation, soil classification, and location
Table 2: Sums and ratios of total carbon, nitrogen, and organic phosphorus at six chronosequence sites
Table 3: P and N availability, nitrous oxide flux, gross N-mineralization and nitrification, and pH of chronosequence sites
Table 4: Initial litter quality of senescent leaves from four chronosequence sites including N, P, Ca, Mg, K, Lignin, and Cellulose percentages
Table 5: Percentage of initial N and P remaining in leaf litter decomposed in situ for 1.5 years
Table 6: Total community basal areas of trees >30 mm in diameter and maximum canopy height
Table 7: Species diversity of the Hawaiian chronosequence sites
  
### Table 1

Table 1 was not modified.

```{r table1}
#| code-fold: true
#| message: false

filename <- 'Crews1995_Table1.csv'
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```

  
### Table 2

Table 2 was modified by condensing the separate Pool sizes and Element ratios sections of the table into a singular section. This was done by including the "Pool sizes (kg/m<sup>2</sup>)" and "Element Ratios" subheadings into the row labels.

```{r table1}
#| code-fold: true
#| message: false

filename <- 'Crews1995_Table2_LK.csv'
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                           show_col_types = FALSE)$caption)
```


{{< include Crews1995_Methods.md >}}

# Transcription comparison

```{bash eval=FALSE}
diff --strip-trailing-cr --suppress-common-lines -y Crews1995_Methods_LK.md Crews1995_Methods__SRSfinal.md
```

# Files

These files are in the Crews1995 data rescue.

- [Readme](README_Crews1995.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Crews1995.bib](Crews1995.bib)
  + citation for article transcribed
- [Crews1995_Methods.bib](Crews1995_Methods.bib)
  + citations for the methods section of the article
- [Crews1995_Methods.md](Crews1995_Methods.md)
  + methods transcribed from primary article
- [Crews1995_Table1.csv](Crews1995_Table1.csv)
  + table 1 from primary article
- [Crews1995_Table2.csv](Crews1995_Table2.csv)
  + table 2 from primary article
- [temp/](temp/)
    + scratch folder that will not be archived includes excel and article pdf document

# Working notes 

## Data Rescue Plan

- [x] describe files in human readable form inside ReadMe
- [x] copy description from Git issue with link
- [x] add contribution yaml
- [x] create excel sheet for tables 1 and 2
- [x] export excel table to csv
- [x] copy over methods section into markdown
- [x] pull down citations in methods section to bib file
- [x] add in citation links to methods
- [x] cross check with second transcriber for tables and methods
- [ ] submit to independent review
- [ ] archive on SoilDRaH


## Citation notes from Methods

Below are the citation notes from the methods section.
Some citations are missing, those are noted here and how best guess were generated.

- [ ] short citation
  + copy from bib
  + manually entered or direct export



