---
title: "Summary of data rescues"
format:
  html:
    toc: true
    code-fold: true
date: last-modified
date-format: YYYY-MMMM
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  #remove ktoddbrown if they did not work on the rescue directly
  - id: ktoddbrown #use the github name as an id
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
timelog:
  - activity:
      description: Parse YAML and Git logs for authorship and repo contribution timeline. Fixed formatting in many of the README qmd files.
      who: ktoddbrown
      start: 2027-07-30
      duration: P3H
  - activity:
      description: Parse yaml time logs. Only minor fixes in the README qmd needed
      who: ktoddbrown
      start: 2027-08-1
      duration: P1H
---


**Purpose**: The purpose of this file is to track the number of data rescues, how often data rescues are contributed, how frequently students (especially undergraduates) contribute data rescues, and how long data rescues take to preform. This is to provide an accurate estimate for future projects rather then tracking metrics for individual people.

# Setting things up

This code base uses the R-packages `tidyverse`, `plyr`, and `yaml` for data processing and YAML parsing.

```{r}
#| label: setup

library(tidyverse)
library(yaml) #used to parse headers
#plyr package not loaded so as not to conflict the name space. 
#...Called out instead using the package::function notate.
```


```{r}
#| label: pullGitLogs

#Pull out the git blame and timestamps
fileGitTouches <- tibble(filenames = list.files(path = '.',
                      recursive = TRUE)) |>
  #remove files in the temp or rendered directories
  filter_out(str_detect(filenames, '/t|Temp/') |
             str_detect(filenames, '_files')) |>
  #slice_sample(n=10) |> #uncomment for code dev
  #pass the system a `git log` request for each file 
  #... (takes a minute, quick stretch break!)
  mutate(gitLogStr = paste0(system(paste0('git log -- ', filenames),
                            intern = TRUE), collapse = '\\n'),
         .by = filenames) |>
  #parse the log file
  reframe(timestamp_str = #when was the commit, many are squashed when merged to main
            str_extract_all(gitLogStr,
                            '(?<=Date:)[[:alnum:]\\s\\-\\:]+') |> unlist(),
         author = #who made the commit
           str_extract_all(gitLogStr, 
                           '(?<=Author:\\s)[^<]*(?=\\s<)') |> unlist(),
         .by = c(filenames, gitLogStr)) |>
  mutate(timestamp_date = 
           parse_date_time(timestamp_str, 
                           orders = '%a %b %d %H:%M:%S %Y %z',
                           truncated = 1), #sometimes no timezone, deal
         studyID = str_extract(filenames, '^\\w+(?=/)'),
         .by = c(filenames, gitLogStr)) 
 
```

```{r}
#| label: pullReadMeYAML


#create a dataframe of readme files
readME.df <- tibble(filenames = list.files(path = '.',
                      recursive = TRUE)) |>
  #define readme based on the appearence of 'readme' at the start of the file name
  #... of files within directories and that end in 'md'
  filter(str_detect(filenames, regex('^\\w+/README_\\w+\\.\\wmd$',
                                     dotall = TRUE, ignore_case = TRUE))) |>
  #pull the directory as study name
  mutate(studyID = str_extract(filenames, '^\\w+(?=/)')) 

readME.ls <- setNames(as.list(readME.df$filenames), readME.df$studyID)
studyID_exclude <- c('CPEAT2025', 'FIA2024', 'NCSS2023', 'ISCN3_2015')

allYAML <- lapply(readME.ls[setdiff(names(readME.ls), studyID_exclude)],
                  function(myfilename){
                    #allYAML <- lapply(readME.ls['Nusslein1999'], function(myfilename){
                    fileLines <- read_lines(file = myfilename) 
                    
                    yamlStartEnd <- which(str_detect(fileLines, '^---\\s*$'))
                    
                    fileHeader <- fileLines[(yamlStartEnd[1]+1):(yamlStartEnd[2]-1)]
                    
                    yamlHeader <- read_yaml(text = fileHeader)
                    
                  })

```

```{r}
#| label: assignStudyType

studyTypes.df <- readME.df |>
  select(studyID) |>
  unique() |>
  mutate(sourceType = 
           case_when(studyID == '_TemplateYYYY' ~ 'template',
                     studyID %in% c('FIA2024', 'NCSS2023') ~ 'monitoring data',
                     studyID %in% c('CPEAT2025', "ISCN3_2015") ~ 'data repository',
                     .default = 'primary literature'
                     
           ))
```

# Scale of data rescued

```{r}
fileStats <- fileGitTouches |>
  filter(filenames != 'ReadMe_DataRescue.qmd',
         !str_detect(filenames, '^_')) |>
  select(filenames_str = filenames, studyID) |>
  unique() |>
  mutate(lineCount = length(read_lines(filenames_str)),
         fileSize = file.size(filenames_str),
         .by = filenames_str) |>
  mutate(file_type =
           case_when(str_detect(filenames_str, '[fF]igure.*\\.csv$') ~ 'figure',
                     str_detect(filenames_str, '[tT]able.*\\.csv$') ~ 'table',
                     str_detect(filenames_str, '\\.bib$') ~ 'bibliography',
                     str_detect(filenames_str, '\\.md$') ~ 'text',
                     str_detect(filenames_str, '\\.qmd$') ~ 'documentation',
                     .default = 'Unknown'),
         .by = filenames_str) |>
  reframe(lines = sum(lineCount),
          files = n(),
          size_kB = sum(fileSize) / 2^10,
          .by = c(studyID, file_type))

studyStats <- fileStats |>
  filter(file_type %in% c('figure', 'table', 'bibliography', 'text')) |>
  reframe(lines = sum(lines),
          files = sum(files),
          size_kB = sum(size_kB),
          .by = studyID)

typeStats <- fileStats |>
  filter(file_type %in% c('figure', 'table', 'bibliography', 'text')) |>
  reframe(lines = sum(lines),
          files = sum(files),
          size_kB = sum(size_kB),
          .by = file_type)
```

```{r}
plot.df <- studyTypes.df |>
  full_join(fileStats) |>
  filter(file_type %in% c('figure', 'table', 'bibliography', 'text')) |>
  filter(sourceType == "primary literature") |>
  pivot_longer(cols = c(lines, files, size_kB),
               names_to = 'sizeType') 
  
ggplot(plot.df) +
  geom_histogram(aes(x=value)) +
  facet_grid(file_type~sizeType, scale='free') +
  labs(title = 'Scale of a data rescue element')
```

```{r}
plot.df <- studyTypes.df |>
  full_join(studyStats) |>
  pivot_longer(cols = c(lines, files, size_kB),
               names_to = 'sizeType') |>
  filter(sourceType == "primary literature")

ggplot(plot.df) +
  geom_histogram(aes(x=value)) +
  facet_wrap(~sizeType, scale='free') +
  labs(title = 'Scale of a data rescue')
```

Overall, there are `r sum(studyStats$size_kB)` kB of data currently rescued across `r nrow(studyStats)` data rescues.

# When and how long does a data rescue take?

```{r}
 
startProjects <- fileGitTouches |>
  select(timestamp_date, studyID) |>
  unique() |>
  reframe(firstCommit = min(timestamp_date),
          lastCommit = max(timestamp_date),
         .by = studyID) |>
  arrange(firstCommit, studyID) |>
  mutate(startingRank = 1:n()) 

```


```{r}
author.df <- plyr::ldply(allYAML, .id = 'studyID',
                         .fun = function(yamlHeader){
 #yamlHeader <- allYAML[[27]]
  #TODO Standardize this across the project
  if('authors' %in% names(yamlHeader)){
    author_str <-'authors'
  }else if('author' %in% names(yamlHeader)){
    author_str <-'author'
  }else{
    stop('bad author key')
  }
  
  ans <- as.data.frame(yamlHeader[[author_str]]) |>
    select(starts_with('id'),
           starts_with('name.given'), starts_with('name.family')) |>
    unique() |>
    pivot_longer(cols = everything()) |>
    mutate(author_order = str_extract(name, '\\d+$')) |>
    mutate(author_order = if_else(is.na(author_order), '0', author_order),
           name = str_remove(name, '\\.\\d+$')) |>
    pivot_wider(names_from = 'name', values_from = 'value') |>
     select(author_order, author_id = id, `name.family`, `name.given`) #|>
    # mutate(studyID = str_extract(tempFilename, pattern = '^[[:alnum:]_]+'))
  return(ans)
}) |>
  mutate(student_status = case_when(
    name.family %in% c("Marupaka", "Whitehead", "Bonner") ~ 'PhD Student',
    name.family %in% c("Green", "Scott", "Blanchard", "Kalandjian", "Nguyen", "Walder", "Saxour", "Velazquez", "Anegondi") ~ 'BS Student',
    .default = 'Other'
  ), 
  UG_cohort = case_when(
    name.family %in% c("Green", "Scott", "Blanchard", "Kalandjian", "Nguyen") ~ ym('2025-1'),
    name.family %in% c("Saxour", "Walder") ~ ym('2025-09'),
    name.family %in% c("Green", "Velazquez", "Anegondi") ~ ym('2026-1'),
    .default = NA
  ))

```

```{r}

loggedTime.df <- plyr::ldply(allYAML, function(list_item){
  plyr::ldply(list_item[['timelog']], function(activity_item){
  return(as.data.frame(activity_item$activity))
}) }, .id = 'studyID') |>
  mutate(duration = lubridate::duration(duration)) 

loggedTime.df |>
  reframe(totalTime = sum(duration)/(60*60),
          description = paste(description, collapse = '; '),
          .by = studyID) |>
  mutate(level1_included = str_detect(description, 'T|transcrip')) |>
  arrange(desc(level1_included), totalTime)
```


```{r}
plot.df <- author.df |>
  left_join(startProjects, by = join_by(studyID)) |>
  filter(author_id != 'ktoddbrown',
         firstCommit > ymd('2025-01-01')) |>
  arrange(name.family, firstCommit) |>
  mutate(cumulative_count = 1:n(),
         time_since_first = firstCommit - min(firstCommit),
         .by = name.family)

ggplot(plot.df |>
         filter(firstCommit > min(firstCommit) ), 
       aes(x=difftime(firstCommit, UG_cohort)/365,
                y = cumulative_count, color = name.family)) +
  guides(color = 'none') +
  geom_point() +
  geom_line() +
  labs(x= 'Years project start', y = 'Cumulative study count',
       title = 'Undergraduate study submits over time')

ggplot(plot.df) +
  geom_point(aes(x=time_since_first/(60*60*24),
                y = cumulative_count)) +
  geom_line(aes(x=time_since_first/(60*60*24),
                y = cumulative_count, color = name.family)) +
  guides(color = 'none') +
  labs(x= 'Days since first commit', y = 'Cumulative study count',
       title = 'Data rescuer study submits over time')

ggplot(plot.df |>
         filter(cumulative_count == max(cumulative_count),
                .by = name.family)) +
  geom_histogram(aes(x =  cumulative_count), binwidth = 1) +
  labs(x = 'Cumulative study count', title = 'Data rescuer productivity')

ggplot(plot.df |>
         filter(cumulative_count == max(cumulative_count),
                .by = name.family),
       mapping = aes(x = time_since_first/(60*60*24*365), y = cumulative_count,
                     color = student_status)) +
  geom_jitter() +
  labs(x= 'Years since first commit', y = 'Cumulative study count',
       title = 'Total data rescuer productivity')

# lm(data = plot.df |>
#          filter(cumulative_count == max(cumulative_count),
#                 .by = name.family),
#   formula =  cumulative_count ~ I(time_since_first/(60*60*24*365)) ) |>
#   summary()
```

```{r}
ggplot(fileGitTouches |>
         full_join(studyTypes.df,
                   by = join_by(studyID)) |>
         full_join(startProjects,
                   by = join_by(studyID)) |>
         filter(studyID != 'ReadMe_DataRescue.qmd')) +
  geom_line(aes(y=startingRank, x = timestamp_date, group = studyID, color = sourceType)) +
  geom_point(aes(y = startingRank, x = timestamp_date, color = sourceType)) +
  geom_line(aes(x=firstCommit, y = startingRank), linetype = 3) +
  labs(x = "Date",
       y = 'Cumulative study count',
       title = 'Data rescue study commits (any change)')
```
