---
title: "Data Rescue: Binkley 2004"
format:
  html:
    toc: true
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - Binkley2004_Methods.bib
  - Binkley2004.bib
author:
  - name:
      given: Katherine
      family: Todd-Brown
      orcid: 0000-0002-3109-8130
      note: https://github.com/ktoddbrown
      affiliations:
        - id: ESS UF
          name: University of Florida
          department: Environmental Engineering Sciences
          city: Gainesville
          state: FL
          country: USA
      role: 
        - "Transcription"
        - "Review"
  - name:
      given: Savannah
      family: Scott
      affiliation:
        - id: UF
          name: University of Florida
          city: Gainesville
          state: FL
          country: USA
      role: "Transcription"
      note: https://github.com/SavaScott
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(kableExtra)
library(bibtex)

methods.file <- 'Binkley2004_Methods.md'
figure1.file <- 'Binkley2004_Figure1.csv'
primaryCitation.file <- 'Binkley2004.bib'
methodsCitation.file <- 'Binkley2004_Methods.bib'
```

This documents the data transcription for @Binkley2004 .
For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/76.

> Binkley, D., Kaye, J., Barry, M. and Ryan, M.G. (2004), First-Rotation Changes in Soil Carbon and Nitrogen in a Eucalyptus Plantation in Hawaii. Soil Sci. Soc. Am. J., 68: 1713-1719. https://doi.org/10.2136/sssaj2004.1713


# Data Summary

@Binkley2004 looks at the changes in soil carbon and nitrogen throughout the full rotation of a Eucalyptus saligna plantation (throughout 8 years). This helped to assess the impact of the change in a natural forest and a plantation on the carbon stock of the soil.

## Fit for purpose: HiCSC

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the HiCSC.

- Location: Given in site description however it's unclear if the lat/long is Hilo or study site.
- Soil carbon stock: Soil carbon fraction and bulk density were measured. Sample declared rock free so no course fraction corrected applied.
- Soil type: Soil type is identified in the methods
- Land use: Detailed land use history provided

# Tables and figures

- Table - No tables

- Figure 
  1) Soil content of C, C derived from C4 plants, C derived from C3 plants and Nitrogen over 8 years
  2) CO2 incubation data
  3) Cumulative soil carbon loss
  
### Figure 1

Figure 1 was captured using [PlotDigitizer](https://plotdigitizer.com/app) and then truncated to represent realistic precision.

```{r figure1}
#| code-fold: true
#| message: false

filename <- 'Binkley2004_Figure1.csv'
plot.df <- read_csv(file = filename,
                   skip = 1) |>
  rename(Layer = `type label`)
caption <- read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption

ggplot(data = plot.df) +
  geom_line(aes(x=x, y=y, color = Layer)) +
  geom_point(aes(x=x, y=y, color = Layer, shape = Layer)) +
  facet_wrap(~`y label`, scales = 'free_y') +
  labs(x='Year', y = '')
```


{{< include Binkley2004_Methods.md >}}

# Transcription comparison

```{bash eval=FALSE}
diff --strip-trailing-cr Binkley2004_Methods_KTB.md Binkley2004_Methods_SRS.md
```

# Files

These files are in the @Binkley2004 data rescue.

- [Readme](README_Binkley2004.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Binkley2004.bib](Binkley2004.bib)
  + citation for article transcribed
- [Binkley2004_Methods.bib](Binkley2004_Methods.bib)
  + citations for the methods section of the article
- [Binkley2004_Methods.Rmd](Binkley2004_Methods.Rmd)
  + methods transcribed from primary article
- [Binkley2004_Figure1.csv](Binkley2004_Figure1.csv)
  + Figure 1 shows the soil content of C, C derive from C$_4$ plants, C derived from C$_3$ and N. 

# Working notes 

## Data Rescue Plan

- [x] describe files in human readable form inside ReadMe
- [x] description with Git issue with link
- [x] add contribution yaml to top of this doc
- [x] use [PlotDigitizer](https://plotdigitizer.com/app) to extract points from Figure 1
- [x] export excel table to csv and reconstruct plot
- [x] copy over methods section into markdown
- [x] pull down citations in methods section to bib file
- [x] add in citation links to methods
- [x] cross check with second transcriber for tables and methods
- [ ] submit to independent review
- [ ] archive on SoilDRaH


## Citation notes from Methods

Below are the citation notes from the methods section.
Some citations are missing, those are noted here and how best guess were generated.

- [ ] short citation
  + copy from bib
  + manually entered or direct export
- [x] 4 -Binkley, D. Rapid changes in soils following Eucalyptus afforestation in Hawaii. Soil Sci. Soc. Am. J.1999 63 222225https://doi.org/10.2136/sssaj1999.03615995006300010032xhttp://gateway.isiknowledge.com/gateway/Gateway.cgi?GWVersion=2&SrcApp=PARTNER_APP&SrcAuth=ag
  + exact export
- [x] 10 - Giardina, C.P. Total belowground carbon allocation in a fast growing Eucalyptus plantation estimated using a carbon balance approach. Ecosystems2002 5 487499https://doi.org/10.1007/s10021-002-0130-8
  + exact export
- [x] 1 - Barnard, H. A test of the hydraulic limitation hypothesis in fast-growing Eucalyptus saligna. Plant Cell Environ.2003 26 12351245https://doi.org/10.1046/j.1365-3040.2003.01046.x
  + exact export
- [x] 26 - Ryan, M.G., D. Binkley, J. Fownes, C. Giardina, and R. Senock. 2004. An experimental test of the causes of age-related decline in forest growth. Ecol. Monogr. (in press).
  + exact export
- [x] 11 - Goncalves, J.L.M., N.F. Barros, E.K.S. Nambiar, and R.F. Novais. 1997. Soil and stand management for short-rotation plantations. p. 379–417. In E.K.S. Nambiar and A. Brown (ed.) Management of soil, nutrients, and water in tropical plantation forests. ACIAR Monograph #43, Canberra.
  + exact export
- [x] 28 - Vitorello, V.A.Organic matter and natural carbon-13 distribution in forested and cultivated Oxisols. Soil Sci. Soc. Am. J.1989 53 773778https://doi.org/10.2136/sssaj1989.03615995005300030024xhttp://gateway.isiknowledge.com/gateway/Gateway.cgi?GWVersion=2&SrcApp=PARTNER_APP&SrcAuth=a
  + exact export
- [x] 2 - Bashkin, M.Changes in soil carbon following afforestation in Hawaii. Ecology1998 79 828833https://doi.org/10.1890/0012-9658(1998)079[0828:CISCFA]2.0.CO;2http://gateway.isiknowledge.com/gateway/Gateway.cgi?GWVersion=2&SrcApp=PARTNER_APP&SrcAuth=a
  + exact export

# Level 0: raw data reads
```{r readLevel0}

data.lvl0.df <- list(citation = 
                    list(primary = 
                           read.bib(file = primaryCitation.file), 
                         methods = 
                           read.bib(file = methodsCitation.file)
                    ),
                  method = read_lines(file = methods.file),
                  data = list(figure1 = 
                                list(caption = 
                                       read_csv(file = figure1.file,
                                                col_types = cols(.default = col_character()),
                                                n_max = 1, col_names = FALSE)$X1[1],
                                     primary = 
                                       read_csv(file = figure1.file,
                                                col_types = cols(.default = col_character()),
                                                skip = 1)
                                )))
```
