---
title: "Data Rescue: Nusslein 1999"
format:
  html:
    toc: true
    eval: false #default to not evaluating the chunks, override this the template is completed
date: 2026-001-22
date-format: YYYY-MMMM
bibliography:
  - Nusslein1999_Methods.bib #contains the citations in the methods section
  - Nusslein1999.bib #contains citation for the data source, generally a single citation of the primary paper
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  - id: ktoddbrown #use the github name as an id
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role: #TODO define these roles and add reference
      - transcription: secondary, review
      - standardization: #lead #review
      - curation: #lead #review
  - id: misheph
    name:
      given: Mirah
      family: Blanchard
    orcid: N/A
    affiliation:
      - ref: uf-mae
    role: #TODO define these roles and add reference
      - transcription: lead 
      - standardization: #lead #review
      - curation: #lead #review

affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
    
timelog:
  - activity: 
      description: Revise based on feedback
      who: @misheph #github username
      #when you started and ended, can be just a year or down to the minute
      end: 2025-01-22
      start: 2025-11
      #alternatively duration can be combined with a being/end
      duration: P30M
version: 'vs202512' #template version, remove before use
---


```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false


library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# data files for level 0
methods.file <- 'Nusslein1999_Methods.md'
table1.file <- 'Nusslein1999_Table1.csv'

# Bibliograph files
primaryCitation.file <- 'Nusslein1999.bib'
methodsCitation.file <- 'Nusslein1999_Methods.bib'

# read function
readsource.file <- '../../../R/readNusslein1999.R'
source('readsource.file')

#semantic files for level 1
of_variable.file <- '../../semantics/SoilDRaHVocabulary_of_variable.csv'
is_type.file <- '../../semantics/SoilDRaHVocabulary_is_type.csv'

```


# Data Summary

> Nüsslein K, Tiedje JM1999.Soil Bacterial Community Shift Correlated with Change from Forest to Pasture Vegetation in a Tropical Soil. Appl Environ Microbiol65:.https://doi.org/10.1128/AEM.65.8.3622-3626.1999

For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/105.

This paper looks at the shift in the soil bacterial community associated with the vegetative change when going from forest to pasture soil. Researchers extracted soil samples from a forest site and a pasture site adjacent to it during three different years, and soil DNA was extracted to analyze differences in G+C content, rRNA copied from soil DNA, and phylogenetics of the most common bacteria. From this, the researchers determined there were distinct shifts in dominant bacterial groups going from forest to pasture soil, due to vegetation changes.
measured and what the data was used for in. This should be <250 words.--->

## Tables and figures

- Table 
  1)  forest vs pasture soil properties (pH, total C, N, CEC, bulk density)
  2)  dominant SSU rDNA phylotypes
  3)  sequence similarities among dominant SSU rDNA clones

- Figure 
  1)  forest vs pasture G+C content profiles
  2)  SSU rDNA phylotype frequency distribution
  3)  pasture rDNA clone ARDRA profiles

## Fit for purpose: HiCSC

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the HiCSC.

- Location: Latitude and longitude for forest site given in methods section; pasture site coordinates not given but described as “immediately adjacent” to forest. Observation years and depth from the surface also given in method section.
- Soil carbon stock: Soil bulk density given in table 1. Soil organic carbon not given specifically, but total carbon % in table 1.
- Soil type: Soil type (Typic Placandept, suborder) given in method section.
- Land use: Current land use and site history detailed in method section for both sites.
  
## Files

These files are in the Nusslein1999 data rescue.

- [Readme](README_Nusslein1999.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Nusslein1999.bib](Nusslein1999.bib)
  + citation for article transcribed
- [Nusslein1999_Methods.bib](Nusslein1999_Methods.bib)
  + citations for the methods section of the article
- [Nusslein1999_Methods.md](Nusslein1999_Methods.md)
  + methods transcribed from primary article
- [Nusslein1999_Table1.csv](Nusslein1999_Table1.csv)
  + table 1 from primary article with forest vs pasture soil properties
- [temp/](temp/)
    + scratch folder that will not be archived on GitHub, it should include local copies of the data resource (likely a PDF article) and any spreadsheets or other tools you used for the transcriptions.

# Data Rescue working notes 

## Data Rescue Plan

+ Lead
  - [x] Read the article or data resource and the discription of the target project
  - [x] Create a fresh fork of the primary repository https://github.com/ktoddbrown/SoilDRaH/tree/main, it might be useful to rename it 'SoilDRaH_AuthorYYYY'
  - [x] Create a copy of this template folder and rename it to Nusslein1999
  - [x] Fill out Data summary section (except for the Files), the Data Rescue Plan (here!), and update the author section of the yml to reflect who is working on the rescue
  - [x] Create a GitHub issue from the Data Rescue template and get feedback from a reviewer on the project (Dr Todd-Brown or a grad student typically).
  - [x] Add the reviewer to the author list in the yml
+ Do
  - [x] Create a spreadsheet with Table1 and export it to csv
  - [x] Transcribe or copy the method section
  - [x] Insert references to bib entries
  - [x] reformat equations and mathematical symbols as [LaTex](https://en.wikibooks.org/wiki/LaTeX/Mathematics)
  - [x] reformat text using [Markdown](https://www.markdownguide.org/cheat-sheet/)
  - [x] Create a bib file with the citations from the methods section
  - [x] Push to your local fork and get a second transcription pushed to your local fork by someone else on the project
+ Measure
  - [x] cross check with second transcriber for tables and methods, and resolve any differences
  - [x] compile ReadMe with all chunks up through the Level 0 data rescue section with `eval: true`
  - [x] push to your local fork and create a pull request to the main branch of the primary repository
  - [ ] address any revisions suggested in your pull request review
  - [ ] celebrate!

## Table 1

Table 1 was modified by moving table notes to same line as the heading.


```{r table1}
#| code-fold: true
#| message: false

read_csv(file = table1.file,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = table1.file, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```

## Methods comparison

```{bash eval=FALSE}
diff --strip-trailing-cr --suppress-common-lines -y Nusslein1999_Methods_KTB.md Nusslein1999_Methods_MB.md
```

## Methods Section Transcription

{{< include Nusslein1999_Methods.md >}}

## Citation notes from Methods

Below are the citation notes from the methods section.
Some citations are missing, those are noted here and how best guess were generated.

- [x] Crews et al., 1995
  + Crews, T. E., K. Kitayama, J. H. Fownes, R. H. Riley, D. A. Herbert, D. Mueller-Dombois, and P. M. Vitousek. 1995. Changes in soil phosphorous fractions and ecosystem dynamics across a long chronosequence in Hawaii. Ecology 76:1407±1424.
  + direct export
- [x] Department of Geography, University of Hawaii, 1983
  + Department of Geography, University of Hawaii. Atlas of Hawaii. 1983. University of Hawaii Press, Honolulu, Hawaii.
  + direct export
- [x] Giambelluca et al., 1986
  + Giambelluca, T. W., M. A. Nullet, and T. A. Schroeder. 1986. Rainfall atlas of Hawaii. Department of Land and Natural Resources, Honolulu, Hawaii.
  + direct export
- [x] Riley and Vitousek, 1995
  + Riley, R. H., and P. M. Vitousek. 1995. Nutrient dynamics and nitrogen trace gas flux during ecosystem development in montane rain forest. Ecology 76:292±304.
  + direct export
- [x] Huston 1994
  + Huston, M. A. 1994. Biological diversity. Cambridge University Press, Cambridge, England.
  + direct export
- [x] Peck et al., 1988
  + Peck, T. R., D. Megel, K. Eik, D. A. Whitney, D. Warncke, R. C. Munter, J. R. Brown, D. Knudsen, W. C. Dahnke, D. J. Eckert, D. Beegle, P. E. Fixen, and E. E. Schulte. 1988. Recommended chemical soil test procedure for the north central region. North Dakota Agricultural Experiment Station, North Dakota State University, Fargo, N. Dak.
  + direct export
- [x] Holben 1994
  + Holben, W. E. 1994. Isolation and puri®cation of bacterial DNA from soil, p. 727±751. In R. Weaver et al. (ed.), Methods of soil analysis, part 2. Microbiological and biochemical properties. Soil Science Society of America book series 5. Soil Science Society of America, Madison, Wis.
  + direct export
- [x] Maidak et al., 1999
  + Maidak, B. L., J. R. Cole, C. T. Parker, Jr., G. M. Garrity, N. Larsen, B. Li, T. G. Lilburn, M. J. McCaughey, G. J. Olsen, R. Overbeek, S. Pramanik, T. M. Schmidt, J. M. Tiedje, and C. R. Woese. 1999. A new version of the RDP (Ribosomal Database Project). Nucleic Acids Res. 27:171±173.
  + direct export
- [x] Holben and Harris 1995
  +  Holben, W. E., and D. Harris. 1995. DNA-based monitoring of total bacterial community structure in environmental samples. Mol. Ecol. 4:627±631.
  + direct export
- [x] Nusslein and Tiedje 1998
  + NuÈsslein, K., and J. M. Tiedje. 1998. Characterization of the dominant and rare members of a young Hawaiian soil bacterial community with small subunit ribosomal DNA ampli®ed from DNA fractionated on the basis of its guanine and cytosine composition. Appl. Environ. Microbiol. 64:1283±1289.
  + direct export
- [x] Weisburg et al., 1991
  + Weisburg, W. W., S. M. Barns, D. A. Pelletier, and D. J. Lane. 1991. 16S ribosomal DNA ampli®cation for phylogenetic study. J. Bacteriol. 173:697±703.
  + direct export
- [x] Zhou et al. 1997
  + Zhou, J., M. E. Davey, J. B. Figueras, E. Rivkina, D. Gilichinsky, and J. M. Tiedje. 1997. Phylogenetic diversity of a bacterial community determined from Siberian tundra soil DNA. Microbiology 143:3913±3919.
  + direct export
- [x] Komatsoulis and Waterman 1997
  + Komatsoulis, G. A., and M. S. Waterman. 1997. A new computational method for detection of chimeric 16S rRNA artifacts generated by PCR amplification from mixed bacterial populations. Appl. Environ. Microbiol. 63:2338±2346.
  + direct export

# Level 0 Data rescue

This level 0 data read parses the transcribed files with minimal cleaning or transformations.

```{r readLevel0}
#This chunk has two purposes. 
#...1) Check the formatting by reading in everything 
#...2) Create a list of everything to process later in level 1

data.lvl0.df <- list(
  #Read in a list of all the bib files
  citation = list(
    #Citation for the article transcriptions are pulled from
    primary = read.bib(file = primaryCitation.file), 
    #Citations for all referenced articles
    methods = read.bib(file = methodsCitation.file)
  ),
  #Read in the text transcription of the article's methods section
  method = read_lines(file = methods.file),
  #Read in the results as tables or figure transcriptions. This includes
  #...the caption as well as the tables themselves
  data = list(
    Table1 = list(
      #Read the caption as a text string. Captions are the first cell on 
      #...the first row.
      caption = read_csv(file = table1.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      #Read in all the data, skipping the first row with the caption and read
      #...in the table as character. This element is a tibble (data.frame).
      primary = read_csv(file = table1.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    ) #End 'Table1' element
    #Repeat for all tables and figures...
   ) #end 'data' element
)

```

```{r}
#| label: checkReadFunction_lvl0

datafunction.lvl0.ls <- readNusslein1999(dataDir = '.',
                                    dataLevel = 'level0')

if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
  print('Read function matches code here.')
}else{
  print('There is a mismatch in the data objects between the code here and the read function.')
}

```


# References

<!---- This space will be filled with the citations when rendered. Leave this blank. ---->

