---
title: "Data Rescue: Ares 2001"
format:
  html:
    toc: true
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - Ares2001_Methods.bib
  - Ares2001.bib
author:
  - name:
      given: Katherine
      family: Todd-Brown
      orcid: 0000-0002-3109-8130
      note: https://github.com/ktoddbrown
      affiliations:
        - id: ESS UF
          name: University of Florida
          department: Environmental Engineering Sciences
          city: Gainesville
          state: FL
          country: USA
      role: 
        -"Transcription"
        -"Review"
        -"Standardization"
  - name:
      given: Colby
      family: Green
      affiliation:
        - id: UF
          name: University of Florida
          city: Gainesville
          state: FL
          country: USA
      role: "Transcription"
      note: https://github.com/colbyGreen1520
  - name:
      given: Savannah
      family: Scott
      affiliation:
        - id: UF
          name: University of Florida
          city: Gainesville
          state: FL
          country: USA
      role: "Transcription"
      note: https://github.com/SavaScott
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(kableExtra)
library(bibtex)

methods.file <- 'Ares2001_Methods.md'
table1.file <- 'Ares2001_Table1.csv'
table3.file <- 'Ares2001_Table3.csv'
primaryCitation.file <- 'Ares2001.bib'
methodsCitation.file <- 'Ares2001_Methods.bib'
```

This documents the data transcription for @Ares2001 .
For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/77 .

> A. Ares and J. H. Fownes. Productivity, resource use, and competitive interactions of fraxinus uhdei in hawaii uplands. Canadian Journal of Forest Research, 31(1):132-142, [https://doi.org/10.1139/x00-156](https://doi.org/10.1139/x00-156). 2001 

# Data Summary

@Ares2001 primarily measures tree growth parameters (including litterfall) but also characterized elemental fractions of soil. They use this data to look at factors affecting a specific invasive tree growth in Hawaii including nitrogen and moisture gradients.

## Fit for purpose: HiCSC

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the HiCSC.

- Location: Location region is identified and lat/lon can likely be recovered from figure 1 by regional expert.
- Soil carbon stock: Soil carbon fraction is given but not bulk density nor coarse fraction.
- Soil type: Soil order and history are identified in methods. High/low activity clay status may need inference by regional expert.
- Land use: Land use and history are clearly identified in methods.

# Tables and figures

- Table 
  1) characterists of stands including oc percent
  2) regression coefficents linking diameter at breast height to biomass and leaf area
  3) climate
  4) elevation vs productivity metrics
  5) fertilizer treatement vs productivity over time
  6) fertilizer vs leaf area index
  7) gradient vs plant efficency metrics

- Figure 
  1) stand map overlayed on soil.
  2) elevation vs productivity metrics
  3) nitrogen correlations
  4) light intercept vs nitrogen and water use effiency metrics
  
### Table 1: Stand characteristics

Table 1 was modified to remove the sub-tables and flatted sub-tables in certain columns.

```{r table1}
#| code-fold: true
#| message: false

filename <- table1.file
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```

### Table 3: Climate

```{r table3}
#| code-fold: true
#| message: false

filename <- table3.file
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```


{{< include Ares2001_Methods.md >}}

# Transcription comparison

```{bash eval=FALSE}
diff Ares2001_Methods_CAG.md Ares2001_Methods.md
```

# Files

These files are in the Ares2001 data rescue.

- [Readme](README_Ares2001.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Ares2001.bib](Ares2001.bib)
  + citation for article transcribed
- [Ares2001_Methods.bib](Ares2001_Methods.bib)
  + citations for the methods section of the article
- [Ares2001_Methods.Rmd](Ares2001_Methods.Rmd)
  + methods transcribed from primary article
- [Ares2001_Table1.csv](Ares2001_Table1.csv)
  + table 1 from primary article with site descriptions
- [Ares2001_Table3.csv](Ares2001_Table3.csv)
  + table 3 from primary article with climate variables


# Working notes (KTB)

## Data Rescue Plan

- [x] describe files in human readable form inside ReadMe
- [x] description with Git issue with link
- [x] add contribution yaml
- [x] create excel sheet with Table 1 and 3
- [x] export excel table to csv
- [x] copy over methods section into markdown 
  - started, PDF doesn't copy but there is an html version which is incomplete and error ridden. Ended up with alot of retyping.
- [x] pull down citations in methods section to bib file
- [x] add in citation links to methods
- [x] cross check with second transcriber for tables and methods
- [ ] submit to independent review
- [ ] archive on SoilDRaH


## Citation notes from Methods

Below are the citation notes from the methods section.
Some citations are missing, those are noted here and how best guess were generated.

- [x] Sato et al. 1973
  + Sato, H.H., Ikeda, W., Paeth, R., Smythe, R., and Takehiro, M. 1973. Soil survey of the island of Hawaii, state of Hawaii. USDA Soil Conservation Service, University of Hawaii Agricultural Experiment Station, U.S. Government Printing Office, Washington D.C.
  + manually entered 
- [x] Harrington and Fownes (1993)
  + Harrington, R.A., and Fownes, J.H. 1993. Allometry and growth of planted versus coppice stands of four fast-growing tropical tree species. For. Ecol. Manage. 56: 315-327.
  + direct export
- [x] Sprugel 1983
  + Sprugel, D.G. 1983. Correcting for bias in log-transformed allometric equations. Ecology, 64: 209-210.
  + direct export
- [x] Avery 1975
  + Avery, T.E. 1975. Natural resources measurements. McGraw-Hill Inc., New York.
  + Manual add as a @book
- [x] Welles and Norman 1991
  + Welles, J.M., and Norman, J.M. 1991. Instrument for indirect measurement of canopy architecture. Agron. J. 83: 818-825.
  + direct export
- [x] Nelson and Sommers ~1972~ 1973
  + Nelson, D.W., and Sommers, L.E. 1972. Determination of total nitrogen in plant material. Agron. J. 65: 109-112.
  + correction to the year, 1973 instead
- [x] Isaac and Johnson 1983
  + Isaac, R.A., and Johnson, W.C. 1983. High speed analysis of agriculture samples using inductively coupled plasma-atomic emission spectroscopy. Spectrochim. Acta, 38B: 277-282.
  + direct export
- [x] Wolf 1974
  + Wolf, B. 1974. Improvements in the azomethine-H-method for the determination of boron. Comm. Soil Sci. Plant Anal. 5: 39-44.
  + direct export
- [x] Vitousek and Sanford 1986
  + Vitousek, PM., and Sanford, R.L. 1986. Nutrient cycling in moist tropical forests. Annu. Rev. Ecol. Syst. 17: 137-167.
  + direct export
  + correct number citation
- [x] Heanes 1984
  + Heanes, D. L. 1984. Determination of total organic C in soils by improved chromic acid digestion and spectrophotometric procedure. Commun. Soil Sci. Plant Anal. 15: 1191-1213.
  + direct export

# Level 0 data read

```{r readLevel0}

data.lvl0.df <- list(citation = 
                    list(primary = 
                           read.bib(file = primaryCitation.file), 
                         methods = 
                           read.bib(file = methodsCitation.file)
                    ),
                  method = read_lines(file = methods.file),
                  data = list(Table1 = 
                                list(caption = 
                                       read_csv(file = table1.file,
                                                col_types = cols(.default = col_character()),
                                                n_max = 1, col_names = FALSE)$X1[1],
                                     primary = 
                                       read_csv(file = table1.file,
                                                col_types = cols(.default = col_character()),
                                                skip = 1)
                                ), 
                               Table3 = 
                                list(caption = 
                                       read_csv(file = table3.file,
                                                col_types = cols(.default = col_character()),
                                                n_max = 1, col_names = FALSE)$X1[1],
                                     primary = 
                                       read_csv(file = table3.file,
                                                col_types = cols(.default = col_character()),
                                                skip = 1)
                                )))
```

# Level 1 data - Hawaii SOC DB

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the HiCSC.

- Location: Location region is identified and lat/lon can likely be recovered from figure 1 by regional expert.
- Soil carbon stock: Soil carbon fraction is given but not bulk density nor coarse fraction.
- Soil type: Soil order and history are identified in methods. High/low activity clay status may need inference by regional expert.
- Land use: Land use and history are clearly identified in methods.

```{r}

#Pull in the soil type and land use from the methods section
#Pull in the soc stock from one of the tables(?)
#Check for geolocation and observation date somewhere???

data.df <- tribble(~of_variable, ~is_type, ~with_entry, ~source,
                   'region', 'value', 'Honaunau Forest on the southwestern slopes of Mauna Loa, island of Hawaii', paste('Method ln5:', paste(data.lvl0.df$method[5], collapse = ' ')),
                   'land_use_type', 'value', 'Tree stands', paste('Method ln14-16:', paste(data.lvl0.df$method[14:16], collapse = ' ')),
                   'inital_planting', 'value', '1959', paste('Method ln14:', data.lvl0.df$method[14]),
                   'observation_year', 'value', '1996', paste('Method ln 30;53;58:', paste0(data.lvl0.df$method[c(30,53,58)], collapse = '... ')),
                   'citation', 'value', format(data.lvl0.df$citation$primary), 'journal citation',
                   'doi', 'value', data.lvl0.df$citation$primary$doi, 'journal citation')

temp1 <- data.lvl0.df$data$Table1$primary |>
  mutate(row_id = paste0('R', 1:n())) |>
  pivot_longer(cols = -row_id,
               names_to = 'column_name', values_to = 'with_entry',
               values_drop_na = TRUE) |>
  mutate(of_variable = case_when(
           column_name == "Stand type" ~ 'stand_type',
           column_name == 'Elevation (m)' ~ 'elevation',
           column_name == 'Soil type' ~ 'soil_class',
           column_name == 'Soil pH' ~ 'soil_ph',
           column_name == 'Soil organic carbon (%)' ~ 'soil_organic_carbon',
           column_name == 'Soil N (%)' ~ 'soil_nitrogen',
           column_name == 'Soil P (mg kg<sup>-1</sup>)' ~ 'soil_phosphorus',
           column_name == 'Stand age (years)' ~ 'stand_age',
           column_name == '*F. uhdei* Stem density (trees / ha)' ~ 'stem_density_F_uhdei',
           column_name == '*F. uhdei* Mean DBH (cm)' ~ 'diameter_at_breast_height_F_uhdei',
           column_name == '*F. uhdei* Mean height(m)' ~ 'height_F_uhdei',
           column_name == '*A. koa* Stem density (trees / ha)' ~ 'stem_density_A_koa',
           column_name == '*A. koa* Mean DBH (cm)' ~ 'diameter_at_breast_height_A_koa',
           column_name == "*A. koa* Mean height(m)"~ 'height_A_koa')) |>
  mutate(source = 'Table 1') 

temp1.meta <-
  temp1 |>
  select(column_name, of_variable) |>
  unique() |>
  mutate(unit = str_extract(column_name, pattern = '(?<=\\().*(?=\\))'),
         source = 'Table 1 column names.') |>
  filter(!is.na(unit)) |>
  select(-column_name) |>
  bind_rows(
  tribble(~of_variable, ~method, ~source,
            'soil_ph', paste0(data.lvl0.df$method[72:73], collapse = ' '), 'Methods ln72-73',
            'soil_organic_carbon', paste0(data.lvl0.df$method[72:73], collapse = ' '), 'Methods ln72-73',
            'soil_nitrogen', paste0(data.lvl0.df$method[72:73], collapse = ' '), 'Methods ln72-73',
            'soil_phosphorus', paste0(data.lvl0.df$method[72:73], collapse = ' '), 'Methods ln72-73',
            'soil_class', paste0(data.lvl0.df$method[10:12], collapse = ' '), 'Methods ln10-12') ) |>
  bind_rows(
    tribble(~of_variable, ~control_vocabulary, ~source,
            'stand_type', '*F. uhdei*: pure stands of Fraxinus uhdei (Wenzig) Lingelsh|Mixed: mixed stands of *Fraxinus uhdei* (Wenzig) Lingelsh and *Acacia koa* Grey', 'Abstract ln1',
            'soil_class', 'Histosol:USDA classification for histosol soil type|Andisols:USDA classification for andisol soil type', 'expert informed')
  ) |>
  pivot_longer(cols = c(unit, method, control_vocabulary),
               names_to = 'is_type',
               values_drop_na = TRUE,
               values_to = 'with_entry')


temp3 <- data.lvl0.df$data$Table3$primary |>
  pivot_longer(cols = -variable,
               names_to = 'row_id', values_to = 'with_entry',
               values_drop_na = TRUE) |>
  #mutate(`elevation_id` = str_extract(elevation_id, '\\d{3,} m')) |>
  pivot_wider(names_from = 'variable', values_from = 'with_entry') |>
  mutate(`Elevation (m)` = str_extract(row_id, '\\d{3,}')) |>
  pivot_longer(cols = -row_id, 
               names_to = 'column_name', values_to = 'with_entry') |>
  mutate(of_variable = case_when(
    column_name == "Mean air temperature (degree C)"~ 'mean_air_temperature',
    column_name == "Total rainfall (mm)" ~ 'total_rainfall',
    column_name == "Mean total solar radiation (MJ m<sup>-2</sup> day<sup>-1</sup>)" ~ 'mean_solar_radiation',
    column_name == "Mean maximum vapor pressure deficit (kPa)" ~ 'mean_max_vapor_pressure_deficit',
    column_name == "Elevation (m)" ~ 'elevation',
    TRUE ~ NA_character_)) |>
  mutate(source = 'Table 3')

temp3.meta <- temp3 |>
  select(column_name, of_variable) |>
  unique() |>
  mutate(unit = str_extract(column_name, pattern = '(?<=\\().*(?=\\))'),
         method = paste('Methods ln72-73:', paste0(data.lvl0.df$method[18:23], collapse = ' '))) |>
  select(-column_name) |>
  pivot_longer(cols = -of_variable,
               names_to = 'is_type', values_to = 'with_entry') |>
  mutate(source = case_when(is_type == 'unit' ~ 'Table 3 column names.',
                            is_type == 'method' ~ 'Methods ln72-73',
                            TRUE ~ NA_character_))


data.lvl1.ls <- list(
  meta = bind_rows(data.df,
                   temp1.meta, 
                   temp3.meta),
  primary = bind_rows(temp1, temp3)|>
    mutate(elevation_id = with_entry[of_variable == 'elevation'],
           is_type = 'value',
           .by = row_id) |>
    arrange(row_id, elevation_id, column_name,
            of_variable, is_type, with_entry, source))

```


# Level 2 data

```{r}

HISOC_variables <- c(
  "citation",
  "doi",
  "region", 
  "land_use_type",
  "observation_year",
  'elevation',
  'mean_air_temperature',
  'total_rainfall',
  "soil_organic_carbon",
  "soil_nitrogen",
  "soil_phosphorus",
  "soil_ph",
  "soil_class",
  "stand_age",
  'stand_type'
)

temp_study <- data.lvl1.ls$meta |>
  filter(of_variable %in% HISOC_variables) |>
select(-source)  |>
  unique()|>
  pivot_wider(names_from = c(of_variable, is_type), names_sep = '::', values_from = with_entry) |>
  mutate(study_id = 'Ares2001')
  
  
temp_plot <- data.lvl1.ls$primary|>
  filter(of_variable %in% HISOC_variables,
         str_detect(of_variable, 'soil') |
           str_detect(of_variable, 'stand'))|>
  select(-c(column_name, source)) |>
  pivot_wider(names_from = c(of_variable, is_type), names_sep = '::', values_from = with_entry) |>
  select(-row_id) |>
  mutate(study_id = 'Ares2001')

temp_evel <- data.lvl1.ls$primary|>
  filter(of_variable %in% HISOC_variables,
         !(str_detect(of_variable, 'soil') |
           str_detect(of_variable, 'stand')))|>
  select(-c(row_id, column_name, source)) |>
  unique() |>
  pivot_wider(names_from = c(of_variable, is_type), names_sep = '::', values_from = with_entry) |>
  mutate(study_id = 'Ares2001')

data.lvl2.df <- full_join(temp_study,
          temp_plot, 
          by = join_by(study_id)) |>
  full_join(temp_evel, 
            by = join_by(elevation_id, study_id)) |>
  mutate(row_id = paste0('ID', 1:n())) |>
  pivot_longer(cols = -c(study_id, elevation_id, row_id),
               values_to = 'with_entry',
               names_sep = '::',
               names_to = c('of_variable', 'is_type'),
               values_drop_na = TRUE) 
  

```

# References