---
title: "Annotations files"
author: "Kathe Todd-Brown"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Annotations files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

annotationsCheckFile <- 'selfDocumentAnnotations.csv'

annotationsDirectory = '../data'
  
selfDoc <- readr::read_delim(file.path(annotationsDirectory, 'selfDocumentAnnotations.csv'), 
                               delim = ';',
                               col_types = readr::cols(.default = readr::col_character()),)
  
  expectedHeaders <- selfDoc$column_id |>
    unique()
  
  data_ref <- '--' #string for reference to annotated data to appear in the `with_entry` column
  
  control_vocab.df <- selfDoc |>
    dplyr::filter(column_id == 'is_type', is_type == 'control_vocabulary') |>
    dplyr::select(with_entry) |>
    tidyr::separate_longer_delim(cols = with_entry, delim = ';') |>
    tidyr::separate_wider_delim(cols = with_entry, delim = '|', names = c('vocabulary', 'definition')) |>
    dplyr::mutate(dplyr::across(.cols = dplyr::everything(), .fns = trimws))
```

The annotations for ingested data files is expected to adhere to the following standards.

# Column names

Column names must contain the following columns and may contain additional columns as needed to provide unique identification.

```{r}
print(expectedHeaders)
```

# Deliminator

Annotations are expected to be `;` deliniated so that it will be comparable with R package data standards.

# Control vocabulary

Control vocabulary entries should be quoted as "term1|human readable definition;term2|second readable definition".
This is currently unenforced by `checkAnnotations`

# data reference

Reference to entries in the primary data set should be noted as `--`.
This is currently unenforced.

# `is_type`

Is type is expected to adhere to the following control vocabulary.

```{r}
knitr::kable(control_vocab.df |>
               dplyr::arrange(vocabulary))
```