---
title: "Data Rescue: Binkley 1999"
format:
  html:
    toc: true
date: 2025-06
date-format: YYYY-MMMM
bibliography:
  - Binkley1999_Methods.bib
  - Binkley1999.bib
author:
  - name:
      given: Katherine
      family: Todd-Brown
      orcid: 0000-0002-3109-8130
      note: https://github.com/ktoddbrown
      affiliations:
        - id: ESS UF
          name: University of Florida
          department: Environmental Engineering Sciences
          city: Gainesville
          state: FL
          country: USA
      role: 
        - "Transcription"
        - "Review"
  - name:
      given: Han
      family: Nguyen
      affiliation:
        - id: UF
          name: University of Florida
          city: Gainesville
          state: FL
          country: USA
      role: "Transcription"
      note: https://github.com/hannguyenuf
  - name:
      given: Savannah
      family: Scott
      affiliation:
        - id: UF
          name: University of Florida
          city: Gainesville
          state: FL
          country: USA
      role: "Transcription"
      note: https://github.com/SavaScott
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(kableExtra)
library(bibtex)

methods.file <- 'Binkley1999_Methods.md'
table1.file <- 'Binkley1999_Table1.csv'
primaryCitation.file <- 'Binkley1999.bib'
methodsCitation.file <- 'Binkley1999_Methods.bib'
```

This documents the data transcription for @Binkley1999 .
For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/76.

> Binkley, D. and Resh, S.C. (1999), Rapid Changes in Soils Following Eucalyptus Afforestation in Hawaii. Soil Science Society of America Journal, 63: 222-225. https://doi.org/10.2136/sssaj1999.03615995006300010032x

# Data Summary

@Binkley1999 looks at soil carbon stocks under ~3 years of Eucalytus plantation from sugarcane. They use this data to conclude that there was no change in carbon stocks but there was a shift from sugarcane derived carbon to Eucalytus.

This paper documents changes in Hawaiian soil after 32 months of development of a eucalyptus plantation.
Soil C was found to not change significantly, suggesting that expectations of C accumulation due to afforestation may need to be reduced.
Rapid changes in soil are examined here as: (1) Soil fertility may affect tree growth rate, and (2) Changes in soil C must be accounted for to predict root production following an efflux in CO2.

## Fit for purpose: HiCSC

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the HiCSC.

- Location: The exact lat/lon is given in the methods section, although it is unclear if this is the location of downtown Hilo or the site. Years of observation is given in text.
- Soil carbon stock: Carbon stock is given in Table 1 at 0 months and 32 months of plantation. Depth of sample included, bulk density determined but unclear where the data is.
- Soil type: Soils of the Kaiwiki thixotropic, isothermic Typic Hydrandepts. Unclear if "Kaiwiki thixotropic" qualifies as soil type here, but the location might be precise enough to extract from soil maps
- Land Use: Detailed soil history and classification provided in methods section. Land use largely characterized by eucalyptus afforestation.

# Tables and figures

- Table 
  1) Soil Properties at Time of Planting and 32 Months after (C, N, pH, Ca, Mg, K+, Al)
  2)  Nutrient Accumulation on Ion-Exchange Resin Bags (age, fertilization, nutrients)
  
- Figure
  1) Loss of Soil C from Sugar Cane (C4) matched with gain of soil C derived from Eucalyptus (C3). bar plot of soil carbon source parition.

### Table 1

Table 1 was modified by filling in blanks with `NA` and down filling the Age column.
There were two rows in the 'K' column that indicated control vs fertilizer, I split this into two separate columns.
The CI for pH was not aggregated to 0-30 but reported for 0-15 and 15-30, I concatenated both values with a ';' separate into one entry.

```{r table1}
#| code-fold: true
#| message: false

filename <- table1.file
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```


{{< include Binkley1999_Methods.md >}}

# Transcription comparison

```{bash eval=FALSE}
diff --strip-trailing-cr Binkley1999_Methods.md Binkley1999_Methods_HN.md
```

# Files

These files are in the @Binkley1999 data rescue.

- [Readme](README_Binkley1999.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Binkley1999.bib](Binkley1999.bib)
  + citation for article transcribed
- [Binkley1999_Methods.bib](Binkley1999_Methods.bib)
  + citations for the methods section of the article
- [Binkley1999_Methods.md](Binkley1999_Methods.md)
  + methods transcribed from primary article
- [Binkley1999_Table1.csv](Binkley1999_Table1.csv)
  + table 1 from primary article with data about the soil properties at time of planting and 32 months after (C, N, pH, Ca, Mg, K+, Al)

# Working notes 

## Data Rescue Plan

- [x] describe files in human readable form inside ReadMe
- [x] description with Git issue with link
- [x] add contribution yaml
- [x] create excel sheet with Table 1
- [x] export excel table to csv
- [x] copy over methods section into markdown
- [x] pull down citations in methods section to bib file
- [x] add in citation links to methods
- [x] cross check with second transcriber for tables and methods
- [ ] submit to independent review
- [ ] archive on SoilDRaH


## Citation notes from Methods

Below are the citation notes from the methods section.
Some citations are missing, those are noted here and how best guess were generated.

- [ ] short citation
  + copy from bib
  + manually entered or direct export

- [x] Binkley, D., K.A. Dunkin, D. Debell, and M.G. Ryan. 1992. Production and nutrient cycling in mixed plantations of *Eucalyptus* and *Albizia* in Hawaii. For. Sci. 38:393-408.
  + copy from bib
- [x] Vitorello, V.A., C.C. Cerri, F. Andreux, C. Feller, and R.L. Victoria. 1989. Organic matter and natural carbon-13 distribution in forested and cultivated Oxisols. Soil Sci. Soc. Am. J. 53:773-778
  + copy from bib
- [x] Bashkin, M., and D. Binkley. 1998. Changes in soil carbon following afforestation in Hawaii. Ecology 79:828-833.
  + copy from bib

# Level 0: raw data reads

The Level 0 data in this package include the tables above (Table 1 and 3) as well as the methods section.
Here we draft the read function that will be transferred to the R scripts.
In the future this draft will be echoed from the R scripts.

```{r readLevel0}

data.lvl0.df <- list(citation = 
                    list(primary = 
                           read.bib(file = primaryCitation.file), 
                         methods = 
                           read.bib(file = methodsCitation.file)
                    ),
                  method = read_lines(file = methods.file),
                  data = list(Table1 = 
                                list(caption = 
                                       read_csv(file = table1.file,
                                                col_types = cols(.default = col_character()),
                                                n_max = 1, col_names = FALSE)$X1[1],
                                     primary = 
                                       read_csv(file = table1.file,
                                                col_types = cols(.default = col_character()),
                                                skip = 1,
                                                na = '-')
                                )
                  ) 
)
```

# Level 1

```{r}

studyMeta <- tribble(~of_variable, ~is_type, ~with_entry, ~source,
                   'region', 'value', '13 km NNE of downtown Hilo, HI', paste('Method ln3:', paste(data.lvl0.df$method[3], collapse = ' ')),
                 #  'land_use_type', 'value', 'Tree stands', paste('Method ln14-16:', paste(data.lvl0.ls$method[14:16], collapse = ' ')),
                  # 'inital_planting', 'value', '1959', paste('Method ln14:', data.lvl0.ls$method[14]),
                  # 'observation_year', 'value', '1996', paste('Method ln 30;53;58:', paste0(data.lvl0.ls$method[c(30,53,58)], collapse = '... ')),
                   'citation', 'value', format(data.lvl0.df$citation$primary), 'journal citation',
                   'doi', 'value', data.lvl0.df$citation$primary$doi, 'journal citation')

```


```

