---
title: "Data Rescue: Dimitrakopoulos 2010"
format:
  html:
    toc: true
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - Dimitrakopoulos2010_Methods.bib #contains the citations in the methods section
  - Dimitrakopoulos2010.bib #contains citation for the data source, generally a single citation of the primary paper
authors: #author scheme from https://quarto.org/docs/journals/authors.html
  - id: ktoddbrown #use the github name as an id
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role: #TODO define these roles and add reference
      - transcription: review #secondary, review
   #   - standardization: lead #review
    #  - curation: lead #review

  - id: SavaScott #use the github name as an id
    name:
      given: Savannah
      family: Scott
    affiliation:
      - ref: uf-ees
    role: #TODO define these roles and add reference
      - transcription: secondary
   #   - standardization: lead #review
    #  - curation: lead #review

  - id: ChaseSaxour #use the github name as an id
    name:
      given: Chase
      family: Saxour
    affiliation:
      - ref: uf-ees
    role: #TODO define these roles and add reference
      - transcription: secondary
   #   - standardization: lead #review
    #  - curation: lead #review

  - id: sierrav-droid #use the github name as an id
    name:
      given: Sierra
      family: Velazquez
    affiliation:
      - ref: uf
    role: #TODO define these roles and add reference
      - transcription: lead
   #   - standardization: lead #review
    #  - curation: lead #review

affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
  - id: uf
    name: University of Florida
    city: Gainesville
    state: Florida
    country: USA
    postal-code: 32611
    url: https://www.ufl.edu/
    
timelog:
  - activity: 
      description: Lead and ReadMe Updated
      who: sierra-droid #github username
      #when you started and ended, can be just a year or down to the minute
      start: 2026-03-25
      end: 2026-04-15
      duration: PT4H30M
  - activity: 
      description: Secondary and ReadMe Updated
      who: ChaseSaxour #github username
      #when you started and ended, can be just a year or down to the minute
      start: 2026-03-25
      end: 2026-04-15
      duration: PT3H30M
  - activity: 
      description: Secondary and ReadMe Updated
      who: SavaScott #github username
      #when you started and ended, can be just a year or down to the minute
      start: 2026-03-25
      end: 2026-04-15
      duration: PT4H15M

---


```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false
#| eval: true


library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# data files for level 0
methods.file <- "Dimitrakopoulos2010_Methods.md"
Table1.file <- 'Dimitrakopoulos2010_Table1.csv'
figure1.file <- 'Dimitrakopoulos2010_Figure1.csv'

# Bibliography files
primaryCitation.file <- 'Dimitrakopoulos2010.bib'
methodsCitation.file <- 'Dimitrakopoulos2010_Methods.bib'

# read function; you will need to create this function after you have your data
#... rescue reviewed and before you finish your data harmonization.
readsource.file <- '../../02_DataHarmonization/readDimitrakopoulos2010.R'
if(file.exists(readsource.file)){
  source(readsource.file)
}

#semantic files for level 1
of_variable.file <- '/semantics/SoilDRaHVocabulary_of_variable.csv'
is_type.file <- '/semantics/SoilDRaHVocabulary_is_type.csv'

```


# Data Summary

> Panayiotis G. Dimitrakopoulos, Influence of evenness on the litter-species richness–decomposition relationship in Mediterranean grasslands, Journal of Plant Ecology, Volume 3, Issue 2, June 2010, Pages 71–78, https://doi.org/10.1093/jpe/rtq009

For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/219.

<!--- After you read through the paper write a brief summary of what the study
measured and what the data was used for in. This should be <250 words.--->

This paper examines how litter-species richness and evenness affect the rate of leaf litter decomposition in a Mediterranean grassland (Lesbos, Greece).
Over 90 days, the researchers measured the percentage dry weight loss across 10 species in various combinations. 
The main goal of the data was to show the positive correlation between species richness and the rate of decay while focusing on how species diversity leads to faster nutrient cycling. 
The data compared the observed mass loss in mixtures against what was expected from single species. 
While the study finds that evenness has some effect on the results, it mainly finds that decomposition rates are driven by the combined effects of both components of litter-species diversity (richness and evenness). 


## Tables and figures

<!--- Write a brief discription of all the tables and figures in this source. This should include full variable names when possible. It is perfectly fine to restate the captions here or copy other text that feels complete. Please keep this to a single sentence for each table/figure.--->

- Table 
  1) Table 1. Species compositions of the different litter mixtures and single-species treatments
  2) Table 2. The number of litterbags within each treatment
  3) Table 3. ANOVA for decomposition rate in the mixture litterbags

- Figure 
  1)  Figure 1. Decomposition rate plotted against species richness in the litterbags with varying levels of evenness
  2) Figure 2. Percent deviation from expected litter mass remaining in the mixtures [100 × (observed − expected)/expected] as a function of litter-species richness for the maximum-evenness (a), intermediate-evenness (b) and low-evenness (c) treatment.

## Fit for purpose: DeAD

This data is identified as a data source for DeAD.
The Decomposition Across Drylands (DeAD) data collection is a new data effort targeting dryland decomposition studies. The purpose of this data collection is to develop better litter decay models for dryland ecosystems.

### Main Variables of Interest

  - Location: Lesbos, Greece (39° N, 27° E, 30 m above sea level)
  - Soil carbon stock: Leaf carbon concentration (LCC) data is found in a previous paper, Dimitrakopoulos 2001 unpublished data
  - Duration: Samples were pulled summer of 2005, and dried and stored. Litter was pulled out again Autumn of 2005 and dried and weighed again after 90 days.
  - Mass Loss: Decomposition rate can be found in Figure 1
  - Solar Exposure: Could not find solar exposure in this paper

### Other Relevant Variables

 - Litter Species: 10 species used were found in Table 1
 - Microsite Data: Section Materials and Methods, subsection Experimental design paragraph 1
 - Litter Chemical Composition: Carbon and nitrogen characteristics are found in another paper. Other chemical compositions could not be found.
 - Litter Bag Materials and Methods: The litter bags were a size 10 cmx10 cm, and made using a nylon mesh with holes of 1.0 mm. The litter bags included a mass of ∼2 g dried plant litter material. The litter bags were 16 m^2 in area.
 - Age: of Ecosystem: >15 years old
 - Land Management: Abandoned arable field
  
## Files

<!--- Include a brief explaination to all the files in the folder--->

These files are in the Dimitrakopoulos2010 data rescue.

- [Readme](README_Dimitrakopoulos2010.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Dimitrakopoulos2010.bib](Dimitrakopoulos2010.bib)
  + citation for article transcribed
- [Dimitrakopoulos2010_Methods.bib](Dimitrakopoulos2010_Methods.bib)
  + citations for the methods section of the article
- [Dimitrakopoulos2010_Methods.md](Dimitrakopoulos2010_Methods.md)
  + methods transcribed from primary article
- [Dimitrakopoulos2010_Table1.csv](Dimitrakopoulos2010_Table1.csv)
  + table 1 from primary article with Species compositions of the different litter mixtures and single-species treatments
- [Dimitrakopoulos2010_Table2.csv](Dimitrakopoulos2010_Table2.csv)
  + table 2 from primary article with The number of litterbags within each treatment
- [Dimitrakopoulos2010_Table3.csv](Dimitrakopoulos2010_Table3.csv)
  + table 3 from primary article with ANOVA for decomposition rate in the mixture litterbags
- [Dimitrakopoulos2010_Figure1.csv](Dimitrakopoulos2010_Figure1.csv)
  + figure 1 from primary article with Decomposition rate plotted against species richness in the litterbags with varying levels of evenness
- [Dimitrakopoulos2010_Figure2.csv](Dimitrakopoulos2010_Figure2.csv)
  + figure 2 from primary article with Percent deviation from expected litter mass remaining in the mixtures as a function of litter-species richness for the eveness treatment.
- [temp/](temp/)
    + scratch folder that will not be archived on GitHub, it should include local copies of the data resource (likely a PDF article) and any spreadsheets or other tools you used for the transcriptions.

# Data Rescue working notes 

## Data Rescue Plan

+ Lead
  - [x] Read the article or data resource and the discription of the target project
  - [x] Create a fresh fork of the primary repository https://github.com/ktoddbrown/SoilDRaH/tree/main, it might be useful to rename it 'SoilDRaH_Dimitrakopoulos2010'
  - [x] Create a copy of this template folder and rename it to Dimitrakopoulos2010
  - [x] Fill out Data summary section (except for the Files), the Data Rescue Plan (here!), and update the Dimitrakopoulos2010 section of the yml to reflect who is working on the rescue
  - [x] Create a GitHub issue from the Data Rescue template and get feedback from a reviewer on the project (Dr Todd-Brown or a grad student typically).
  - [x] Add the reviewer to the author list in the yml
+ Do
  - [x] Create a spreadsheet with Table1 and export it to csv
  - [x] Extract points from FigureM and export it to csv
  - [x] Transcribe or copy the method section
  - [x] Insert references to bib entries
  - [x] reformat equations and mathematical symbols as [LaTex](https://en.wikibooks.org/wiki/LaTeX/Mathematics)
  - [x] reformat text using [Markdown](https://www.markdownguide.org/cheat-sheet/)
  - [x] Create a bib file with the citations from the methods section
  - [x] Push to your local fork and get a second transcription pushed to your local fork by someone else on the project
+ Measure
  - [x] cross check with second transcriber for tables and methods, and resolve any differences
  - [x] compile or render this ReadMe with all chunks up through the Level 0 data rescue section with `eval: true`
  - [x] push to your local fork and create a pull request to the main branch of the primary repository
  - [x] address any revisions in your pull request review
  - [x] celebrate!

## Table 1

Table 1 was modified to included markdown notation and put the footnotes in the captions.

```{r}
#| label: table1Check


read_csv(file = Table1.file, #read in the specific file
         skip = 1, #skip the first line with the caption
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = Table1.file, #read in that file again but...
                           n_max = 1, #only read the first line
                           col_names = 'caption', #call it the caption
                           #read everything as a character
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE #do not return messages
                           )$caption #and take it out of the table
  )
```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareTable1
#| eval: true

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y temp/Dimitrakopolous2010_Table1_CS.csv Dimitrakopoulos2010_Table1.csv
```

## FIgure 1

Figure 1 was modified to included markdown notation and put the footnotes in the captions.

```{r}
#| label: figure1Check

library(ggplot2)
library(readr)
library(dplyr)

fig1Caption <- read_csv(file = figure1.file,
                        n_max = 1, col_names = 'caption',
                        show_col_types = FALSE)$caption

fig1Data <- read_csv(file = figure1.file,
                     skip = 1,
                     show_col_types = FALSE)

fig1Data$`x value` <- factor(fig1Data$`x value`,
                             levels = c("1", "3", "6"))

fig1Data$`evenness` <- factor(fig1Data$`evenness`,
                             levels = c("Maximum", "Intermediate", "Low"))

fig1Data$`y value SE` <- as.numeric(fig1Data$`y value SE`)
fig1Data <- fig1Data %>%
  mutate(error_var = `y value SE` - `y value`,
  ymax = `y value` + `error_var`)

ggplot(fig1Data, aes(x = `x value`, y = `y value`, fill = `evenness`)) + scale_y_continuous(breaks = c(30, 40, 50, 60, 70, 80)) + geom_bar(stat = "identity", position = position_dodge(width = 0.95)) + geom_errorbar(aes(
  ymin = `y value`,
  ymax = ymax),
  position = position_dodge(width = 0.95),
width = 0.9, color = "black") + labs(
  title = "Figure 1",
  x = "Species richness",
  y = "Litter mixture mass loss (%)",
  caption = fig1Caption)


```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareFigure1
#| eval: true

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y temp/Dimitrakopoulos2010_Figure1_SRS.csv Dimitrakopoulos2010_Figure1.csv
```


## Methods

Any different to the cross check for the methods section will be output after this line.

```{bash}
#| label: compareMethods
#| eval: true
# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.
diff --strip-trailing-cr --suppress-common-lines -y temp/Dimitrakopoulos2010_SRS.md Dimitrakopoulos2010_Methods.md
```

### Methods  - Text

The final text transcription from the Methods section is rendered below.

```{r}
#| label: readMethods
#| output: asis
#| eval: true

# This revised section will append two ### to any existing headers here so that it renders nicely with the other headers in the document.

methodsLines <- readLines(methods.file) |>
  str_replace_all(pattern = '^#', replacement = '####')

cat(paste0(methodsLines, collapse = '\n'))

```

<!----
## Citation notes from Methods

Below are the citation notes from the methods section.
This generally doesn't read well in the final report so commenting it out here but it is very useful for tracking citations when you are doing the transcriptions.
Some citations are missing, those are noted here and how best guess were generated.
Think of this as a lab notebook documenting how you found each citation.

- [ ] short citation
  + copy from bib
  + manually entered or direct export
  
- [x] Dimitrakopoulos PG, Galanidis A, Siamantziouras A-SD, et al., Short-term invasibility patterns in burnt and unburnt experimental Mediterranean grassland communities of varying diversities, Oecologia, 2005, vol. 143 (pg. 428-37)
  + copy from bib

- [x] Dimitrakopoulos PG, Siamantziouras A-SD, Galanidis A, et al., The interactive effects of fire and diversity on short-term responses of ecosystem processes in experimental Mediterranean grasslands, Environ Manage, 2006, vol. 37 (pg. 826-39)

- [x] Schmid B, Hector A, Huston MA, et al.Loreau M, Naeem S, Inchausti P., The design and analysis of biodiversity experiments, Biodiversity and Ecosystem Functioning: Synthesis and Perspectives, 2002OxfordOxford University Press(pg. 61-78)
  + copy from bib

- [x] Dimitrakopoulos PG., Interactive effects of fire and diversity on ecosystem functioning in experimental Mediterranean grassland communities: testing the insurance hypothesis, Ph.D. Thesis, 2001Mytilene, GreeceUniversity of the Aegean (in Greek)
  + copy from bib

- [x] Garnier E, Cortez J, Billès G, et al., Plant functional markers capture ecosystem properties during secondary succession, Ecology, 2004, vol. 85 (pg. 2630-7)
  + copy from bib

- [x] Lecerf A, Risnoveanu G, Popescu C, et al., Decomposition of diverse litter mixtures in streams, Ecology, 2007, vol. 88 (pg. 219-27)
  + copy from bib

- [x] Moore TN, Fairweather PG., Decay of multiple species of seagrass detritus is dominated by species identity, with an important influence of mixing litters, Oikos, 2006, vol. 114 (pg. 329-37)
  + copy from bib
  
- [x] Loreau M., Separating sampling and other effects in biodiversity experiments, Oikos, 1998, vol. 82 (pg. 600-2)
  + copy from bib
---->

# Level 0 Data rescue

This level 0 data read parses the transcribed files with minimal cleaning or transformations.

```{r}
#| label: readLevel0
#| eval: true
#This chunk has two purposes. 
#...1) Check the formatting by reading in everything 
#...2) Create a list of everything to process later in level 1

data.lvl0.ls <- list(
  #Read in a list of all the bib files
  citation = list(
    #Citation for the article transcriptions are pulled from
    primary = read.bib(file = primaryCitation.file), 
    #Citations for all referenced articles
    methods = read.bib(file = methodsCitation.file)
  ),
  method = read_lines(file = methods.file),
  data = list(
    Table1 = list(
      caption = read_csv(file = Table1.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      primary = read_csv(file = Table1.file,
                         col_types = cols(.default = col_character()),
                        skip = 1)
    ),
     Figure1 = list(
      caption = read_csv(file = figure1.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      primary = read_csv(file = figure1.file,
                         col_types = cols(.default = col_character()),
                        skip = 1)
     )
   ) #end 'data' element
)

```

The code is also in `readsource.file` and returned with the `dataLevel = 'level0'` argument.
Below verifies that the data from the function matches what is coded above.

```{r}
#| label: checkReadFunction_lvl0
#| eval: false

if(file.exists(readsource.file)){
  datafunction.lvl0.ls <- readDimitrakopoulos2010(dataDir = '.',
                                         dataLevel = 'level0')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function matches code here.')
  }else{
    print('There is a mismatch in the data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}
```

# Level 1 data - CURATION PROJECT

This data is identified as a data source for the ...
Broad classes of information that we are looking for include:

<!--- Include description of where the data is located in the source.--->
- Location: 
- Soil carbon stock: 
- Soil type: 
- Land use: 

Information from the methods and other sections are pulled into tables with line by line source citations.
This information can be values, units, methods, or other useful notes and is manually extracted from the transcriptions loaded into the level 0 data above.

The final format of the data frame here will include the following 

  + grouping ids
  + `of_variable`
  + `is_type`
  + `with_entry`
  + `from_source`

The `of_variable` being used here are:

 + citation
 
 <!--- include other variables being used, this should match vocabulary in the SoilDRaHVocabulary resources --->

```{r}
#| label: createLvl1
#| eval: true

#### Method-section only data ####
#Create a table for the data in the method-section that is not directly 
#...associated with any table. Often this is study level information.

studyMeta <- tribble(~of_variable, ~is_type, ~with_entry, ~from_source,
                   'region', 'value', 'Honaunau Forest on the southwestern slopes of Mauna Loa, island of Hawaii', paste('Method ln5:', paste(data.lvl0.ls$method[5], collapse = ' ')))

#### Table/Figure N ####
# Process the primary data from Table/Figure M by maintaining data groups and 
#...standardizing the variable name so that we can link this to the study 
#...metadata and data in the other tables when needed. Link in the methods 
#...and other information from the method-section as needed.

Table1Primary <- data.lvl0.ls$data$Table1$primary |>
  #Add a row id preserving row grouping -- not unique
  mutate(row_id = paste0('R', 1:n())) |>
  #Transform the table to long data by row group and column names.
  #...This allows us to then create the cross links between the tables.
  pivot_longer(cols = -row_id,
               names_to = 'column_name', values_to = 'with_entry',
               values_drop_na = TRUE) |> #drop missing values
  #Link the variable names to the column names using a switch statement
  mutate(of_variable = case_when(
           #Soil PH is given as an example below
           column_name == 'SOC' ~ 'soil_organic_carbon')) |>
  #Flag the source of this data as Table 1, this table is merged later with
  #...other data so this keeps track of where the data came from
  mutate(is_type = 'value') |>
  mutate(from_source = 'Table 1')

## Create the metadata for table one from the header information and cited methods
Table1Meta <- Table1Primary |>
  select(column_name, of_variable) |>
  unique() |>
  #Grab everything between the parentheses as units and attribute the source as the column names.
  mutate(unit = str_extract(column_name, pattern = '(?<=\\().*(?=\\))'),
         from_source = 'Table 1 column names.') |>
  #If there aren't units then drop the row
  filter(!is.na(unit)) |>
  #Don't keep the column name now that you have the units, cross link via the variable
  select(-column_name) |>
  #Stack into the table the methods for each variable that we can find
  bind_rows(
  tribble(~of_variable, ~method, ~from_source,
            'soil_organic_carbon', paste0(data.lvl0.ls$method[72:73], collapse = ' '), 'Methods ln72-73') ) |>
  #Stack onto the tables the controlled vocabulary used
  bind_rows(
    tribble(~of_variable, ~control_vocabulary, ~from_source,
            'soil_class', 'Histosol:USDA classification for histosol soil type|Andisols:USDA classification for andisol soil type', 'expert informed')
  ) |>
  # Push the dimensions of the variable into a single column. Making this a long table.
  pivot_longer(cols = c(unit, method, control_vocabulary),
               names_to = 'is_type',
               values_drop_na = TRUE,
               values_to = 'with_entry')

#### Create level 1 #####
## Pull everything together into by stacking the meta and primary data tables
data.lvl1.ls <- list(
  meta = bind_rows(studyMeta, Table1Meta),
  primary = bind_rows(Table1Primary)|>
    arrange(row_id, column_name,
            of_variable, is_type, with_entry, from_source)
  )

```

## Check for new variables

`of_variable` in the data files should be defined in the SoilDRaH Vocabulary files.

```{r}
#| label: check_of_variable
#| eval: false

#Read in the list of current 'of_variable' entries
SoilDRaH_variable_list <- read_csv(file =of_variable.file,
                                   show_col_types = FALSE)

#compare the variables in the level 1 tables to the SoilDRaH variables
new_variables <- lapply(data.lvl1.ls, function(xx){
  setdiff(unique(xx$of_variable), 
          unique(SoilDRaH_variable_list$of_variable))
}) |>
  unlist()

#Check that there are no new variables
if(length(new_variables) == 0){
  print('All of_variables are in the SoilDRaH variable list. No additions required')
}else{
  cat(paste('Please start a issue for new variables [',
               new_variables, ']', sep = '', collapse = '\n'))
}
```

## Check for is_type

`is_type` in the data files should be defined in the SoilDRaH Vocabulary files.

```{r}
#| label: check_is_type
#| eval: false

#read in the current types and what variable they are associated with
SoilDRaH_type_list <- read_csv(file = is_type.file,
                                   show_col_types = FALSE) |>
  #remove the project definition and only keep the variable and type
  select(of_variable, is_type) |>
  #variables are pushed together in the type defined table. Separate the
  #...entries here
  separate_longer_delim(cols = 'of_variable', delim = ',')

#Pull all the of_variable-is_type pairs from the source data
source_pairs <- tibble()

#If there are a large number of entries, consider replacing this for-loop.
#...However this avoids having to install the plyr or other libraries, and 
#...there are generally no more then 3-5 tables
for(dataTable in data.lvl1.ls){
  source_pairs <- dataTable |>
    select(of_variable, is_type) |> 
    unique() |>
    bind_rows(source_pairs)
}

undefined_data_type <- source_pairs |>
  #subset for variables that are not 'incomplete'
  semi_join(SoilDRaH_variable_list |>
              filter(project_status != 'incomplete'),
            by = join_by(of_variable)) |>
  #check for of_variable-is_type tuples that are not in the SoilDRaH type definitions
  anti_join(SoilDRaH_type_list,
            by = join_by(of_variable, is_type))

#Tell the user what you found.
if(nrow(undefined_data_type) == 0){
  print('All of_variables-is_types are in the SoilDRaH type list. No additions required')
}else{
  print(paste0('Please start a issue for new of_variable-is_type [', paste(paste(undefined_data_type$of_variable, undefined_data_type$is_type, sep = '-'), collapse = ','), ']'))
}
```

## Check function match

Check for an exact match with the read function.

```{r}
#| label: checkReadFunction_lvl1
#| eval: false

if(file.exists(readsource.file)){
  datafunction.lvl1.ls <- readDimitrakopoulos2010(dataDir = '.',
                                         dataLevel = 'level1')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function output matches ReadMe code for level 1.')
  }else{
    print('There is a mismatch in the level 1 data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}

```


# References

<!----This space will be filled with the citations when rendered. Leave this blank. ---->