---
title: "Data Rescue: Elmore 2006"
format:
  html:
    toc: true
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - Elmore2006_Methods.bib
  - Elmore2006.bib
authors:
  - id: ktoddbrown
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role:
      - transcription: secondary
      - standardization: review
    #  - curation: lead #review
  
  - id: SavaScott
    name:
      given: Savannah
      family: Scott
    affiliation:
      - ref: uf-ees
    role:
      - transcription: lead
      - standardization: lead 
   #   - curation: lead #review

affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
  - id: uf
    name: University of Florida
    city: Gainesville
    state: Florida
    country: USA
    postal-code: 32611
    url: https://www.ufl.edu/
  
timelog:
  - activity: 
      description: Secondary Transcription and Review
      who: ktoddbrown
      start: 2025-05-31
      end: 2025-06-09
      duration: PT2H30M
      
  - activity: 
      description: Primary and ReadMe Update
      who: SavaScott
      start: 2025-05-31
      end: 2025-06-09
      duration: PT5H
---

```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false

library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# data files for level 0
methods.file <- "Elmore2006_Methods.md"
table1.file <- 'Elmore2006_Table1.csv'

# Bibliography files
primaryCitation.file <- 'Elmore2006.bib'
methodsCitation.file <- 'Elmore2006_Methods.bib'

# read function; you will need to create this function after you have your data
#... rescue reviewed and before you finish your data harmonization.
readsource.file <- '../../02_DataHarmonization/readElmore2006.R'
if(file.exists(readsource.file)){
  source(readsource.file)
}

#semantic files for level 1
of_variable.file <- '../../semantics/SoilDRaHVocabulary_of_variable.csv'
is_type.file <- '../../semantics/SoilDRaHVocabulary_is_type.csv'


```

This documents the data transcription for @Elmore2006.
For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/93.

> ELMORE, A.J. and ASNER, G.P. (2006), Effects of grazing intensity on soil carbon stocks following deforestation of a Hawaiian dry tropical forest. Global Change Biology, 12: 1761-1772. https://doi.org/10.1111/j.1365-2486.2006.01198.x

# Data Summary

This study investigates the impact of livestock grazing on soil carbon (C) levels following the clearing of dry tropical forests in Hawaii.
The findings show that long-term grazing reduces soil organic carbon, with the extent of these changes increasing at higher elevations.

# Tables and figures

- Table 
  1) The fraction of SOC derived from forest sources increased with elevation (R=0.62; P<0.001), but total soil C stocks and pasture-derived SOC were not significantly influenced by elevation.
  2) Soil measurements made from forest soils at two elevations
  3) Model results using data from both transects

- Figure 
  1) A portion of Pu'u Wa'awa'a Ranch on the island of Hawai'i, showing the location of field transects.
  2) Mean changes in vegetation and soil properties along transects of grazing intensity (constant elevation and precipitation): (a) remotely sensed fractional cover of photosynthetic and nonphotosynthetic vegetation cover and base substrate; (b) soil organic C (SOC); (c) SOC derived from forest soils; (d) SOC derived from pasture grasses.
  3)  Mean changes in species cover along transects of grazing intensity.
  4) The relationship between remotely sensed nonphotosynthetic vegetation fraction and Pennisetum Setaceum cover.
  5)  Mean changes in plant and soil properties along elevation transects: (a) remotely sensed fractional cover of photosynthetic and nonphotosynthetic vegetation cover and base substrate; (b) soil organic C (SOC); (c) SOC derived from forest soils; (d) SOC derived from pasture grasses.
  6) The relationship between total soil organic carbon (SOC) from soil cores to 10cm depth and nonphotosynthetic vegetation from imaging spectroscopy.

## Fit for purpose: HiCSC

This source is identified as a data source for the Hawai'i Soil Organic Carbon Database as part of the HiCSC.

- Location:  Pu'u Wa'awa'a Ranch on the north side of Hualalai volcano, in-depth description found in the materials and methods section, subsection Site description.
- Soil carbon stock: The depth of the soil and SOC cores were taken, and total SOC was determined using a pool modeling model.
- Soil type: Volcanic ash (about ~200 to 3500-year-old basaltic lava flows), with forest converted to pasture lands
- Land use:  Land use of the soil was primary pasture, converted from dry tropical forest 100 years prior. In-depth land use is in the methods and materials section, subsection Site description.
  
# Files

These files are in the @Elmore2006 data rescue.

- [Readme](README_Elmore2006.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Elmore2006.bib](Elmore2006.bib)
  + citation for article transcribed
- [Elmore2006_Methods.bib](Elmore2006_Methods.bib)
  + citations for the methods section of the article
- [Elmore2006_Methods.Rmd](Elmore2006_Methods.Rmd)
  + methods transcribed from primary article
- [Elmore2006_Table1.csv](Elmore2006_Table1.csv)
  + table N from primary article with  Soil data collected along gradients in grazing intensity (G) and elevation (E).

# Data Rescue working notes 

## Data Rescue Plan

+ Lead
  - [x] Read the article or data resource and the discription of the target project
  - [x] Create a fresh fork of the primary repository https://github.com/ktoddbrown/SoilDRaH/tree/main, it might be useful to rename it 'SoilDRaH_AuthorYYYY'
  - [x] Create a copy of this template folder and rename it to AuthorYYYY
  - [x] Fill out Data summary section (except for the Files), the Data Rescue Plan (here!), and update the author section of the yml to reflect who is working on the rescue
  - [x] Create a GitHub issue from the Data Rescue template and get feedback from a reviewer on the project (Dr Todd-Brown or a grad student typically).
  - [x] Add the reviewer to the author list in the yml
+ Do
  - [x] Create a spreadsheet with TableN and export it to csv
  - [x] Extract points from FigureM and export it to csv
  - [x] Transcribe or copy the method section
  - [x] Insert references to bib entries
  - [x] reformat equations and mathematical symbols as [LaTex](https://en.wikibooks.org/wiki/LaTeX/Mathematics)
  - [x] reformat text using [Markdown](https://www.markdownguide.org/cheat-sheet/)
  - [x] Create a bib file with the citations from the methods section
  - [x] Push to your local fork and get a second transcription pushed to your local fork by someone else on the project
+ Measure
  - [x] cross check with second transcriber for tables and methods, and resolve any differences
  - [x] compile or render this ReadMe with all chunks up through the Level 0 data rescue section with `eval: true`
  - [x] push to your local fork and create a pull request to the main branch of the primary repository
  - [x] address any revisions in your pull request review
  - [x] celebrate!

## Table 1

Table 1 was modified to included markdown format. 

```{r}
#| label: table1Check
#| message: false

filename <- table1.file
read_csv(file = filename,
                   skip = 1,
                   col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = filename, 
                                 n_max = 1, col_names = 'caption', 
                                 show_col_types = FALSE)$caption)
```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareTable1
#| eval: false


diff --strip-trailing-cr --suppress-common-lines -y temp/AuthorYYYY_Table1_KTB.md AuthorYYYY_Table1.md
```

## Methods

Any different to the cross check for the methods section will be output after this line.

```{bash}
#| label: compareMethods
#| eval: false

diff --strip-trailing-cr Elmore2006_Methods.md Elmore2006_Methods_KTB.Rmd

```

### Methods  - Text

The final text transcription from the Methods section is rendered below.

```{r}
#| label: readMethods
#| output: asis
#| eval: true

# This revised section will append two ### to any existing headers here so that it renders nicely with the other headers in the document.

methodsLines <- readLines(methods.file) |>
  str_replace_all(pattern = '^#', replacement = '####')

cat(paste0(methodsLines, collapse = '\n'))

```

<!----
## Citation notes from Methods

Below are the citation notes from the methods section.
Some citations are missing, those are noted here and how best guess were generated.

- [x] Wolfe EW, Morris J (1996) Geologic map of the island of Hawaii. US Geological Survey. Map I-2524-A. US Department of the Interior.
  + direct export
- [x] MacCaughey V (1917) Vegetation of Hawaiian lava flows. Botanical Gazette, 64, 386–420.
  + direct export
- [x] Eggler WA (1971) Quantitative studies of vegetation on sixteen young lava flows on the island of Hawaii. Tropical Ecology, 12, 66–100.
  + direct export
- [x] Blackmore M, Vitousek PM (2000) Cattle grazing, forest loss, and fuel loading in a dry forest ecosystem at Pu'u Wa'aWa'a ranch, Hawai'i. Biotropica, 32, 625–632.
  + direct export
- [x] State of Hawaii. (2003) Management plan for the Ahupua'a of Pu'u Wa'awa'a and the Makai Lands of Pu'u Anahulu. Department of Land and Natural Resources, Division of Forestry and Wildlife, Hilo, Hawaii.
  + direct export
- [x] Savory A (1992) Livestock and Desertification. BioScience, 42, 244–244.
  + direct export
- [x] Asner GP, Elmore AJ, Hughes FMR et al. (2005) Ecosystem Structure along Bioclimatic Gradients in Hawai'i from Imaging Spectroscopy. Remote Sensing of Environment, 96, 497–508.
  + direct export
- [x] Adams JB, Smith MO, Johnson PE (1986) Spectral mixture modeling: a new analysis of rock and soil types at the Viking Lander I Site. Journal of Geophysical Research, 91, 8098–8112.
  + direct export
- [x] Smith MO, Ustin SL, Adams JB et al. (1990) Vegetation in deserts: I. A regional measure of abundance from multispectral images. Remote Sensing Environment, 31, 1–26.
  + direct export
- [x] Elmore AJ, Mustard JF, Manning SJ et al. (2000) Quantifying vegetation change in semiarid environments: precision and accuracy of spectral mixture analysis and the Normalized Difference Vegetation Index. Remote Sensing of Environment, 73, 87–102.
  + direct export
- [x] Asner GP, Bateson CA, Privette JL et al. (1998) Estimating vegetation structural effects on carbon uptake using satellite data fusion and inverse modeling. Journal of Geophysical Research, 103, 28,839–28,853.
  + direct export
- [x] Asner GP, Heidebrecht KB (2002) Spectral unmixing of vegetation, soil and dry carbon cover in arid regions: comparing multispectral and hyperspectral observations. International Journal of Remote Sensing, 23, 3939–3958.
  + direct export
- [x] Canfield RH (1941) Application of the line interception method. In sampling range vegetation. Journal of Forestry, 39, 388–304.
  + direct export
- [x] Daniel JA, Potter K, Altom W et al. (2002) Long-term grazing density impacts on soil compaction. Transactions of the ASAE, 45, 1911–1915.
  + direct export
- [x] Veldkamp E (1994) Organic-Carbon Turnover in three tropical soils under pasture after deforestation. Soil Science Society of America Journal, 58, 175–180.
  + direct export
- [x] Murty D, Kirschbaum MUF, McMurtrie RE et al. (2002) Does conversion of forest to agricultural land change soil carbon and nitrogen? a review of the literature. Global Change Biology, 8, 105–123.
  + direct export
- [x] Neill C, Melillo JM, Steudler PA et al. (1997) Soil carbon and nitrogen stocks following forest clearing for pasture in the southwestern Brazilian Amazon. Ecological Applications, 7, 1216–1225.
  + direct export
- [x] Vitorello B, Cerri CC, Andreux F et al. (1989) Organic matter and natural carbon-13 distribution in forested and cultivated Oxisols. Soil Science Society of America Journal, 53, 773–778.
  + direct export
- [x] Townsend AR, Vitousek PM, Trumbore SE (1995) Soil organic-matter dynamics along gradients in temperature and land-use on the Island of Hawaii. Ecology, 76, 721–733.
  + direct export
---->

# Level 0: raw reads

The Level 0 data in this package include the tables above (Table 1) as well as the methods section.
Here we draft the read function that will be transferred to the R scripts.
In the future this draft will be echoed from the R scripts.

```{r readLevel0}

data.lvl0.df <- list(
  citation = 
    list(primary = 
           read.bib(file = primaryCitation.file), 
         methods = 
           read.bib(file = methodsCitation.file)
    ), #end citation
  method = read_lines(file = methods.file),
  data = list(Table1 = 
                list(caption = 
                       read_csv(file = table1.file,
                                col_types = cols(.default = col_character()),
                                n_max = 1, col_names = FALSE)$X1[1],
                     primary = 
                       read_csv(file = table1.file,
                                col_types = cols(.default = col_character()),
                                skip = 1,
                                na = '-')
                ) #end Table 1
  ) #end data
)
```


The code is also in `readsource.file` and returned with the `dataLevel = 'level0'` argument.
Below verifies that the data from the function matches what is coded above.

```{r}
#| label: checkReadFunction_lvl0

if(file.exists(readsource.file)){
  datafunction.lvl0.df <- readAuthorYYYY(dataDir = '.',
                                         dataLevel = 'level0')
  
  if(isTRUE(all.equal(datafunction.lvl0.df, data.lvl0.df))){
    print('Read function matches code here.')
  }else{
    print('There is a mismatch in the data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}
```


# Level 1 data - CURATION HiCSC

This data is identified as a data source for the HiCSC.
Broad classes of information that we are looking for include:

- Location: Pu'u Wa'awa'a Ranch on the north side of Hualalai volcano, in-depth description found in the materials and methods section, subsection Site description.
- Soil carbon stock: The depth of the soil and SOC cores were taken, and total SOC was determined using a pool modeling model.
- Soil type: volcanic ash (about ~200 to 3500-year-old basaltic lava flows)
- Land use: Land use of the soil was primary pasture, converted from dry tropical forest 100 years prior. In-depth land use is in the methods and materials section, subsection Site description.

Information from the methods and other sections are pulled into tables with line by line source citations.
This information can be values, units, methods, or other useful notes and is manually extracted from the transcriptions loaded into the level 0 data above.

The final format of the data frame here will include the following 

  + grouping ids
  + `of_variable`
  + `is_type`
  + `with_entry`
  + `from_source`

The `of_variable` being used here are:

 + citation
 + doi
 + elevation
 + carbon_percent
 + total_organic_carbon
 + dry_bulk_density
 + delta_13_carbon
 + soil_class
 + soil_age
 + precipitation
 + land_use
 + core_depth_interval
 + land_use
 
```{r}
#| lable: constructLevel1

#### Method section only data ####
#Create a table for the data in the methods section that is not directly 
#...associated with any table.
studyMeta <- tribble(~of_variable, ~is_type, ~with_entry, ~from_source,
                    'citation', 'value', format(data.lvl0.df$citation$primary), 'journal citation',
                   'doi', 'value', data.lvl0.df$citation$primary$doi, 'journal citation',
                   'region', 'value', "portion of Pu'u Wa'awa'a Ranch on the north side of Hualalai volcano, on the Island of Hawai'i", paste('Method ln1:', paste(data.lvl0.df$method[1], collapse = ' ')),
                   'elevation', 'min', '850', paste('Method ln8:', paste(data.lvl0.df$method[8], collapse = ' ')),
                   'elevation', 'max', '1100', paste('Method ln8:', paste(data.lvl0.df$method[8], collapse = ' ')),
                   'elevation', 'unit', 'm', paste('Method ln8:', paste(data.lvl0.df$method[8], collapse = ' ')),
                   'soil_age', 'min', '2500', paste('Method ln6:', paste(data.lvl0.df$method[6], collapse = ' ')),
                   'soil_age', 'max', '3500', paste('Method ln6:', paste(data.lvl0.df$method[6], collapse = ' ')),
                   'soil_age', 'unit', 'yrs', paste('Method ln6:', paste(data.lvl0.df$method[6], collapse = ' ')),
                   'precipitation', 'min', '625', paste('Method ln7:', paste(data.lvl0.df$method[7], collapse = ' ')),
                   'precipitation', 'max', '900', paste('Method ln7:', paste(data.lvl0.df$method[7], collapse = ' ')),
                   'precipitation', 'unit', 'mm', paste('Method ln7:', paste(data.lvl0.df$method[7], collapse = ' ')),
                   'land_use', 'value', 'Cattle grazing', paste('Method ln21:', paste(data.lvl0.df$method[21], collapse = ' ')),
                   'land_use', 'interval', '1900',  paste('Method ln21:', paste(data.lvl0.df$method[21], collapse = ' ')),
                   'land_use', 'interval_format', 'YYYY', NA)

#### Table 1 ####
Table1Primary <- data.lvl0.df$data$Table1$primary |>
  mutate(row_id = paste0('R', 1:n())) |>
  pivot_longer(cols = -row_id,
               names_to = 'column_name', values_to = 'with_entry',
 values_drop_na = TRUE) |>
  
   mutate(of_variable = case_when(
           column_name == "Sample #" ~ "sample_id",
           column_name == "Distance (m)" ~ "grazing_distance",
           column_name == "Elevation (m)" ~ "elevation",
           column_name == '% C' ~ "carbon_percent",
           column_name == 'delta$^{13}$ C' ~ "delta_13_carbon",
           column_name == "rho (g cm$^{-3}$)" ~ "dry_bulk_density",
           column_name == "TOC (kg C m$^{-2}$)" ~ "total_organic_carbon")) |>
    mutate(from_source = 'Table 1')

Table1Meta <- Table1Primary |>
 select(column_name, of_variable) |>
  unique() |>
 mutate(unit = str_extract(column_name, pattern = '(?<=\\().*(?=\\))'),
  from_source = 'Table 1 column names.') |>
   filter(!is.na(unit)) |>
   bind_rows(
    tribble(~of_variable, ~method, ~from_source,
            'delta_13_carbon', paste0(data.lvl0.df$method[c(99,103:106)], collapse = ' '), 'Methods ln99;103-106',
            'dry_bulk_density', paste0(data.lvl0.df$method[96], collapse = ' '), 'Methods ln96',
           'total_organic_carbon',  paste0(data.lvl0.df$method[c(94, 96, 98)], collapse = ' '), 'Methods ln96;96;98') |>
  
   left_join(Table1Primary |>
                  select(column_name, of_variable) |>
                  unique(),
                by = join_by(of_variable))) |>
  
   pivot_longer(cols = c(unit, method),
               names_to = 'is_type',
               values_drop_na = TRUE,
               values_to = 'with_entry')

#### Create level 1 #####
data.lvl1.df <- list(
  study = studyMeta,
  primary_meta = dplyr::bind_rows(Table1Meta),
  primary = bind_rows(Table1Primary)|>
    mutate(elevation_id = with_entry[of_variable == 'elevation'],
           is_type = 'value',
           .by = row_id) |>
    arrange(row_id, elevation_id, column_name,
            of_variable, is_type, with_entry, from_source)
  )

```
 
 