---
title: "Carbon in Peat on EArth through Time"
format:
  html:
    toc: true
date: last-modified
date-format: YYYY-MMMM
authors:
  - name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    note: https://github.com/ktoddbrown
    affiliation:
      - ref: uf-ees
  - name:
      given: Vaasuki
      family: Marupaka
    affiliation:
      - ref: uf-ees
affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: FL
    country: USA
    url: https://essie.ufl.edu/ees/
---

The purpose of this document is to summarize the portions of the Carbon in Peat on EArth through Time (C-PEAT) Database that are relevant to data collections in the SoilDRaH project and walk through the data ingestion.
Here you will find links to documentation from the data provider, links to where you can access the data, data model description, data processing workflows, and visuals for the collection relevant variables.

# What is C-PEAT

The version of the Carbon in Peat on EArth through Time (C-PEAT) Database described here is the public facing version of the PANGAEA repository (https://www.pangaea.de).

> **C-PEAT Database**
> Peatlands played a key role in the global carbon cycle during the Holocene and previous interglacials. High-latitude and tropical peatlands have acted as net long-term atmospheric sinks for carbon dioxide (CO2). However, key uncertainties remain regarding many fundamental patterns and processes. C-PEAT aims to address these uncertainties.

Quote from [https://pastglobalchanges.org/science/wg/former/peat-carbon/intro](https://pastglobalchanges.org/science/wg/former/peat-carbon/intro) (Accessed 17-June-2024)

# Where is the data

This data was downloaded from [https://www.pangaea.de/?q=project:label:PAGES_C-PEAT](https://www.pangaea.de/?q=project:label:PAGES_C-PEAT) using the project label `PAGES_C-PEAT` and the search function of the package `pangaear` to identify 850 unique DOIs.
We have noticed that these DOIs have reduced over time, some of these were merged data sets (merging age and geochemistry from the same core into a single core, 10.1594/PANGAEA.890358 and 10.1594/PANGAEA.928557), others appear to be newly registered data, and others seem to disappear from new quires.
We have provided a list of doi's retrieved at various points in the project as a `echo: false` chunk in the qmd code for documentation and the commented out `pangaear::pg_search` code.

There is some one-off data collections for ingest that means we can not ensure that future data contributions will be consistent with what we present here.
This should be a good starting point however for adding new contributions.

```{r}
#| label: setup

library(tidyverse)
#library(pangaear)

download.dir <- '../../../temp/CPEAT'
#list.files(download.dir) |> length()
```

```{r}
#| label: doi_list
#| echo: false

#### Long doi list ####

downloadUrl.arr <- c(
  "https://doi.pangaea.de/10.1594/PANGAEA.889936?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.889937?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890186?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890189?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890198?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890199?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890200?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890201?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890202?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890203?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890204?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890205?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890206?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890208?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890212?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890213?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890214?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890215?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890216?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890217?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890218?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890219?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890220?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890222?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890345?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890346?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890347?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890348?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890349?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890350?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890351?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890352?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890353?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890354?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890355?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890356?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890357?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890358?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890380?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890381?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890382?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890383?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890384?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890385?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890386?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890387?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890388?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890389?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890390?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890391?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890392?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890393?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890397?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890398?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890399?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890400?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890401?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890402?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890403?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890404?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890405?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890406?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890407?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890408?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890409?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890410?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890411?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890412?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890413?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890414?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890415?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890416?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890417?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890418?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890419?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890420?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890421?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890423?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890424?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890425?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890426?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890427?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890428?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890429?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890430?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890431?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890432?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890433?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890434?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890436?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890437?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890438?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890439?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890446?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890447?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890448?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890471?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890472?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890473?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890474?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890475?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890476?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890477?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890478?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890479?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890480?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890481?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890482?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890483?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890484?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890485?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890486?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890487?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890488?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890489?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890490?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890494?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890495?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890496?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890497?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890498?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890499?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890500?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890501?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890502?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890503?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890504?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890505?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890506?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890507?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890508?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890509?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890510?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890511?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890512?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890513?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890527?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890528?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890529?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890530?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890531?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890532?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890533?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890534?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890535?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890536?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890537?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890538?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890539?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890540?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890543?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890544?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890545?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890546?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890547?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890548?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890549?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890550?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890551?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890552?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890553?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890554?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890555?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.890556?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926348?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926350?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926351?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926352?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926353?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926356?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926361?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926363?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926364?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926365?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926402?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926404?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926405?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926408?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926409?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926410?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926411?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926412?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926413?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926415?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926421?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926423?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926424?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926425?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926426?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926427?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926432?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.926434?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927337?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927383?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927384?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927385?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927394?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927398?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927399?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927404?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927405?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927406?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927407?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927408?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927409?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927410?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927411?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927412?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927414?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927453?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927454?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927455?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927456?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927457?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927461?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927462?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927463?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927465?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927531?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927532?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927533?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927534?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927535?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927536?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927537?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927538?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927539?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927655?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927656?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927658?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927659?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927661?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927662?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927688?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927689?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927692?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927693?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927695?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927698?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927699?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927700?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927702?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927710?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927712?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927715?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927716?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927717?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927763?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927765?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927766?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927767?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927769?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927773?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927777?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927780?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927781?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927782?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927783?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927784?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927785?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927836?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927837?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927839?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927840?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927841?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927843?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927844?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927845?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927846?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927847?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927848?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927849?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927850?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927872?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927873?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927875?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927877?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927878?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927883?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927889?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927907?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927920?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927921?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927966?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927967?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927968?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927969?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927970?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927972?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927973?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927974?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927975?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927978?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927979?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927980?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.927981?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928052?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928053?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928055?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928057?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928058?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928060?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928061?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928062?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928063?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928064?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928065?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928066?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928067?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928068?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928070?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928071?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928072?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928074?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928075?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928076?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928077?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928144?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928146?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928147?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928149?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928172?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928175?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928177?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928180?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928181?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928182?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928183?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928184?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928185?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928186?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928187?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928188?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928189?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928190?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928249?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928250?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928252?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928253?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928254?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928255?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928256?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928257?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928258?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928259?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928261?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928262?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928264?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928265?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928268?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928296?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928298?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928299?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928301?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928303?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928304?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928305?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928306?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928307?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928308?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928309?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928310?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928338?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928347?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928348?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928375?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928376?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928377?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928378?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928379?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928380?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928382?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928383?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928384?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928385?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928386?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928387?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928388?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928424?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928425?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928426?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928428?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928431?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928433?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928435?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928436?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928437?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928439?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928440?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928442?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928443?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928552?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928553?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928554?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928555?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928556?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928557?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928558?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928559?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928560?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928561?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928562?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928583?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928584?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928585?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928586?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928622?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928623?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928624?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928625?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928626?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928627?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928629?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928630?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928631?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928632?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928634?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928635?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928636?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928637?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928638?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928640?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928641?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.928643?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929003?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929011?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929064?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929068?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929223?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929224?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929225?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929226?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929227?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929228?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929268?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929269?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929271?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929274?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929278?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929300?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929302?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929306?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929309?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929310?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929311?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929313?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929336?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929337?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929338?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929356?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929357?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929358?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929364?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929365?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929366?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929367?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929368?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929370?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929372?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929373?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929388?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929395?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929397?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929398?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929399?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929407?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929415?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929426?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929429?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929432?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929444?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929446?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929448?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929449?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929452?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929489?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929494?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929496?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929498?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929501?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929506?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929507?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929511?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929531?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929536?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929537?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929538?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929539?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929540?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929541?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929542?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929543?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929545?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929546?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929547?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929566?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929567?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929573?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929575?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929577?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929579?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929581?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929611?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929644?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929646?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929647?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929648?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929651?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929652?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929653?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929654?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929655?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929656?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929679?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929681?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929682?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929685?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929686?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929687?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929688?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929689?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929699?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929702?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929703?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929705?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929707?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929708?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929710?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929723?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929724?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929725?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929727?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929731?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929732?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929734?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929736?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929737?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929744?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929745?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929746?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929761?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929762?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929766?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929774?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929776?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929778?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929830?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929835?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929843?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929844?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929845?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929846?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929847?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929848?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929849?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929851?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929852?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929854?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929855?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929856?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929858?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929860?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929861?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929862?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929864?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929867?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929868?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929870?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929871?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929875?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929877?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929883?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929884?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929889?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929891?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929894?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929896?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929898?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929900?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929901?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929905?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929906?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929908?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929910?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929911?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929958?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929971?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929972?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929974?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929980?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929981?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929988?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929995?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929998?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.929999?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930000?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930001?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930002?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930012?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930081?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930106?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930107?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930108?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930109?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930110?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930112?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930115?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930116?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930117?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930118?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930121?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930122?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930123?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930124?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930127?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930129?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930130?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930133?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930139?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930145?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930146?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930148?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930168?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930169?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930170?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930171?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930174?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930179?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930181?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930184?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930193?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930481?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930482?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930489?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930490?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930491?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930492?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930501?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930502?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.930503?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.933369?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.933371?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.933374?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.933375?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.933507?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.933508?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.933509?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.933511?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934062?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934064?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934065?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934066?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934067?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934068?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934069?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934070?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934071?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934072?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934111?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934112?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934113?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934115?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934118?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934120?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934122?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934123?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934124?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934126?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934137?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934138?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934139?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934140?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934141?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934149?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934150?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934151?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934152?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934153?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934154?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934155?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934156?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934157?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934161?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934162?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934164?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934165?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934166?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934167?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934256?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934257?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934259?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934260?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934261?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934262?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934263?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934264?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934265?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934266?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934267?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934274?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934275?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934276?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934277?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934278?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934279?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934280?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934281?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934282?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934283?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934284?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934330?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934333?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934334?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934335?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934336?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934337?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934339?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934341?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934342?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934343?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934389?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934390?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934391?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934392?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934393?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934394?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934395?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934396?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934397?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934398?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934438?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934439?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934440?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934442?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934443?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934444?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934445?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934448?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934449?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934453?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934618?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934619?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934620?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934621?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934622?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934623?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934624?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934627?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934628?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934631?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934633?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934634?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934635?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934636?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934637?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934638?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934639?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934640?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934641?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934642?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934721?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934722?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934723?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934724?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934725?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934726?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934727?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934728?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934729?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934730?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934766?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934768?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934769?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934770?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934771?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934772?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934774?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934775?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934776?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934777?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934829?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934831?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934832?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934833?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934834?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934835?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934836?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934837?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934838?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934839?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934908?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934909?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934910?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934911?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934912?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934913?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934914?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934915?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934916?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934917?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934930?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.934933?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935014?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935015?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935016?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935017?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935018?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935019?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935020?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935021?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935022?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935024?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935059?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935060?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935061?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935062?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935063?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935064?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935065?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935066?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935067?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935068?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935069?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935086?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935088?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935089?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935090?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935092?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935093?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935094?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935095?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935096?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935097?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935721?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935722?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935723?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935724?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935725?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935726?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935727?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935728?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935729?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.935730?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.936499?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.936500?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.936501?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.936502?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.936503?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.936520?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.936521?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.936522?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.936523?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.936524?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.936579?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.936581?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.936582?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.936583?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.936584?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.936585?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.936586?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.941079?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.941094?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.941098?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.941132?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.941134?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.941139?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.941140?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.971569?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.971570?format=textfile",
  "https://doi.pangaea.de/10.1594/PANGAEA.971576?format=textfile")

#### Check pangaea search ####

# temp.df <- pangaear::pg_search(query = 'project:label:PAGES_C-PEAT', count = 500) |>
#   bind_rows(
#     pangaear::pg_search(query = 'project:label:PAGES_C-PEAT', offset = 500, count = 500))
# 
# #is there anything new?
# setdiff(paste0("https://doi.pangaea.de/",temp.df$doi, "?format=textfile"),
#         downloadUrl.arr)
# #is there anything "lost" from the search
# setdiff(downloadUrl.arr,
#         paste0("https://doi.pangaea.de/",temp.df$doi, "?format=textfile"))
```

# Level 0 data

```{r}
#| label: downloadCPEAT

basenames.arr <- downloadUrl.arr |> 
  # start with:
  # "https://doi.pangaea.de/10.1594/PANGAEA.889936?format=textfile"
  str_extract(pattern = '(?<=pangaea.de/).*(?=.format)') |> 
  # extract the string between the doi url and the format information:
  # 10.1594/PANGAEA.889936
  str_replace(patter = '/', replacement = '_') |>
  # replase the forward slash so that it is not treated as a folder
  # 10.1594_PANGAEA.889936
  paste0('.txt')
  # add the extension

download.flag <- !file.exists(file.path(download.dir, basenames.arr))
if(any(download.flag)){
  download.file(url = downloadUrl.arr[download.flag], 
                destfile = file.path(download.dir,
                                     basenames.arr[download.flag]))
}
```

```{r}
#| label: loadLevel0

#file.str <- readr::read_file(file.path(download.dir, basenames.arr[1]))
#
basenames.ls <- setNames(object = as.list(basenames.arr), str_remove(basenames.arr, '.txt'))
all.text <- plyr::llply(.data = basenames.ls, 
                        .fun = function(basename.str){
  readr::read_file(file.path(download.dir, basename.str))
})

all.meta <- plyr::ldply(.data = all.text,
                        .fun = function(file.str){
  #### Parse the first level of the metadata ####
  #file.str <- all.text$`10.1594_PANGAEA.928061`
  header.str <- file.str |>
    str_extract(regex('(?<=/\\* ).*(?=\\*/)', multiline=TRUE, dotall=TRUE))
  
  meta.df <- header.str |>
    str_split_1('(?<=\\n)[^\\t]*\\:') |>
    as.list() |>
    setNames(c('meta', str_extract_all(header.str, '(?<=\\n)[^\\t]*\\:') |> 
                 unlist() |>
                 str_remove('\\:$'))) |>
    as_tibble_row()
  return(meta.df)
}, .id = 'doi') 

all.primary <- plyr::ldply(.data = all.text, .fun = function(file.str){
  
#### Parse the primary data file ####

  header.str <- file.str |> #all.text[388] |> #
  str_extract(regex('(?<=\\*/\\n)[^\\n]*', multiline=TRUE, dotall=TRUE)) |>
    str_split(patter = '\\t') |>
    unlist()
  
  #"10.1594_PANGAEA.928439" index 388 has an extra 3 at the end of the file randomly, this throws a warning but does not affect how the data is read in
primary.df <- file.str |> #all.text[388] |> #
  str_extract(regex('(?<=\\*/\\n).*', multiline=TRUE, dotall=TRUE)) |>
  read_tsv(col_types = cols(.default = col_character()),
           skip = 1, col_names = FALSE) |>
  mutate(across(.cols = everything(), .fns = trimws)) |>
  mutate(row_id = paste0('R',1:n())) |>
  pivot_longer(cols = -row_id, 
               names_to = 'column_index', values_to = 'with_entry') |>
  full_join(tibble(column_index = paste0('X', 1:length(header.str)),
                   column_name = header.str),
            by = join_by(column_index))

  return(primary.df)
}, .id = 'doi')


```


```{r}
#| label: processLevel1

study.df <- all.meta |>
  select(doi, citation = Citation, 
         replaced_by = `Replaced by`, related_to = `Related to`,
         further_details = `Further details`,
         projects = `Project(s)`,
         change_history = `Change history`,
         license = License,
         size = Size, abstract = Abstract,
         keyword = `Keyword(s)`, status = Status) |>
  mutate(across(everything(), trimws)) |>
  mutate(across(c(projects, license, keyword, status), as.factor)) |>
  mutate(replaced_by_doi = str_extract(replaced_by, '10\\.1594/PANGAEA\\.\\d*') |>
           str_replace('/', '_'))

core.df <- all.meta |>
  select(doi, coverage = Coverage,
         event = `Event(s)`, parameters = `Parameter(s)`) |>
  separate_wider_delim(cols = parameters, 
                       delim = '\t',
                       names = paste0('X', 0:14),
                       too_few = 'align_start')

#identfy and process the sub-patterns
structures.df <- all.meta |>
  select(Coverage, `Event(s)`, Size, `Parameter(s)`) |>
  mutate(
    Coverage_pattern = 
      str_replace_all(Coverage, '\\-?\\d*\\.?\\d+', '%d'),
    Events_pattern = 
      str_replace(`Event(s)`, '^[^\\*]*\\*', '%s') |>
      str_replace(':[^\\*]*$', ': %s') |>
      str_replace_all(':[^\\*]*\\*', ': %s'),
    Size_pattern = 
      str_replace_all(Size, '\\d+', '%d')) |>
  ##Size: Convert to numeric with unit descriptions
  ##Coverage: Separate variables based on * and \n, then by [name] : [entry], then extract unit from entry. Strip white space
  ##Events: Separate variables based on * , then by [name] : [entry], process Comments via ; split, then by [name] : [entry]
  ##Parameters: Separate and process the parameter descriptions by str subtraction
  separate_wider_delim(cols = `Parameter(s)`, 
                       delim = '\t',
                       names = paste0('X', 0:14),
                       too_few = 'align_start') |>
  select(-X0) |>
  pivot_longer(cols = starts_with('X'),
               names_to = 'column_index',
               values_drop_na = TRUE) |>
  mutate(column_name_match = str_detect(value, '^.+ \\*( GEOCODE \\*)? PI:.+( \\* METHOD/DEVICE:.+)?( \\* COMMENT:.+)?$'))
  # move from the end with comments and strip out the values in reverse order

```

# Restructuring plan

**Code as Documentation**



3) Collapse the header list into a SoilDRaH data tuple `doi-is_type-with_entry-source` and `doi-column_name-column_number-is_type-with_entry-source` 
4) Collapse the primary data table into a SoilDRaH data tuple `doi-column_name-column_number-row_number-is_type-with_entry-source`
5) Append the three tables `study`, `core`, `layer` across the doi's as a level 2 data product

# 

```{r}

```


# OLD


The annotations table draws heavily from the parameters described in the orginal metadata and also from the point of contact for C-PEAT collection of datasets (Dr. Julie Loisel). 

The C-PEAT database on pangaea is presented in a complex nested format with the primary study level and metadata information presented as a graphical database structure while the layer level information or the data table are presented in regular relational database format. We have worked with merging this complex data structure and collapsing it into one single table format. For example, when you call the function $str$ to find out the structure of the object, the data is presented in a nested list format for the original CPEAT database structure. Our goal is to shoestring this data into one long table along with the annotations. 

# Data processing

```{r setup, echo=TRUE, warning=FALSE, message=FALSE}

library(readr) # read in the csv tables
library(tibble) # use tibbles instead of a data frame
library(plyr) # transform a list into a data frame for the bind
library(dplyr) # work with data tables filters/joins/reframes
library(tidyr) # work with data table pivots
library(ggplot2) # make plots
library(stringr) # extract text from descriptions
library(knitr) # make prettier tables
library(kableExtra) # make tables scrollable
#library(pangaear) # read in data from p

#locate the read script
readCPEAT_file <- 'readCPEAT.R'

#locate the data directory we will download to
dataDir <- 'CPEAT'
dataAnnotations <- 'annotations_CPEAT.csv'
```

```{r maintainer_filenames}
#| echo: FALSE
#| eval: TRUE

#These maintainer_X chunks are intended for folks maintaining this document, 
#  updating the workflow to modify annotations or include more cores, or
#  running this locally. All maintainer chunks except for this one are set to
#. eval and echo FALSE

#Change this file to run locally for you
dataDir  <- '../../temp/CPEAT'

readCPEAT_file <- file.path('../../R', readCPEAT_file)
dataAnnotations <- file.path('../../data', dataAnnotations)
```

```{r sourceReadFunction}
source(readCPEAT_file, local = knitr::knit_global())
```

```{r longCPEAT_read}
#| cache: TRUE
#| warning: FALSE
#| message: FALSE

CPEAT.ls <- readCPEAT(dataDir, 
                  annotationFilename = dataAnnotations, 
                  #when first running, switch this to true 
                  verbose = FALSE, 
                  fileCount = Inf,
                  format = 'long') # 2,640,892 obs for all datasets

#read the same example we take below
singleCPEAT <- readCPEAT(dataDir, 
                  annotationFilename = dataAnnotations, 
                  #when first running, switch this to true 
                  verbose = FALSE, 
                  fileCount = 1,
                  format = 'original')
```

## Orginal data structure

Each core in CPEAT has it's own doi-identified data package with it's own entry on the Pangaea repository.
Each data package includes graphical metadata at the front of a tab separated data table.
The graphical metadata is listed between `/*` and `*/`.
Identifiers in this graphical format are followed by `:` with aspects of this item shown inline with a ` * `-separation followed by another `:`-denoted key-vale, and child items are preceded by a tab.

See the example below.

```{r showDataStr}
#| echo: false
#| code-fold: true
#| code-summary: 'Example Pangaea File'
#| comment: ''
examplefile <- readLines(list.files(file.path(dataDir), 
                                    full.names = TRUE)[10])
cat(paste0(examplefile, collapse = '\n'))
```

This can be represented as a graph (without the value entries) as follows:

```{mermaid CPEAT_diagram}
flowchart LR
  A[Citation]
  B[Related to]
  C[Further details]
  D["Project(s)"]
  E[Coverage] --> EA[LATITUDE] & EB[LONGITUDE] & EC["MINIMUM DEPTH, sediment/rock"] & I["MAXIMUM DEPTH, sediment/rock"]
  J["Events(s)"] --> K[" "] & L[LATITUDE] & M[LONGITUDE] & N[Recovery] & O[Location] & P["METHOD/DEVISE"] & Q[COMMENT]
  Q --> QA["Coring year"] & S["core length"] & T["peatland type"] & U[basal age depth] & V[basal age] & W[basal age] & X[n dates] & Y[core microtopography] & Z[carbon rate site] & AA[peat properties site]
```


# Reading the data

**TODO This entire section needs to be rewritten for the new algorithm**

This read function is a bit longer and more complicated then others in this collection.
Here we choose to walk through key parts of the code base and explain our approach.
Broadly we use the data DOIs in the annotation file to download the data then load this into R using our own parser.
For the long format, then walked through the list of individual cores using 
`dlply` and generated a single data table in our long format.
Note that the data annotations for this collection are minimal because much of 
the data traditionally encoded as metadata is encoded in the downloaded data.


```{r loadReferenceCode}
#| echo: false
CPEAT_code <- read_lines(readCPEAT_file)
```

First we find the data on the Pangaea repository using `pangaear::pg_search`.
Search results are limited to the top 500 so we need to repeat this to get the total 870.
Note that if the number of cores in this collection change we need to revisit 
this function to ensure we still have the correct cleaning algorithms.

```{r seach_Pangaear}
#| eval: false
#| echo: false
#| class-output: R
#| filename: "ln 34-45: readCPEAT.R"
#| comment: ''
cat(paste0(CPEAT_code[34:45], collapse = '\n'))
```


We then download the data and read it in at the same time using the 
`pangaear::pg_data` function.
According to the documentation we should be able to pass a list of doi's but
this does not appear to work right now -- instead we load each doi separately and then piece them back together as a list
using the `plyr::dlply` function.
We also were unable to stop fresh downloads at each run -- as a result
we recommend that you cache the results when possible instead of rerunning this
read function.
 
```{r load_Pangaear}
#| eval: false
#| echo: false
#| filename: "ln 69-89: readCPEAT.R"
#| comment: ''
#| class-output: R
#| code-block-border-left: true
cat(paste0(CPEAT_code[69:89], collapse = '\n'))
```

If the original format is requested we then load in the annotations file and
return the results here.
This results in a list of lists as structured below.

```{r readSingle}
#| cache: true
#| echo: false
str(singleCPEAT)
```

Otherwise we move onto transforming this list of lists into a single long data table by iterating through using `plyr::ldply` after filtering the annotations for those
that apply to the primary data (denoted by `--` in the `with_entry` column).


```{r transformLong}
#| eval: false
#| echo: false
#| filename: "ln 108-117: readCPEAT.R"
#| comment: ''
#| class-output: R
#| code-block-border-left: true
cat(paste0(CPEAT_code[108:117], collapse = '\n'))
```

### Lists to table

To each of `pg_read` lists items we apply the following:

First we pull the headers used in the associated data and apply an ordering to them using a `V`-index pattern.
This was done because not all the headers match those documented in the 
parameter descriptions, likely due to an export truncation at some point.
We are instead relying on ordering to match up the column names with the 
parameter descriptions.

```{r parseHeader}
#| eval: false
#| echo: false
#| filename: "ln 172-197: readCPEAT.R"
#| comment: ''
#| class-output: R
#| code-block-border-left: true
cat(paste0(CPEAT_code[172:197], collapse = '\n'))
```

Then we add the descriptions from the `parameter` and combin that with
`of_variable` tag
from the annotations table.

```{r parseParameter}
#| eval: false
#| echo: false
#| filename: "ln 199-220: readCPEAT.R"
#| comment: ''
#| code-output: R
# code-block-border-left: true
cat(paste0(CPEAT_code[199:220], collapse = '\n'))
```

Next the `data` element contains the layer level observations.
The column names have been truncated in some cases, so first we replace them using a column number with a `V` prepend and rely on the ordering to match these up with the column descriptions and other information constructed above.
There is no clear row/observation identifier (some tables contain multiple observations for the same layer), so we add a row number with a `R` prepend here to create a unique ID.
Then the table is joined to prior information and pivoted to create a long format.

```{r parseDataTable}
#| eval: false
#| echo: false
#| filename: "ln 223-254: readCPEAT.R"
#| comment: ''
#| class-output: R
#| code-block-border-left: true
cat(paste0(CPEAT_code[223:254], collapse = '\n'))
```

Finally we construct the secondary information about the study stored in elements which are not `parameters` nor `data` that we have not worked with yet


```{r parseOther}
#| eval: false
#| echo: false
#| filename: "ln 347-385: readCPEAT.R"
#| comment: ''
#| class-output: R
#| code-block-border-left: true
cat(paste0(CPEAT_code[347:385], collapse = '\n'))
```


The extrated data frames are then bound the rows together.
And then returned to the `plyr::ldply` function to be appended to other core data sets.

```{r bindTables}
#| eval: false
#| echo: false
#| filename: "ln 387-388: readCPEAT.R"
#| comment: ''
#| class-output: R
#| code-block-border-left: true
cat(paste0(CPEAT_code[387:388], collapse = '\n'))
```

### Long format output

We then clean up the output and return the resulting long format.

```{r finalCleaningLong}
#| eval: FALSE
#| echo: false
#| filename: "ln 393-413: readCPEAT.R"
#| comment: ''
#| class-output: R
#| code-block-border-left: true
cat(paste0(CPEAT_code[393:413], collapse = '\n'))
```

# CPEAT long format

This long format has the following structure.

```{r exampleLong}


CPEAT.ls$long %>%  # 2,640,892 entries, 7 total columns
  slice_head(n=100) %>%
  knitr::kable() %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "300px")
```

## `table_name`

Often `table_name` will reflect the file name in a set of inter-related csv files, in this case we use it to refer to what `list` the data was extracted from in the `pangaear::pg_data` return.

```{r listTables}
CPEAT.ls$long |>
  reframe(count = n(),
          .by = table_name) |>
  arrange(desc(count))
```

## `is_type`

The `is_type` entry usually falls into one of the following categories: {`value`, `definition`, `unit`, `method`, `control_vocabulary`}. 
This entry describes the type of information associated with `of_variable` found in `with_entry`. 
Note that we use `value` as a general term for simplicity, regardless of whether the data is numerical or text. 

```{r listTypes}

CPEAT.ls$long |>
  reframe(count = n(),
          .by = is_type) |>
  arrange(desc(count))
```

## `of_variable`

The of\_variable column describes the variables linked to the column\_id, such as bulk density, organic carbon, vegetation, other covariates. It's important to maintain the original intent of the data providers when creating these annotations. While the of\_variable entries are highly dependent on the dataset, is\_type should have a more restricted range of entries.

```{r listVariables}
CPEAT.ls$long |>
  reframe(count = n(),
          .by = of_variable) |>
  arrange(desc(count))
```


```{r maintainer_annotationComplete}
#| eval: FALSE
#| echo: FALSE

#Check to see if there are NA variables implying that the annotations is incomplete

badReads <- CPEAT.ls$long |>
  filter(is.na(of_variable))
```

# Subseting for soil carbon

There is a lot of data here but not all of it is of interest for all purposes.
In this example we are interested in layer-level soil carbon related measurements and the geolocation.
Note that we are NOT checking any of the quality control flags here which you would want to do if using this for research purposes.

```{r}
CPEAT_soc <- CPEAT.ls$long %>% 
  #identify the variables that we are interested in
  mutate(new_table = case_when(
    of_variable %in% c(#variables that identify the source data
             'doi', "data_doi", "download_url", "data_local_filepath",
             "project", "core_name", "citation",
             "related_objects", "related_citation", 
             "core_further_details", "current_status",
             "curation_status", "change_log",
             "keyword",
             "core_comment", "comment", "data_license",
             "principle_investigator") ~ 'data_source',
    of_variable %in% c(#geolocation of the core
             "latitude", "longitude", "region", "elevation", "elevation_end",
             "elevation_start", "bounded_coverage") ~ 'geolocation',
    of_variable %in% c(# related to the sampling of the core
            "sampling_device", 
            "core_minimum_depth", "core_maximum_depth",
            "core_penetration", "core_recovery",
            "peat_description") ~ 'sample_collection',
    of_variable %in% c(#depth related variables 
             "layer_mid", "layer_thickness", "depth_thickness",
             "depth_mid", "depth_mid_loi", "depth_top", "depth_bottom") ~ 'soil_depth',
    of_variable %in% c(#related to the carbon content
            "carbon", "total_carbon",
            "organic_carbon_density", "total_organic_carbon",
            "organic_matter", "organic_matter_density", "loss_on_ignition",
            "total_inorganic_carbon", "inorganic_carbon") ~ 'soil_carbon',
    of_variable %in% c(#related to bulk density
            "bulk_density") ~ 'bulk_density')) |>
  filter(!is.na(new_table)) |>
  mutate(of_variable = case_when(str_detect(doi, '10.1594_PANGAEA.927849') &
                                   str_detect(of_variable, 'depth_mid') &
                                   str_detect(column_id, '^V3$')
                                   ~ 'depth_mid_loi',
                                 TRUE ~ of_variable))

CPEAT_soc %>%
  slice_head(n=100) %>%
  knitr::kable() %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "300px")
```

## Make small tables

Often this data is presented in three tables: a site level table with lat/lon and year, a layer level table with the depth and physiochemical information, and a meta table with information on the units and descriptions.
We given an example below of how this might be organized.

### Study information

```{r}
CPEAT_study <- CPEAT_soc %>% 
  #identify the variables that we are interested in
  filter(new_table == 'data_source',
         !is.na(with_entry),
         #description only holds Comment (Comment), removing
         !str_detect(is_type, pattern = 'description')) |>
  select(doi, of_variable, is_type, with_entry) |>
  pivot_wider(names_from = of_variable, values_from = with_entry,
              values_fn = ~ paste(.x, collapse = '; ')) |>
  select(-is_type) |>
  select(core_name, doi, everything()) |>
  dplyr::mutate(across(all_of(c( "core_name", "principle_investigator",
                                 "project", "data_license", 'keyword', 
                                 'curation_status')), as.factor)) |>
  arrange(core_name)
```

```{r}
CPEAT_study |>
  knitr::kable() |>
  kable_paper() |>
  scroll_box(width = "100%", height = "300px")
```



```{r}
CPEAT_study |>
  select(!c(doi, core_name)) |>
  select(where(is.factor)) |>
  dplyr::mutate(across(everything(), as.character)) |>
  pivot_longer(cols = everything(), 
               names_to = 'of_variable', values_to = 'with_entry') |>
  reframe(count = n(),
          .by = c(of_variable, with_entry)) |>
  arrange(-count, of_variable, with_entry) |>
  knitr::kable() |>
  kable_paper() |>
  scroll_box(width = "100%", height = "300px")
```


### Profile information

```{r}
CPEAT_profile <- CPEAT_soc |>
  filter(new_table %in% c('geolocation', 'sample_collection') |
           of_variable %in% c('doi', 'core_name')) |>
  select(doi, of_variable, with_entry) |>
  unique() |> #remove duplicate lat/lon generated by metadata
  pivot_wider(names_from = of_variable, values_from= with_entry) |>
  pivot_longer(cols = !c(doi, core_name), 
               names_to = 'of_variable', values_to = 'with_entry') |>
  mutate(value = str_remove(with_entry, '\\s\\D+$'),
         unit = str_extract(with_entry, '(?<=\\d\\s)\\D+$')) |>
  filter(!is.na(value)) |>
  mutate(value=as.numeric(value)) |>
  select(core_name, doi, of_variable, value, unit)

```

```{r}
ggplot(CPEAT_profile) +
  geom_histogram(aes(x=value)) +
  facet_wrap(~of_variable, scales = 'free')
```

```{r}

world_map <- map_data("world")

plot.df <- CPEAT_profile |>
  filter(of_variable %in% c('latitude', 'longitude')) |>
  select(core_name, doi, of_variable, value) |>
  pivot_wider(names_from = of_variable, values_from = value)



ggplot(data=world_map) + 
   geom_polygon(colour="lightgrey", fill="white", aes(x=long, y=lat, group=group)) +
  geom_point(data = plot.df, 
             aes(x=longitude, y = latitude),
             color = 'blue', alpha = .1) +
  theme(axis.title = element_blank(),
        axis.text = element_blank())
```

### Layer information

```{r}


#we want to extract the units from the `descriptions` that are between '[]'
CPEAT_units <- CPEAT_soc |>
  filter(new_table %in% c("soil_depth", "bulk_density", "soil_carbon")|
           of_variable %in% c('doi', 'core_name'),
         is_type == 'description') |>
  select(of_variable, description = with_entry) |>
  unique() |>
  mutate(unit = str_extract(description, '(?<=\\[).*(?=\\])')) |>
  mutate(description = if_else(of_variable == 'depth_mid_loi',
                               'DEPTH, sediment/rock for loi [m] (Depth sed loi)',
                               description))
  

CPEAT_layer <- CPEAT_soc |>
  filter(new_table %in% c("soil_depth", "bulk_density", "soil_carbon")) |>
  filter(!is.na(observation_id)) |>
  #if there is anything in the doi that is not depth, keep that doi
  filter(any(!str_detect(of_variable, 'depth' )), .by = doi) |>
  select(doi, observation_id, of_variable, with_entry) |>
  mutate(value = as.numeric(with_entry)) |> #two entries '?0' being replaced with NA
  left_join(CPEAT_units, by = join_by(of_variable)) |>
  left_join(CPEAT_study |>
              select(doi, core_name) |>
              unique(), by = join_by(doi)) 
```

#### Sample of table

```{r}
CPEAT_layer |>
  select(core_name, doi, observation_id, description, value) |>
  pivot_wider(names_from = description, values_from = value) |>
  slice_head(n = 100) |>
  knitr::kable() |>
  kable_paper() |>
  scroll_box(width = "100%", height = "300px")
  

```

#### Measured histograms

```{r}

ggplot(CPEAT_layer) +
  geom_histogram(aes(x=value)) +
  facet_wrap(~description, scales='free', ncol = 2)

```

#### Bulk density vs carbon %

```{r}

plot.df <- CPEAT_layer |>
         filter(of_variable %in% c('bulk_density', 'organic_matter')) |>
         select(doi, observation_id, of_variable, value) |>
         pivot_wider(names_from = of_variable, values_from = value)

ggplot(plot.df) +
  geom_point(aes(x=bulk_density, y = organic_matter))

```

#### Geolocations of OC

```{r}
world_map <- map_data("world")

plot.df <- CPEAT_profile |>
  filter(of_variable %in% c('latitude', 'longitude')) |>
  select(core_name, doi, of_variable, value) |>
  filter(doi %in% CPEAT_layer$doi) |>
  pivot_wider(names_from = of_variable, values_from = value)



ggplot(data=world_map) + 
   geom_polygon(colour="lightgrey", fill="white", aes(x=long, y=lat, group=group)) +
  geom_point(data = plot.df, 
             aes(x=longitude, y = latitude),
             color = 'blue', alpha = .3) +
  theme(axis.title = element_blank(),
        axis.text = element_blank())
```

# Appendix

## readCPEAT.R

```{r}
#| file: '../R/readCPEAT.R'
#| code-line-numbers: TRUE
#| echo: TRUE
#| eval: FALSE
```

